### 2022-12-13 Rmd/runCond.Rmd
1st draft

---
title: "**Density Dependence**"
subtitle: "Summary"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/cond/",
               fig.path  ="../tex/figs/cond/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(ggplotFL)
library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

# Mackerel

```{r, mac.data}
load("/home/laurie/Dropbox/pelagics/data/om/mac.RData")
ts =as.data.frame(read_excel("/home/laurie/Desktop/projects/ddmse/data/inputs/dat.xlsx",3))

wt =transform(as.data.frame(stock.wt(mac),drop=TRUE),cohort=year-as.numeric(age))
wt =merge(wt,ts,by="year")
wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
wt$age=factor(wt$age,levels=sort(unique(wt$age)))

mat=transform(as.data.frame(mat(mac),drop=TRUE),cohort=year-as.numeric(age))
mat=merge(mat,ts)
mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
mat$age=factor(mat$age,levels=sort(unique(mat$age)))

m  =transform(as.data.frame(m(mac),drop=TRUE),cohort=year-as.numeric(age))  
m  =merge(m,ts)
m  =ddply(m,.(age), transform, rsdl =data/mean(data))
m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
m$age=factor(m$age,levels=sort(unique(m$age))) 
```

```{r, mac.ts, fig.width=8,fig.height=4}
ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
  geom_line(aes(year,data,col=quant))+
  theme(legend.position="bottom")+
  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
  xlab("Year")+ylab("")
```

**Figure `r iFig=iFig+1; iFig`** Time series

```{r, mac.wt, fig.width=8,fig.height=4}
ggplot(wt)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r, mac.m, fig.width=8,fig.height=4}
ggplot(mat)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

```{r, mac.mat, fig.width=8,fig.height=4}
ggplot(m)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** M-at-age

```{r, mac.wt2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))

a=ggplot(wt)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")

b=ggplot(subset(wt,rsdl<3&rsdl>0))+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.)+
  xlab("Biomass")+ylab("Residual")

ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
          theme_bw(16)+
          theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary

```{r, mac.wt.tb, fig.height=8,fig.width=8}
ggplot(subset(wt,age!=0),aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass

```{r, mac.mat2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))

a=ggplot(mat)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")

b=ggplot(mat)+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.)+
  xlab("Biomass")+ylab("Residual")

ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
          theme_bw(16)+
          theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary

```{r, mac.mat.tb, fig.height=5,fig.width=8}
ggplot(subset(mat,age%in%(1:5)),aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Maturity-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v Biomass

```{r}
gmr=gammV(data~s(tb, bs="tp")+age, data=wt)

summary(gmr)

plot(gmr)
```

# Blue whiting

```{r, whb.data}
load("/home/laurie/Dropbox/pelagics/data/om/whb.RData")
ts=model.frame(FLQuants(whb,tb=biomass,ssb=ssb),drop=T)
ts$tb=ts$tb/1e6

wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
wt =merge(wt,ts,by="year")
wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
wt$age=factor(wt$age,levels=sort(unique(wt$age)))

mat=transform(as.data.frame(mat(whb),drop=TRUE),cohort=year-as.numeric(age))
mat=merge(mat,ts)
mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
mat$age=factor(mat$age,levels=sort(unique(mat$age)))

m  =transform(as.data.frame(m(whb),drop=TRUE),cohort=year-as.numeric(age))  
m  =merge(m,ts)
m  =ddply(m,.(age), transform, rsdl =data/mean(data))
m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
m$age=factor(m$age,levels=sort(unique(m$age)))
```

```{r, whb.ts, fig.width=8,fig.height=4}
ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
  geom_line(aes(year,data,col=quant))+
  theme(legend.position="bottom")+
  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
  xlab("Year")+ylab("")
```

**Figure `r iFig=iFig+1; iFig`** Time series

```{r, whb.wt, fig.width=8,fig.height=4}
ggplot(wt)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r, whb.mat, fig.width=8,fig.height=4}
ggplot(mat)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

```{r, whb.m, fig.width=8,fig.height=4}
ggplot(m)+  
  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
  geom_line(aes(year,data,group=age,col=age))+
  geom_point(aes(year,data,col=age))+
  xlab("Year")+ylab("Mass-at-age")+
  theme(legend.position="bottom")+
  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
```

**Figure `r iFig=iFig+1; iFig`** M-at-age

```{r, whb.wt2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))

a=ggplot(wt)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")

b=ggplot(subset(wt,rsdl<3&rsdl>0))+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.,ncol=3)+
  xlab("Biomass")+ylab("Residual")

ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
          theme_bw(16)+
          theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary

```{r, whb.wt.tb, fig.height=8,fig.width=8}
ggplot(wt,aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass

```{r, whb.mat2, fig.width=8,fig.height=5}
#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))

a=ggplot(mat)+
  geom_boxplot(aes(age,data))+
  xlab("Mass-at-age")+ylab("Age")

b=ggplot(mat)+
  geom_point(aes(tb,rsdl))+
  geom_smooth(aes(tb,rsdl),method="lm")+
  facet_wrap(age~.,ncol=3)+
  xlab("Biomass")+ylab("Residual")

ggarrange(a,b ,label.y=1,  
          widths = c(4,6), heights = c(1,1,1),
          nrow   = 1,          ncol    = 2)+
          theme_bw(16)+
          theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank(),
          axis.title.y = element_blank())
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary

```{r, whb.mat.tb, fig.height=5,fig.width=8}
ggplot(subset(mat,age%in%1:5),aes(tb,data))+
  geom_point()+
  geom_smooth(se=F,method="lm")+
  facet_wrap(~age,scale="free",ncol=3)+
  xlab("Biomass")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass

```{r}
gmr=gammV(data~s(tb)+age, data=wt)

summary(gmr)

plot(gmr)

```

# Funding

# References

### 2022-12-13 Rmd/runCond.Rmd
1st Draft

@@
 ```{r, mac.data}
-load("/home/laurie/Dropbox/pelagics/data/om/mac.RData")
-ts =as.data.frame(read_excel("/home/laurie/Desktop/projects/ddmse/data/inputs/dat.xlsx",3))
+load(file.path(dropboxdir,"data/om/mac.RData"))
+ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
 
@@
 ```{r, whb.data}
-load("/home/laurie/Dropbox/pelagics/data/om/whb.RData")
+load(file.path(dropboxdir,"data/om/whb.RData"))
 ts=model.frame(FLQuants(whb,tb=biomass,ssb=ssb),drop=T)

### 2022-12-15 Rmd/runCond.Rmd
added dropbox dir

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: pdf_document
+output: html_document
 mathjax: TRUE
@@
 theme_set(theme_bw(16))
+
+source("../R/get_dropbox.R")
+
 ```
@@
 ```{r, dbox}
-dropboxdir<-try(file.path(get_dropbox(), "dd"))
+dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
 
@@
 load(file.path(dropboxdir,"data/om/mac.RData"))
-ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
+ts =as.data.frame(read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),3))
 

### 2022-12-15 Rmd/runDD.Rmd
1st push

---
title: "**Density Dependence**"
subtitle: "Simulations"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
mathjax: TRUE

tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
# bibliography: refs.bib
---

This document provides an example of how to model density dependence in mass-at-age and M-at-age.

The life history parameters of mackerel are used to construct an `FLBRP` object that represents the equilibrium dynamics, this is then used to construct an `FLStock` object that models the time series dynamics.

The Lorenzen relationship between mass-at-age and M is used to link M and weight-at-age. For example, lorenzen M = a*300^-0.288, so if you want M=0.2 at 300g then a= 0.2 /(300^-0.288)

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/dd/",
               fig.path  ="../tex/figs/dd/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

# Mackerel

```{r, mac.data}
load(file.path(dropboxdir,"data/om/mac.RData"))
ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
```

```{r eq}
par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
eq =lhEql(par,m="lorenzen")

plot(eq,refpt="msy")
```

**Figure `r iFig=iFig+1; iFig`** Equilibrium dynamics

```{r}
om=as(eq,"FLStock")
om=fwd(om,fbar=fbar(om)[,-1],sr=eq)

plot(om)
```

**Figure `r iFig=iFig+1; iFig`** Time series dynamics

```{r}
plot(qmax((biomass(eq)%/%biomass(eq)[,c(14)])^0.3,0.5))+
  scale_y_continuous(labels = scales::percent)+
  xlab("Biomass")+ylab("% Change in mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Density dependence multiplier

```{r}
ddWt<-function(wt,biomass, bmsy)
  wt%*%qmax((biomass/bmsy)^0.3,0.5)

wt=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"])

ggplot(as.data.frame(wt))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("NMass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in mass-at-age

```{r}
m=par["m1"]%*%(wt%^%par["m2"])

ggplot(as.data.frame(m))+           
    geom_line(aes(age,data,col=factor(year)))+
    theme(legend.position="none")+
    scale_x_continuous(limits=c(0,15))+
    xlab("Age")+ylab("M-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in M-at-age

```{r}
x=propagate(eq,length(dimnames(eq)$year))
fbar(x)=fbar(x)[,1]
fbar(x)=c(fbar(eq))

for (i in seq(20)){
stock.wt(x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
x=brp(x)}

ggplot()+
  geom_line(aes(biomass,catch),
            data=model.frame(FLQuants(x,biomass=function(x) biomass(x) ,catch=function(x) catch(x)),drop=T))+
  geom_line(aes(biomass,catch),
            data=model.frame(FLQuants(eq,biomass=function(x) biomass(x) ,catch=function(x) catch(x)),drop=T),col="red")
```

**Figure `r iFig=iFig+1; iFig`** Comparison of equilibrium curves with (black) and without (red) density dependence.

```{r}
ggplot()+
  geom_line(aes(ssb,catch),
            data=model.frame(FLQuants(x,ssb=function(x) ssb(x) ,catch=function(x) catch(x)),drop=T))+
  geom_line(aes(ssb,catch),
            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x) ,catch=function(x) catch(x)),drop=T),col="red")
```

**Figure `r iFig=iFig+1; iFig`** Comparison of equilibrium curves with (black) and without (red) density dependence.

# Funding

# References

### 2022-12-15 Rmd/runCond.Rmd
1st

@@
 library(knitr)
-
 opts_chunk$set(cache     =TRUE, 
@@
                dev       ="png")
-
 iFig=0
@@
 library(ggpubr)
-
 library(rjson) # install.packages("rjson"); used for dropbox function
 library(RJSONIO)
-
 library(plyr)
@@
 library(reshape)
-
 library(readxl)
@@
 theme_set(theme_bw(16))
+```
 
+```{r, eval=FALSE}
 source("../R/get_dropbox.R")
-
 ```
@@
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
-
 if ("try-error"%in%is(dropboxdir))
@@
 ts =as.data.frame(read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),3))
-
 wt =transform(as.data.frame(stock.wt(mac),drop=TRUE),cohort=year-as.numeric(age))
@@
 wt$age=factor(wt$age,levels=sort(unique(wt$age)))
-
 mat=transform(as.data.frame(mat(mac),drop=TRUE),cohort=year-as.numeric(age))
@@
 mat$age=factor(mat$age,levels=sort(unique(mat$age)))
-
 m  =transform(as.data.frame(m(mac),drop=TRUE),cohort=year-as.numeric(age))  
@@
 #ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
-
 a=ggplot(wt)+
@@
   xlab("Mass-at-age")+ylab("Age")
-
 b=ggplot(subset(wt,rsdl<3&rsdl>0))+
@@
   xlab("Biomass")+ylab("Residual")
-
 ggarrange(a,b ,label.y=1,  
@@
 #ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
-
 a=ggplot(mat)+
@@
   xlab("Mass-at-age")+ylab("Age")
-
 b=ggplot(mat)+
@@
 gmr=gammV(data~s(tb, bs="tp")+age, data=wt)
-
 summary(gmr)
-
 plot(gmr)
@@
 ts$tb=ts$tb/1e6
-
 wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
@@
 wt$age=factor(wt$age,levels=sort(unique(wt$age)))
-
 mat=transform(as.data.frame(mat(whb),drop=TRUE),cohort=year-as.numeric(age))
@@
 mat$age=factor(mat$age,levels=sort(unique(mat$age)))
-
 m  =transform(as.data.frame(m(whb),drop=TRUE),cohort=year-as.numeric(age))  
@@
 gmr=gammV(data~s(tb)+age, data=wt)
-
 summary(gmr)
-
 plot(gmr)
-
 ```

### 2022-12-16 Rmd/runDD.Rmd
tidy up

@@
 
-This document provides an example of how to model density dependence in mass-at-age and M-at-age.
+This document provides an example of how to model density dependence in mass and M-at-age.
 
@@
 
-The Lorenzen relationship between mass-at-age and M is used to link M and weight-at-age. For example, lorenzen M = a*300^-0.288, so if you want M=0.2 at 300g then a= 0.2 /(300^-0.288)
+The Lorenzen relationship between mass-at-age and M can model M as a function of weight-at-age. For example, lorenzen M = a*300^-0.288, so if you want M=0.2 at 300g then a= 0.2 /(300^-0.288)
 
@@
 
-# Mackerel
+# Sumulations
 

### 2022-12-16 Rmd/runDD.Rmd
plot

@@
   scale_y_continuous(labels = scales::percent)+
+  geom_vline(aes(xintercept=14),col="red")+
   xlab("Biomass")+ylab("% Change in mass-at-age")
@@
 
-**Figure `r iFig=iFig+1; iFig`** Density dependence multiplier
+**Figure `r iFig=iFig+1; iFig`** Density dependence multiplier, red is $B_{MSY}$
 
@@
     scale_x_continuous(limits=c(0,15))+
-    xlab("Age")+ylab("NMass-at-age")
+    xlab("Age")+ylab("Mass-at-age")
 ```

### 2022-12-16 Rmd/runCond.Rmd
started documenting the RMD file

--- 2022-12-15 Rmd/runCond.Rmd
@@
 
+**Biomass trends**
+
+[Just to check: what is meant by ICES biomass? ]
+
 ```{r, mac.ts, fig.width=8,fig.height=4}
@@
 
+**Weight at age**
+
+Yearclasses connected via dotted lines.
+
 ```{r, mac.wt, fig.width=8,fig.height=4}
@@
   theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age)))) +
+  guides(color = guide_legend(nrow = 1))
+
 ```
@@
 **Figure `r iFig=iFig+1; iFig`** Mass-at-age
+
+**Maturity at age**
+
+Most of the variability is in the maturity at age 2 and, to a lesser extent, age 3.
 
@@
   geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
+  xlab("Year")+ylab("Maturity-at-age")+
   theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age)))) +
+  guides(color = guide_legend(nrow = 1))
+
 ```
@@
 **Figure `r iFig=iFig+1; iFig`** Maturity-at-age
+
+**Natural mortality at age**
+
+Constant by age and by year
 
@@
   geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
+  xlab("Year")+ylab("M-at-age")+
   theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age)))) +
+  guides(color = guide_legend(nrow = 1))
+
 ```
@@
 **Figure `r iFig=iFig+1; iFig`** M-at-age
+
+**Weight at age vs. biomass**
+
+Boxplot of weight at age and [xxxx]
 

### 2022-12-28 Rmd/runCond.Rmd
with simulations

--- 2022-12-15 Rmd/runCond.Rmd
@@
 source("../R/get_dropbox.R")
+```
+
+```{r}
+matFn<-function(wt,x) 
+   1/(1+exp(-x[1]*(wt-x[2])))
 ```
@@
 
+$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
+
+```{r}
+dat=model.frame(FLQuants(mac,"wt"=stock.wt,"mat"=function(x) mat(x)))
+dat$age=factor(dat$age)
+
+fn<-function(x,dat) 
+   sum((dat$mat-matFn(x,dat$wt))^2)
+
+par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
+
+par$par
+
+ggplot(dat,aes(wt,mat))+
+  geom_point(aes(wt,mat,col=age))+
+  geom_smooth(se=F,span=0.2)+
+  xlab("Age")+ylab("Matutrity")+
+  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age
+
 ```{r}
@@
 
+$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
+
+```{r}
+dat=model.frame(FLQuants(whb,"wt"=stock.wt,"mat"=function(x) mat(x)))
+dat$age=factor(dat$age)
+
+fn<-function(x,dat) 
+   sum((dat$mat-matFn(dat$wt,x))^2)
+
+par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
+
+ggplot(dat,aes(wt,mat))+
+  geom_point(aes(wt,mat,col=age))+
+  geom_smooth(se=F,span=0.2)+
+  xlab("Age")+ylab("Matutrity")+
+  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
+```
+
 ```{r}

### 2022-12-28 Rmd/runDD.Rmd
with simulations

--- 2022-12-16 Rmd/runDD.Rmd
@@
 title: "**Density Dependence**"
-subtitle: "Simulations"
+subtitle: "Example of modellimng density dependence in mass, maturity and M-at-age"
 author: "L Kell"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: pdf_document
+output: html_document
 mathjax: TRUE
@@
 license: Creative Commons Attribution-ShareAlike 4.0 International Public License
-# bibliography: refs.bib
+bibliography: refs.bib
 ---
 
-This document provides an example of how to model density dependence in mass and M-at-age.
+Mackerel life history parameters are used to construct an `FLBRP` object to represent the equilibrium dynamics and then an `FLStock` object models the time series dynamics.
 
-The life history parameters of mackerel are used to construct an `FLBRP` object that represents the equilibrium dynamics, this is then used to construct an `FLStock` object that models the time series dynamics.
+## Natural mortality
 
-The Lorenzen relationship between mass-at-age and M can model M as a function of weight-at-age. For example, lorenzen M = a*300^-0.288, so if you want M=0.2 at 300g then a= 0.2 /(300^-0.288)
+@lorenzen2000allometry noted that theoretical and empirical studies point to the existence of an allometric relationship between natural mortality $M_a$ and body mass ($W_a$) at-age in fish of the form:
+
+  $M_a \propto {W_a}^{d}$
+
+i.e.
+
+  $M_W=M_uW^d$
+
+where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37, e.g.
+if $M_t=0.2$ at a weight when $50\%$ of individuals reach maturity $W_{50}$ then
+
+  $M_u = 0.2/({W_{50}}^{d})$
+
+Since weight is approximately proportional to the third power of length  values of $d$ of approximately –1/3 imply that $M$ is inversely proportional to body length.
+@10.1093/icesjms/fst226 recognised that for ($L$) if the von Bertalanffy and length–weight parameters are known, then the mortality schedule can be defined by
+
+$L_t=L_{\infty}{(1-e^{(1-k(t-t_0))})}^{b(-0.305)}$
+
+where $L_{\infty}$ is the asmypotopic length,  $k$ growth rate, and $t_0$ is the age at which length is zero of the von Bertalanffy growth equation,  and the allometric relationship $W = a L^b$.
+
+This gives
+
+$M_t = M_{\infty}{(W_t/W_{\infty})}^{-0.3}$
+
+where $M_t$ is the instantaneous $M$ rate per year at age $t$ and $M_{\infty}$ is the asymptotic $M$ rate at the asymptotic weight $W_{\infty}$, derived from life history characteristics @hoenig1992empirical
+
+## Multispecies modelling of predation mortality (M2)
+
+@pope2006modelling (doi:10.1016/j.icesjms.2006.04.015) proposed a way of modelling predation mortality (M2) based on size spectra, e.g.
+
+$N(L) = \Omega L^{\kappa}$
+
+where $N$ is numbers at length $L$ and $\Omega$ is a prefactor and $\kappa$ the scaling by length
+
+This was further developed by @POPE2021105952 (https://doi.org/10.1016/j.fishres.2021.105952) by assuming that predation by all fish in the spectrum is mediated through a feeding relationship with mean $\mu$ and standard deviation $\sigma$, both common to all predators and that UM2 is a constant that relates predation pressure to mortality, and where $\lambda$ is the scaling of the feeding rate of predators by length then the predation mortality on a prey species of length $L^{\prime}$ can be written as
+
+$M2(L^\prime) = \Omega \times UM2(L^\prime)^{(\kappa + \lambda)} e^{(\mu((\kappa + \lambda)/3))} \times e^{({\sigma}^2((\kappa+\lambda)^2)/18)}$
+
+<!-- Another insight derived from size spectra studies comes from the work of Andersen et al.
+(2009).
+They derive a formulation for the predation mortality to be expected from a size spectrum with a given slope and intercept in terms of mass.
+However, the CSM is a rather different model to theirs since it only considers the fish part of the ecosystem and is framed in terms of length rather than mass.
+Moreover, it treats predation mortality as a top-down process rather than in terms of mass balance.
+Predation mortality is defined on a simple UM2*predation pressure basis.
+Here, UM2 is defined as a constant M2 per unit of predation pressure.
+The predation pressure is summed over all predator lengths with a weighting for the feeding relationship between specific prey and predator lengths.
+For a particular length L of predator, predation pressure is taken to be their numbers times their length to a power $\gamma$ (if $\gamma=3$), this would be equivalent to biomass except that the condition factor linking $L^3$ to mass is included into the UM2 constant to simplify the equations).
+Hence, assuming a numbers size spectrum described by -->
 
@@
 
+```{r}
+ddWt<-function(wt,biomass, bmsy)
+  wt%*%qmax((biomass/bmsy)^0.3,0.5)
+
+matFn<-function(wt,x) 
+   1/(1+exp(-x[1]*(wt-x[2])))
+
+ddFn<-function(year,x,par){
+  lsr=landings.wt(x)[,year]/stock.wt(x)[,year]
+  dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
+
+  stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bmsy"])
+  landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
+  discards.wt(x)[,year]=stock.wt(x)[,year]*dsr
+
+  mat(x)[,year]=matFn(stock.wt(x)[,year],par[c("k","w50")])
+  m(  x)[,year]=par["m1"]%*%(stock.wt(x)[,year]%^%par["m2"])
+
+  fwd(x,f=f[,year],sr=eq)}
+```
+
 ```{r, dbox}
@@
 ```{r}
-ddWt<-function(wt,biomass, bmsy)
-  wt%*%qmax((biomass/bmsy)^0.3,0.5)
-
 wt=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"])
@@
 **Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in mass-at-age
+
+```{r}
+parMat=FLPar(c(k=0.1, w5=111.0)) 
+mat=1/(1+exp(-parMat[1]*(wt-parMat[2])))
+
+ggplot(as.data.frame(mat))+           
+    geom_line(aes(age,data,col=factor(year)))+
+    theme(legend.position="none")+
+    scale_x_continuous(limits=c(0,15))+
+    xlab("Age")+ylab("M-at-age")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in maturity-at-age
 
@@
 ```{r}
+save(eq,file="/home/laurie/Desktop/tmp/t.RData")
+```
+
+```{r}
 x=propagate(eq,length(dimnames(eq)$year))
@@
 
+lsr=landings.wt(eq)/stock.wt(eq)
+dsr=discards.wt(eq)/stock.wt(eq)
+# iterate
 for (i in seq(20)){
-stock.wt(x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
+  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
+  landings.wt(x)=stock.wt(x)*lsr
+  discards.wt(x)=stock.wt(x)*dsr
+  x=brp(x)}
+
+ggplot()+
+  geom_line(aes(ssb,catch),
+            data=model.frame(FLQuants(x, ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T))+
+  geom_line(aes(ssb,catch),
+            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T),col="red")+
+  xlab("SSB")+ylab("Catch")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age: Comparison of equilibrium curves with (black) and without (red) density dependence.
+
+```{r}
+x=propagate(eq,length(dimnames(eq)$year))
+fbar(x)=fbar(x)[,1]
+fbar(x)=c(fbar(eq))
+
+lsr=landings.wt(eq)/stock.wt(eq)
+dsr=discards.wt(eq)/stock.wt(eq)
+# iterate
+for (i in seq(20)){
+  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
+  landings.wt(x)=stock.wt(x)*lsr
+  discards.wt(x)=stock.wt(x)*dsr
+
+    mat(     x)=1/(1+exp(-parMat[1]*(stock.wt(x)-parMat[2])))
+  x=brp(x)}
+
+ggplot()+
+  geom_line(aes(ssb,catch),
+            data=model.frame(FLQuants(x, ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T))+
+  geom_line(aes(ssb,catch),
+            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T),col="red")+
+  xlab("SSB")+ylab("Catch")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass and maturity-at-age: Comparison of equilibrium curves with (black) and without (red) density dependence.
+
+```{r}
+x=propagate(eq,length(dimnames(eq)$year))
+fbar(x)=fbar(x)[,1]
+fbar(x)=c(fbar(eq))
+
+lsr=landings.wt(eq)/stock.wt(eq)
+dsr=discards.wt(eq)/stock.wt(eq)
+# iterate
+for (i in seq(20)){
+  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
+  landings.wt(x)=stock.wt(x)*lsr
+  discards.wt(x)=stock.wt(x)*dsr
+
+  mat(     x)=1/(1+exp(-parMat[1]*(stock.wt(x)-parMat[2])))
   m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
@@
 ggplot()+
-  geom_line(aes(biomass,catch),
-            data=model.frame(FLQuants(x,biomass=function(x) biomass(x) ,catch=function(x) catch(x)),drop=T))+
-  geom_line(aes(biomass,catch),
-            data=model.frame(FLQuants(eq,biomass=function(x) biomass(x) ,catch=function(x) catch(x)),drop=T),col="red")
+  geom_line(aes(ssb,catch),
+            data=model.frame(FLQuants(x, ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T))+
+  geom_line(aes(ssb,catch),
+            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T),col="red")+
+  xlab("SSB")+ylab("Catch")
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Comparison of equilibrium curves with (black) and without (red) density dependence.
+**Figure `r iFig=iFig+1; iFig`** Mass, maturity and M-at-age: Comparison of equilibrium curves with (black) and without (red) density dependence.
 
 ```{r}
-ggplot()+
-  geom_line(aes(ssb,catch),
-            data=model.frame(FLQuants(x,ssb=function(x) ssb(x) ,catch=function(x) catch(x)),drop=T))+
-  geom_line(aes(ssb,catch),
-            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x) ,catch=function(x) catch(x)),drop=T),col="red")
+save(eq,parMat,par,sr,file="/home/laurie/Desktop/tmp/t.RData")
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Comparison of equilibrium curves with (black) and without (red) density dependence.
+```{r}
+load("/home/laurie/Desktop/tmp/t.RData")
+
+par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","ssb"]),b=0.3,k=0.1, w50=111.0))
+
+f=FLQuant(c(rep(seq(0,1,length.out=101),each=51),rep(seq(1,5,length.out=101),each=51)[-1]),
+          dimnames=list(year=dimnames(om)$year,iter=seq(101)))%*%refpts(eq)["msy","harvest"]
+fbar(eq)=f
+
+eq=brp(eq)
+om=as(eq,"FLStock")
+om=fwd(om,f=f[,-1],sr=eq)
+
+om2=om
+for (year in dimnames(om2)$year[-1]){
+#  lsr=landings.wt(om2)[,year]/stock.wt(om2)[,year]
+#  dsr=discards.wt(om2)[,year]/stock.wt(om2)[,year]
+
+#  stock.wt(   om2)[,year]=ddWt(stock.wt(om2)[,year],biomass(om2)[,ac(an(year)-1)],par["bmsy"])
+#  landings.wt(om2)[,year]=stock.wt(om2)[,year]*lsr
+#  discards.wt(om2)[,year]=stock.wt(om2)[,year]*dsr
+
+#  mat(om2)[,year]=1/(1+exp(-par["k"]*(stock.wt(om2)[,year]-par["w50"])))
+#  m(  om2)[,year]=par["m1"]%*%(stock.wt(om2)[,year]%^%par["m2"])
+
+#  om2=fwd(om2,f=f[,year],sr=eq)
+
+om2=ddFn(year,om2,par)}
+
+plot(window(FLStocks("OM"=om,DD=om2),start=11))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass, maturity and M-at-age: Comparison of projections with  and without density dependence.
 

### 2022-12-28 Rmd/runCond.Rmd

@@
+<<<<<<< HEAD
 ---
@@
 # References
+
+=======
+---
+title: "**Density Dependence**"
+subtitle: "Summary"
+author: "L Kell"
+date: "`r format(Sys.time(), '%d %B, %Y')`"
+output: html_document
+mathjax: TRUE
+
+tags: FLR FLCore introduction
+license: Creative Commons Attribution-ShareAlike 4.0 International Public License
+# bibliography: refs.bib
+---
+
+```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
+knitr::opts_chunk$set(echo = FALSE)
+# getwd()
+library(knitr)
+
+opts_chunk$set(cache     =TRUE, 
+               comment   =NA, 
+               warning   =FALSE, 
+               message   =FALSE, 
+               error     =FALSE, 
+               echo      =FALSE, 
+               eval      =TRUE,cache   =TRUE,
+               cache.path="../cache/cond/",
+               fig.path  ="../tex/figs/cond/",
+               fig.width =8,
+               fig.height=6,
+               dev       ="png")
+
+iFig=0
+iTab=0
+```
+
+```{r, pkgs}
+library(FLCore)
+library(ggplotFL)
+library(GGally)
+library(ggpubr)
+
+library(rjson) # install.packages("rjson"); used for dropbox function
+library(RJSONIO)
+
+library(plyr)
+library(dplyr)
+library(reshape)
+
+library(readxl)
+
+library(mgcv)
+library(mgcViz)
+
+theme_set(theme_bw(16))
+
+source("../R/get_dropbox.R")
+
+```
+
+```{r, dbox}
+dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
+
+if ("try-error"%in%is(dropboxdir))
+  dropboxdir="~/Dropbox/pelagics"
+```
+
+# Mackerel
+
+```{r, mac.data}
+load(file.path(dropboxdir,"data/om/mac.RData"))
+ts =as.data.frame(read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),3))
+
+wt =transform(as.data.frame(stock.wt(mac),drop=TRUE),cohort=year-as.numeric(age))
+wt =merge(wt,ts,by="year")
+wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
+wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
+wt$age=factor(wt$age,levels=sort(unique(wt$age)))
+
+mat=transform(as.data.frame(mat(mac),drop=TRUE),cohort=year-as.numeric(age))
+mat=merge(mat,ts)
+mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
+mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
+mat$age=factor(mat$age,levels=sort(unique(mat$age)))
+
+m  =transform(as.data.frame(m(mac),drop=TRUE),cohort=year-as.numeric(age))  
+m  =merge(m,ts)
+m  =ddply(m,.(age), transform, rsdl =data/mean(data))
+m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
+m$age=factor(m$age,levels=sort(unique(m$age))) 
+```
+
+**Biomass trends**
+
+[Just to check: what is meant by ICES biomass? ]
+
+```{r, mac.ts, fig.width=8,fig.height=4}
+ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
+  geom_line(aes(year,data,col=quant))+
+  theme(legend.position="bottom")+
+  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
+  xlab("Year")+ylab("")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Time series
+
+**Weight at age**
+
+Yearclasses connected via dotted lines.
+
+```{r, mac.wt, fig.width=8,fig.height=4}
+ggplot(wt)+  
+  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
+  geom_line(aes(year,data,group=age,col=age))+
+  geom_point(aes(year,data,col=age))+
+  xlab("Year")+ylab("Mass-at-age")+
+  theme(legend.position="bottom")+
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age)))) +
+  guides(color = guide_legend(nrow = 1))
+
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age
+
+**Maturity at age**
+
+Most of the variability is in the maturity at age 2 and, to a lesser extent, age 3.
+
+```{r, mac.m, fig.width=8,fig.height=4}
+ggplot(mat)+  
+  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
+  geom_line(aes(year,data,group=age,col=age))+
+  geom_point(aes(year,data,col=age))+
+  xlab("Year")+ylab("Maturity-at-age")+
+  theme(legend.position="bottom")+
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age)))) +
+  guides(color = guide_legend(nrow = 1))
+
+```
+
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
+
+**Natural mortality at age**
+
+Constant by age and by year
+
+```{r, mac.mat, fig.width=8,fig.height=4}
+ggplot(m)+  
+  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
+  geom_line(aes(year,data,group=age,col=age))+
+  geom_point(aes(year,data,col=age))+
+  xlab("Year")+ylab("M-at-age")+
+  theme(legend.position="bottom")+
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age)))) +
+  guides(color = guide_legend(nrow = 1))
+
+```
+
+**Figure `r iFig=iFig+1; iFig`** M-at-age
+
+**Weight at age vs. biomass**
+
+Boxplot of weight at age and [xxxx]
+
+```{r, mac.wt2, fig.width=8,fig.height=5}
+#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
+#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
+
+a=ggplot(wt)+
+  geom_boxplot(aes(age,data))+
+  xlab("Mass-at-age")+ylab("Age")
+
+b=ggplot(subset(wt,rsdl<3&rsdl>0))+
+  geom_point(aes(tb,rsdl))+
+  geom_smooth(aes(tb,rsdl),method="lm")+
+  facet_wrap(age~.)+
+  xlab("Biomass")+ylab("Residual")
+
+ggarrange(a,b ,label.y=1,  
+          widths = c(4,6), heights = c(1,1,1),
+          nrow   = 1,          ncol    = 2)+
+          theme_bw(16)+
+          theme(axis.text.y = element_blank(),
+          axis.ticks.y = element_blank(),
+          axis.title.y = element_blank())
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary
+
+```{r, mac.wt.tb, fig.height=8,fig.width=8}
+ggplot(subset(wt,age!=0),aes(tb,data))+
+  geom_point()+
+  geom_smooth(se=F,method="lm")+
+  facet_wrap(~age,scale="free",ncol=3)+
+  xlab("Biomass")+ylab("Mass-at-age")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
+
+```{r, mac.mat2, fig.width=8,fig.height=5}
+#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
+#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
+
+a=ggplot(mat)+
+  geom_boxplot(aes(age,data))+
+  xlab("Mass-at-age")+ylab("Age")
+
+b=ggplot(mat)+
+  geom_point(aes(tb,rsdl))+
+  geom_smooth(aes(tb,rsdl),method="lm")+
+  facet_wrap(age~.)+
+  xlab("Biomass")+ylab("Residual")
+
+ggarrange(a,b ,label.y=1,  
+          widths = c(4,6), heights = c(1,1,1),
+          nrow   = 1,          ncol    = 2)+
+          theme_bw(16)+
+          theme(axis.text.y = element_blank(),
+          axis.ticks.y = element_blank(),
+          axis.title.y = element_blank())
+```
+
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary
+
+```{r, mac.mat.tb, fig.height=5,fig.width=8}
+ggplot(subset(mat,age%in%(1:5)),aes(tb,data))+
+  geom_point()+
+  geom_smooth(se=F,method="lm")+
+  facet_wrap(~age,scale="free",ncol=3)+
+  xlab("Biomass")+ylab("Maturity-at-age")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v Biomass
+
+```{r}
+gmr=gammV(data~s(tb, bs="tp")+age, data=wt)
+
+summary(gmr)
+
+plot(gmr)
+```
+
+# Blue whiting
+
+```{r, whb.data}
+load(file.path(dropboxdir,"data/om/whb.RData"))
+ts=model.frame(FLQuants(whb,tb=biomass,ssb=ssb),drop=T)
+ts$tb=ts$tb/1e6
+
+wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
+wt =merge(wt,ts,by="year")
+wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
+wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
+wt$age=factor(wt$age,levels=sort(unique(wt$age)))
+
+mat=transform(as.data.frame(mat(whb),drop=TRUE),cohort=year-as.numeric(age))
+mat=merge(mat,ts)
+mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
+mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
+mat$age=factor(mat$age,levels=sort(unique(mat$age)))
+
+m  =transform(as.data.frame(m(whb),drop=TRUE),cohort=year-as.numeric(age))  
+m  =merge(m,ts)
+m  =ddply(m,.(age), transform, rsdl =data/mean(data))
+m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
+m$age=factor(m$age,levels=sort(unique(m$age)))
+```
+
+```{r, whb.ts, fig.width=8,fig.height=4}
+ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
+  geom_line(aes(year,data,col=quant))+
+  theme(legend.position="bottom")+
+  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
+  xlab("Year")+ylab("")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Time series
+
+```{r, whb.wt, fig.width=8,fig.height=4}
+ggplot(wt)+  
+  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
+  geom_line(aes(year,data,group=age,col=age))+
+  geom_point(aes(year,data,col=age))+
+  xlab("Year")+ylab("Mass-at-age")+
+  theme(legend.position="bottom")+
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age
+
+```{r, whb.mat, fig.width=8,fig.height=4}
+ggplot(mat)+  
+  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
+  geom_line(aes(year,data,group=age,col=age))+
+  geom_point(aes(year,data,col=age))+
+  xlab("Year")+ylab("Mass-at-age")+
+  theme(legend.position="bottom")+
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
+
+```{r, whb.m, fig.width=8,fig.height=4}
+ggplot(m)+  
+  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
+  geom_line(aes(year,data,group=age,col=age))+
+  geom_point(aes(year,data,col=age))+
+  xlab("Year")+ylab("Mass-at-age")+
+  theme(legend.position="bottom")+
+  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
+```
+
+**Figure `r iFig=iFig+1; iFig`** M-at-age
+
+```{r, whb.wt2, fig.width=8,fig.height=5}
+#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
+#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
+
+a=ggplot(wt)+
+  geom_boxplot(aes(age,data))+
+  xlab("Mass-at-age")+ylab("Age")
+
+b=ggplot(subset(wt,rsdl<3&rsdl>0))+
+  geom_point(aes(tb,rsdl))+
+  geom_smooth(aes(tb,rsdl),method="lm")+
+  facet_wrap(age~.,ncol=3)+
+  xlab("Biomass")+ylab("Residual")
+
+ggarrange(a,b ,label.y=1,  
+          widths = c(4,6), heights = c(1,1,1),
+          nrow   = 1,          ncol    = 2)+
+          theme_bw(16)+
+          theme(axis.text.y = element_blank(),
+          axis.ticks.y = element_blank(),
+          axis.title.y = element_blank())
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary
+
+```{r, whb.wt.tb, fig.height=8,fig.width=8}
+ggplot(wt,aes(tb,data))+
+  geom_point()+
+  geom_smooth(se=F,method="lm")+
+  facet_wrap(~age,scale="free",ncol=3)+
+  xlab("Biomass")+ylab("Mass-at-age")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
+
+```{r, whb.mat2, fig.width=8,fig.height=5}
+#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
+#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
+
+a=ggplot(mat)+
+  geom_boxplot(aes(age,data))+
+  xlab("Mass-at-age")+ylab("Age")
+
+b=ggplot(mat)+
+  geom_point(aes(tb,rsdl))+
+  geom_smooth(aes(tb,rsdl),method="lm")+
+  facet_wrap(age~.,ncol=3)+
+  xlab("Biomass")+ylab("Residual")
+
+ggarrange(a,b ,label.y=1,  
+          widths = c(4,6), heights = c(1,1,1),
+          nrow   = 1,          ncol    = 2)+
+          theme_bw(16)+
+          theme(axis.text.y = element_blank(),
+          axis.ticks.y = element_blank(),
+          axis.title.y = element_blank())
+```
+
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary
+
+```{r, whb.mat.tb, fig.height=5,fig.width=8}
+ggplot(subset(mat,age%in%1:5),aes(tb,data))+
+  geom_point()+
+  geom_smooth(se=F,method="lm")+
+  facet_wrap(~age,scale="free",ncol=3)+
+  xlab("Biomass")+ylab("Mass-at-age")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
+
+```{r}
+gmr=gammV(data~s(tb)+age, data=wt)
+
+summary(gmr)
+
+plot(gmr)
+
+```
+
+# Funding
+
+# References
+>>>>>>> 0396807443278301a395fdf2e6e0f8bbf2867d49

### 2022-12-28 Rmd/runCond.Rmd
with mat

--- 2022-12-28 Rmd/runCond.Rmd
@@
 fn<-function(x,dat) 
-   sum((dat$mat-matFn(x,dat$wt))^2)
+   sum((dat$mat-matFn(dat$wt,x))^2)
 
 par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
-
-par$par
 

### 2022-12-28 Rmd/runDD.Rmd
with mat

@@
 Mackerel life history parameters are used to construct an `FLBRP` object to represent the equilibrium dynamics and then an `FLStock` object models the time series dynamics.
+
+Mass-at-age is assumed to vary with total biomass, we assume that estimates of mass-at-age ({W_t}^\prime) are based on observations around $B_{MSY}$ and if total biomass increased density dependence results in a decline in the expected value of $W_a$, i.e.
+$W_t = {W_t}^\prime a(B_t/B_{MSY})^b$
+
+Maturity at age is modelled by a logistic relationship $f(x) = \frac{1}{1+e^{-r(W-$W_50$)}}$ where $W_{50}$ is the weight when $50%$ are mature, and $r$ a parameter that determines how fast individuals mature.
 

### 2022-12-28 Rmd/runDD.Rmd
with eqns

@@
 
-Mackerel life history parameters are used to construct an `FLBRP` object to represent the equilibrium dynamics and then an `FLStock` object models the time series dynamics.
+Mackerel life history parameters are used to construct an example of modelling density dependence, first an `FLBRP` object is used  represent the equilibrium dynamics and then an `FLStock` object is constructed to models the time series dynamics.
 

### 2023-01-10 Rmd/mseMac.Rmd
cleaned up

---
title: "**Density Dependence**"
subtitle: "Example of MSE for Mackerel"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mseMac/",
               fig.path  ="../tex/figs/mseMac/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLAssess)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r}
source("../R/dd.R")
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

```{r, mac.data}
load(file.path(dropboxdir,"data/om/mac.RData"))

par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bmsy=3500000,b=0.3,k=30.1, w50=0.2))

stock.wt(   mac)[1]=0.01
catch.wt(   mac)[1]=0.01
landings.wt(mac)[1]=0.01
discards.wt(mac)[1]=0.01

om=ddFn2(mac,par)
om=om+VPA(om)
om=ddFn2(om,par)

sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
eq=FLBRP(om,sr=sr)

om=fwdWindow(om,end=2080,eq)
f =fbar(om)[,ac(2021:2080)]%=%refpts(eq)["msy","harvest"]
om=fwd(om,fbar=f,sr=eq)

om2=om
om3=om
om4=om
for (iYear in ac(2021:2080)){
  om2=ddFn(iYear,om2,par)
  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)

  om3=ddFn(iYear,om3,par)
  om3=fwd(om3,fbar=f[,iYear],sr=eq)

  om4=ddFn(iYear,om4,par)
  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
  }

plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
```

```{r}
ggplot(FLQuants("ICES"=apply(stock.wt(mac[,"2021"]),1,mean),
                "OM"  =stock.wt(om[ ,"2080"]),
                "0.8" =stock.wt(om2[,"2080"]),
                "1.0" =stock.wt(om3[,"2080"]),
                "1.2" =stock.wt(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r}
ggplot(FLQuants("ICES"=m(mac[,"2021"]), 
                "OM"  =m(om[ ,"2080"]),
                "0.8" =m(om2[,"2080"]),
                "1.0" =m(om3[,"2080"]),
                "1.2" =m(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("M-at-age")
```

**Figure `r iFig=iFig+1; iFig`** M-at-age

```{r}
ggplot(FLQuants("ICES"=apply(mat(mac[,"2021"]),1,mean),
                "OM"  =mat(om[ ,"2080"]),
                "0.8" =mat(om2[,"2080"]),
                "1.0" =mat(om3[,"2080"]),
                "1.2" =mat(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("Maturity-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

### 2023-01-10 Rmd/mseWhb.Rmd
cleaned up

---
title: "**Density Dependence**"
subtitle: "Example of MSE for Blue whiting"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE
tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mseWhb/",
               fig.path  ="../tex/figs/mseWhb/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLAssess)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

theme_set(theme_bw(16))
```

```{r}
source("../R/dd.R")
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

```{r, whb.data}
load(file.path(dropboxdir,"data/om/whb.RData"))

par=FLPar(c(m1=0.08/(0.2^-0.288),m2=-0.28,bmsy=6e6,k=63,w50=0.08))

stock.wt(   whb)[1]=0.01
catch.wt(   whb)[1]=0.01
landings.wt(whb)[1]=0.01
discards.wt(whb)[1]=0.01

om=ddFn2(whb,par)
om=om+VPA(om)
om=ddFn2(om,par)

sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
eq=FLBRP(om,sr=sr)

om=fwdWindow(om,end=2080,eq)
f =fbar(om)[,ac(2021:2080)]%=%refpts(eq)["msy","harvest"]
om=fwd(om,fbar=f,sr=eq)

om2=om
om3=om
om4=om
for (iYear in ac(2021:2080)){
  om2=ddFn(iYear,om2,par)
  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)

  om3=ddFn(iYear,om3,par)
  om3=fwd(om3,fbar=f[,iYear],sr=eq)

  om4=ddFn(iYear,om4,par)
  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
  }

plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
```

```{r}
ggplot(FLQuants("ICES"=stock.wt(om[ ,"2080"]), 
                "0.8" =stock.wt(om2[,"2080"]), 
                "1.0" =stock.wt(om3[,"2080"]),
                "1.2" =stock.wt(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("Mass-at-age")
```

**Figure `r iFig=iFig+1; iFig`** Mass-at-age

```{r}
ggplot(FLQuants("ICES"=m(om[ ,"2080"]),  
                "0.8" =m(om2[,"2080"]), 
                "1.0" =m(om3[,"2080"]),
                "1.2" =m(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("M") 
```

**Figure `r iFig=iFig+1; iFig`** M-at-mass

```{r}
ggplot(FLQuants("ICES"=mat(om[ ,"2080"]),  
               "0.8"  =mat(om2[,"2080"]),
               "1.0"  =mat(om3[,"2080"]), 
               "1.2"  =mat(om4[,"2080"])))+
  geom_line(aes(age,data,col=qname))+
  xlab("Age")+ylab("Maturity-a-age")
```

**Figure `r iFig=iFig+1; iFig`** Maturity-at-age

```{r}
ggplot(rbind(cbind(What="ICES",model.frame(FLQuants(Mat=mat(om[  ,"2080"]),Mass=stock.wt(om[ ,"2080"])))),  
             cbind(What= "0.8",model.frame(FLQuants(Mat=mat(om2[ ,"2080"]),Mass=stock.wt(om2[,"2080"])))),
             cbind(What= "1.0",model.frame(FLQuants(Mat=mat(om3[ ,"2080"]),Mass=stock.wt(om3[,"2080"])))),
             cbind(What= "1.2",model.frame(FLQuants(Mat=mat(om4[ ,"2080"]),Mass=stock.wt(om4[,"2080"]))))))+
  geom_line(aes(Mass,Mat,col=What))+
  geom_point(aes(Mass,Mat,col=What))
```

**Figure `r iFig=iFig+1; iFig`** M-at-mass

### 2023-01-10 Rmd/runCond.Rmd
cleaned up

--- 2022-12-28 Rmd/runCond.Rmd
@@
 wt$age=factor(wt$age,levels=sort(unique(wt$age)))
+
 mat=transform(as.data.frame(mat(mac),drop=TRUE),cohort=year-as.numeric(age))
@@
 mat$age=factor(mat$age,levels=sort(unique(mat$age)))
+
 m  =transform(as.data.frame(m(mac),drop=TRUE),cohort=year-as.numeric(age))  
@@
   geom_smooth(se=F,span=0.2)+
-  xlab("Age")+ylab("Matutrity")+
+  xlab("Mass-at-age")+ylab("Matutrity-at-age")+
   geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
@@
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
 
 ```{r}
-gmr=gammV(data~s(tb, bs="tp")+age, data=wt)
+wt=subset(wt,age!=0)
+gmr=gammV(log(data)~s(tb/mean(tb), bs="tp")+age, data=wt)
 summary(gmr)
 plot(gmr)
+coefficients(gmr)
+
+wt=cbind(wt,hat=predict(gmr))
+
+plot(exp(coefficients(gmr)[2:12]))
+wt=merge(wt,data.frame(age=1:12,mean=exp(c(0,coefficients(gmr)[2:12]))))
+
+ggplot(wt)+
+  geom_line(aes(tb,exp(hat)/mean,col=age))+
+  geom_point(aes(tb,data/mean,col=age))+
+  xlab("Biomass")+ylab("Mass-Mean")
+
+summary(lm(hat~tb,data=transform(wt,tb=tb/mean(tb))))
 ```
@@
 load(file.path(dropboxdir,"data/om/whb.RData"))
-ts=model.frame(FLQuants(whb,tb=biomass,ssb=ssb),drop=T)
+ts=model.frame(FLQuants(whb,tb=stock,ssb=ssb),drop=T)
 ts$tb=ts$tb/1e6
@@
 wt$age=factor(wt$age,levels=sort(unique(wt$age)))
+
 mat=transform(as.data.frame(mat(whb),drop=TRUE),cohort=year-as.numeric(age))
@@
 mat$age=factor(mat$age,levels=sort(unique(mat$age)))
+
 m  =transform(as.data.frame(m(whb),drop=TRUE),cohort=year-as.numeric(age))  
@@
 
+**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
+
 ```{r}
-gmr=gammV(data~s(tb)+age, data=wt)
+gmr=gammV(log(data)~s(tb, bs="tp")+s(year, bs="tp")+age, data=wt)
 summary(gmr)
 plot(gmr)
+coefficients(gmr)
+
+wt=cbind(wt,hat=predict(gmr))
+
+plot(coefficients(gmr)[2:10])
+wt=merge(wt,data.frame(age=1:10,mean=c(0,coefficients(gmr)[2:10])))
+
+p=ggplot(wt)+
+  geom_line(aes(tb,exp(hat-mean),col=age))+
+  geom_point(aes(tb,data-exp(mean),col=age))+
+  xlab("Biomass")+ylab("Mass-Mean")
+```
+
+```{r}
+wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
+wt =merge(wt,ts,by="year")
+wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
+wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
+wt$age=factor(wt$age,levels=sort(unique(wt$age)))
+
+wt=subset(wt,age!=0)
+gmr=gammV(log(data)~s(tb/mean(tb), bs="tp")+age, data=wt)
+summary(gmr)
+plot(gmr)
+coefficients(gmr)
+
+wt=cbind(wt,hat=predict(gmr))
+
+plot(exp(coefficients(gmr)[2:12]))
+wt=merge(wt,data.frame(age=1:12,mean=exp(c(0,coefficients(gmr)[2:12]))))
+
+ggplot(wt)+
+  geom_line(aes(tb,exp(hat)/mean,col=age))+
+  geom_point(aes(tb,data/mean,col=age))+
+  xlab("Biomass")+ylab("Mass-Mean")
+
+summary(lm(hat~tb,data=transform(wt,tb=tb/mean(tb))))
 ```

### 2023-01-10 Rmd/runDD.Rmd
cleaned up

--- 2022-12-28 Rmd/runDD.Rmd
@@
 title: "**Density Dependence**"
-subtitle: "Example of modellimng density dependence in mass, maturity and M-at-age"
+subtitle: "Example of modelling density dependence in mass, maturity and M-at-age"
 author: "L Kell"
@@
 
-Mackerel life history parameters are used to construct an `FLBRP` object to represent the equilibrium dynamics and then an `FLStock` object models the time series dynamics.
+Mackerel life history parameters are used to construct an example of modelling density dependence. An `FLBRP` object represents the equilibrium and an `FLStock` models the time series dynamics.
 
-## Natural mortality
+### Mass-at-age
 
-@lorenzen2000allometry noted that theoretical and empirical studies point to the existence of an allometric relationship between natural mortality $M_a$ and body mass ($W_a$) at-age in fish of the form:
+Mass-at-age ($W_a$) is modelled as
 
-  $M_a \propto {W_a}^{d}$
+$W_a = {\alpha}_a B^{{\beta}_a}$
 
-i.e.
+Where $\alpha$ and $\beta$ are taken from a linear regression of Mass~Total biomass, i.e. we model mass-at-age at each age based on empirical data.
+
+$W_a$ is also constrained so sizes do not become unfeasibly small or large
+
+### Predation mortality-at-age (M2)
+
+#### Pope
+
+@pope2006modelling (doi:10.1016/j.icesjms.2006.04.015) proposed a way of modelling predation mortality (M2) based on size spectra, by modelled $M$ as a function of length i.e.
+
+$M2(L)=\gamma(F) e^{\lambda(F) L}$
+
+where Where $\gamma=2.0753+0.9914F$ and $\lambda = −0.0276−0.1088F$ and $F$ is fishing mortality
+
+However, the model is in mass-at-age, so we convert $L$ to $W$
+
+$M2(L)=\gamma(F) e^{\lambda(F) (W/a)^{1/b}}$
+
+$\gamma$ is scaled so that M at at reference length is equal to a desired value.
+
+A problem is that this will formulation respond immediately to a change in F.
+However, the size spectra depends on the historical not the current $F$.
+For example, for a specific value of $F$ in the current year, if the stock as been subject to a period of heavy exploitation the size spectra will be very different to that of a virgin fishery in the first year of exploitation.
+This will create problems in simulations of management as in an MSE.
+
+#### Lorenzen
+
+Theoretical and empirical studies [@lorenzen2000allometry] support the existence of an allometric relationship between natural mortality ($M_a$) and body mass ($W_a$) at age of the form:
 
@@
 
-where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37, e.g.
-if $M_t=0.2$ at a weight when $50\%$ of individuals reach maturity $W_{50}$ then
+where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
 
-  $M_u = 0.2/({W_{50}}^{d})$
+If $M=0.2$ at $W_{50}$ then $M_u = 0.2/({W_{50}}^{d})$
 
-Since weight is approximately proportional to the third power of length  values of $d$ of approximately –1/3 imply that $M$ is inversely proportional to body length.
-@10.1093/icesjms/fst226 recognised that for ($L$) if the von Bertalanffy and length–weight parameters are known, then the mortality schedule can be defined by
+$d$ is approximately equal to $–1/3$, which implies that $M$ is inversely proportional to body length., since mass is approximately proportional to the third power of length.
 
-$L_t=L_{\infty}{(1-e^{(1-k(t-t_0))})}^{b(-0.305)}$
+### Maturity-at-weight
 
-where $L_{\infty}$ is the asmypotopic length,  $k$ growth rate, and $t_0$ is the age at which length is zero of the von Bertalanffy growth equation,  and the allometric relationship $W = a L^b$.
+Modelled as a logistic function of mass-at-age e.g.
 
-This gives
+$O(W) = 1/(1+e^(k(W-W50)))$
 
-$M_t = M_{\infty}{(W_t/W_{\infty})}^{-0.3}$
+## Sparholt, empirical
 
-where $M_t$ is the instantaneous $M$ rate per year at age $t$ and $M_{\infty}$ is the asymptotic $M$ rate at the asymptotic weight $W_{\infty}$, derived from life history characteristics @hoenig1992empirical
+Mackerel
 
-## Multispecies modelling of predation mortality (M2)
+$M = 0.0232 * TB + 0.0524$
 
-@pope2006modelling (doi:10.1016/j.icesjms.2006.04.015) proposed a way of modelling predation mortality (M2) based on size spectra, e.g.
+Blue whiting:
 
-$N(L) = \Omega L^{\kappa}$
-
-where $N$ is numbers at length $L$ and $\Omega$ is a prefactor and $\kappa$ the scaling by length
-
-This was further developed by @POPE2021105952 (https://doi.org/10.1016/j.fishres.2021.105952) by assuming that predation by all fish in the spectrum is mediated through a feeding relationship with mean $\mu$ and standard deviation $\sigma$, both common to all predators and that UM2 is a constant that relates predation pressure to mortality, and where $\lambda$ is the scaling of the feeding rate of predators by length then the predation mortality on a prey species of length $L^{\prime}$ can be written as
-
-$M2(L^\prime) = \Omega \times UM2(L^\prime)^{(\kappa + \lambda)} e^{(\mu((\kappa + \lambda)/3))} \times e^{({\sigma}^2((\kappa+\lambda)^2)/18)}$
-
-<!-- Another insight derived from size spectra studies comes from the work of Andersen et al.
-(2009).
-They derive a formulation for the predation mortality to be expected from a size spectrum with a given slope and intercept in terms of mass.
-However, the CSM is a rather different model to theirs since it only considers the fish part of the ecosystem and is framed in terms of length rather than mass.
-Moreover, it treats predation mortality as a top-down process rather than in terms of mass balance.
-Predation mortality is defined on a simple UM2*predation pressure basis.
-Here, UM2 is defined as a constant M2 per unit of predation pressure.
-The predation pressure is summed over all predator lengths with a weighting for the feeding relationship between specific prey and predator lengths.
-For a particular length L of predator, predation pressure is taken to be their numbers times their length to a power $\gamma$ (if $\gamma=3$), this would be equivalent to biomass except that the condition factor linking $L^3$ to mass is included into the UM2 constant to simplify the equations).
-Hence, assuming a numbers size spectrum described by -->
+$M = 0.0221 * TB  + 0.0719$
 
@@
 ```{r}
-ddWt<-function(wt,biomass, bmsy)
-  wt%*%qmax((biomass/bmsy)^0.3,0.5)
-
-matFn<-function(wt,x) 
-   1/(1+exp(-x[1]*(wt-x[2])))
-
-ddFn<-function(year,x,par){
-  lsr=landings.wt(x)[,year]/stock.wt(x)[,year]
-  dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
-
-  stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bmsy"])
-  landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
-  discards.wt(x)[,year]=stock.wt(x)[,year]*dsr
-
-  mat(x)[,year]=matFn(stock.wt(x)[,year],par[c("k","w50")])
-  m(  x)[,year]=par["m1"]%*%(stock.wt(x)[,year]%^%par["m2"])
-
-  fwd(x,f=f[,year],sr=eq)}
+source("../R/dd.R")
 ```
@@
 
-# Sumulations
+# Simulations
 
@@
 ```{r}
-plot(qmax((biomass(eq)%/%biomass(eq)[,c(14)])^0.3,0.5))+
+plot(qmin(qmax(0.96*biomass(eq)%/%biomass(eq)[,14],0.5),1.5))+
   scale_y_continuous(labels = scales::percent)+
@@
 ```{r}
+save(eq,file="~/Desktop/eq.RData")
+```
+
+```{r}
 wt=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"]) 
-
 ggplot(as.data.frame(wt))+           
@@
 ```{r}
+ggplot(ddM(stock.wt(eq),fbar(eq)))+
+    geom_line(aes(age,data,col=factor(year)))+
+    theme(legend.position="none")+
+    scale_x_continuous(limits=c(0,15))+
+    xlab("Age")+ylab("M-at-age") 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Pope: simulation of density dependence in M-at-age
+
+```{r}
 m=par["m1"]%*%(wt%^%par["m2"])
@@
 
-**Figure `r iFig=iFig+1; iFig`** Simulation of density dependence in M-at-age
+**Figure `r iFig=iFig+1; iFig`** Lorenzen: Simulation of density dependence in M-at-age
 
@@
 
-```{r}
+```{r, dd-eq-mass}
 x=propagate(eq,length(dimnames(eq)$year))
 fbar(x)=fbar(x)[,1]
-fbar(x)=c(fbar(eq))
+fbar(x)[]=c(fbar(eq))*2
 
@@
 # iterate
-for (i in seq(20)){
-  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"])
+for (i in seq(10)){
+  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"],alpha=-0.17)
   landings.wt(x)=stock.wt(x)*lsr
@@
 
-```{r}
+```{r, dd-eq-mass-mat}
 x=propagate(eq,length(dimnames(eq)$year))
+fbar(x)[,1]=c(seq(0,max(fbar(eq)),length.out=51),seq(1,50,length.out=51)*max(fbar(eq)))[-52]
 fbar(x)=fbar(x)[,1]
-fbar(x)=c(fbar(eq))
 
@@
 
-```{r}
+```{r, dd-eq-mass-mat-m}
 x=propagate(eq,length(dimnames(eq)$year))
+fbar(x)[,1]=c(seq(0,max(fbar(eq)),length.out=51),seq(1,50,length.out=51)*max(fbar(eq)))[-52]
 fbar(x)=fbar(x)[,1]
-fbar(x)=c(fbar(eq))
 
@@
 om2=om
+lsr=landings.wt(om2)/stock.wt(om2)
+dsr=discards.wt(om2)/stock.wt(om2)
+
 for (year in dimnames(om2)$year[-1]){
-#  lsr=landings.wt(om2)[,year]/stock.wt(om2)[,year]
-#  dsr=discards.wt(om2)[,year]/stock.wt(om2)[,year]
 
-#  stock.wt(   om2)[,year]=ddWt(stock.wt(om2)[,year],biomass(om2)[,ac(an(year)-1)],par["bmsy"])
-#  landings.wt(om2)[,year]=stock.wt(om2)[,year]*lsr
-#  discards.wt(om2)[,year]=stock.wt(om2)[,year]*dsr
+  #stock.wt(   om2)[,year]=ddWt(stock.wt(om2)[,year],biomass(om2)[,ac(an(year))],par["bmsy"])
+  #landings.wt(om2)[,year]=stock.wt(om2)[,year]*lsr[,year]
+  #discards.wt(om2)[,year]=stock.wt(om2)[,year]*dsr[,year]
 
-#  mat(om2)[,year]=1/(1+exp(-par["k"]*(stock.wt(om2)[,year]-par["w50"])))
-#  m(  om2)[,year]=par["m1"]%*%(stock.wt(om2)[,year]%^%par["m2"])
+  #mat(om2)[,year]=1/(1+exp(-par["k"]*(stock.wt(om2)[,year]-par["w50"])))
+  #m(  om2)[,year]=par["m1"]%*%(stock.wt(om2)[,year]%^%par["m2"])
 
-#  om2=fwd(om2,f=f[,year],sr=eq)
+  #om2=fwd(om2,f=f[,year],sr=eq)}
 
@@
 # References
+
+## Mackerel
+
+Mass-at-age
+
+$W_a = \alpha_aB+\beta$
+
+<!-- $M = ~~0.0232 \times B + 0.0524$ -->
+
+Predation mortality-at-age
+
+$M_2(L) =ae^{bL}$
+
+where $a=2.0753+0.9914F$ and $b=-0.0276-0.1088F$
+
+Maturity-at-weight
+
+$O(W) = \frac{1}{1+e^{-1.67(W-28.2)}}$
+
+## Blue whiting:
+
+$M = 0.0221 \times B   + 0.0719$
+
+## Mass-at-age
+
+Individual mass-at-age is assumed to vary with total biomass, estimates of mass-at-age (${W_t}^\prime$) are based on observations around $B_{MSY}$, and if total biomass increases density dependence results in a decline in the expected value of $W_a$, i.e.
+
+$W_t = {W_t}^\prime a(B_t/B_{MSY})^b$
+
+## Maturity-at-age
+
+Maturity at age is modelled by a logistic relationship
+
+$f(x) = \frac{1}{1+e^{-r(W-W_{50})}}$
+
+where $W_{50}$ is the mass-at-age when $50%$ of individuals are mature, and $r$ the logistic growth rate or steepness of the curve.
+
+## Natural mortality
+
+Theoretical and empirical studies [@lorenzen2000allometry] support the existence of an allometric relationship between natural mortality ($M_a$) and body mass ($W_a$) at age of the form:
+
+  $M_W=M_uW^d$
+
+where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
+
+If $M=0.2$ at $W_{50}$ then $M_u = 0.2/({W_{50}}^{d})$
+
+$d$ is approximately equal to $–1/3$, which implies that $M$ is inversely proportional to body length., since mass is approximately proportional to the third power of length.
+
+@10.1093/icesjms/fst226 recognised that if the von Bertalanffy growth equation
+
+$L_t=L_{\infty}{(1-e^{(1-k(t-t_0))})}^{b(-0.305)}$
+
+and the allometric length weight relationship
+
+$W = a L^b$.
+
+are combined since
+
+$M_a = M_{\infty}{(W_a/W_{\infty})}^{-0.3}$
+
+then mortality schedule can be defined by
+
+$M_a=M_{\infty}(1−e^{(−K(t−t_0))})^{−b(0.3)}$
+
+where $M_{\infty}$ can be derived from life history characteristics @hoenig1992empirical
+
+## Multispecies modelling of predation mortality (M2)
+
+@pope2006modelling (doi:10.1016/j.icesjms.2006.04.015) proposed a way of modelling predation mortality (M2) based on size spectra, e.g.
+
+$N(L) = \Omega L^{\kappa}$
+
+where $N$ is numbers at length $L$ and $\Omega$ is a prefactor and $\kappa$ the scaling by length
+
+This was further developed by @POPE2021105952 (https://doi.org/10.1016/j.fishres.2021.105952) by assuming that predation by all fish in the spectrum is mediated through a feeding relationship with mean $\mu$ and standard deviation $\sigma$, both common to all predators and that UM2 is a constant that relates predation pressure to mortality, and where $\lambda$ is the scaling of the feeding rate of predators by length then the predation mortality on a prey species of length $L^{\prime}$ can be written as
+
+$M2(L^\prime) = \Omega \times UM2 (L^\prime)^{(\kappa + \lambda)}\times e^{\mu((\kappa + \lambda)/3)} \times e^{{\sigma}^2((\kappa+\lambda)^2)/18}$
+
+This simplifies to
+
+$M2(L) = aL^b$
+
+and is therefore equivalent to
+
+  $M_W=M_uW^d$
+
+depending on the choice of $a$ and $b$
+
+<!-- Another insight derived from size spectra studies comes from the work of Andersen et al.
+(2009).
+They derive a formulation for the predation mortality to be expected from a size spectrum with a given slope and intercept in terms of mass.
+However, the CSM is a rather different model to theirs since it only considers the fish part of the ecosystem and is framed in terms of length rather than mass.
+Moreover, it treats predation mortality as a top-down process rather than in terms of mass balance.
+Predation mortality is defined on a simple UM2*predation pressure basis.
+Here, UM2 is defined as a constant M2 per unit of predation pressure.
+The predation pressure is summed over all predator lengths with a weighting for the feeding relationship between specific prey and predator lengths.
+For a particular length L of predator, predation pressure is taken to be their numbers times their length to a power $\gamma$ (if $\gamma=3$), this would be equivalent to biomass except that the condition factor linking $L^3$ to mass is included into the UM2 constant to simplify the equations).
+Hence, assuming a numbers size spectrum described by -->

### 2023-01-10 R/dd.R

@@ -1,0 +1,39 @@
+ddWt<-function(wt,biomass,bmsy,alpha=-0.25)
+  wt%*%qmin(qmax((biomass%/%bmsy)^alpha,0.1),10)
+
+ddMat<-function(wt,x) 
+  1/(1+exp(-x[1]%*%(wt%-%x[2])))
+
+ddM<-function(wt, F, gamma=2.0753+0.9914*F, lambda=-0.0276-0.1088*F, refWt=310, refM=0.10, M1=0.05){
+  
+  ref=refM/(gamma%*%exp(lambda*(refWt/0.0003)^(1/3)))
+  
+  exp(((wt/0.0001)^(1/3))%*%lambda)%*%gamma%*%ref+M1}
+
+ddFn<-function(year,x,par){
+  lsr=landings.wt(x)[,year]/stock.wt(x)[,year]
+  dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
+  dsr[!is.finite(dsr)]=0
+  
+  stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bmsy"],alpha=-0.25)
+  landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
+  discards.wt(x)[,year]=stock.wt(x)[,year]*dsr
+  
+  mat(x)[,year]=ddMat(stock.wt(x)[,year],par[c("k","w50")])
+  m(  x)[,year]=par["m1"]%*%(stock.wt(x)[,year]%^%par["m2"])
+  
+  x}
+
+ddFn2<-function(x,par){
+  lsr=landings.wt(x)/stock.wt(x)
+  dsr=discards.wt(x)/stock.wt(x)
+  dsr[!is.finite(dsr)]=0
+  
+  stock.wt(   x)=ddWt(stock.wt(x),biomass(x),par["bmsy"],alpha=-0.25)
+  landings.wt(x)=stock.wt(x)*lsr
+  discards.wt(x)=stock.wt(x)*dsr
+  
+  mat(x)=ddMat(stock.wt(x),par[c("k","w50")])
+  m(  x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
+  
+  x}

### 2023-02-02 Rmd/runDD.Rmd
with eqns

@@
 title: "**Density Dependence**"
-subtitle: "Example of modelling density dependence in mass, maturity and M-at-age"
+subtitle: "Modelling density dependence in mass, maturity and M-at-age"
 author: "L Kell"
@@
 
-Mackerel life history parameters are used to construct an example of modelling density dependence. An `FLBRP` object represents the equilibrium and an `FLStock` models the time series dynamics.
+Mackerel life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
+The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
 
@@
 
-Where $\alpha$ and $\beta$ are taken from a linear regression of Mass~Total biomass, i.e. we model mass-at-age at each age based on empirical data.
-
-$W_a$ is also constrained so sizes do not become unfeasibly small or large
-
-### Predation mortality-at-age (M2)
-
-#### Pope
-
-@pope2006modelling (doi:10.1016/j.icesjms.2006.04.015) proposed a way of modelling predation mortality (M2) based on size spectra, by modelled $M$ as a function of length i.e.
-
-$M2(L)=\gamma(F) e^{\lambda(F) L}$
-
-where Where $\gamma=2.0753+0.9914F$ and $\lambda = −0.0276−0.1088F$ and $F$ is fishing mortality
-
-However, the model is in mass-at-age, so we convert $L$ to $W$
-
-$M2(L)=\gamma(F) e^{\lambda(F) (W/a)^{1/b}}$
-
-$\gamma$ is scaled so that M at at reference length is equal to a desired value.
-
-A problem is that this will formulation respond immediately to a change in F.
-However, the size spectra depends on the historical not the current $F$.
-For example, for a specific value of $F$ in the current year, if the stock as been subject to a period of heavy exploitation the size spectra will be very different to that of a virgin fishery in the first year of exploitation.
-This will create problems in simulations of management as in an MSE.
-
-#### Lorenzen
-
-Theoretical and empirical studies [@lorenzen2000allometry] support the existence of an allometric relationship between natural mortality ($M_a$) and body mass ($W_a$) at age of the form:
-
-  $M_W=M_uW^d$
-
-where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
-
-If $M=0.2$ at $W_{50}$ then $M_u = 0.2/({W_{50}}^{d})$
-
-$d$ is approximately equal to $–1/3$, which implies that $M$ is inversely proportional to body length., since mass is approximately proportional to the third power of length.
+Where $\alpha$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data.
+$W_a$ is constrained so sizes do not become unfeasible large or small.
 
@@
 
-Modelled as a logistic function of mass-at-age e.g.
+Maturity is modelled as a logistic function of mass-at-age i.e.
 
-$O(W) = 1/(1+e^(k(W-W50)))$
+$O(W_a) = 1/(1+e^{k(W_a-W_{50})})$
 
-## Sparholt, empirical
+### Natural mortality
 
-Mackerel
+Is assumed to vary at unit length ($L_1$) e.g.
 
-$M = 0.0232 * TB + 0.0524$
+$ln(M_1) =\alpha + \beta ln(L_{\infty}) + \gamma ln(k)$
+
+where $\alpha=0.65$, $\beta=0.91$, $\gamma=0.87$
+
+Length-at-age ($L_a$) can be obtained from the length-weight relationship
+
+$L_a=(W/a)^{1/b}$
+
+giving
+
+$ln(M_1) =(W/a)^{1/b}(\alpha + \beta ln(L_{\infty}) + \gamma ln(k))$
+
+For DD there is a relationship between $M_a$ and total biomass ($B$)
+
+Mackerel:
+
+$M = 0.0232B + 0.0524$
 
@@
 
-$M = 0.0221 * TB  + 0.0719$
+$M = 0.0221B + 0.0719$
+
+Combining these provides values of $M_a$
 
@@
 
-## Mackerel
+<!-- ## Mackerel -->
 
-Mass-at-age
+<!-- Mass-at-age -->
 
-$W_a = \alpha_aB+\beta$
+<!-- $W_a = \alpha_aB+\beta$ -->
 
-<!-- $M = ~~0.0232 \times B + 0.0524$ -->
+<!-- <!-- $M = ~~0.0232 \times B + 0.0524$ --> -->
 
-Predation mortality-at-age
+<!-- Predation mortality-at-age -->
 
-$M_2(L) =ae^{bL}$
+<!-- $M_2(L) =ae^{bL}$ -->
 
-where $a=2.0753+0.9914F$ and $b=-0.0276-0.1088F$
+<!-- where $a=2.0753+0.9914F$ and $b=-0.0276-0.1088F$ -->
 
-Maturity-at-weight
+<!-- Maturity-at-weight -->
 
-$O(W) = \frac{1}{1+e^{-1.67(W-28.2)}}$
+<!-- $O(W) = \frac{1}{1+e^{-1.67(W-28.2)}}$ -->
 
-## Blue whiting:
+<!-- ## Blue whiting: -->
 
-$M = 0.0221 \times B   + 0.0719$
+<!-- $M = 0.0221 \times B + 0.0719$ -->
 
-## Mass-at-age
+<!-- ## Mass-at-age -->
 
-Individual mass-at-age is assumed to vary with total biomass, estimates of mass-at-age (${W_t}^\prime$) are based on observations around $B_{MSY}$, and if total biomass increases density dependence results in a decline in the expected value of $W_a$, i.e.
+<!-- Individual mass-at-age is assumed to vary with total biomass, estimates of mass-at-age (${W_t}^\prime$) are based on observations around $B_{MSY}$, and if total biomass increases density dependence results in a decline in the expected value of $W_a$, i.e.
+-->
 
-$W_t = {W_t}^\prime a(B_t/B_{MSY})^b$
+<!-- $W_t = {W_t}^\prime a(B_t/B_{MSY})^b$ -->
 
-## Maturity-at-age
+<!-- ## Maturity-at-age -->
 
-Maturity at age is modelled by a logistic relationship
+<!-- Maturity at age is modelled by a logistic relationship -->
 
-$f(x) = \frac{1}{1+e^{-r(W-W_{50})}}$
+<!-- $f(x) = \frac{1}{1+e^{-r(W-W_{50})}}$ -->
 
-where $W_{50}$ is the mass-at-age when $50%$ of individuals are mature, and $r$ the logistic growth rate or steepness of the curve.
+<!-- where $W_{50}$ is the mass-at-age when $50%$ of individuals are mature, and $r$ the logistic growth rate or steepness of the curve. -->
 
-## Natural mortality
+<!-- ## Natural mortality -->
 
-Theoretical and empirical studies [@lorenzen2000allometry] support the existence of an allometric relationship between natural mortality ($M_a$) and body mass ($W_a$) at age of the form:
+<!-- Theoretical and empirical studies [@lorenzen2000allometry] support the existence of an allometric relationship between natural mortality ($M_a$) and body mass ($W_a$) at age of the form: -->
 
-  $M_W=M_uW^d$
+<!-- $M_W=M_uW^d$ -->
 
-where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
+<!-- where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from --0.3 to --0.37. -->
 
-If $M=0.2$ at $W_{50}$ then $M_u = 0.2/({W_{50}}^{d})$
+<!-- If $M=0.2$ at $W_{50}$ then $M_u = 0.2/({W_{50}}^{d})$ -->
 
-$d$ is approximately equal to $–1/3$, which implies that $M$ is inversely proportional to body length., since mass is approximately proportional to the third power of length.
+<!-- $d$ is approximately equal to $–1/3$, which implies that $M$ is inversely proportional to body length., since mass is approximately proportional to the third power of length. -->
 
-@10.1093/icesjms/fst226 recognised that if the von Bertalanffy growth equation
+<!-- @10.1093/icesjms/fst226 recognised that if the von Bertalanffy growth equation -->
 
-$L_t=L_{\infty}{(1-e^{(1-k(t-t_0))})}^{b(-0.305)}$
+<!-- $L_t=L_{\infty}{(1-e^{(1-k(t-t_0))})}^{b(-0.305)}$ -->
 
-and the allometric length weight relationship
+<!-- and the allometric length weight relationship -->
 
-$W = a L^b$.
+<!-- $W = a L^b$. -->
 
-are combined since
+<!-- are combined since -->
 
-$M_a = M_{\infty}{(W_a/W_{\infty})}^{-0.3}$
+<!-- $M_a = M_{\infty}{(W_a/W_{\infty})}^{-0.3}$ -->
 
-then mortality schedule can be defined by
+<!-- then mortality schedule can be defined by -->
 
-$M_a=M_{\infty}(1−e^{(−K(t−t_0))})^{−b(0.3)}$
+<!-- $M_a=M_{\infty}(1−e^{(−K(t−t_0))})^{−b(0.3)}$ -->
 
-where $M_{\infty}$ can be derived from life history characteristics @hoenig1992empirical
+<!-- where $M_{\infty}$ can be derived from life history characteristics @hoenig1992empirical -->
 
-## Multispecies modelling of predation mortality (M2)
+<!-- ## Multispecies modelling of predation mortality (M2) -->
 
-@pope2006modelling (doi:10.1016/j.icesjms.2006.04.015) proposed a way of modelling predation mortality (M2) based on size spectra, e.g.
+<!-- @pope2006modelling (<doi:10.1016/j.icesjms.2006.04.015>) proposed a way of modelling predation mortality (M2) based on size spectra, e.g. -->
 
-$N(L) = \Omega L^{\kappa}$
+<!-- $N(L) = \Omega L^{\kappa}$ -->
 
-where $N$ is numbers at length $L$ and $\Omega$ is a prefactor and $\kappa$ the scaling by length
+<!-- where $N$ is numbers at length $L$ and $\Omega$ is a prefactor and $\kappa$ the scaling by length -->
 
-This was further developed by @POPE2021105952 (https://doi.org/10.1016/j.fishres.2021.105952) by assuming that predation by all fish in the spectrum is mediated through a feeding relationship with mean $\mu$ and standard deviation $\sigma$, both common to all predators and that UM2 is a constant that relates predation pressure to mortality, and where $\lambda$ is the scaling of the feeding rate of predators by length then the predation mortality on a prey species of length $L^{\prime}$ can be written as
+<!-- This was further developed by @POPE2021105952 (<https://doi.org/10.1016/j.fishres.2021.105952>) by assuming that predation by all fish in the spectrum is mediated through a feeding relationship with mean $\mu$ and standard deviation $\sigma$, both common to all predators and that UM2 is a constant that relates predation pressure to mortality, and where $\lambda$ is the scaling of the feeding rate of predators by length then the predation mortality on a prey species of length $L^{\prime}$ can be written as -->
 
-$M2(L^\prime) = \Omega \times UM2 (L^\prime)^{(\kappa + \lambda)}\times e^{\mu((\kappa + \lambda)/3)} \times e^{{\sigma}^2((\kappa+\lambda)^2)/18}$
+<!-- $M2(L^\prime) = \Omega \times UM2 (L^\prime)^{(\kappa + \lambda)}\times e^{\mu((\kappa + \lambda)/3)} \times e^{{\sigma}^2((\kappa+\lambda)^2)/18}$ -->
 
-This simplifies to
+<!-- This simplifies to -->
 
-$M2(L) = aL^b$
+<!-- $M2(L) = aL^b$ -->
 
-and is therefore equivalent to
+<!-- and is therefore equivalent to -->
 
-  $M_W=M_uW^d$
+<!-- $M_W=M_uW^d$ -->
 
-depending on the choice of $a$ and $b$
+<!-- depending on the choice of $a$ and $b$ -->
 
-<!-- Another insight derived from size spectra studies comes from the work of Andersen et al.
+<!-- <!-- Another insight derived from size spectra studies comes from the work of Andersen et al.
 (2009).
@@
 For a particular length L of predator, predation pressure is taken to be their numbers times their length to a power $\gamma$ (if $\gamma=3$), this would be equivalent to biomass except that the condition factor linking $L^3$ to mass is included into the UM2 constant to simplify the equations).
-Hence, assuming a numbers size spectrum described by -->
+Hence, assuming a numbers size spectrum described by --> -->
+
+<!-- ### Predation mortality-at-age (M2) -->
+
+<!-- #### Pope -->
+
+<!-- @pope2006modelling (<doi:10.1016/j.icesjms.2006.04.015>) proposed a way of modelling predation mortality (M2) based on size spectra, by modelled $M$ as a function of length i.e. -->
+
+<!-- $M2(L)=\gamma(F) e^{\lambda(F) L}$ -->
+
+<!-- where Where $\gamma=2.0753+0.9914F$ and $\lambda = −0.0276−0.1088F$ and $F$ is fishing mortality -->
+
+<!-- However, the model is in mass-at-age, so we convert $L$ to $W$ -->
+
+<!-- $M2(L)=\gamma(F) e^{\lambda(F) (W/a)^{1/b}}$ -->
+
+<!-- $\gamma$ is scaled so that M at at reference length is equal to a desired value. -->
+
+<!-- A problem is that this will formulation respond immediately to a change in F.
+However, the size spectra depends on the historical not the current $F$.
+For example, for a specific value of $F$ in the current year, if the stock as been subject to a period of heavy exploitation the size spectra will be very different to that of a virgin fishery in the first year of exploitation.
+This will create problems in simulations of management as in an MSE.
+-->
+
+<!-- #### Lorenzen -->
+
+<!-- Theoretical and empirical studies [@lorenzen2000allometry] support the existence of an allometric relationship between natural mortality ($M_a$) and body mass ($W_a$) at age of the form: -->
+
+<!-- $M_W=M_uW^d$ -->
+
+<!-- where $M_W$ is $M$ at weight $W$, $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from --0.3 to --0.37. -->
+
+<!-- If $M=0.2$ at $W_{50}$ then $M_u = 0.2/({W_{50}}^{d})$ -->
+
+<!-- $d$ is approximately equal to $–1/3$, which implies that $M$ is inversely proportional to body length., since mass is approximately proportional to the third power of length. -->

### 2023-02-02 Rmd/mseMac.Rmd

@@
 ```{r, pkgs}
-library(FLCore)
-library(FLBRP)
-library(FLasher)
-library(FLAssess)
-library(FLife)
+
+library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
+library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
+library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
+library(FLAssess)  # install.packages("FLAssess", repos="http://flr-project.org/R")
+library(FLife)     # install.packages("FLife", repos="http://flr-project.org/R")
 library(ggplotFL)
@@
 source("../R/dd.R")
+source("../R/get_dropbox.r")
 ```

### 2023-02-07 Rmd/runDD.Rmd
tidy up

@@
 
-plot(eq,refpt="msy")
+p1=plot(eq,refpt="msy")
 ```
-
-**Figure `r iFig=iFig+1; iFig`** Equilibrium dynamics
 
@@
 
-plot(om)
+p2=plot(om)
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Time series dynamics
-
-```{r}
+```{r, eval=FALSE}
 plot(qmin(qmax(0.96*biomass(eq)%/%biomass(eq)[,14],0.5),1.5))+

### 2023-02-08 Rmd/condMac.Rmd
cond split buy stock

--- 2023-01-10 Rmd/runCond.Rmd
@@
 ---
-title: "**Density Dependence**"
-subtitle: "Summary"
-author: "L Kell"
+title: "**Nort East Atlantic Mackerel**"
+subtitle: "Emprirical evidence for density dependence"
+author: "L Kell & M Pastoors"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: word_document
 mathjax: TRUE
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/cond/",
-               fig.path  ="../tex/figs/cond/",
+               cache.path="../cache/mac/cond/",
+               fig.path  ="../tex/figs/mac/cond",
                fig.width =8,
@@
 
-# Blue whiting
-
-```{r, whb.data}
-load(file.path(dropboxdir,"data/om/whb.RData"))
-ts=model.frame(FLQuants(whb,tb=stock,ssb=ssb),drop=T)
-ts$tb=ts$tb/1e6
-wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
-wt =merge(wt,ts,by="year")
-wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
-wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
-wt$age=factor(wt$age,levels=sort(unique(wt$age)))
-
-mat=transform(as.data.frame(mat(whb),drop=TRUE),cohort=year-as.numeric(age))
-mat=merge(mat,ts)
-mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
-mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
-mat$age=factor(mat$age,levels=sort(unique(mat$age)))
-
-m  =transform(as.data.frame(m(whb),drop=TRUE),cohort=year-as.numeric(age))  
-m  =merge(m,ts)
-m  =ddply(m,.(age), transform, rsdl =data/mean(data))
-m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
-m$age=factor(m$age,levels=sort(unique(m$age)))
-```
-
-```{r, whb.ts, fig.width=8,fig.height=4}
-ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
-  geom_line(aes(year,data,col=quant))+
-  theme(legend.position="bottom")+
-  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
-  xlab("Year")+ylab("")
-```
-
-**Figure `r iFig=iFig+1; iFig`** Time series
-
-```{r, whb.wt, fig.width=8,fig.height=4}
-ggplot(wt)+  
-  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
-  geom_line(aes(year,data,group=age,col=age))+
-  geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
-  theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age
-
-```{r, whb.mat, fig.width=8,fig.height=4}
-ggplot(mat)+  
-  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
-  geom_line(aes(year,data,group=age,col=age))+
-  geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
-  theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
-
-```{r, whb.m, fig.width=8,fig.height=4}
-ggplot(m)+  
-  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
-  geom_line(aes(year,data,group=age,col=age))+
-  geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
-  theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
-```
-
-**Figure `r iFig=iFig+1; iFig`** M-at-age
-
-```{r, whb.wt2, fig.width=8,fig.height=5}
-#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
-#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
-a=ggplot(wt)+
-  geom_boxplot(aes(age,data))+
-  xlab("Mass-at-age")+ylab("Age")
-b=ggplot(subset(wt,rsdl<3&rsdl>0))+
-  geom_point(aes(tb,rsdl))+
-  geom_smooth(aes(tb,rsdl),method="lm")+
-  facet_wrap(age~.,ncol=3)+
-  xlab("Biomass")+ylab("Residual")
-ggarrange(a,b ,label.y=1,  
-          widths = c(4,6), heights = c(1,1,1),
-          nrow   = 1,          ncol    = 2)+
-  theme_bw(16)+
-  theme(axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.y = element_blank())
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary
-
-```{r, whb.wt.tb, fig.height=8,fig.width=8}
-ggplot(wt,aes(tb,data))+
-  geom_point()+
-  geom_smooth(se=F,method="lm")+
-  facet_wrap(~age,scale="free",ncol=3)+
-  xlab("Biomass")+ylab("Mass-at-age")
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
-
-```{r, whb.mat2, fig.width=8,fig.height=5}
-#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
-#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
-a=ggplot(mat)+
-  geom_boxplot(aes(age,data))+
-  xlab("Mass-at-age")+ylab("Age")
-b=ggplot(mat)+
-  geom_point(aes(tb,rsdl))+
-  geom_smooth(aes(tb,rsdl),method="lm")+
-  facet_wrap(age~.,ncol=3)+
-  xlab("Biomass")+ylab("Residual")
-ggarrange(a,b ,label.y=1,  
-          widths = c(4,6), heights = c(1,1,1),
-          nrow   = 1,          ncol    = 2)+
-  theme_bw(16)+
-  theme(axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.y = element_blank())
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary
-
-```{r, whb.mat.tb, fig.height=5,fig.width=8}
-ggplot(subset(mat,age%in%1:5),aes(tb,data))+
-  geom_point()+
-  geom_smooth(se=F,method="lm")+
-  facet_wrap(~age,scale="free",ncol=3)+
-  xlab("Biomass")+ylab("Mass-at-age")
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
-
-$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
-
-```{r}
-dat=model.frame(FLQuants(whb,"wt"=stock.wt,"mat"=function(x) mat(x)))
-dat$age=factor(dat$age)
-
-fn<-function(x,dat) 
-   sum((dat$mat-matFn(dat$wt,x))^2)
-
-par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
-
-ggplot(dat,aes(wt,mat))+
-  geom_point(aes(wt,mat,col=age))+
-  geom_smooth(se=F,span=0.2)+
-  xlab("Mass-at-age")+ylab("Matutrity-at-age")+
-  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
-
-```{r}
-gmr=gammV(log(data)~s(tb, bs="tp")+s(year, bs="tp")+age, data=wt)
-summary(gmr)
-plot(gmr)
-coefficients(gmr)
-
-wt=cbind(wt,hat=predict(gmr))
-
-plot(coefficients(gmr)[2:10])
-wt=merge(wt,data.frame(age=1:10,mean=c(0,coefficients(gmr)[2:10])))
-
-p=ggplot(wt)+
-  geom_line(aes(tb,exp(hat-mean),col=age))+
-  geom_point(aes(tb,data-exp(mean),col=age))+
-  xlab("Biomass")+ylab("Mass-Mean")
-```
-
-```{r}
-wt =transform(as.data.frame(stock.wt(whb),drop=TRUE),cohort=year-as.numeric(age))
-wt =merge(wt,ts,by="year")
-wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
-wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
-wt$age=factor(wt$age,levels=sort(unique(wt$age)))
-
-wt=subset(wt,age!=0)
-gmr=gammV(log(data)~s(tb/mean(tb), bs="tp")+age, data=wt)
-summary(gmr)
-plot(gmr)
-coefficients(gmr)
-
-wt=cbind(wt,hat=predict(gmr))
-
-plot(exp(coefficients(gmr)[2:12]))
-wt=merge(wt,data.frame(age=1:12,mean=exp(c(0,coefficients(gmr)[2:12]))))
-
-ggplot(wt)+
-  geom_line(aes(tb,exp(hat)/mean,col=age))+
-  geom_point(aes(tb,data/mean,col=age))+
-  xlab("Biomass")+ylab("Mass-Mean")
-
-summary(lm(hat~tb,data=transform(wt,tb=tb/mean(tb))))
-```
-
 # Funding

### 2023-02-08 Rmd/condWhb.Rmd
cond split buy stock

--- 2023-01-10 Rmd/runCond.Rmd
@@
 ---
-title: "**Density Dependence**"
-subtitle: "Summary"
-author: "L Kell"
+title: "**Blue Whiting**"
+subtitle: "Emprirical evidence for density dependence"
+author: "L Kell & M Pastoors"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: word_document
 mathjax: TRUE
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/cond/",
-               fig.path  ="../tex/figs/cond/",
+               cache.path="../cache/whb/cond/",
+               fig.path  ="../tex/figs/whb/cond",
                fig.width =8,
@@
 ```
-
-# Mackerel
-
-```{r, mac.data}
-load(file.path(dropboxdir,"data/om/mac.RData"))
-ts =as.data.frame(read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),3))
-wt =transform(as.data.frame(stock.wt(mac),drop=TRUE),cohort=year-as.numeric(age))
-wt =merge(wt,ts,by="year")
-wt=ddply(wt,.(age), transform, rsdl =data/mean(data))
-wt=ddply(wt,.(age), transform, rsdl_=rsdl/var(data)^0.5)
-wt$age=factor(wt$age,levels=sort(unique(wt$age)))
-
-mat=transform(as.data.frame(mat(mac),drop=TRUE),cohort=year-as.numeric(age))
-mat=merge(mat,ts)
-mat=ddply(mat,.(age), transform, rsdl =data/mean(data))
-mat=ddply(mat,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
-mat$age=factor(mat$age,levels=sort(unique(mat$age)))
-
-m  =transform(as.data.frame(m(mac),drop=TRUE),cohort=year-as.numeric(age))  
-m  =merge(m,ts)
-m  =ddply(m,.(age), transform, rsdl =data/mean(data))
-m  =ddply(m,.(age), transform, rsdl_=rsdl/var(data)^0.5) 
-m$age=factor(m$age,levels=sort(unique(m$age))) 
-```
-
-```{r, mac.ts, fig.width=8,fig.height=4}
-ggplot(transform(melt(ts,id="year"),year=year,quant=variable,data=value))+
-  geom_line(aes(year,data,col=quant))+
-  theme(legend.position="bottom")+
-  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
-  xlab("Year")+ylab("")
-```
-
-**Figure `r iFig=iFig+1; iFig`** Time series
-
-```{r, mac.wt, fig.width=8,fig.height=4}
-ggplot(wt)+  
-  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+
-  geom_line(aes(year,data,group=age,col=age))+
-  geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
-  theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age
-
-```{r, mac.m, fig.width=8,fig.height=4}
-ggplot(mat)+  
-  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
-  geom_line(aes(year,data,group=age,col=age))+
-  geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
-  theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
-
-```{r, mac.mat, fig.width=8,fig.height=4}
-ggplot(m)+  
-  geom_line(aes(year,data,group=cohort),linetype=3,colour="grey10")+ 
-  geom_line(aes(year,data,group=age,col=age))+
-  geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
-  theme(legend.position="bottom")+
-  scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
-```
-
-**Figure `r iFig=iFig+1; iFig`** M-at-age
-
-```{r, mac.wt2, fig.width=8,fig.height=5}
-#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
-#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
-a=ggplot(wt)+
-  geom_boxplot(aes(age,data))+
-  xlab("Mass-at-age")+ylab("Age")
-b=ggplot(subset(wt,rsdl<3&rsdl>0))+
-  geom_point(aes(tb,rsdl))+
-  geom_smooth(aes(tb,rsdl),method="lm")+
-  facet_wrap(age~.)+
-  xlab("Biomass")+ylab("Residual")
-ggarrange(a,b ,label.y=1,  
-          widths = c(4,6), heights = c(1,1,1),
-          nrow   = 1,          ncol    = 2)+
-  theme_bw(16)+
-  theme(axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.y = element_blank())
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary
-
-```{r, mac.wt.tb, fig.height=8,fig.width=8}
-ggplot(subset(wt,age!=0),aes(tb,data))+
-  geom_point()+
-  geom_smooth(se=F,method="lm")+
-  facet_wrap(~age,scale="free",ncol=3)+
-  xlab("Biomass")+ylab("Mass-at-age")
-```
-
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
-
-```{r, mac.mat2, fig.width=8,fig.height=5}
-#wt$age=factor(wt$age,levels=rev(levels(wt$age)))
-#ggbivariate(subset(wt,rsdl<5&rsdl>0), outcome="data", explanatory=c("age", "tb"))
-a=ggplot(mat)+
-  geom_boxplot(aes(age,data))+
-  xlab("Mass-at-age")+ylab("Age")
-b=ggplot(mat)+
-  geom_point(aes(tb,rsdl))+
-  geom_smooth(aes(tb,rsdl),method="lm")+
-  facet_wrap(age~.)+
-  xlab("Biomass")+ylab("Residual")
-ggarrange(a,b ,label.y=1,  
-          widths = c(4,6), heights = c(1,1,1),
-          nrow   = 1,          ncol    = 2)+
-  theme_bw(16)+
-  theme(axis.text.y = element_blank(),
-        axis.ticks.y = element_blank(),
-        axis.title.y = element_blank())
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary
-
-```{r, mac.mat.tb, fig.height=5,fig.width=8}
-ggplot(subset(mat,age%in%(1:5)),aes(tb,data))+
-  geom_point()+
-  geom_smooth(se=F,method="lm")+
-  facet_wrap(~age,scale="free",ncol=3)+
-  xlab("Biomass")+ylab("Maturity-at-age")
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v Biomass
-
-$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
-
-```{r}
-dat=model.frame(FLQuants(mac,"wt"=stock.wt,"mat"=function(x) mat(x)))
-dat$age=factor(dat$age)
-
-fn<-function(x,dat) 
-   sum((dat$mat-matFn(dat$wt,x))^2)
-
-par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
-
-ggplot(dat,aes(wt,mat))+
-  geom_point(aes(wt,mat,col=age))+
-  geom_smooth(se=F,span=0.2)+
-  xlab("Mass-at-age")+ylab("Matutrity-at-age")+
-  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
-```
-
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
-
-```{r}
-wt=subset(wt,age!=0)
-gmr=gammV(log(data)~s(tb/mean(tb), bs="tp")+age, data=wt)
-summary(gmr)
-plot(gmr)
-coefficients(gmr)
-
-wt=cbind(wt,hat=predict(gmr))
-
-plot(exp(coefficients(gmr)[2:12]))
-wt=merge(wt,data.frame(age=1:12,mean=exp(c(0,coefficients(gmr)[2:12]))))
-
-ggplot(wt)+
-  geom_line(aes(tb,exp(hat)/mean,col=age))+
-  geom_point(aes(tb,data/mean,col=age))+
-  xlab("Biomass")+ylab("Mass-Mean")
-
-summary(lm(hat~tb,data=transform(wt,tb=tb/mean(tb))))
-```
-
-# Blue whiting
 

### 2023-02-08 Rmd/condMac.Rmd
with figs as "A.x"

@@
                cache.path="../cache/mac/cond/",
-               fig.path  ="../tex/figs/mac/cond",
+               fig.path  ="../tex/figs/mac/cond/",
                fig.width =8,
@@
 
-**Figure `r iFig=iFig+1; iFig`** Time series
+**Figure A.`r iFig=iFig+1; iFig`** Time series
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age
+**Figure A.`r iFig=iFig+1; iFig`** Mass-at-age
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
+**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** M-at-age
+**Figure A.`r iFig=iFig+1; iFig`** M-at-age
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age, summary
+**Figure A.`r iFig=iFig+1; iFig`** Mass-at-age, summary
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age v Biomass
+**Figure A.`r iFig=iFig+1; iFig`** Mass-at-age v Biomass
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age, summary
+**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age, summary
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v Biomass
+**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v Biomass
 
@@
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
+**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
 

### 2023-02-08 Rmd/condWhb.Rmd
with figs as "A.x"

@@
                cache.path="../cache/whb/cond/",
-               fig.path  ="../tex/figs/whb/cond",
+               fig.path  ="../tex/figs/whb/cond/",
                fig.width =8,

### 2023-02-09 Rmd/condMac.Rmd
conditional; updated

@@
 theme_set(theme_bw(16))
-```
 
-```{r, eval=FALSE}
 source("../R/get_dropbox.R")
-```
 
-```{r}
 matFn<-function(wt,x) 
    1/(1+exp(-x[1]*(wt-x[2])))
-```
 
-```{r, dbox}
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
   dropboxdir="~/Dropbox/pelagics"
+
 ```
@@
 ```{r, mac.data}
+
 load(file.path(dropboxdir,"data/om/mac.RData"))
-ts =as.data.frame(read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),3))
+ts =as.data.frame(readxl::read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),sheet="ts"))
 wt =transform(as.data.frame(stock.wt(mac),drop=TRUE),cohort=year-as.numeric(age))
@@
 
+\newpage
+
 ```{r, mac.wt, fig.width=8,fig.height=4}
@@
 
+\newpage
+
 ```{r, mac.m, fig.width=8,fig.height=4}
@@
   geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
+  xlab("Year")+ylab("Maturity-at-age")+
   theme(legend.position="bottom")+
@@
 
+\newpage
+
 ```{r, mac.mat, fig.width=8,fig.height=4}
@@
   geom_point(aes(year,data,col=age))+
-  xlab("Year")+ylab("Mass-at-age")+
+  xlab("Year")+ylab("M-at-age")+
   theme(legend.position="bottom")+
@@
 **Figure A.`r iFig=iFig+1; iFig`** M-at-age
+
+\newpage
 
@@
 
+\newpage
+
 ```{r, mac.wt.tb, fig.height=8,fig.width=8}
@@
 
+\newpage
+
 ```{r, mac.mat2, fig.width=8,fig.height=5}
@@
   geom_boxplot(aes(age,data))+
-  xlab("Mass-at-age")+ylab("Age")
+  xlab("Maturity-at-age")+ylab("Age")
 b=ggplot(mat)+
@@
 
+\newpage
+
 ```{r, mac.mat.tb, fig.height=5,fig.width=8}
@@
 **Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v Biomass
+
+\newpage
 
@@
 **Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
+
+\newpage
 

### 2023-02-09 Rmd/ddMac.Rmd
checked

--- 2023-02-07 Rmd/runDD.Rmd
@@
 title: "**Density Dependence**"
-subtitle: "Modelling density dependence in mass, maturity and M-at-age"
+subtitle: "Mackerel: density dependence in mass, maturity and M-at-age"
 author: "L Kell"
@@
 
-Mackerel life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
+Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
 The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
@@
 
-$W_a = {\alpha}_a B^{{\beta}_a}$
+$W_a = {\overline{W_a}} B^{{\beta}_a}$
 
-Where $\alpha$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data.
+Where $\overline{W_a}$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data.
 $W_a$ is constrained so sizes do not become unfeasible large or small.
@@
 
-Maturity is modelled as a logistic function of mass-at-age i.e.
+Maturity is then  modelled as a logistic function of mass-at-age i.e.
 
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/dd/",
-               fig.path  ="../tex/figs/dd/",
+               cache.path="../cache/mac/dd/",
+               fig.path  ="../tex/figs/mac/dd/",
                fig.width =8,
@@
 
-```{r}
+```{r, cache=FALSE}
 source("../R/dd.R")
@@
 
-```{r, mac.data}
-load(file.path(dropboxdir,"data/om/mac.RData"))
-ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
-```
-
-```{r eq}
+```{r, eq}
 par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
@@
 eq =lhEql(par,m="lorenzen")
-
-p1=plot(eq,refpt="msy")
 ```
 
-```{r}
+```{r, stk}
 om=as(eq,"FLStock")
 om=fwd(om,fbar=fbar(om)[,-1],sr=eq)
-
-p2=plot(om)
 ```
 
-```{r, eval=FALSE}
-plot(qmin(qmax(0.96*biomass(eq)%/%biomass(eq)[,14],0.5),1.5))+
-  scale_y_continuous(labels = scales::percent)+
-  geom_vline(aes(xintercept=14),col="red")+
-  xlab("Biomass")+ylab("% Change in mass-at-age")
+```{r, prdFn}
+ggplot(model.frame(FLQuants(Biomass=biomass(eq),Yield=catch(eq))))+
+  geom_line(aes(Biomass,Yield))
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Density dependence multiplier, red is $B_{MSY}$
+**Figure `r iFig=iFig+1; iFig`** Yield v Biomass at equilibrium
 
 ```{r}
-save(eq,file="~/Desktop/eq.RData")
-```
+dat=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"]) 
+dat=merge(as.data.frame(dat,drop=T),as.data.frame(biomass(eq)%/%refpts(eq)["msy","biomass"],drop=T),by="year")
+names(dat)[3:4]=c("mass","biomass")
 
-```{r}
-wt=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"]) 
-ggplot(as.data.frame(wt))+           
-    geom_line(aes(age,data,col=factor(year)))+
-    theme(legend.position="none")+
-    scale_x_continuous(limits=c(0,15))+
-    xlab("Age")+ylab("Mass-at-age") 
+ggplot(dat)+           
+    geom_line(aes(age,mass,col=biomass,group=year))+
+    scale_x_continuous(limits=c(0,6))+
+    scale_y_continuous(limits=c(0,700))+
+    geom_line(aes(age,data),data=as.data.frame(stock.wt(eq)),col="red",size=2)+
+    xlab("Age")+ylab("Mass-at-age")+
+    theme(legend.position="none")
 ```
@@
 ```{r}
-parMat=FLPar(c(k=0.1, w5=111.0))
-mat=1/(1+exp(-parMat[1]*(wt-parMat[2])))
+parMat=FLPar(c(k=0.1, w5=par["a"]*par["l50"]^par["b"]))
+dat=transform(dat,mat=1/(1+exp(-parMat[1]*(mass-parMat[2]))))
 
-ggplot(as.data.frame(mat))+           
-    geom_line(aes(age,data,col=factor(year)))+
-    theme(legend.position="none")+
-    scale_x_continuous(limits=c(0,15))+
-    xlab("Age")+ylab("M-at-age")
+ggplot(dat)+           
+    geom_line(aes(age,mat,col=biomass,group=year))+
+    scale_x_continuous(limits=c(0,6))+
+    geom_line(aes(age,data),data=as.data.frame(mat(eq)),col="red",size=2)+
+    xlab("Age")+ylab("Proportion mature-at-age")+
+    theme(legend.position="none")
 ```
@@
 ```{r}
-ggplot(ddM(stock.wt(eq),fbar(eq)))+
-    geom_line(aes(age,data,col=factor(year)))+
+dat=transform(dat,m=c(par["m1"])*(mass^c(par["m2"])))
+
+ggplot(dat)+           
+    geom_line(aes(age,m,col=biomass,group=year))+
+    scale_x_continuous(limits=c(0,15))+
+    geom_line(aes(age,data),data=as.data.frame(m(eq)),col="red",size=2)+
     theme(legend.position="none")+
-    scale_x_continuous(limits=c(0,15))+
-    xlab("Age")+ylab("M-at-age") 
-```
-
-**Figure `r iFig=iFig+1; iFig`** Pope: simulation of density dependence in M-at-age
-
-```{r}
-m=par["m1"]%*%(wt%^%par["m2"])
-
-ggplot(as.data.frame(m))+           
-    geom_line(aes(age,data,col=factor(year)))+
-    theme(legend.position="none")+
-    scale_x_continuous(limits=c(0,15))+
     xlab("Age")+ylab("M-at-age") 
@@
 **Figure `r iFig=iFig+1; iFig`** Lorenzen: Simulation of density dependence in M-at-age
-
-```{r}
-save(eq,file="/home/laurie/Desktop/tmp/t.RData")
-```
 
@@
 ```{r, dd-eq-mass-mat-m}
-x=propagate(eq,length(dimnames(eq)$year))
-fbar(x)[,1]=c(seq(0,max(fbar(eq)),length.out=51),seq(1,50,length.out=51)*max(fbar(eq)))[-52]
-fbar(x)=fbar(x)[,1]
+x=propagate(eq,151)
+fbar(x)=FLQuant(c(seq(0,max(fbar(eq)),length.out=51),seq(1,100,length.out=101)*max(fbar(eq)))[-152],dimnames=list(iter=seq(151)))
 
@@
 
-```{r}
-save(eq,parMat,par,sr,file="/home/laurie/Desktop/tmp/t.RData")
-```
-
-```{r}
-load("/home/laurie/Desktop/tmp/t.RData")
-
-par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","ssb"]),b=0.3,k=0.1, w50=111.0))
+```{r, check1}
+par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","ssb"]),k=0.1,w50=par["a"]*par["l50"]^par["b"]))
 
@@
 om2=om
-lsr=landings.wt(om2)/stock.wt(om2)
-dsr=discards.wt(om2)/stock.wt(om2)
 
@@
 
+catch(om2)=computeCatch(om2,slot="all")
 plot(window(FLStocks("OM"=om,DD=om2),start=11))
@@
 
-**Figure `r iFig=iFig+1; iFig`** Mass, maturity and M-at-age: Comparison of projections with and without density dependence.
+**Figure `r iFig=iFig+1; iFig`** Check 1.
+
+```{r, check2}
+save(om2,file="~/Desktop/tmp/t.RData")
+load("~/Desktop/tmp/t.RData")
+ggplot(model.frame(FLQuants(om2[,dim(m(om2))[2]],"Biomass"=biomass,"Yield"=catch)))+
+  geom_line(aes(Biomass,Yield),col="red")+
+  geom_line(aes(Biomass,Yield),data=model.frame(FLQuants(om[,dim(m(om))[2]],"Biomass"=biomass,"Yield"=catch)))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Check 2.
+
+I think we need 4 OMs and 2 MPs
+
+    OMs are
+        OM with no DD,
+        DD in mass,
+        DD in mass + mat,
+        DD in mass, mat + m
+    MPs are HCR with
+        refpts without DD (i.e. OM 1) and
+        same as OM (for 2,3&4)
+
+So we match/mismatch, in that we get the DD hypothesis right or use current practice
+
+We condition all the OMs for M-at-age varying, and reported catch. An issue is the short-cut since if we assume no DD in the MP but there is DD in the OM we have to replicate the bias.
+
+The remaining issue is random variation in recs, mass etc.
 

### 2023-02-09 Rmd/condWhb.Rmd

@@
 
-```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
+```{r echo=FALSE, warning=FALSE, cache=FALSE, knitr, eval=TRUE}
 knitr::opts_chunk$set(echo = FALSE)
@@
 theme_set(theme_bw(16))
-```
 
-```{r, eval=FALSE}
 source("../R/get_dropbox.R")
-```
-
-```{r}
-matFn<-function(wt,x) 
-   1/(1+exp(-x[1]*(wt-x[2])))
-```
-
-```{r, dbox}
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
   dropboxdir="~/Dropbox/pelagics"
+
+matFn<-function(wt,x) 
+   1/(1+exp(-x[1]*(wt-x[2])))
+
 ```
 
-```{r, whb.data}
+```{r echo=FALSE}
+
 load(file.path(dropboxdir,"data/om/whb.RData"))
@@
   theme(legend.position="bottom")+
-  scale_color_manual("Quantity",labels=c("SSB","Total Biomass","ICES Biomass"),values=rainbow(3))+
+  scale_color_manual("Quantity",labels=c("VPA SSB","VPA Total Biomass","ICES Total Biomass"),values=rainbow(3))+
   xlab("Year")+ylab("")
@@
 **Figure A.`r iFig=iFig+1; iFig`** Time series
+
+\newpage
 
@@
 
+\newpage
+
 ```{r, whb.mat, fig.width=8,fig.height=4}
@@
 
+\newpage
+
 ```{r, whb.m, fig.width=8,fig.height=4}
@@
 
+\newpage
+
 ```{r, whb.wt.tb, fig.height=8,fig.width=8}
@@
 
+\newpage
+
 ```{r, whb.mat2, fig.width=8,fig.height=5}
@@
 
+\newpage
+
 ```{r, whb.mat.tb, fig.height=5,fig.width=8}
@@
   facet_wrap(~age,scale="free",ncol=3)+
-  xlab("Biomass")+ylab("Mass-at-age")
+  xlab("Biomass")+ylab("Maturity-at-age")
 ```
@@
 **Figure A.`r iFig=iFig+1; iFig`** Mass-at-age v Biomass
+
+\newpage
 

### 2023-02-09 R/dd.R
slight mod

@@ -1,5 +1,8 @@
-ddWt<-function(wt,biomass,bmsy,alpha=-0.25)
-  wt%*%qmin(qmax((biomass%/%bmsy)^alpha,0.1),10)
+ddWt<-function(wt,biomass,bmsy,alpha=-0.2,min=0.5,max=1.5){
+  adj=(biomass%/%bmsy)^alpha
+  
+  wt%*%qmin(qmax(adj,min),max)}
+
 
 ddMat<-function(wt,x) 
   1/(1+exp(-x[1]%*%(wt%-%x[2])))

### 2023-02-09 Rmd/ddMac.Rmd
cleaning up

@@
 ```{r, check2}
-save(om2,file="~/Desktop/tmp/t.RData")
-load("~/Desktop/tmp/t.RData")
+# save(om2,file="~/Desktop/tmp/t.RData")
+# load("~/Desktop/tmp/t.RData")
 ggplot(model.frame(FLQuants(om2[,dim(m(om2))[2]],"Biomass"=biomass,"Yield"=catch)))+

### 2023-02-09 Rmd/ddMac.Rmd
fixed ssb issue

--- 2023-02-09 Rmd/ddMac.Rmd
@@
 for (i in seq(10)){
-  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"],alpha=-0.17)
+  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"],alpha=-0.2)
   landings.wt(x)=stock.wt(x)*lsr
@@
 ggplot()+
-  geom_line(aes(ssb,catch),
-            data=model.frame(FLQuants(x, ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T))+
-  geom_line(aes(ssb,catch),
-            data=model.frame(FLQuants(eq,ssb=function(x) ssb(x), catch=function(x) catch(x)),drop=T),col="red")+
+  geom_line(aes(biomass,catch),
+            data=model.frame(FLQuants(x, biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T))+
+  geom_line(aes(biomass,catch),
+            data=model.frame(FLQuants(eq,biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T),col="red")+
   xlab("SSB")+ylab("Catch")
@@
 ```{r, check1}
-par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","ssb"]),k=0.1,w50=par["a"]*par["l50"]^par["b"]))
+par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","biomass"]),k=0.1,w50=par["a"]*par["l50"]^par["b"]))
 
-f=FLQuant(c(rep(seq(0,1,length.out=101),each=51),rep(seq(1,5,length.out=101),each=51)[-1]),
+f=FLQuant(c(rep(seq(0,1,length.out=51),each=101),rep(seq(1,25,length.out=61)[-1],each=101)),
           dimnames=list(year=dimnames(om)$year,iter=seq(101)))%*%refpts(eq)["msy","harvest"]
@@
 
-  om2=ddFn(year,om2,par)}
+  om2=ddFn(year,om2,par)
+  om2=fwd(om2,f=f[,year],sr=eq)
+  }
 
 catch(om2)=computeCatch(om2,slot="all")
-plot(window(FLStocks("OM"=om,DD=om2),start=11))
+plot(window(FLStocks("OM"=om,DD=om2),start=400))
 ```
@@
 ```{r, check2} 
-save(om2,file="~/Desktop/tmp/t.RData")
-load("~/Desktop/tmp/t.RData")
 ggplot(model.frame(FLQuants(om2[,dim(m(om2))[2]],"Biomass"=biomass,"Yield"=catch)))+  
-  geom_line(aes(Biomass,Yield),col="red")+
-  geom_line(aes(Biomass,Yield),data=model.frame(FLQuants(om[,dim(m(om))[2]],"Biomass"=biomass,"Yield"=catch)))
+  geom_line(aes(Biomass,Yield),col="black")+ 
+  geom_line(aes(Biomass,Yield),data=model.frame(FLQuants(om[,dim(m(om))[2]],"Biomass"=biomass,"Yield"=catch)),col="red")
 ```

### 2023-02-09 Rmd/runDD.Rmd

@@
 source("../R/dd.R")
-```
+source("../R/get_dropbox.R")
 
-```{r, dbox}
-dropboxdir<-try(file.path(get_dropbox(), "dd"))
+dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
 
@@
   dropboxdir="~/Dropbox/pelagics"
+
+load(file.path(dropboxdir,"data/om/mac.RData"))
+ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
+
 ```
@@
 
-```{r, mac.data}
-load(file.path(dropboxdir,"data/om/mac.RData"))
-ts =as.data.frame(read_excel("../data/inputs/dat.xlsx",3))
-```
-
 ```{r eq}
-par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
-par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
-eq =lhEql(par,m="lorenzen")
+par=FLife::teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
+par=FLife::lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
+eq =FLife::lhEql(par,m="lorenzen")
 
@@
 ```{r}
-save(eq,file="~/Desktop/eq.RData")
+# save(eq,file="~/Desktop/eq.RData")
 ```
@@
 ```{r}
-save(eq,file="/home/laurie/Desktop/tmp/t.RData")
+# save(eq,file="/home/laurie/Desktop/tmp/t.RData")
 ```
@@
 ```{r}
-save(eq,parMat,par,sr,file="/home/laurie/Desktop/tmp/t.RData")
+# save(eq,parMat,par,sr,file="/home/laurie/Desktop/tmp/t.RData")
 ```
@@
 ```{r}
-load("/home/laurie/Desktop/tmp/t.RData")
+# load("/home/laurie/Desktop/tmp/t.RData")
 

### 2023-02-09 R/dd.R
redon dd

@@ -18,7 +18,7 @@ ddFn<-function(year,x,par){
   dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
   dsr[!is.finite(dsr)]=0
   
-  stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bmsy"],alpha=-0.25)
+  stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bmsy"])
   landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
   discards.wt(x)[,year]=stock.wt(x)[,year]*dsr
   

### 2023-02-09 Rmd/ddMac.Rmd
hopefully now works

@@
   }
+catch(om2)=computeCatch(om2,slot="all")
+```
 
-catch(om2)=computeCatch(om2,slot="all")
-plot(window(FLStocks("OM"=om,DD=om2),start=400))
+```{r, check2}
+plot(window(FLStocks("OM"=om,DD=om2),start=40))
 ```
@@
 
-```{r, check2} 
+```{r, check3} 
 ggplot(model.frame(FLQuants(om2[,dim(m(om2))[2]],"Biomass"=biomass,"Yield"=catch)))+  

### 2023-02-09 Rmd/ddMac.Rmd
works!

@@
 for (year in dimnames(om2)$year[-1]){
-
-  #stock.wt(   om2)[,year]=ddWt(stock.wt(om2)[,year],biomass(om2)[,ac(an(year))],par["bmsy"])
-  #landings.wt(om2)[,year]=stock.wt(om2)[,year]*lsr[,year]
-  #discards.wt(om2)[,year]=stock.wt(om2)[,year]*dsr[,year]
-
-  #mat(om2)[,year]=1/(1+exp(-par["k"]*(stock.wt(om2)[,year]-par["w50"])))
-  #m(  om2)[,year]=par["m1"]%*%(stock.wt(om2)[,year]%^%par["m2"])
-
-  #om2=fwd(om2,f=f[,year],sr=eq)}
-
   om2=ddFn(year,om2,par)

### 2023-02-09 Rmd/ddWhb.Rmd
copied from ddMac

--- 2023-02-09 Rmd/ddMac.Rmd
@@
 title: "**Density Dependence**"
-subtitle: "Mackerel: density dependence in mass, maturity and M-at-age"
+subtitle: "Blue Whiting: density dependence in mass, maturity and M-at-age"
 author: "L Kell"
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/mac/dd/",
-               fig.path  ="../tex/figs/mac/dd/",
+               cache.path="../cache/whb/dd/",
+               fig.path  ="../tex/figs/whbdata/dd/",
                fig.width =8,
@@
 ```{r, eq}
-par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
+par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Micromesistius poutassou"]
 par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)

### 2023-02-09 Rmd/report.R
started on report.R to generate all plots

@@ -1,0 +1,172 @@
+# ------------------------------------------------------------------------------
+# report.R
+#
+# Generating DDMSE results to be included into Rmd files
+#
+# Laurie Kell and Martin Pastoors
+#
+# 09/02/2023 First coding
+# ------------------------------------------------------------------------------
+
+rm(list=ls())
+
+iFig=0
+iTab=0
+
+library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
+library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
+library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
+library(FLAssess)  # install.packages("FLAssess", repos="http://flr-project.org/R")
+library(FLife)     # install.packages("FLife", repos="http://flr-project.org/R")
+library(ggplotFL)
+
+library(popbio)
+
+library(GGally)
+library(ggpubr)
+
+library(rjson) # install.packages("rjson"); used for dropbox function
+library(RJSONIO)
+
+library(plyr)
+library(dplyr)
+library(reshape)
+
+library(readxl)
+  
+library(mgcv)
+library(mgcViz)
+
+library(tidyverse)
+
+theme_set(theme_bw(16))
+
+source("R/dd.R")
+source("R/get_dropbox.r")
+source("R/pm.R")
+
+dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
+
+if ("try-error" %in% is(dropboxdir))
+  dropboxdir="~/Dropbox/pelagics"
+
+myparams <- data.frame(
+  stock = c("mac"            ,  "whb"            ),
+  m1    = c(0.15/(0.2^-0.288),  0.2/(300^-0.288) ),
+  m2    = c(-0.28            ,  -0.28            ),
+  m3    = c(NA               ,  NA               ),
+  blim  = c(2000000          , 1500000           ),
+  btrig = c(2580000          , 2250000           ),
+  bmsy  = c(3500000          ,  NA               ),
+  b     = c(0.3              ,  NA               ),
+  k     = c(30.1             ,  NA               ), 
+  w50   = c(0.2              ,  NA               ),
+  wt1   = c(0.01             ,  NA               ))
+
[…]

### 2023-02-09 Rmd/report whb.Rmd
markdown reports mac and whb

---
output:
  word_document:
    reference_docx: ../report_template.dotx
mathjax: TRUE
---

```{r, knitr, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

knitr::opts_chunk$set(
               cache     =TRUE, 
               comment   ="", 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               crop      =TRUE,
               cache.path="../cache/figs/cond/",
               fig.path  ="../tex/figs",
               fig.width =10,
               fig.height=8,
               dev       ="png")

rm(list=ls())

iFig=0
iTab=0

library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
library(FLAssess)  # install.packages("FLAssess", repos="http://flr-project.org/R")
library(FLife)     # install.packages("FLife", repos="http://flr-project.org/R")
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

library(tidyverse)

theme_set(theme_bw(16))

source("../R/dd.R")
source("../R/get_dropbox.r")
source("../R/pm.R")

dropboxdir<-try(file.path(get_dropbox(), "pelagics"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"

# mystk     <- "whb"; mystkname <- "Blue whiting"
# mystk     <- "her"; mystkname <- "North Sea herring"
mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; maxyear <- 2050

```

# Evaluating impacts of Density Dependent processes on `r mystkname` MSE

**L. Kell, M.A. Pastoors

`r format(Sys.time(), '%d/%m/%Y')`

## Introduction

## OM conditioning

Stock assessments with SAM, Density dependent SAM and Density dependent VPA

```{r, mac.data, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

load(file.path(dropboxdir,"data/om/mac.RData"))
combY <- as.data.frame(pmYear(window(mac, start=2001)),drop=T) %>% mutate(stk=mystk, type="SAM")

par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bmsy=3500000,b=0.3,k=30.1, w50=0.2))

stock.wt(   mac)[1]=0.01
catch.wt(   mac)[1]=0.01
landings.wt(mac)[1]=0.01
discards.wt(mac)[1]=0.01

om_orig=ddFn2(mac,par); 
combY <- bind_rows(combY,
                   as.data.frame(pmYear(window(om_orig, start=2001)),drop=T) %>% mutate(stk=mystk, type="DD SAM"))

om=om_orig+VPA(om_orig)
om=ddFn2(om,par)
combY <- bind_rows(combY, 
                   as.data.frame(pmYear(window(om, start=2001)),drop=T) %>% mutate(stk=mystk, type="DD VPA") )

combY %>% 
  filter(qname %in% c("catch","ssb","rec","fbar")) %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  geom_ribbon(aes(fill=type, ymin=lowq, ymax=uppq, xmin=year, xmax=year), alpha=0.3) +
  geom_line(aes(colour=type)) +
  facet_grid(qname~., scales="free_y")

```

\newpage

Forward projections with F=fmsy (fwd), 0.8xfmsy?, 1.0xfmsy, 1.2xfmsy?

```{r, mac.data2, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
eq=FLBRP(om,sr=sr)

om=fwdWindow(om,end=maxyear,eq)
f =fbar(om)[,ac(2021:maxyear)]%=%refpts(eq)["msy","harvest"]
om=fwd(om,fbar=f,sr=eq)

om2=om
om3=om
om4=om

for (iYear in ac(2021:maxyear)){
  om2=ddFn(iYear,om2,par)
  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)

  om3=ddFn(iYear,om3,par)
  om3=fwd(om3,fbar=f[,iYear],sr=eq)

  om4=ddFn(iYear,om4,par)
  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
  }

plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))

```

## Simulation results

## Discussion and conclusions

ICES
Fmsy
HCR1 (no bounds)

```{r, runWhb1, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar", "rec")
scenarios <- c("ICES", "Fmsy", "HCR1")
rp        <- refs  %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data)) 
# box <- 
#   bind_rows(
#     df   %>% dplyr::select(.id, qname, stk, data=median),
#     ices %>% dplyr::select(qname, stk, data=median) %>% mutate(.id="ICES")
#   ) %>% 
#   group_by(qname, stk) %>% 
#   summarise(max = max(data, na.rm=TRUE), min = min(data, na.rm=TRUE)) %>% 
#   mutate(min = ifelse(min<0, min, 0)) %>%

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median),
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq),
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices,
            aes(year,median),
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp,
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

```

**Figure `r mystk`.2**  `r mystkname`. Comparison of different HCR’s without TAC bounds

\newpage

#### `r mystkname`: comparing different HCR’s with and without TAC bounds

Fmsy
HCR0 (no bounds)
HCR1 bnd (-25%/+20% TAC change bounds apply when stock is above MSY Btrigger)
HCR2 bnd (-25%/+20% TAC change bounds apply when stock is above Blim)

```{r, runWhb2, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("Fmsy", "HCR1",  "HCR1 bnd", "HCR2 bnd")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
    mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

```

**Figure `r mystk`.3**  `r mystkname`. Comparison of different HCR’s with and without TAC bounds

\newpage

#### `r mystkname`: comparing HCR’s with and without a cap on TAC bounds

HCR1 bnd
HCR1 bnd cap
HCR2 bnd
HCR2 bnd cap

```{r, runWhb3, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("HCR1 bnd", "HCR1 bnd cap", "HCR2 bnd", "HCR2 bnd cap")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
    mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +
  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

```

**Figure `r mystk`.4**  `r mystkname`. Comparison of different HCR’s with and without a cap on TAC bounds

\newpage

#### `r mystkname`: comparing HCR’s with and without implementation error (without bounds on TAC change)

HCR0
HCR1 10%
HCR1 20%
HCR1 30%
HCR1 10% recent
HCR1 10% random recent

```{r, runWhb4, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("HCR1", "HCR1 +10%", "HCR1 +20%", "HCR1 +30%", "HCR1 +10% recent", 
               "HCR1 +random recent")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

```

**Figure `r mystk`.5**  `r mystkname`. Comparison of different HCR’s with and without implementation error (without bounds on TAC change)

\newpage

#### `r mystkname`: comparing HCR’s with and without implementation error in combination with bounds on TAC change

HCR1 bnd
HCR1 10% bnd
HCR1 10% recent bnd
HCR1 10% random bnd
HCR1 10% random recent bnd

```{r, runWhb5, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch","cCum", "ssb","fbar")
scenarios <- c("HCR1 bnd", 
               "HCR1 +10% bnd",
               "HCR1 +10% bnd recent",
               "HCR1 +random bnd", 
               "HCR1 +random bnd recent")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  # vline at 2010
  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +

  # iters
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +

  # median and ribbon
  geom_line(aes(y=median), 
            linewidth=1, colour="red") +
  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
              alpha=0.2, fill="red") +

  # ICES
  geom_line(data= ices, aes(year,median), 
            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +

  # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +

  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

```

**Figure `r mystk`.6**  `r mystkname`. Comparison of different HCR’s with and without implementation error in combination with bounds on TAC change

\newpage

#### `r mystkname`: checking the working of different constraints

```{r, whbcheck, echo=FALSE, fig.asp=.5, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("ssb")
# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10%", "HCR1 +10% bnd", "HCR1 +10% bnd cap")
scenarios <- c("HCR1", "HCR1 bnd", "HCR2", "HCR2 bnd")
#scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10% bnd", "HCR1 +20% bnd", "HCR1 +30% bnd")
rp        <- refs %>% filter(stk == mystk, refpoint %in% c("msybtrigger", "blim"))
ices      <- combY %>% filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% dplyr::select(- .id)
df        <- combY %>% filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  left_join(dplyr::select(rp,
                          stk, year, qname, refpoint, reftype, rp=data), 
            by=c("stk", "year", "qname")) %>% 
  mutate(trigger = ifelse(median >= rp, "above","below"))
iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 10) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  left_join(dplyr::select(rp,
                          stk, year, qname, refpoint, reftype, rp=data), 
            by=c("stk", "year", "qname")) %>% 
  mutate(trigger = ifelse(data >= rp, "above","below"))
df %>% 
  ggplot(aes(x=year, y=median)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
  geom_line(data=iters,
            aes(y=data, group=iter),
            linewidth=0.4, colour="darkgray") +
  geom_point(data=filter(iters,
                         trigger=="above"),
            aes(y=data, group=iter),
            size=1.0, colour="black") +

  geom_line(linewidth=1, colour="red") +
  geom_point(data=filter(df,
                        trigger=="above"),
            colour="red", size=2) +
    # refpoints
  geom_line(data=rp, 
            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
            linewidth=1) +
  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
  labs(y="ssb (tonnes)") +
  expand_limits(y=0) +
  facet_grid(qname ~ .id, scales="free_y")

myvar     <- c("catch")
# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10%", "HCR1 +10% bnd", "HCR1 +10% bnd cap")
# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10% bnd", "HCR1 +20% bnd", "HCR1 +30% bnd")
rp        <- refs %>% 
  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
  mutate(qname = factor(qname, levels=myvar))
ices      <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
  dplyr::select(- .id) %>% 
  mutate(qname = factor(qname, levels=myvar))
df        <- combY %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(qname = factor(qname, levels=myvar)) %>% 
  group_by(stk, .id) %>% 
  mutate(perc_diff = (median-lag(median))/lag(median))

iters        <- combYI %>% 
  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  group_by(stk, .id, iter) %>% 
  mutate(perc_diff = (data-lag(data))/lag(data))

df %>% 
  ggplot(aes(x=year, y=perc_diff)) +
  theme_bw() +
  theme(panel.grid.minor = element_blank(),  
        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+

  geom_line(data=iters,
            aes(x=year, y=perc_diff, group=iter),
            linewidth=.4, colour="darkgray", inherit.aes = FALSE) +

  geom_line(linewidth=1, colour="red") +

  geom_hline(yintercept = 0.25, colour="black", linewidth=1, linetype=2) +
  geom_hline(yintercept = -0.2, colour="black", linewidth=1, linetype=2) +
  expand_limits(y=0) +
  scale_y_continuous(labels=scales::percent, limits=c(-2,2)) +
  facet_grid(qname ~ .id, scales="free_y")

```

**Figure `r mystk`.6**  `r mystkname`. Comparison of the working of constraints on TAC change relative to stock size and reference point MSY Btrigger. Black dots indicate estimated stock size on individual runs above MSY Btrigger (i.e. TAC contstraint applies). Red dots indicate median stock size above MSY Btrigger.

\newpage

#### `r mystkname`: comparing all scenario's on median catch, IAV and AAV

```{r, whbsmry, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

myvar     <- c("catch.50", "iav.50", "aav")
scenarios <- c("ICES","Fmsy","HCR1","HCR1 bnd","HCR1 bnd cap", "HCR1 +10% bnd","HCR1 +10% bnd cap",
               "HCR1 +10% bnd recent","HCR1 +random bnd",",HCR1 +random bnd recent","HCR1 +10%",
               "HCR1 +10% recent","HCR1 +random","HCR1 +random recent","HCR1 +20% bnd", 
               "HCR1 +20% bnd cap","HCR1 +20% bnd recent","HCR1 +20%","HCR1 +20% recent",
               "HCR1 +30% bnd","HCR1 +30% bnd cap","HCR1 +30% bnd recent","HCR1 +30%",
               "HCR1 +30% recent",
               "HCR2","HCR2 bnd","HCR2 bnd cap","HCR2 +10% bnd","HCR2 +10% bnd cap",
               "HCR2 +10% bnd recent","HCR2 +random bnd","HCR2 +random bnd recent","HCR2 +10%",
               "HCR2 +10% recent","HCR2 +random","HCR2 +random recent","HCR2 +20% bnd",
               "HCR2 +20% bnd cap","HCR2 +20% bnd recent","HCR2 +20%","HCR2 +20% recent",
               "HCR2 +30% bnd","HCR2 +30% bnd cap","HCR2 +30% bnd recent","HCR2 +30%",
               "HCR2 +30% recent")

df        <- combI %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", myvar) %>% 
  filter(stk == mystk,  .id %in% scenarios) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(variable = factor(variable, levels=myvar))

calc_boxplot_stat <- function(x) {
  coef <- 1.5
  n <- sum(!is.na(x))
  # calculate quantiles
  stats <- quantile(x, probs = c(0.0, 0.25, 0.5, 0.75, 1.0))
  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
  iqr <- diff(stats[c(2, 4)])
  # set whiskers
  outliers <- x < (stats[2] - coef * iqr) | x > (stats[4] + coef * iqr)
  if (any(outliers)) {
    stats[c(1, 5)] <- range(c(stats[2:4], x[!outliers]), na.rm = TRUE)
  }
  return(stats)
}

df %>% 
  ggplot(aes(x=.id, y=data)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +
  expand_limits(y=0) +
  # scale_y_continuous(limits = quantile(t$data, c(0.1, 0.9))) +
  labs(x="", y="") +
  facet_grid(variable~., scales="free", space="free_x")

```

**Figure `r mystk`.7**  `r mystkname`. Comparison of all scenario's on median catch, median IAV (Interannual variability in catch) and AAV (Average Annual Variability in catch)

```{r smry2, eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, whbsmry, echo=FALSE}

mystk     <- c("whb","her","mac")
myvar     <- c("catch.50", "iav.50", "aav")
scenarios <- c("ICES","Fmsy","HCR1","HCR1 bnd","HCR1 bnd cap", "HCR1 +10% bnd","HCR1 +10% bnd cap",
               "HCR1 +10% bnd recent","HCR1 +random bnd",",HCR1 +random bnd recent","HCR1 +10%",
               "HCR1 +10% recent","HCR1 +random","HCR1 +random recent","HCR1 +20% bnd", 
               "HCR1 +20% bnd cap","HCR1 +20% bnd recent","HCR1 +20%","HCR1 +20% recent",
               "HCR1 +30% bnd","HCR1 +30% bnd cap","HCR1 +30% bnd recent","HCR1 +30%",
               "HCR1 +30% recent",
               "HCR2","HCR2 bnd","HCR2 bnd cap","HCR2 +10% bnd","HCR2 +10% bnd cap",
               "HCR2 +10% bnd recent","HCR2 +random bnd","HCR2 +random bnd recent","HCR2 +10%",
               "HCR2 +10% recent","HCR2 +random","HCR2 +random recent","HCR2 +20% bnd",
               "HCR2 +20% bnd cap","HCR2 +20% bnd recent","HCR2 +20%","HCR2 +20% recent",
               "HCR2 +30% bnd","HCR2 +30% bnd cap","HCR2 +30% bnd recent","HCR2 +30%",
               "HCR2 +30% recent")

df        <- combI %>% 
  tidyr::pivot_longer(names_to = "variable", values_to = "data", myvar) %>% 
  filter(stk %in% mystk,  .id %in% scenarios) %>% 
  mutate(.id = factor(.id, levels=scenarios)) %>% 
  mutate(variable = factor(variable, levels=myvar)) %>% 
  mutate(stk = factor(stk, levels=mystk))

calc_boxplot_stat <- function(x) {
  coef <- 1.5
  n <- sum(!is.na(x))
  # calculate quantiles
  stats <- quantile(x, probs = c(0.0, 0.25, 0.5, 0.75, 1.0))
  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
  iqr <- diff(stats[c(2, 4)])
  # set whiskers
  outliers <- x < (stats[2] - coef * iqr) | x > (stats[4] + coef * iqr)
  if (any(outliers)) {
    stats[c(1, 5)] <- range(c(stats[2:4], x[!outliers]), na.rm = TRUE)
  }
  return(stats)
}

df %>% 
  ggplot(aes(x=.id, y=data)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +
  expand_limits(y=0) +
  # scale_y_continuous(limits = quantile(t$data, c(0.1, 0.9))) +
  labs(x="", y="") +
  facet_grid(variable~stk, scales="free", space="free_x")

```

### 2023-02-10 Rmd/report mac.Rmd
working on report.r

---
output:
  word_document:
    reference_docx: ../report_template.dotx
mathjax: TRUE
---

```{r, knitr, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
library(knitr)

knitr::opts_chunk$set(
               cache     =TRUE, 
               comment   ="", 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,
               crop      =TRUE,
               cache.path="../cache/figs/cond/",
               fig.path  ="../tex/figs",
               fig.width =10,
               fig.height=8,
               dev       ="png")

rm(list=ls())

iFig=0
iTab=0

library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
library(FLAssess)  # install.packages("FLAssess", repos="http://flr-project.org/R")
library(FLife)     # install.packages("FLife", repos="http://flr-project.org/R")
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

library(tidyverse)

theme_set(theme_bw(16))

library(captioner)     # captions
tab_nums <- captioner::captioner(prefix = "Table", levels=3, type=c("n","n","n"), infix=".")
fig_nums <- captioner::captioner(prefix = "Figure", levels=3, type=c("n","n","n"), infix=".")

source("../R/dd.R")
source("../R/get_dropbox.r")
source("../R/pm.R")

dropboxdir<-try(file.path(get_dropbox(), "pelagics"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"

mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
# mystk     <- "whb"; mystkname <- "Blue whiting";                mylatin <- "Micromesistius poutassou" 
maxyear <- 2050

# ------------------------------------------------------------------------------
# Operating model
# ------------------------------------------------------------------------------
resultsdir <- file.path(dropboxdir, "results", mystk, "figures")

```

<!-- ## Mackerel report -->

**Evaluating impacts of Density Dependent processes on `r mystkname` via MSE simulations**

**L. Kell, M.A. Pastoors**

`r format(Sys.time(), '%d/%m/%Y')`

Commissioned by: University of Copenhagen (Henrik Sparholt)

# Executive summary

[summary here]

# Introduction

The purpose of this work is to support the project “Fmsy-project for six high profile fish stocks”. The project focuses on the current management approach for fish stocks based on single-species Fmsy Values where the only density-dependent process that is included is recruitment. The project includes three other density-dependent population dynamics: growth (weight at age), sexual maturity and natural mortality.

This report deals with the evaluation of density dependent processes for `r mystkname` using Management Strategy Evaluation (MSE). Four alternatives MSE's have been included:

1. No density dependent processes
2. DD in growth
3. DD in growth and maturity
4. DD in growth, maturity and natural mortality

All MSE's have been developed within the FLR framework [ref] and all the code is stored on github [ref].

All data and results are available on DropBox [ref]

# On density-dependent processes

Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age. The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.

## Mass-at-age [SHOULD WE CALL IT WEIGHT AT AGE, SINCE WE ARE USING Wa? ]

Mass-at-age ($W_a$) is modelled as

$W_a = {\overline{W_a}} B^{{\beta}_a}$

Where $\overline{W_a}$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data. $W_a$ is constrained so sizes do not become unfeasible large or small.

```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "sim_dd_mass", level = 1, display = FALSE,
  caption = "Simulation of density dependence in mass-at-age for different levels of biomass relative to Bmsy")

knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_mass.jpg")))

```

_`r fig_nums("sim_dd_mass")`_

## Maturity-at-weight

Maturity is then  modelled as a logistic function of mass-at-age i.e.

$O(W_a) = 1/(1+e^{k(W_a-W_{50})})$

```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "sim_dd_mat", level = 1, display = FALSE,
  caption = "Simulation of density dependence in maturity-at-age for different levels of biomass relative to Bmsy")

knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_maturity.jpg")))

```

_`r fig_nums("sim_dd_mat")`_

## Natural mortality

Is assumed to vary at unit length ($L_1$) e.g.

$ln(M_1) =\alpha + \beta ln(L_{\infty}) + \gamma ln(k)$

where $\alpha=0.65$, $\beta=0.91$, $\gamma=0.87$

Length-at-age ($L_a$) can be obtained from the length-weight relationship

$L_a=(W/a)^{1/b}$

giving

$ln(M_1) =(W/a)^{1/b}(\alpha + \beta ln(L_{\infty}) + \gamma ln(k))$

For DD there is a relationship between Mass at age $M_a$ and total biomass ($B$)

Mackerel:

$M = 0.0232B + 0.0524$

```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "sim_dd_m", level = 1, display = FALSE,
  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")

knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))

```

_`r fig_nums("sim_dd_m")`_

Comparison of equilibrium yield plots for simulation with and without density-dependent processes

```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

fig_nums(
  name    = "sim_yield", level = 1, display = FALSE,
  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")

knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))

```

_`r fig_nums("sim_dd_m")`_

# OM conditioning

Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA. The VPA reconstructions of the SAM assessment, shows a strongly divergent pattern in recruitment and also some differences in the trend of SSB. Mean fishing mortality is somewhat more variable in the VPA reconstruction compared to the SAM assessment. Density-dependent processes (on weight, maturity and M) tend to lead to lower SSB in the latter part of the time series.

```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_om.jpg")))

```

\newpage

Forward projections with F=fmsy (fwd), 0.8xfmsy?, 1.0xfmsy, 1.2xfmsy?

```{r, mac.data2, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}

# sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
# eq=FLBRP(om,sr=sr)
# 
# om=fwdWindow(om,end=maxyear,eq)
# f =fbar(om)[,ac(2021:maxyear)]%=%refpts(eq)["msy","harvest"]
# om=fwd(om,fbar=f,sr=eq)
# 
# om2=om
# om3=om
# om4=om
# 
# for (iYear in ac(2021:maxyear)){
#   om2=ddFn(iYear,om2,par)
#   om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)
# 
#   om3=ddFn(iYear,om3,par)
#   om3=fwd(om3,fbar=f[,iYear],sr=eq)
#   
#   om4=ddFn(iYear,om4,par)
#   om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
#   }
# 
# plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))

```

# Simulation results

# Discussion and conclusions

The quality shall live up to that demanded by the International Council for the Exploration of the Sea (ICES) and potentially be used in ICES advice to fisheries managers. The Consultant shall be interacting with the leadership of the project Henrik Sparholt, or his representative, in the process of developing the software. The software shall be made publicly available via Github or on alternative platform if so directed by Henrik Sparholt or his representative.

# Funding

# References

# Annex 1: OM conditioning

# Annex 2: Density dependent processes

### 2023-02-10 Rmd/report.R

@@ -61,17 +61,22 @@ myparams <- data.frame(
   b     = c(0.3              ,  NA               ),
   k     = c(30.1             ,  NA               ), 
   w50   = c(0.2              ,  NA               ),
-  wt1   = c(0.01             ,  NA               ))
+  wt1   = c(0.01             ,  NA               ),
+  maxage= c(12               ,  10               ))
+
+save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
 mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scomber scombrus" 
 # mystk     <- "whb"; mystkname <- "Blue whiting";                mylatin <- "Micromesistius poutassou" 
 maxyear <- 2050
 
+resultsdir <- file.path(dropboxdir, "results", mystk, "figures")
 
 # ------------------------------------------------------------------------------
 # Operating model
 # ------------------------------------------------------------------------------
-resultsdir <- file.path(dropboxdir, "results", mystk, "figures")
+
+maxage <-  myparams[myparams$stock==mystk,"maxage"]
 
 btrig_blim <- myparams[myparams$stock==mystk,"btrig"]/myparams[myparams$stock==mystk,"blim"] #mac 1.29; whb 1.5
   
@@ -132,8 +137,186 @@ ggsave(p,
        width=10, height=10, units="in")
 
 # ------------------------------------------------------------------------------
+# Reference points
+# ------------------------------------------------------------------------------
+
+# ------------------------------------------------------------------------------
 # Density dependence: assumptions and estimation
 # ------------------------------------------------------------------------------
+
+# HOW DO THESE PARAMETERS RELATE TO MYPARAMS; e.g. b??
+
+# Get theoretical values from literature
+par=FLife::teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter==mylatin]
+
+# Uses life history theory to derive parameters for biological relationships
+par=FLife::lhPar(par,
[…]

### 2023-02-13 Rmd/report mac.Rmd
Report.R & report MAC.rmd

@@
 
+<!-- =============================================================================================================== -->
 <!-- ## Mackerel report                                                                                              -->
+<!-- =============================================================================================================== -->
 
@@
 
+<!-- =============================================================================================================== -->
+<!-- Section 1: introduction                                                                                         -->
+<!-- =============================================================================================================== -->
+
 # Introduction
@@
 
-This report deals with the evaluation of density dependent processes for `r mystkname` using Management Strategy Evaluation (MSE). Four alternatives MSE's have been included:
+This report deals with the evaluation of density dependent processes for `r mystkname` using Management Strategy Evaluation (MSE). Steps in the analysis:
+
+1. Conditioning and summaries based on the 2022 ICES assessment (section 2)
+2. Estimating Density Dependent effects based on equilibrium analyses and explore DD before setting up the OMs (section 3)
+3. Setting up the Operating Models (OM) (section 4)
+4. Running MSE's (section 5)
+
+Four alternatives OMs have been included in the analyses:
 
@@
 
-All MSE's have been developed within the FLR framework [ref] and all the code is stored on github [ref].
+All MSE's have been developed within the FLR framework (https://flr-project.org/) and all the code is stored on github (https://github.com/martinpastoors/ddmse).
 
-All data and results are available on DropBox [ref]
+All data and results are available on DropBox (https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0)
 
-# On density-dependent processes
+<!-- =============================================================================================================== -->
+<!-- Section 2: conditioning and summaries of ICES assessment                                                        -->
+<!-- =============================================================================================================== -->
+
+# Conditioning and summaries based on the 2022 ICES assessment
+
+<!-- =============================================================================================================== -->
+<!-- Section 3: Estimating Density Dependent effects based on equilibrium analyses                                   -->
+<!-- =============================================================================================================== -->
+
+# Estimating Density Dependent effects based on equilibrium analyses
 
@@
 
-knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))
+knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_eq_yield.jpg")))
 
@@
 
-# OM conditioning
+<!-- =============================================================================================================== -->
+<!-- Section 4: Setting up the Operating Models (OM)                                                                 -->
+<!-- =============================================================================================================== -->
+
+# Setting up the Operating Models (OM)
 
@@
 
-```{r, mac.data2, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-# sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
-# eq=FLBRP(om,sr=sr)
-# 
-# om=fwdWindow(om,end=maxyear,eq)
-# f =fbar(om)[,ac(2021:maxyear)]%=%refpts(eq)["msy","harvest"]
-# om=fwd(om,fbar=f,sr=eq)
-# 
-# om2=om
-# om3=om
-# om4=om
-# 
-# for (iYear in ac(2021:maxyear)){
-#   om2=ddFn(iYear,om2,par)
-#   om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)
-# 
-#   om3=ddFn(iYear,om3,par)
-#   om3=fwd(om3,fbar=f[,iYear],sr=eq)
-#   
-#   om4=ddFn(iYear,om4,par)
-#   om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
-#   }
-# 
-# plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-# Simulation results
+<!-- =============================================================================================================== -->
+<!-- Section 5: Running MSE's                                                                                        -->
+<!-- =============================================================================================================== -->
+
+# Running MSE's
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+```
+
+<!-- =============================================================================================================== -->
+<!-- Section 6: Discussion and conclusions                                                                           -->
+<!-- =============================================================================================================== -->
 
@@
 
+<!-- =============================================================================================================== -->
+<!-- Funding                                                                                                         -->
+<!-- =============================================================================================================== -->
+
 # Funding
+
+<!-- =============================================================================================================== -->
+<!-- References                                                                                                      -->
+<!-- =============================================================================================================== -->
 
@@
 
+<!-- =============================================================================================================== -->
+<!-- Annex 1: OM conditioning                                                                           -->
+<!-- =============================================================================================================== -->
+
 # Annex 1: OM conditioning
 
+<!-- =============================================================================================================== -->
+<!-- Annex 2: Density dependent processes                                                                           -->
+<!-- =============================================================================================================== -->
+
 # Annex 2: Density dependent processes

### 2023-02-13 Rmd/report.R

@@ -224,9 +224,8 @@ ggsave(ggplot(dat)+
        width=10, height=10, units="in")
 
 
-# Comparison of equilibrium yields curves without DD, with DD mass, DD mass+maturity and DD mass+mat+M 
+# Comparison of equilibrium yields curves without DD and with DD mass
 
-# First by mass
 x         = propagate(eq,length(dimnames(eq)$year))
 fbar(x)   = fbar(x)[,1]
 fbar(x)[] = c(fbar(eq))*2
@@ -248,7 +247,8 @@ df <-
     model.frame(FLQuants(x, biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T) %>% mutate(scen="DD mass")
   )
 
-# then by mass and maturity
+# Comparison of equilibrium yields curves with DD mass+maturity
+
 x=propagate(eq,length(dimnames(eq)$year))
 fbar(x)[,1]=c(seq(0,max(fbar(eq)),length.out=51),seq(1,50,length.out=51)*max(fbar(eq)))[-52]
 fbar(x)=fbar(x)[,1]
@@ -271,7 +271,8 @@ df <-
     model.frame(FLQuants(x, biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T) %>% mutate(scen="DD mass+mat")
   )
 
-# then by mass and maturity and M
+# Comparison of equilibrium yields curves with DD mass+mat+M 
+
 x=propagate(eq,length(dimnames(eq)$year))
 fbar(x)[,1]=c(seq(0,max(fbar(eq)),length.out=51),seq(1,50,length.out=51)*max(fbar(eq)))[-52]
 fbar(x)=fbar(x)[,1]

### 2023-02-13 Rmd/condMac.Rmd
clean

@@
 title: "**Nort East Atlantic Mackerel**"
-subtitle: "Emprirical evidence for density dependence"
+subtitle: "Mackerel: Emprirical evidence for density dependence"
 author: "L Kell & M Pastoors"
@@
 theme_set(theme_bw(16))
+```
 
+```{r, dropbox}
 source("../R/get_dropbox.R")
-
-matFn<-function(wt,x) 
-   1/(1+exp(-x[1]*(wt-x[2])))
 
@@
   dropboxdir="~/Dropbox/pelagics"
-
 ```
@@
 ```{r}
+matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))
+
 dat=model.frame(FLQuants(mac,"wt"=stock.wt,"mat"=function(x) mat(x)))
@@
 
-fn<-function(x,dat) 
-   sum((dat$mat-matFn(dat$wt,x))^2)
+fn<-function(x,dat) {
+  sum((dat$mat-matFn(dat$wt,x))^2)}
 
-par=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
+matPar=optim(par=c(k=0.5,w50=.2), fn,dat=dat)
 
@@
   xlab("Mass-at-age")+ylab("Matutrity-at-age")+
-  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,par$par),wt=dat$wt))
+  geom_line(aes(wt,mat),col="red",data=data.frame(mat=matFn(dat$wt,matPar$par),wt=dat$wt))
 ```
 
-**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r par$par[1]` $W_{50}$=`r par$par[2]`)
+**Figure A.`r iFig=iFig+1; iFig`** Maturity-at-age v mass-at-age  (k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`)
 

### 2023-02-13 Rmd/condWhb.Rmd
clean

@@
   dropboxdir="~/Dropbox/pelagics"
-
-matFn<-function(wt,x) 
-   1/(1+exp(-x[1]*(wt-x[2])))
-
 ```
@@
 ```{r}
+matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))
+
 dat=model.frame(FLQuants(whb,"wt"=stock.wt,"mat"=function(x) mat(x)))
@@
 
-fn<-function(x,dat) 
-   sum((dat$mat-matFn(dat$wt,x))^2)
+fn<-function(x,dat) {
+  sum((dat$mat-matFn(dat$wt,x))^2)}
 

### 2023-02-13 Rmd/ddMac.Rmd
clean

@@
 
+```{r, dd-par}
+par=rbind(par,
+          FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,
+                  bref =c(refpts(eq)["msy","biomass"]),
+                  delta=-0.2,
+                  matk = 0.2,
+                  w50  =par["a"]*par["l50"]^par["b"],
+                  alpha=-0.2)))
+```
+
 ```{r, prdFn}
@@
 ```{r}
-parMat=FLPar(c(k=0.1, w5=par["a"]*par["l50"]^par["b"]))
-dat=transform(dat,mat=1/(1+exp(-parMat[1]*(mass-parMat[2]))))
+dat=transform(dat,mat=1/(1+exp(-par["matk"]*(mass-par["w50"]))))
 
@@
 
-    mat(     x)=1/(1+exp(-parMat[1]*(stock.wt(x)-parMat[2])))
+    mat(     x)=1/(1+exp(-par["matk"]*(stock.wt(x)-par["w50"])))
   x=brp(x)}
@@
 dsr=discards.wt(eq)/stock.wt(eq)
+
 # iterate
@@
 
-  mat(     x)=1/(1+exp(-parMat[1]*(stock.wt(x)-parMat[2])))
+  mat(     x)=1/(1+exp(-par["matk"]*(stock.wt(x)-par["w50"])))
   m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
@@
 ```{r, check1}
-par=FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,bmsy=c(refpts(eq)["msy","biomass"]),k=0.1,w50=par["a"]*par["l50"]^par["b"]))
-
 f=FLQuant(c(rep(seq(0,1,length.out=51),each=101),rep(seq(1,25,length.out=61)[-1],each=101)),
@@
 om2=om
-
 for (year in dimnames(om2)$year[-1]){
@@
 **Figure `r iFig=iFig+1; iFig`** Check 2.
+
+```{r}
+save(par,file="../data/om/parMac.RData")
+```
 

### 2023-02-13 Rmd/ddWhb.Rmd
clean

@@
                cache.path="../cache/whb/dd/",
-               fig.path  ="../tex/figs/whbdata/dd/",
+               fig.path  ="../tex/figs/whb/dd/",
                fig.width =8,
@@
 
+```{r, dd-par}
+par=rbind(par,
+          FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,
+                  bref =c(refpts(eq)["msy","biomass"]),
+                  delta=-0.2,
+                  matk = 0.1,
+                  w50  =par["a"]*par["l50"]^par["b"],
+                  alpha=-0.2)))
+```
+
 ```{r, prdFn}
@@
 **Figure `r iFig=iFig+1; iFig`** Check 2.
+
+```{r}
+save(par,file="../data/om/parWhb.RData")
+```
 

### 2023-02-13 Rmd/omMac.Rmd
clean

---
title: "**Density Dependence**"
subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
author: "L Kell"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
mathjax: TRUE

tags: FLR FLCore introduction
license: Creative Commons Attribution-ShareAlike 4.0 International Public License
bibliography: refs.bib
---

```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# getwd()
library(knitr)

opts_chunk$set(cache     =TRUE, 
               comment   =NA, 
               warning   =FALSE, 
               message   =FALSE, 
               error     =FALSE, 
               echo      =FALSE, 
               eval      =TRUE,cache   =TRUE,
               cache.path="../cache/mac/om/",
               fig.path  ="../tex/figs/mac/om/",
               fig.width =8,
               fig.height=6,
               dev       ="png")

iFig=0
iTab=0
```

```{r, pkgs}
library(FLCore)
library(FLBRP)
library(FLasher)
library(FLife)
library(ggplotFL)

library(popbio)

library(GGally)
library(ggpubr)

library(rjson) # install.packages("rjson"); used for dropbox function
library(RJSONIO)

library(plyr)
library(dplyr)
library(reshape)

library(readxl)

library(mgcv)
library(mgcViz)

library(compositions)

theme_set(theme_bw(16))
```

```{r, dbox}
dropboxdir<-try(file.path(get_dropbox(), "dd"))

if ("try-error"%in%is(dropboxdir))
  dropboxdir="~/Dropbox/pelagics"
```

```{r}
source("~/Desktop/projects/ddmse/R/dd.R")
```

```{r}
load(file.path(dropboxdir,"data/om/mac.RData"))
```

```{r, ices}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["mac.27.nea"]]
ices=window(ices,start=1991)

sr  =as.FLSR(ices,model=bevholtSV)
sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(ices))),control=list(silent=TRUE))
eq  =FLBRP(ices,ab(sr))
ices=fwdWindow(ices,eq,end=2050)

F  =propagate(window(fbar(ices),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])

control=as(FLQuants("f"=F),"fwdControl")
ices   =fwd(ices,control=control,sr=eq)

dices=model.frame(FLQuants(ices[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                         "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
ggplot(dices)+
  geom_line(aes(SSB,Yield))
```

**Figure `r iFig=iFig+1; iFig`** .

```{r, vpa}
load(file.path(dropboxdir,"data/inputs/ices.RData"))
ices=ices[["mac.27.nea"]]
ices=window(ices,start=1991)
vpa =ices+FLAssess:::VPA(ices)

sr  =as.FLSR(vpa,model=bevholtSV)
sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(vpa))),control=list(silent=TRUE))
eq  =FLBRP(vpa,ab(sr))
vpa =fwdWindow(vpa,eq,end=2050)

F  =propagate(window(fbar(vpa),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])

control=as(FLQuants("f"=F),"fwdControl")
vpa    =fwd(vpa,control=control,sr=eq)

dvpa=model.frame(FLQuants(vpa[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                       "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
pvpa=ggplot(dvpa)+
  geom_line(aes(SSB,Yield))
```

**Figure `r iFig=iFig+1; iFig`** .

```{r, vpaM}
load("../data/om/parMac.RData")
par["m1"]  =par["m1"]/10
par["w50"] =par["w50"]/1000
par["matk"]=30

load(file.path(dropboxdir,"data/inputs/ices.RData"))
vpaM=ices[["mac.27.nea"]]
vpaM=window(vpaM,start=1991)
m(vpaM)=par["m1"]%*%(stock.wt(vpaM)%^%par["m2"])
vpaM=vpaM+FLAssess:::VPA(vpaM)

sr  =as.FLSR(vpaM,model=bevholtSV)
sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(vpa))),control=list(silent=TRUE))
eq  =FLBRP(vpaM,ab(sr))
vpaM=fwdWindow(vpaM,eq,end=2050)

F  =propagate(window(fbar(vpaM),start=2020),101)
F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.5,length.out=51)[-1]),each=dim(F)[2])

control=as(FLQuants("f"=F),"fwdControl")
vpaM  =fwd(vpaM,control=control,sr=eq)

d1=model.frame(FLQuants(vpaM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                      "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
d=rbind(cbind("What"="ICES",dices),
        cbind("What"="VPA",dvpa),
        cbind("What"="M",d1))
ggplot(d)+
  geom_line(aes(SSB,Yield,col=What))
```

```{r, vpaDDM}
source("~/Desktop/projects/ddmse/R/dd.R")

##################################
par["bref"]=subset(d1,Yield==max(Yield))[,"Biomass"]
##################################

vpaDDM=vpaM

for (i in ac(2020:2050)) {
  vpaDDM =ddFn(i,vpaDDM,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDM =fwd(vpaDDM,control=control,sr=eq)}

p1=ggplot(stock.wt(vpaDDM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mass-at-age")
p2=ggplot(mat(vpaDDM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(vpaDDM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("M-at-age")

d2=model.frame(FLQuants(vpaDDM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                        "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d2)+
  geom_line(aes(SSB,Yield))

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
```

```{r, vpaDDMM}
vpaDDMM =vpaM

for (i in ac(2020:2050)) {
  vpaDDMM =ddFn(i,vpaDDMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMM =fwd(vpaDDMM,control=control,sr=eq)}

p1=ggplot(stock.wt(vpaDDMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mass-at-age")
p2=ggplot(mat(vpaDDMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(vpaDDMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("M-at-age")

d3=model.frame(FLQuants(vpaDDMM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                         "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d3)+
  geom_line(aes(SSB,Yield))

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
```

**Figure `r iFig=iFig+1; iFig`** .

```{r, vpaDDMMM}
vpaDDMMM=vpaM

for (i in ac(2020:2050)) {
  vpaDDMMM =ddFn(i,vpaDDMMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
  control=as(FLQuants("f"=F[,i]),"fwdControl")
  vpaDDMMM =fwd(vpaDDMMM,control=control,sr=eq)}

p1=ggplot(stock.wt(vpaDDMMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(stock.wt(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mass-at-age")
p2=ggplot(mat(vpaDDMMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(mat(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("Mat-at-age")
p3=ggplot(m(vpaDDMMM[,"2030"]))+geom_line(aes(age,data,group=iter))+
  geom_line(aes(age,data),data=as.data.frame(m(vpaM[,"2031"])),col="red")+xlab("Age")+ylab("M-at-age")

d4=model.frame(FLQuants(vpaDDMMM[,"2050"],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                          "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
p4=ggplot(d4)+
  geom_line(aes(SSB,Yield))

ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
```

**Figure `r iFig=iFig+1; iFig`** Blue whiting: simple projection.

```{r}
d=rbind(cbind("What"="M",d1),
        cbind("What"="DD-M",d2),
        cbind("What"="DD-MM",d3),
        cbind("What"="DD-MMM",d4))
ggplot(d)+
  geom_line(aes(SSB,Yield,col=What))
```

# Funding

# References

### 2023-02-13 Rmd/omWhb.Rmd
clean

--- 2023-02-13 Rmd/omMac.Rmd
@@
 title: "**Density Dependence**"
-subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
+subtitle: "Blue whiting: Operating Model with density dependence in mass, maturity and M-at-age"
 author: "L Kell"
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/mac/om/",
-               fig.path  ="../tex/figs/mac/om/",
+               cache.path="../cache/whb/om/",
+               fig.path  ="../tex/figs/whb/om/",
                fig.width =8,
@@
 ```{r}
-load(file.path(dropboxdir,"data/om/mac.RData"))
+load(file.path(dropboxdir,"data/om/whb.RData"))
 ```
@@
 load(file.path(dropboxdir,"data/inputs/ices.RData"))
-ices=ices[["mac.27.nea"]]
+ices=ices[["whb.27.1-91214"]]
 ices=window(ices,start=1991)
@@
 F[]=rep(c(seq(0,                             c(refpts(eq)["msy",  "harvest"]),length.out=51),
-          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])
+          seq(c(refpts(eq)["msy","harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])
 
@@
 ```{r, vpaM}
-load("../data/om/parMac.RData")
+load("../data/om/parWhb.RData")
 par["m1"]  =par["m1"]/10
-par["w50"] =par["w50"]/1000
-par["matk"]=30
+par["w50"] =par["w50"]/100
+par["w50"] =0.08
+par["matk"]=100
 
 load(file.path(dropboxdir,"data/inputs/ices.RData"))
-vpaM=ices[["mac.27.nea"]]
+vpaM=ices[["whb.27.1-91214"]]
 vpaM=window(vpaM,start=1991)
@@
 F  =propagate(window(fbar(vpaM),start=2020),101)
-F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
-          seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.5,length.out=51)[-1]),each=dim(F)[2])
+F[]=rep(c(seq(0,                             c(refpts(eq)["msy",  "harvest"]),  length.out=51),
+          seq(c(refpts(eq)["msy","harvest"]),c(refpts(eq)["crash","harvest"])*2,length.out=51)[-1]),each=dim(F)[2])
 
@@
 
-##################################
+####################################################
 par["bref"]=subset(d1,Yield==max(Yield))[,"Biomass"]
-##################################
+####################################################
 
@@
   vpaDDM =ddFn(i,vpaDDM,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+  mat(vpaDDM)[1]=0
   control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
   vpaDDMM =ddFn(i,vpaDDMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
+  mat(vpaDDMM)[1]=0
   control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
   vpaDDMMM =ddFn(i,vpaDDMMM,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+  mat(vpaDDMMM)[1]=0
   control=as(FLQuants("f"=F[,i]),"fwdControl")

### 2023-02-13 R/dd.R

@@ -1,5 +1,5 @@
-ddWt<-function(wt,biomass,bmsy,alpha=-0.2,min=0.5,max=1.5){
-  adj=(biomass%/%bmsy)^alpha
+ddWt<-function(wt,biomass,bref,alpha=-0.2,min=0.5,max=1.5){
+  adj=(biomass%/%bref)^alpha
   
   wt%*%qmin(qmax(adj,min),max)}
 
@@ -7,36 +7,24 @@ ddWt<-function(wt,biomass,bmsy,alpha=-0.2,min=0.5,max=1.5){
 ddMat<-function(wt,x) 
   1/(1+exp(-x[1]%*%(wt%-%x[2])))
 
-ddM<-function(wt, F, gamma=2.0753+0.9914*F, lambda=-0.0276-0.1088*F, refWt=310, refM=0.10, M1=0.05){
+ddM<-function(wt,par){
+    par["m1"]%*%(wt%^%par["m2"])}
   
-  ref=refM/(gamma%*%exp(lambda*(refWt/0.0003)^(1/3)))
+ddFn<-function(year,x,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE){
   
-  exp(((wt/0.0001)^(1/3))%*%lambda)%*%gamma%*%ref+M1}
+  if (massFlag){
+    lsr=landings.wt(x)[,year]/stock.wt(x)[,year]
+    dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
+    dsr[!is.finite(dsr)]=0
+    
+    stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bref"])
+    landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
+    discards.wt(x)[,year]=stock.wt(x)[,year]*dsr}
+    
+  if (matFlag) 
+    mat(x)[,year]=ddMat(stock.wt(x)[,year],par[c("matk","w50")])
 
-ddFn<-function(year,x,par){
-  lsr=landings.wt(x)[,year]/stock.wt(x)[,year]
-  dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
-  dsr[!is.finite(dsr)]=0
-  
-  stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bmsy"])
[…]

### 2023-02-13 R/hcrICESV3.R

@@ -1,0 +1,213 @@
+#' hcrICES
+#' 
+#' @title hcrICES 
+#' 
+#' @description Harvest Control Rule, calculates Total Allowable Catch (TAC) based on a hockey stock harvest control rule.
+#' @author Laurence Kell, Sea++
+#'  
+#' @usage hcrICES(object,eql,sr_deviances,params, 
+#'          start=max(dimnames(object)$year)-10, end=start+10, interval=1,
+#'          err=NULL,bndTac=c(0,Inf),...)
+#'          
+#' @param object an object of class \code{FLStock} 
+#' @param eql \code{FLBRP} with a stock recruitment relationship used for projection
+#' @param sr_deviances \code{FLQuant} recuitment deviates on the log scale, i.e. multiplicative
+#' @param params \code{FLPar} HCR parameters, specifying blim, btrig, ftar and fmin
+#' @param start \code{numeric} first year for simulation
+#' @param end \code{numeric}   last year for simukation
+#' @param interval  \code{numeric} time step, 1 year by default
+#' @param err \code{FLQuant} assessment error on SSB for year used for HCR
+#' @param bndTac \code{numeric} bounds on TAC, by default these are turned off, for 20 percent constraint set to c(0.8,1.2)
+#' @param ... any additional arguments
+#' 
+#' @aliases hcrICES hcrICES-method hcrICES,FLStock,FLBRP-method
+#' 
+#' @export hcrICES
+#' @docType methods
+#' 
+#' @rdname hcrICES
+#' 
+#' @return returns a \code{list} with \code{FLStock} and \code{FLPar} objects for the stock and HCR 
+#'
+#' @import FLCore 
+#' @import FLBRP
+#' 
+#' @export
+#' 
+#' @examples
+#' \dontrun{
+#' data(pl4)
+#' }
+#' 
+#' 
+setGeneric('hcrICES', function(object,eql,...) standardGeneric('hcrICES'))
+
+setMethod('hcrICES', signature(object="FLStock",eql='FLBRP'),
+          function(object,eql,sr_deviances,params,
+                   start   =max(dimnames(object)$year)-10,
+                   end     =start+10,
+                   interval=1,
+                   lag     =1,
+                   err     =NULL,
+                   implErr =0,
+                   bndTac  =c(0,Inf),
[…]

### 2023-02-13 Rmd/ddMac.Rmd
m1 corrected

@@
 par=rbind(par,
-          FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,
+          FLPar(c(m1=0.15/(300^-0.288),m2=-0.28,
                   bref =c(refpts(eq)["msy","biomass"]),

### 2023-02-13 Rmd/ddWhb.Rmd
m1 corrected

@@
 par=rbind(par,
-          FLPar(c(m1=0.2/(300^-0.288),m2=-0.28,
+          FLPar(c(m1=0.2/(150^-0.288),m2=-0.28,
                   bref =c(refpts(eq)["msy","biomass"]),

### 2023-02-13 Rmd/omMac.Rmd
added Prod Curve to DD plots

@@
 p4=ggplot(d2)+
-  geom_line(aes(SSB,Yield))
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
 
@@
 p4=ggplot(d3)+
-  geom_line(aes(SSB,Yield))
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
 
@@
 p4=ggplot(d4)+
-  geom_line(aes(SSB,Yield))
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
 

### 2023-02-13 Rmd/omWhb.Rmd
added Prod Curve to DD plots

@@
 ggplot(d)+
-  geom_line(aes(SSB,Yield,col=What))
+  geom_line(aes(SSB,Yield,col=What))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+
 ```

### 2023-02-14 run/runHcr.R
cleaned up and backed up

@@ -1,0 +1,243 @@
+require(FLasher)
+require(FLBRP)
+require(FLAssess)
+require(mpb)
+require(ggplotFL)
+require(plyr)
+require(magrittr)
+
+#require(rdrop2)
+#drop_auth()
+
+dropboxdir="~/Dropbox/pelagics"
+
+#setwd("/home/laurie/Desktop/projects/pelagics/R")
+
+load(file.path(dropboxdir,paste("data/runs/",stk,".RData",sep="")))
+
+source("../R/hcrICESV2.R")
+
+stkid=c("whb.27.1-91214","mac.27.nea","her.27.3a47d")
+
+#stk     ="whb"
+
+nits    =100
+start   =2001
+end     =2021
+interval=1
+
+load(paste(file.path(dropboxdir,paste("data/om/",stk,"OM.RData",sep=""))))
+
+ices=get(paste(stk,"Ctc", sep=""))
+fmsy=get(paste(stk,"Fmsy",sep=""))
+err =get(paste(stk,"Err", sep=""))
+ftar=get(paste(stk,"Ftar",sep=""))
+refs=get(paste(stk,"Refs",sep=""))
+eql =FLBRP(ices)
+
+par=as(model.frame(mcf(FLQuants(ftar=ftar,
+                                btrig=refs[["msybtrigger"]][,ac(start:end)],
+                                fmin=ftar%=%0.05,
+                                blim=refs[["blim"]][,ac(start:end)])),drop=T)[,-1],"FLPar")
+par["btrig"][is.na(par["btrig"])]=min(par["btrig"],na.rm=T)
+par["blim"][ is.na(par["blim"])] =min(par["blim"],na.rm=T)
+
+if (refCurrent)
+  par[]=par[,21]
+
+err     =rlnorm(nits,FLQuant(0,dimnames=list(year=start:end)),var(err[-1,1])^0.5)
+
+implErr=FLQuants(NULL)
+implErr[["10%"]]=propagate(FLQuant(0.10,dimnames=list(year=(start-1):end)),nits)
+implErr[["20%"]]=propagate(FLQuant(0.20,dimnames=list(year=(start-1):end)),nits)
+implErr[["30%"]]=propagate(FLQuant(0.30,dimnames=list(year=(start-1):end)),nits)
+
[…]

### 2023-02-14 Rmd/ddMac.Rmd
updates omMac and ddMac

--- 2023-02-13 Rmd/ddMac.Rmd
@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: word_document
 mathjax: TRUE
@@
 
-```{r, pkgs}
+```{r include=FALSE}
 library(FLCore)
@@
 ```{r, dbox}
-dropboxdir<-try(file.path(get_dropbox(), "dd"))
+
+source("../R/get_dropbox.R")
+
+dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
 
@@
                   alpha=-0.2)))
+
+```
+
+```{r}
+
+as.data.frame(par) %>% pander::pandoc.table(style="simple")
+
 ```
@@
 ```{r}
+
 dat=ddWt(stock.wt(eq),biomass(eq),refpts(eq)["msy","biomass"]) 
@@
 ```{r, dd-eq-mass}
+
+# Start the clock
+ptm <- proc.time()
+
 x=propagate(eq,length(dimnames(eq)$year))
@@
   xlab("SSB")+ylab("Catch")
+
+# Stop the clock
+proc.time() - ptm
+
 ```
@@
 ```{r, dd-eq-mass-mat}
+
+# Start the clock
+ptm <- proc.time()
+
 x=propagate(eq,length(dimnames(eq)$year))
@@
 ```{r, dd-eq-mass-mat-m}
+
+# Start the clock
+ptm <- proc.time()
+
 x=propagate(eq,151)
@@
 
-```{r, check1}
+```{r message=FALSE, warning=FALSE}
+
+# Start the clock
+ptm <- proc.time()
+
 f=FLQuant(c(rep(seq(0,1,length.out=51),each=101),rep(seq(1,25,length.out=61)[-1],each=101)),
@@
 catch(om2)=computeCatch(om2,slot="all")
+
+# Stop the clock
+proc.time() - ptm
+
 ```
 
-```{r, check2}
+```{r message=FALSE, warning=FALSE}
 plot(window(FLStocks("OM"=om,DD=om2),start=40))  
@@
 ```{r}
-save(par,file="../data/om/parMac.RData")
+save(par,file=file.path(dropboxdir,"data/om/parMac.RData"))
+
 ```

### 2023-02-14 Rmd/omMac.Rmd
updates omMac and ddMac

@@
 ---
-title: "**Density Dependence**"
+title: "**Operating model**"
 subtitle: "Mackerel: Operating Model with density dependence in mass, maturity and M-at-age"
@@
 license: Creative Commons Attribution-ShareAlike 4.0 International Public License
-bibliography: refs.bib
+# bibliography: refs.bib
 ---
 
-```{r, knitr, eval=TRUE, echo=FALSE, warning=FALSE, cache=FALSE}
+```{r warning=FALSE, cache=FALSE, include=FALSE, knitr, eval=TRUE}
+
 knitr::opts_chunk$set(echo = FALSE)
@@
                echo      =FALSE, 
-               eval      =TRUE,cache   =TRUE,
+               eval      =TRUE,
+               cache     =TRUE,
                cache.path="../cache/mac/om/",
@@
 iTab=0
-```
 
-```{r, pkgs}
 library(FLCore)
@@
 theme_set(theme_bw(16))
-```
 
-```{r, dbox}
-dropboxdir<-try(file.path(get_dropbox(), "dd"))
+source("../R/get_dropbox.R")
 
-if ("try-error"%in%is(dropboxdir))
+dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
+
+if ("try-error"%in% is(dropboxdir))
   dropboxdir="~/Dropbox/pelagics"
@@
 
-```{r}
-source("~/Desktop/projects/ddmse/R/dd.R")
+```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
+
+# source("~/Desktop/projects/ddmse/R/dd.R")
+source("../R/dd.R")
+
 ```
 
-```{r}
+```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
+
 load(file.path(dropboxdir,"data/om/mac.RData"))
+
 ```
 
-```{r, ices}
+```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
+
+# Start the clock
+ptm <- proc.time()
+
 load(file.path(dropboxdir,"data/inputs/ices.RData"))
@@
   geom_line(aes(SSB,Yield))
+
+# Stop the clock
+proc.time() - ptm
+
 ```
@@
 ```{r, vpa}
+
+# Start the clock
+ptm <- proc.time()
+
 load(file.path(dropboxdir,"data/inputs/ices.RData"))
@@
   geom_line(aes(SSB,Yield))
+
+# Stop the clock
+proc.time() - ptm
 ```
@@
 ```{r, vpaM}
-load("../data/om/parMac.RData")
+
+# Start the clock
+ptm <- proc.time()
+
+load(file.path(dropboxdir, "data/om/parMac.RData"))
 par["m1"]  =par["m1"]/10
@@
   geom_line(aes(SSB,Yield,col=What))
+
+# Stop the clock
+proc.time() - ptm
 ```
 
+\newpage
+
+```{r}
+
+as.data.frame(par) %>% pander::pandoc.table(style="simple")
+
+```
+
+\newpage
+
 ```{r, vpaDDM}
-source("~/Desktop/projects/ddmse/R/dd.R")
+
+# source("~/Desktop/projects/ddmse/R/dd.R")
+
+# Start the clock
+ptm <- proc.time()
 
@@
 ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+
+# Stop the clock
+proc.time() - ptm
 ```
 
+\newpage
+
 ```{r, vpaDDMM}
+
+# Start the clock
+ptm <- proc.time()
+
 vpaDDMM =vpaM
@@
 ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+
+# Stop the clock
+proc.time() - ptm
 ```
@@
 
+\newpage
+
 ```{r, vpaDDMMM}
+
+# Start the clock
+ptm <- proc.time()
+
 vpaDDMMM=vpaM
@@
 **Figure `r iFig=iFig+1; iFig`** Blue whiting: simple projection.
+
+\newpage
 

### 2023-02-14 Rmd/report mac.Rmd
report.R and report mac.Rmd

@@
 library(captioner)     # captions
-tab_nums <- captioner::captioner(prefix = "Table", levels=3, type=c("n","n","n"), infix=".")
-fig_nums <- captioner::captioner(prefix = "Figure", levels=3, type=c("n","n","n"), infix=".")
+tab_nums <- captioner::captioner(prefix = "Table", levels=2, type=c("n","n"), infix=".")
+fig_nums <- captioner::captioner(prefix = "Figure", levels=2, type=c("n","n"), infix=".")
 
 source("../R/dd.R")
-source("../R/get_dropbox.r")
 source("../R/pm.R")
 
+source("../R/get_dropbox.r")
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
 
+load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
+
 # ------------------------------------------------------------------------------
@@
 # ------------------------------------------------------------------------------
-resultsdir <- file.path(dropboxdir, "results", mystk, "figures")
+figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
+tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
 
@@
 <!-- =============================================================================================================== -->
+
+```{r}
+
+tab_nums(name    = "section1_tab", level = 1, display = FALSE)
+fig_nums(name    = "section1_fig", level = 1, display = FALSE)
+
+```
 
@@
 
+**Weight at age**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_waa", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Weight at age by year and cohort"))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_waa")`_
+
+\newpage
+
+**Maturity at age**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_mat", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". Maturity at age by year and cohort"))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_mat")`_
+
+\newpage
+
+**Natural mortality at age (fixed)**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_natmor", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". Natural mortality at age by year and cohort (all ages and years the same M)"))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "m.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_natmor")`_
+
+\newpage
+
+**Density dependence in Weight at age vs. Total biomass**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_ddwaa", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". Density dependence and weight at age."))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_waa.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_ddwaa")`_
+
+\newpage
+
+**GAMM at age**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_gammwaa", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". GAMM analysis of density dependence and weight at age."))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "gamm.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_gammwaa")`_
+
+\newpage
+
+**GAMM summary**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "cond_gammsummary", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". GAMM summary of density dependence and weight at age."))
+
+cat(readLines(file.path(tablesdir, paste(mystk, "gamm summary.txt", sep="_"))), sep = '\n')
+
+```
+
+_`r tab_nums("cond_gammsummary")`_
+
+\newpage
+
+**GAMM summary**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "cond_gammcoefficients", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))
+
+cat(readLines(file.path(tablesdir, paste(mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')
+
+```
+
+_`r tab_nums("cond_gammcoefficients")`_
+
+\newpage
+
+**LM of density dependence in weight at age**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_lmwaa", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". LM analysis of density dependence and weight at age."))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "lm_waa.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_lmwaa")`_
+
+\newpage
+
+**LM summary**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "cond_lmsummary", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))
+
+cat(readLines(file.path(tablesdir, paste(mystk, "lm summary.txt", sep="_"))), sep = '\n')
+
+```
+
+_`r tab_nums("cond_gammcoefficients")`_
+
+\newpage
+
+**DD analysis of maturity at age**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_ddmat", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". Analysis of density dependence and maturity at age."))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_mat.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_ddmat")`_
+
+\newpage
+
+**Estimated maturity at weight**
+
+$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
+
+Estimated parameters: k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_matogive", level = 2, display = FALSE,
+  caption = paste0(mystkname, ". Estimated maturity ogive at weight"))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat_ogive.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_matogive")` _
+
 <!-- =============================================================================================================== -->
@@
 
-knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_mass.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_mass.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_maturity.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_maturity.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_eq_yield.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_eq_yield.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_om.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_om.jpg")))
 

### 2023-02-14 Rmd/report.R

@@ -1,4 +1,4 @@
-# ------------------------------------------------------------------------------
+ # ------------------------------------------------------------------------------
 # report.R
 #
 # Generating DDMSE results to be included into Rmd files
@@ -6,6 +6,7 @@
 # Laurie Kell and Martin Pastoors
 #
 # 09/02/2023 First coding
+# 13/02/2023 Implemented all codes from cond, DD and OM
 # ------------------------------------------------------------------------------
 
 rm(list=ls())
@@ -55,17 +56,17 @@ myparams <- data.frame(
   stock    = c("mac"                       , "whb"            ),
   mystkname=c("Northeast Atlantic mackerel", "Blue whiting"   ),
   mylatin = c("Scomber scombrus"           ,"Micromesistius poutassou"),
-  m1      = c(0.15/(300^-0.288)            , 0.2/(150^-0.288) ),
-  m2      = c(-0.288                        , -0.288            ),
+  m1      = c(0.15/(300^-0.288)            , 0.2/(150^-0.288) ), # 300 g MAC, 150 g WHB 
+  m2      = c(-0.288                        , -0.288          ),
   m3      = c(NA                           , NA               ),
   m       = c("lorenzen"                   , "lorenzen"       ),
   blim    = c(2000000                      , 1500000          ),
   btrig   = c(2580000                      , 2250000          ),
-  bmsy    = c(3500000                      , 4000000          ),
+  bmsy    = c(3500000                      , 4000000          ), # guesstimates
   alpha   = c(-0.2                         , -0.2             ),
   delta   = c(-0.2                         , -0.2             ),
-  matk    = c(0.2                          , 0.1              ), 
-  wt1     = c(0.01                         , NA               ),  # for resetting weight at age 1
+  matk    = c(30                           , 0.1              ), # source: conditioning
+  wt1     = c(0.01                         , NA               ), # for resetting weight at age 1
[…]

### 2023-02-14 Rmd/ddMac.Rmd
ddMac.Rmd

@@
 par=teleost[c("linf","k","t0","l50","a","b"),dimnames(teleost)$iter=="Scomber scombrus"]
-par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
+# par=lhPar(par,m1=0.2/(300^-0.288),m2=-0.28,m3=NA)
+par=lhPar(par,m1=0.15/(300^-0.288),m2=-0.288,m3=NA)
 eq =lhEql(par,m="lorenzen")
@@
 par=rbind(par,
-          FLPar(c(m1=0.15/(300^-0.288),m2=-0.28,
+          FLPar(c(
+                  # m1=0.15/(300^-0.288),m2=-0.28,
                   bref =c(refpts(eq)["msy","biomass"]),
@@
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 
@@
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 dat=transform(dat,mat=1/(1+exp(-par["matk"]*(mass-par["w50"]))))
@@
 
-```{r}
+```{r message=FALSE, warning=FALSE}
 dat=transform(dat,m=c(par["m1"])*(mass^c(par["m2"])))
@@
   m(       x)=par["m1"]%*%(stock.wt(x)%^%par["m2"])
-  x=brp(x)}
+  x=brp(x)
+}
 

### 2023-02-15 Rmd/report mac.Rmd
report.R and Rmd

@@
 library(captioner)     # captions
-tab_nums <- captioner::captioner(prefix = "Table", levels=2, type=c("n","n"), infix=".")
-fig_nums <- captioner::captioner(prefix = "Figure", levels=2, type=c("n","n"), infix=".")
+tab_nums <- captioner::captioner(prefix = "Table", levels=1, type=c("n"), infix=".")
+fig_nums <- captioner::captioner(prefix = "Figure", levels=1, type=c("n"), infix=".")
 
@@
 
-# Executive summary
+**Executive summary**
 
@@
 
-tab_nums(name    = "section1_tab", level = 1, display = FALSE)
-fig_nums(name    = "section1_fig", level = 1, display = FALSE)
+# captioner::bump(fig_nums, level=1)
+# fig_nums(name    = "section1_fig", caption = "section1_fig", level = 1, display = FALSE)
+# tab_nums(name    = "section1_tab", caption = "section1_tab", level = 1, display = FALSE)
 
@@
 
+```{r}
+
+# fig_nums(name    = "section2_fig", caption = "section2_fig", level = 1, display = FALSE)
+# tab_nums(name    = "section2_tab", caption = "section2_tab", level = 1, display = FALSE)
+
+```
+
 **Weight at age**
@@
 fig_nums(
-  name    = "cond_mat", level = 2, display = FALSE,
+  name    = "cond_mat", level = 1, display = FALSE,
   caption = paste0(mystkname, ". Maturity at age by year and cohort"))
@@
 fig_nums(
-  name    = "cond_natmor", level = 2, display = FALSE,
+  name    = "cond_natmor", level = 1, display = FALSE,
   caption = paste0(mystkname, ". Natural mortality at age by year and cohort (all ages and years the same M)"))
@@
 fig_nums(
-  name    = "cond_ddwaa", level = 2, display = FALSE,
+  name    = "cond_ddwaa", level = 1, display = FALSE,
   caption = paste0(mystkname, ". Density dependence and weight at age."))
@@
 fig_nums(
-  name    = "cond_gammwaa", level = 2, display = FALSE,
+  name    = "cond_gammwaa", level = 1, display = FALSE,
   caption = paste0(mystkname, ". GAMM analysis of density dependence and weight at age."))
@@
 tab_nums(
-  name    = "cond_gammcoefficients", level = 2, display = FALSE,
+  name    = "cond_gammcoefficients", level = 1, display = FALSE,
   caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))
@@
 fig_nums(
-  name    = "cond_lmwaa", level = 2, display = FALSE,
+  name    = "cond_lmwaa", level = 1, display = FALSE,
   caption = paste0(mystkname, ". LM analysis of density dependence and weight at age."))
@@
 tab_nums(
-  name    = "cond_lmsummary", level = 2, display = FALSE,
+  name    = "cond_lmsummary", level = 1, display = FALSE,
   caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))
@@
 fig_nums(
-  name    = "cond_ddmat", level = 2, display = FALSE,
+  name    = "cond_ddmat", level = 1, display = FALSE,
   caption = paste0(mystkname, ". Analysis of density dependence and maturity at age."))
@@
 fig_nums(
-  name    = "cond_matogive", level = 2, display = FALSE,
+  name    = "cond_matogive", level = 1, display = FALSE,
   caption = paste0(mystkname, ". Estimated maturity ogive at weight"))
@@
 
-# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_mass.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_mass.jpg")))
 
@@
 
-# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_maturity.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_maturity.jpg")))
 
@@
 
-# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_m.jpg")))
 
@@
 
-# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_eq_yield.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_eq_yield.jpg")))
 
@@
 
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+[to be done]
 
-# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_om.jpg")))
+**Density dependence in mass-at-age**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_VPADDM", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in Mass-at-age")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDM_4panels.jpg")))
 
@@
 
+_`r fig_nums("om_VPADDM")`_
+
 \newpage
 
-Forward projections with F=fmsy (fwd), 0.8xfmsy?, 1.0xfmsy, 1.2xfmsy?
+**Density dependence in mass-at-age and maturity at age**
 
-```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_VPADDMM", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in Mass-at-age and Maturity at age")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMM_4panels.jpg")))
 
 ```
+
+_`r fig_nums("om_VPADDMM")`_
+
+\newpage
+
+**Density dependence in mass-at-age, maturity at age and M at age**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_VPADDMMM", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMMM_4panels.jpg")))
+
+```
+
+_`r fig_nums("om_VPADDMMM")`_
+
+\newpage
+
+**Comparison of equilibrium curves for density dependence analyses: biomass against yield**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_eqyield", level = 1, display = FALSE,
+  caption = "Equilibrium curves for density dependence analyses: biomass against yield")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_om_biomass_yield.jpg")))
+
+```
+
+_`r fig_nums("om_eqyield")`_
+
+\newpage
+
+**Comparison of equilibrium curves for density dependence analyses: biomass against F**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_eqf", level = 1, display = FALSE,
+  caption = "Equilibrium curves for density dependence analyses: biomass against F")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_om_biomass_F.jpg")))
+
+```
+
+_`r fig_nums("om_eqf")`_
+
+\newpage
 

### 2023-02-15 Rmd/report.R

@@ -1,4 +1,4 @@
- # ------------------------------------------------------------------------------
+# ------------------------------------------------------------------------------
 # report.R
 #
 # Generating DDMSE results to be included into Rmd files
@@ -65,7 +65,7 @@ myparams <- data.frame(
   bmsy    = c(3500000                      , 4000000          ), # guesstimates
   alpha   = c(-0.2                         , -0.2             ),
   delta   = c(-0.2                         , -0.2             ),
-  matk    = c(30                           , 0.1              ), # source: conditioning
+  matk    = c(0.2                           , 0.1              ), # source: conditioning
   wt1     = c(0.01                         , NA               ), # for resetting weight at age 1
   minage  = c(0                            , 1                ),
   maxage  = c(12                           , 10               ),
@@ -151,10 +151,10 @@ mystk     <- "mac";
     # guides(colour = guide_legend(nrow = 1)) +
     scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
   
-  ggsave(p, 
-         filename=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")), 
-         device="jpeg", 
-         width=10, height=6, units="in")
+  jpeg(filename=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")),
+       width=10, height=6, units="in", res=300)
+  print(p) 
+  dev.off()
 
   # plot maturity at age -------------------------------------------------------
   p <-
@@ -170,12 +170,13 @@ mystk     <- "mac";
     # guides(colour = guide_legend(nrow = 1)) +
     scale_color_manual("Age",values=rainbow(length(unique(mat$age))))
   
-  ggsave(p, 
-         filename=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")), 
-         device="jpeg", 
-         width=10, height=6, units="in")
 
-
+  jpeg(filename=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")),
+       width=10, height=6, units="in", res=300)
+  print(p) 
+  dev.off()
+  
+  
[…]

### 2023-02-15 Rmd/report mac.Rmd
report.R & Rmd updated

@@
 
+**Stock assessments**
+
 Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA.
@@
 
-[to be done]
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_assessments", level = 1, display = FALSE,
+  caption = "Comparison of different assessments. VPA-M is used going forward")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_assessments.jpg")))
+
+```
+
+_`r fig_nums("om_VPADDM")`_
+
+\newpage
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_om_biomass_yield.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd.jpg")))
 
@@
 
-**Comparison of equilibrium curves for density dependence analyses: biomass against F**
-
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "om_eqf", level = 1, display = FALSE,
-  caption = "Equilibrium curves for density dependence analyses: biomass against F")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_om_biomass_F.jpg")))
-
-```
-
-_`r fig_nums("om_eqf")`_
-
-\newpage
-
 <!-- =============================================================================================================== -->
-<!-- Section 5: Running MSE's                                                                                        -->
+<!-- Section 5: Running forward projections                                                                          -->
 <!-- =============================================================================================================== -->
 
-# Running MSE's
+# Forward projections without feedback
 

### 2023-02-15 Rmd/report.R

@@ -69,7 +69,10 @@ myparams <- data.frame(
   wt1     = c(0.01                         , NA               ), # for resetting weight at age 1
   minage  = c(0                            , 1                ),
   maxage  = c(12                           , 10               ),
-  maxyear = c(2050                         , 2050             ))
+  maxyear = c(2050                         , 2050             ),
+  m1scaler= c(8                            , 8                ),
+  w50scaler= c(1000                        , 1000             ),
+  matk2   = c(30                           , 30               ) )
 
 save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
@@ -631,6 +634,13 @@ mystk     <- "mac";
   ices=window(stk,start=1991)
   name(ices) <- mystkname
 
+  assess_df <-
+    model.frame(FLQuants(ices, "Biomass"=function(x) biomass(x),
+                               "SSB"=function(x) ssb(x),
+                               "F"=function(x) fbar(x),
+                               "Yield"=function(x) catch(x),
+                               "Rec"=function(x) rec(x)),drop=T) %>% mutate(scen="sam")
+  
   # Start with the SAM assessment in forward projection
   sr  = as.FLSR(ices,model=bevholtSV)
   sr  = fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(ices))),control=list(silent=TRUE))
@@ -638,8 +648,9 @@ mystk     <- "mac";
   sam = fwdWindow(ices,eq,end=2050) 
   
   F  =propagate(window(fbar(sam),start=2020),101)
-  F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
-            seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])
+  # F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
[…]

### 2023-02-15 Rmd/report whb.Rmd
report.R&Rmd for WHB

@@
 
+library(captioner)     # captions
+tab_nums <- captioner::captioner(prefix = "Table", levels=1, type=c("n"), infix=".")
+fig_nums <- captioner::captioner(prefix = "Figure", levels=1, type=c("n"), infix=".")
+
 source("../R/dd.R")
-source("../R/get_dropbox.r")
 source("../R/pm.R")
 
+source("../R/get_dropbox.r")
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
 
-# mystk     <- "whb"; mystkname <- "Blue whiting"
-# mystk     <- "her"; mystkname <- "North Sea herring"
-mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; maxyear <- 2050
+# mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
+mystk     <- "whb"; mystkname <- "Blue whiting";                mylatin <- "Micromesistius poutassou" 
+maxyear <- 2050
+
+load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
+
+# ------------------------------------------------------------------------------
+# Operating model
+# ------------------------------------------------------------------------------
+figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
+tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
 
@@
 
-# Evaluating impacts of Density Dependent processes on `r mystkname` MSE
+<!-- =============================================================================================================== -->
+<!-- ## Mackerel report                                                                                              -->
+<!-- =============================================================================================================== -->
 
-**L. Kell, M.A. Pastoors
+**Evaluating impacts of Density Dependent processes on `r mystkname` via MSE simulations**
+
+**L. Kell, M.A. Pastoors**
 
@@
 
-## Introduction
+Commissioned by: University of Copenhagen (Henrik Sparholt)
 
-## OM conditioning
+**Executive summary**
 
-Stock assessments with SAM, Density dependent SAM and Density dependent VPA
+[summary here]
 
-```{r, mac.data, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+<!-- =============================================================================================================== -->
+<!-- Section 1: introduction                                                                                         -->
+<!-- =============================================================================================================== -->
 
-load(file.path(dropboxdir,"data/om/mac.RData"))
-combY <- as.data.frame(pmYear(window(mac, start=2001)),drop=T) %>% mutate(stk=mystk, type="SAM")
+```{r}
 
-par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bmsy=3500000,b=0.3,k=30.1, w50=0.2))
-
-stock.wt(   mac)[1]=0.01
-catch.wt(   mac)[1]=0.01
-landings.wt(mac)[1]=0.01
-discards.wt(mac)[1]=0.01
-
-om_orig=ddFn2(mac,par); 
-combY <- bind_rows(combY,
-                   as.data.frame(pmYear(window(om_orig, start=2001)),drop=T) %>% mutate(stk=mystk, type="DD SAM"))
-
-om=om_orig+VPA(om_orig)
-om=ddFn2(om,par)
-combY <- bind_rows(combY, 
-                   as.data.frame(pmYear(window(om, start=2001)),drop=T) %>% mutate(stk=mystk, type="DD VPA") )
-
-combY %>% 
-  filter(qname %in% c("catch","ssb","rec","fbar")) %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  geom_ribbon(aes(fill=type, ymin=lowq, ymax=uppq, xmin=year, xmax=year), alpha=0.3) +
-  geom_line(aes(colour=type)) +
-  facet_grid(qname~., scales="free_y")
+# captioner::bump(fig_nums, level=1)
+# fig_nums(name    = "section1_fig", caption = "section1_fig", level = 1, display = FALSE)
+# tab_nums(name    = "section1_tab", caption = "section1_tab", level = 1, display = FALSE)
 
@@
 
-\newpage
+# Introduction
 
-Forward projections with F=fmsy (fwd), 0.8xfmsy?, 1.0xfmsy, 1.2xfmsy?
+The purpose of this work is to support the project “Fmsy-project for six high profile fish stocks”.
+The project focuses on the current management approach for fish stocks based on single-species Fmsy Values where the only density-dependent process that is included is recruitment.
+The project includes three other density-dependent population dynamics: growth (weight at age), sexual maturity and natural mortality.
 
-```{r, mac.data2, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+This report deals with the evaluation of density dependent processes for `r mystkname` using Management Strategy Evaluation (MSE). Steps in the analysis:
 
-sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
-eq=FLBRP(om,sr=sr)
+1. Conditioning and summaries based on the 2022 ICES assessment (section 2)
+2. Estimating Density Dependent effects based on equilibrium analyses and explore DD before setting up the OMs (section 3)
+3. Setting up the Operating Models (OM) (section 4)
+4. Running MSE's (section 5)
 
-om=fwdWindow(om,end=maxyear,eq)
-f =fbar(om)[,ac(2021:maxyear)]%=%refpts(eq)["msy","harvest"]
-om=fwd(om,fbar=f,sr=eq)
+Four alternatives OMs have been included in the analyses:
 
-om2=om
-om3=om
-om4=om
+1. No density dependent processes
+2. DD in growth
+3. DD in growth and maturity
+4. DD in growth, maturity and natural mortality
 
-for (iYear in ac(2021:maxyear)){
-  om2=ddFn(iYear,om2,par)
-  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)
+All MSE's have been developed within the FLR framework (https://flr-project.org/) and all the code is stored on github (https://github.com/martinpastoors/ddmse).
 
-  om3=ddFn(iYear,om3,par)
-  om3=fwd(om3,fbar=f[,iYear],sr=eq)
+All data and results are available on DropBox (https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0)
 
-  om4=ddFn(iYear,om4,par)
-  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
-  }
+<!-- =============================================================================================================== -->
+<!-- Section 2: conditioning and summaries of ICES assessment                                                        -->
+<!-- =============================================================================================================== -->
 
-plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
+# Conditioning and summaries based on the 2022 ICES assessment
+
+```{r}
+
+# fig_nums(name    = "section2_fig", caption = "section2_fig", level = 1, display = FALSE)
+# tab_nums(name    = "section2_tab", caption = "section2_tab", level = 1, display = FALSE)
 
@@
 
-## Simulation results
+**Weight at age**
 
-## Discussion and conclusions
+[text here]
 
-ICES
-Fmsy
-HCR1 (no bounds)
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-```{r, runWhb1, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+fig_nums(
+  name    = "cond_waa", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Weight at age by year and cohort"))
 
-myvar     <- c("catch","cCum", "ssb","fbar", "rec")
-scenarios <- c("ICES", "Fmsy", "HCR1")
-rp        <- refs  %>% 
-  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-ices      <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
-  dplyr::select(- .id) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-df        <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  group_by(stk, .id, iter) %>% 
-  mutate(perc_diff = (data-lag(data))/lag(data)) 
-# box <- 
-#   bind_rows(
-#     df   %>% dplyr::select(.id, qname, stk, data=median),
-#     ices %>% dplyr::select(qname, stk, data=median) %>% mutate(.id="ICES")
-#   ) %>% 
-#   group_by(qname, stk) %>% 
-#   summarise(max = max(data, na.rm=TRUE), min = min(data, na.rm=TRUE)) %>% 
-#   mutate(min = ifelse(min<0, min, 0)) %>%
-
-df %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-
-  # vline at 2010
-  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +
-
-  # iters
-  geom_line(data=iters,
-            aes(y=data, group=iter),
-            linewidth=0.4, colour="darkgray") +
-
-  # median and ribbon
-  geom_line(aes(y=median),
-            linewidth=1, colour="red") +
-  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq),
-              alpha=0.2, fill="red") +
-
-  # ICES
-  geom_line(data= ices,
-            aes(year,median),
-            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +
-
-  # refpoints
-  geom_line(data=rp,
-            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
-            linewidth=1) +
-  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
-  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
-
-  expand_limits(y=0) +
-  facet_grid(qname ~ .id, scales="free_y")
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")))
 
@@
 
-**Figure `r mystk`.2**  `r mystkname`. Comparison of different HCR’s without TAC bounds
+_`r fig_nums("cond_waa")`_
 
@@
 
-#### `r mystkname`: comparing different HCR’s with and without TAC bounds
+**Maturity at age**
 
-Fmsy
-HCR0 (no bounds)
-HCR1 bnd (-25%/+20% TAC change bounds apply when stock is above MSY Btrigger)
-HCR2 bnd (-25%/+20% TAC change bounds apply when stock is above Blim)
+[text here]
 
-```{r, runWhb2, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-myvar     <- c("catch","cCum", "ssb","fbar")
-scenarios <- c("Fmsy", "HCR1",  "HCR1 bnd", "HCR2 bnd")
-rp        <- refs %>% 
-  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
-    mutate(qname = factor(qname, levels=myvar))
-ices      <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
-  dplyr::select(- .id) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-df        <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  group_by(stk, .id, iter) %>% 
-  mutate(perc_diff = (data-lag(data))/lag(data))
+fig_nums(
+  name    = "cond_mat", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Maturity at age by year and cohort"))
 
-df %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-
-  # vline at 2010
-  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +
-
-  # iters
-  geom_line(data=iters,
-            aes(y=data, group=iter),
-            linewidth=0.4, colour="darkgray") +
-
-  # median and ribbon
-  geom_line(aes(y=median), 
-            linewidth=1, colour="red") +
-  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
-              alpha=0.2, fill="red") +
-
-  # ICES
-  geom_line(data= ices, aes(year,median), 
-            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +
-
-  # refpoints
-  geom_line(data=rp, 
-            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
-            linewidth=1) +
-  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
-  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
-
-  expand_limits(y=0) +
-  facet_grid(qname ~ .id, scales="free_y")
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")))
 
@@
 
-**Figure `r mystk`.3**  `r mystkname`. Comparison of different HCR’s with and without TAC bounds
+_`r fig_nums("cond_mat")`_
 
@@
 
-#### `r mystkname`: comparing HCR’s with and without a cap on TAC bounds
+**Natural mortality at age (fixed)**
 
-HCR1 bnd
-HCR1 bnd cap
-HCR2 bnd
-HCR2 bnd cap
+[text here]
 
-```{r, runWhb3, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-myvar     <- c("catch","cCum", "ssb","fbar")
-scenarios <- c("HCR1 bnd", "HCR1 bnd cap", "HCR2 bnd", "HCR2 bnd cap")
-rp        <- refs %>% 
-  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
-    mutate(qname = factor(qname, levels=myvar))
-ices      <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
-  dplyr::select(- .id) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-df        <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  group_by(stk, .id, iter) %>% 
-  mutate(perc_diff = (data-lag(data))/lag(data))
+fig_nums(
+  name    = "cond_natmor", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Natural mortality at age by year and cohort (all ages and years the same M)"))
 
-df %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-
-  # vline at 2010
-  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +
-
-  # iters
-  geom_line(data=iters,
-            aes(y=data, group=iter),
-            linewidth=0.4, colour="darkgray") +
-
-  # median and ribbon
-  geom_line(aes(y=median), 
-            linewidth=1, colour="red") +
-  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
-              alpha=0.2, fill="red") +
-  # ICES
-  geom_line(data= ices, aes(year,median), 
-            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +
-
-  # refpoints
-  geom_line(data=rp, 
-            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
-            linewidth=1) +
-  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
-  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
-
-  expand_limits(y=0) +
-  facet_grid(qname ~ .id, scales="free_y")
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "m.jpg", sep="_")))
 
@@
 
-**Figure `r mystk`.4**  `r mystkname`. Comparison of different HCR’s with and without a cap on TAC bounds
+_`r fig_nums("cond_natmor")`_
 
@@
 
-#### `r mystkname`: comparing HCR’s with and without implementation error (without bounds on TAC change)
+**Density dependence in Weight at age vs. Total biomass**
 
-HCR0
-HCR1 10%
-HCR1 20%
-HCR1 30%
-HCR1 10% recent
-HCR1 10% random recent
+[text here]
 
-```{r, runWhb4, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-myvar     <- c("catch","cCum", "ssb","fbar")
-scenarios <- c("HCR1", "HCR1 +10%", "HCR1 +20%", "HCR1 +30%", "HCR1 +10% recent", 
-               "HCR1 +random recent")
-rp        <- refs %>% 
-  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-ices      <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
-  dplyr::select(- .id) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-df        <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  group_by(stk, .id, iter) %>% 
-  mutate(perc_diff = (data-lag(data))/lag(data))
+fig_nums(
+  name    = "cond_ddwaa", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Density dependence and weight at age."))
 
-df %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-
-  # vline at 2010
-  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +
-
-  # iters
-  geom_line(data=iters,
-            aes(y=data, group=iter),
-            linewidth=0.4, colour="darkgray") +
-
-  # median and ribbon
-  geom_line(aes(y=median), 
-            linewidth=1, colour="red") +
-  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
-              alpha=0.2, fill="red") +
-
-  # ICES
-  geom_line(data= ices, aes(year,median), 
-            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +
-
-  # refpoints
-  geom_line(data=rp, 
-            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
-            linewidth=1) +
-  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
-  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
-
-  expand_limits(y=0) +
-  facet_grid(qname ~ .id, scales="free_y")
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_waa.jpg", sep="_")))
 
@@
 
-**Figure `r mystk`.5**  `r mystkname`. Comparison of different HCR’s with and without implementation error (without bounds on TAC change)
+_`r fig_nums("cond_ddwaa")`_
 
@@
 
-#### `r mystkname`: comparing HCR’s with and without implementation error in combination with bounds on TAC change
+**GAMM at age**
 
-HCR1 bnd
-HCR1 10% bnd
-HCR1 10% recent bnd
-HCR1 10% random bnd
-HCR1 10% random recent bnd
+[text here]
 
-```{r, runWhb5, echo=FALSE, fig.asp=1.15, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-myvar     <- c("catch","cCum", "ssb","fbar")
-scenarios <- c("HCR1 bnd", 
-               "HCR1 +10% bnd",
-               "HCR1 +10% bnd recent",
-               "HCR1 +random bnd", 
-               "HCR1 +random bnd recent")
-rp        <- refs %>% 
-  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-ices      <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
-  dplyr::select(- .id) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-df        <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  group_by(stk, .id, iter) %>% 
-  mutate(perc_diff = (data-lag(data))/lag(data))
+fig_nums(
+  name    = "cond_gammwaa", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". GAMM analysis of density dependence and weight at age."))
 
-df %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-
-  # vline at 2010
-  geom_vline(xintercept = 2010, colour="darkgray", linewidth=1) +
-
-  # iters
-  geom_line(data=iters,
-            aes(y=data, group=iter),
-            linewidth=0.4, colour="darkgray") +
-
-  # median and ribbon
-  geom_line(aes(y=median), 
-            linewidth=1, colour="red") +
-  geom_ribbon(aes(xmin=year, xmax=year, ymin=lowq, ymax=uppq), 
-              alpha=0.2, fill="red") +
-
-  # ICES
-  geom_line(data= ices, aes(year,median), 
-            col="blue", linewidth=1, linetype=2, inherit.aes=FALSE) +
-
-  # refpoints
-  geom_line(data=rp, 
-            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
-            linewidth=1) +
-  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
-  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
-
-  expand_limits(y=0) +
-  facet_grid(qname ~ .id, scales="free_y")
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "gamm.jpg", sep="_")))
 
@@
 
-**Figure `r mystk`.6**  `r mystkname`. Comparison of different HCR’s with and without implementation error in combination with bounds on TAC change
+_`r fig_nums("cond_gammwaa")`_
 
@@
 
-#### `r mystkname`: checking the working of different constraints
+**GAMM summary**
 
-```{r, whbcheck, echo=FALSE, fig.asp=.5, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+[text here]
 
-myvar     <- c("ssb")
-# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10%", "HCR1 +10% bnd", "HCR1 +10% bnd cap")
-scenarios <- c("HCR1", "HCR1 bnd", "HCR2", "HCR2 bnd")
-#scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10% bnd", "HCR1 +20% bnd", "HCR1 +30% bnd")
-rp        <- refs %>% filter(stk == mystk, refpoint %in% c("msybtrigger", "blim"))
-ices      <- combY %>% filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% dplyr::select(- .id)
-df        <- combY %>% filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  left_join(dplyr::select(rp,
-                          stk, year, qname, refpoint, reftype, rp=data), 
-            by=c("stk", "year", "qname")) %>% 
-  mutate(trigger = ifelse(median >= rp, "above","below"))
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 10) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  left_join(dplyr::select(rp,
-                          stk, year, qname, refpoint, reftype, rp=data), 
-            by=c("stk", "year", "qname")) %>% 
-  mutate(trigger = ifelse(data >= rp, "above","below"))
-df %>% 
-  ggplot(aes(x=year, y=median)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-  geom_line(data=iters,
-            aes(y=data, group=iter),
-            linewidth=0.4, colour="darkgray") +
-  geom_point(data=filter(iters,
-                         trigger=="above"),
-            aes(y=data, group=iter),
-            size=1.0, colour="black") +
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-  geom_line(linewidth=1, colour="red") +
-  geom_point(data=filter(df,
-                        trigger=="above"),
-            colour="red", size=2) +
-    # refpoints
-  geom_line(data=rp, 
-            aes(x=year, y=data, group=reftype, col=reftype, linetype=reftype),
-            linewidth=1) +
-  scale_linetype_manual(values = c("longdash","dashed"), breaks = c("lim", "msy") ) +
-  scale_colour_manual  (values = c("black","darkgray"), breaks = c("lim", "msy") ) +
-  labs(y="ssb (tonnes)") +
-  expand_limits(y=0) +
-  facet_grid(qname ~ .id, scales="free_y")
+tab_nums(
+  name    = "cond_gammsummary", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". GAMM summary of density dependence and weight at age."))
 
-myvar     <- c("catch")
-# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10%", "HCR1 +10% bnd", "HCR1 +10% bnd cap")
-# scenarios <- c("ICES", "HCR1 bnd", "HCR1 +10% bnd", "HCR1 +20% bnd", "HCR1 +30% bnd")
-rp        <- refs %>% 
-  filter(stk == mystk, refpoint %in% c("fmsy", "blim", "msybtrigger")) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-ices      <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% "ICES",  year %in% 2000:2021) %>% 
-  dplyr::select(- .id) %>% 
-  mutate(qname = factor(qname, levels=myvar))
-df        <- combY %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(qname = factor(qname, levels=myvar)) %>% 
-  group_by(stk, .id) %>% 
-  mutate(perc_diff = (median-lag(median))/lag(median))
-
-iters        <- combYI %>% 
-  filter(stk == mystk, qname %in% c(myvar),  .id %in% scenarios, year %in% 2000:2021, iter <= 5) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  group_by(stk, .id, iter) %>% 
-  mutate(perc_diff = (data-lag(data))/lag(data))
-
-df %>% 
-  ggplot(aes(x=year, y=perc_diff)) +
-  theme_bw() +
-  theme(panel.grid.minor = element_blank(),  
-        panel.border     = element_rect(linetype = 1, colour="black", fill=NA))+
-
-  geom_line(data=iters,
-            aes(x=year, y=perc_diff, group=iter),
-            linewidth=.4, colour="darkgray", inherit.aes = FALSE) +
-
-  geom_line(linewidth=1, colour="red") +
-
-  geom_hline(yintercept = 0.25, colour="black", linewidth=1, linetype=2) +
-  geom_hline(yintercept = -0.2, colour="black", linewidth=1, linetype=2) +
-  expand_limits(y=0) +
-  scale_y_continuous(labels=scales::percent, limits=c(-2,2)) +
-  facet_grid(qname ~ .id, scales="free_y")
+cat(readLines(file.path(tablesdir, paste(mystk, "gamm summary.txt", sep="_"))), sep = '\n')
 
@@
 
-**Figure `r mystk`.6**  `r mystkname`.
-Comparison of the working of constraints on TAC change relative to stock size and reference point MSY Btrigger.
-Black dots indicate estimated stock size on individual runs above MSY Btrigger (i.e.
-TAC contstraint applies).
-Red dots indicate median stock size above MSY Btrigger.
+_`r tab_nums("cond_gammsummary")`_
 
@@
 
-#### `r mystkname`: comparing all scenario's on median catch, IAV and AAV
+**GAMM summary**
 
-```{r, whbsmry, echo=FALSE, fig.asp=1.2, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+[text here]
 
-myvar     <- c("catch.50", "iav.50", "aav")
-scenarios <- c("ICES","Fmsy","HCR1","HCR1 bnd","HCR1 bnd cap", "HCR1 +10% bnd","HCR1 +10% bnd cap",
-               "HCR1 +10% bnd recent","HCR1 +random bnd",",HCR1 +random bnd recent","HCR1 +10%",
-               "HCR1 +10% recent","HCR1 +random","HCR1 +random recent","HCR1 +20% bnd", 
-               "HCR1 +20% bnd cap","HCR1 +20% bnd recent","HCR1 +20%","HCR1 +20% recent",
-               "HCR1 +30% bnd","HCR1 +30% bnd cap","HCR1 +30% bnd recent","HCR1 +30%",
-               "HCR1 +30% recent",
-               "HCR2","HCR2 bnd","HCR2 bnd cap","HCR2 +10% bnd","HCR2 +10% bnd cap",
-               "HCR2 +10% bnd recent","HCR2 +random bnd","HCR2 +random bnd recent","HCR2 +10%",
-               "HCR2 +10% recent","HCR2 +random","HCR2 +random recent","HCR2 +20% bnd",
-               "HCR2 +20% bnd cap","HCR2 +20% bnd recent","HCR2 +20%","HCR2 +20% recent",
-               "HCR2 +30% bnd","HCR2 +30% bnd cap","HCR2 +30% bnd recent","HCR2 +30%",
-               "HCR2 +30% recent")
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-df        <- combI %>% 
-  tidyr::pivot_longer(names_to = "variable", values_to = "data", myvar) %>% 
-  filter(stk == mystk,  .id %in% scenarios) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(variable = factor(variable, levels=myvar))
+tab_nums(
+  name    = "cond_gammcoefficients", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))
 
-calc_boxplot_stat <- function(x) {
-  coef <- 1.5
-  n <- sum(!is.na(x))
-  # calculate quantiles
-  stats <- quantile(x, probs = c(0.0, 0.25, 0.5, 0.75, 1.0))
-  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
-  iqr <- diff(stats[c(2, 4)])
-  # set whiskers
-  outliers <- x < (stats[2] - coef * iqr) | x > (stats[4] + coef * iqr)
-  if (any(outliers)) {
-    stats[c(1, 5)] <- range(c(stats[2:4], x[!outliers]), na.rm = TRUE)
-  }
-  return(stats)
-}
-
-df %>% 
-  ggplot(aes(x=.id, y=data)) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +
-  expand_limits(y=0) +
-  # scale_y_continuous(limits = quantile(t$data, c(0.1, 0.9))) +
-  labs(x="", y="") +
-  facet_grid(variable~., scales="free", space="free_x")
+cat(readLines(file.path(tablesdir, paste(mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')
 
@@
 
-**Figure `r mystk`.7**  `r mystkname`. Comparison of all scenario's on median catch, median IAV (Interannual variability in catch) and AAV (Average Annual Variability in catch)
+_`r tab_nums("cond_gammcoefficients")`_
 
-```{r smry2, eval=FALSE, fig.align="center", fig.asp=1.2, message=FALSE, warning=FALSE, cache=FALSE, include=FALSE, whbsmry, echo=FALSE}
+\newpage
 
-mystk     <- c("whb","her","mac")
-myvar     <- c("catch.50", "iav.50", "aav")
-scenarios <- c("ICES","Fmsy","HCR1","HCR1 bnd","HCR1 bnd cap", "HCR1 +10% bnd","HCR1 +10% bnd cap",
-               "HCR1 +10% bnd recent","HCR1 +random bnd",",HCR1 +random bnd recent","HCR1 +10%",
-               "HCR1 +10% recent","HCR1 +random","HCR1 +random recent","HCR1 +20% bnd", 
-               "HCR1 +20% bnd cap","HCR1 +20% bnd recent","HCR1 +20%","HCR1 +20% recent",
-               "HCR1 +30% bnd","HCR1 +30% bnd cap","HCR1 +30% bnd recent","HCR1 +30%",
-               "HCR1 +30% recent",
-               "HCR2","HCR2 bnd","HCR2 bnd cap","HCR2 +10% bnd","HCR2 +10% bnd cap",
-               "HCR2 +10% bnd recent","HCR2 +random bnd","HCR2 +random bnd recent","HCR2 +10%",
-               "HCR2 +10% recent","HCR2 +random","HCR2 +random recent","HCR2 +20% bnd",
-               "HCR2 +20% bnd cap","HCR2 +20% bnd recent","HCR2 +20%","HCR2 +20% recent",
-               "HCR2 +30% bnd","HCR2 +30% bnd cap","HCR2 +30% bnd recent","HCR2 +30%",
-               "HCR2 +30% recent")
+**LM of density dependence in weight at age**
 
-df        <- combI %>% 
-  tidyr::pivot_longer(names_to = "variable", values_to = "data", myvar) %>% 
-  filter(stk %in% mystk,  .id %in% scenarios) %>% 
-  mutate(.id = factor(.id, levels=scenarios)) %>% 
-  mutate(variable = factor(variable, levels=myvar)) %>% 
-  mutate(stk = factor(stk, levels=mystk))
+[text here]
 
-calc_boxplot_stat <- function(x) {
-  coef <- 1.5
-  n <- sum(!is.na(x))
-  # calculate quantiles
-  stats <- quantile(x, probs = c(0.0, 0.25, 0.5, 0.75, 1.0))
-  names(stats) <- c("ymin", "lower", "middle", "upper", "ymax")
-  iqr <- diff(stats[c(2, 4)])
-  # set whiskers
-  outliers <- x < (stats[2] - coef * iqr) | x > (stats[4] + coef * iqr)
-  if (any(outliers)) {
-    stats[c(1, 5)] <- range(c(stats[2:4], x[!outliers]), na.rm = TRUE)
-  }
-  return(stats)
-}
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-df %>% 
-  ggplot(aes(x=.id, y=data)) +
-  theme_bw() +
-  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
-  stat_summary(fun.data = calc_boxplot_stat, geom="boxplot") +
-  expand_limits(y=0) +
-  # scale_y_continuous(limits = quantile(t$data, c(0.1, 0.9))) +
-  labs(x="", y="") +
-  facet_grid(variable~stk, scales="free", space="free_x")
+fig_nums(
+  name    = "cond_lmwaa", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". LM analysis of density dependence and weight at age."))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "lm_waa.jpg", sep="_")))
 
 ```
+
+_`r fig_nums("cond_lmwaa")`_
+
+\newpage
+
+**LM summary**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "cond_lmsummary", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))
+
+cat(readLines(file.path(tablesdir, paste(mystk, "lm summary.txt", sep="_"))), sep = '\n')
+
+```
+
+_`r tab_nums("cond_gammcoefficients")`_
+
+\newpage
+
+**DD analysis of maturity at age**
+
+[text here]
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_ddmat", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Analysis of density dependence and maturity at age."))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_mat.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_ddmat")`_
+
+\newpage
+
+**Estimated maturity at weight**
+
+$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
+
+Estimated parameters: k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "cond_matogive", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Estimated maturity ogive at weight"))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat_ogive.jpg", sep="_")))
+
+```
+
+_`r fig_nums("cond_matogive")` _
+
+<!-- =============================================================================================================== -->
+<!-- Section 3: Estimating Density Dependent effects based on equilibrium analyses                                   -->
+<!-- =============================================================================================================== -->
+
+# Estimating Density Dependent effects based on equilibrium analyses
+
+Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
+The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
+
+## Mass-at-age [SHOULD WE CALL IT WEIGHT AT AGE, SINCE WE ARE USING Wa? ]
+
+Mass-at-age ($W_a$) is modelled as
+
+$W_a = {\overline{W_a}} B^{{\beta}_a}$
+
+Where $\overline{W_a}$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data.
+$W_a$ is constrained so sizes do not become unfeasible large or small.
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "sim_dd_mass", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in mass-at-age for different levels of biomass relative to Bmsy")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_mass.jpg")))
+
+```
+
+_`r fig_nums("sim_dd_mass")`_
+
+## Maturity-at-weight
+
+Maturity is then  modelled as a logistic function of mass-at-age i.e.
+
+$O(W_a) = 1/(1+e^{k(W_a-W_{50})})$
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "sim_dd_mat", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in maturity-at-age for different levels of biomass relative to Bmsy")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_maturity.jpg")))
+
+```
+
+_`r fig_nums("sim_dd_mat")`_
+
+## Natural mortality
+
+Is assumed to vary at unit length ($L_1$) e.g.
+
+$ln(M_1) =\alpha + \beta ln(L_{\infty}) + \gamma ln(k)$
+
+where $\alpha=0.65$, $\beta=0.91$, $\gamma=0.87$
+
+Length-at-age ($L_a$) can be obtained from the length-weight relationship
+
+$L_a=(W/a)^{1/b}$
+
+giving
+
+$ln(M_1) =(W/a)^{1/b}(\alpha + \beta ln(L_{\infty}) + \gamma ln(k))$
+
+For DD there is a relationship between Mass at age $M_a$ and total biomass ($B$)
+
+Mackerel:
+
+$M = 0.0232B + 0.0524$
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "sim_dd_m", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_m.jpg")))
+
+```
+
+_`r fig_nums("sim_dd_m")`_
+
+Comparison of equilibrium yield plots for simulation with and without density-dependent processes
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "sim_yield", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_eq_yield.jpg")))
+
+```
+
+_`r fig_nums("sim_dd_m")`_
+
+<!-- =============================================================================================================== -->
+<!-- Section 4: Setting up the Operating Models (OM)                                                                 -->
+<!-- =============================================================================================================== -->
+
+# Setting up the Operating Models (OM)
+
+**Stock assessments**
+
+Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA.
+The VPA reconstructions of the SAM assessment, shows a strongly divergent pattern in recruitment and also some differences in the trend of SSB.
+Mean fishing mortality is somewhat more variable in the VPA reconstruction compared to the SAM assessment.
+Density-dependent processes (on weight, maturity and M) tend to lead to lower SSB in the latter part of the time series.
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_assessments", level = 1, display = FALSE,
+  caption = "Comparison of different assessments. VPA-M is used going forward")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_assessments.jpg")))
+
+```
+
+_`r fig_nums("om_VPADDM")`_
+
+\newpage
+
+**Density dependence in mass-at-age**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_VPADDM", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in Mass-at-age")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDM_4panels.jpg")))
+
+```
+
+_`r fig_nums("om_VPADDM")`_
+
+\newpage
+
+**Density dependence in mass-at-age and maturity at age**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_VPADDMM", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in Mass-at-age and Maturity at age")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMM_4panels.jpg")))
+
+```
+
+_`r fig_nums("om_VPADDMM")`_
+
+\newpage
+
+**Density dependence in mass-at-age, maturity at age and M at age**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_VPADDMMM", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMMM_4panels.jpg")))
+
+```
+
+_`r fig_nums("om_VPADDMMM")`_
+
+\newpage
+
+**Comparison of equilibrium curves for density dependence analyses: biomass against yield**
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_eqyield", level = 1, display = FALSE,
+  caption = "Equilibrium curves for density dependence analyses: biomass against yield")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd.jpg")))
+
+```
+
+_`r fig_nums("om_eqyield")`_
+
+\newpage
+
+<!-- =============================================================================================================== -->
+<!-- Section 5: Running forward projections                                                                          -->
+<!-- =============================================================================================================== -->
+
+# Forward projections without feedback
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+```
+
+<!-- =============================================================================================================== -->
+<!-- Section 6: Discussion and conclusions                                                                           -->
+<!-- =============================================================================================================== -->
+
+# Discussion and conclusions
+
+The quality shall live up to that demanded by the International Council for the Exploration of the Sea (ICES) and potentially be used in ICES advice to fisheries managers.
+The Consultant shall be interacting with the leadership of the project Henrik Sparholt, or his representative, in the process of developing the software.
+The software shall be made publicly available via Github or on alternative platform if so directed by Henrik Sparholt or his representative.
+
+<!-- =============================================================================================================== -->
+<!-- Funding                                                                                                         -->
+<!-- =============================================================================================================== -->
+
+# Funding
+
+<!-- =============================================================================================================== -->
+<!-- References                                                                                                      -->
+<!-- =============================================================================================================== -->
+
+# References
+
+<!-- =============================================================================================================== -->
+<!-- Annex 1: OM conditioning                                                                           -->
+<!-- =============================================================================================================== -->
+
+# Annex 1: OM conditioning
+
+<!-- =============================================================================================================== -->
+<!-- Annex 2: Density dependent processes                                                                           -->
+<!-- =============================================================================================================== -->
+
+# Annex 2: Density dependent processes

### 2023-02-15 Rmd/report.R

@@ -70,13 +70,13 @@ myparams <- data.frame(
   minage  = c(0                            , 1                ),
   maxage  = c(12                           , 10               ),
   maxyear = c(2050                         , 2050             ),
-  m1scaler= c(8                            , 8                ),
-  w50scaler= c(1000                        , 1000             ),
-  matk2   = c(30                           , 30               ) )
+  m1scaler= c(8                            , 10               ),
+  w50scaler= c(1000                        , 500              ),
+  matk2   = c(30                           , 100               ) )
 
 save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-mystk     <- "mac";  
+mystk     <- "whb";  
 # mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   

### 2023-02-16 Rmd/report mac.Rmd
First full draft of report mac.Rmd

@@
 load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
+load(file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-# ------------------------------------------------------------------------------
-# Operating model
-# ------------------------------------------------------------------------------
 figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
@@
 
-```{r}
+The 2022 ICES update assessment of mackerel (ICES 2022) has been carried out by fitting the state-space assessment model SAM (Nielsen and Berg, 2014).
+The update is based on the benchmark protocol as established in 2019 (ICES 2019).
+The assessment model is fit to catch-at-age data for ages 0 to 12 (plus group) for the period 1980 to 2021 (with a strong down-weighting of the catches for the period 1980-1999) and three surveys: 1) SSB estimates from the triennial Mackerel Egg survey (every three years in the period 1992-2022), 2) a recruitment index from the western Europe bottom trawl IBTS Q1 and Q4 surveys (1998-2020, could not be updated for 2021) and 3) the abundance estimates for ages 3 to 11 from the IESSNS survey (2010, 2012-2022).
+The model also incorporates tagging-recapture data from the Norwegian tagging program (for fish recaptured between 1980 and 2005 for the steel tags time series, and fish recaptured between 2014 and 2021 (age 5 and older at release) for the radio frequency tags time series).
+Fishing mortality-at-age and recruitment are modelled as random walks, and there is a process error term on abundances at ages 1-11 (ICES 2022).
 
-# fig_nums(name    = "section2_fig", caption = "section2_fig", level = 1, display = FALSE)
-# tab_nums(name    = "section2_tab", caption = "section2_tab", level = 1, display = FALSE)
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "assessment", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Summary of ICES assessment (ICES 2022)"))
+
+knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "assessment_summary.jpg", sep="_")))
 
@@
 
+_`r fig_nums("assessment")`_
+
+To be able to handle the outputs of the SAM assessment in the FLR framework, a conversion from the SAM object to a FLStock object has been applied.
+Trends in weight at age (mass at age), maturity at age and natural mortality are shown below, with a specific focus on detecting potential density-dependent relationships.
+
 **Weight at age**
 
-[text here]
+As noted in the ICES assessment report, "there is a trend towards lighter weight-at-age for the most age classes (except 0 to 2 years old) starting around 2005, continuing until 2013 (Figure 8.5.2.1).
+This decrease in the catch mean weight-at-age seems to have stopped since 2013 and values for the last seven years do not show any particular trend for the older ages (age 6 and older) and are slightly increasing for younger ages (ages 1 to 5)." (ICES 2022)
 
@@
 
-[text here]
+The maturity ogive for mackerel is calculated as the average of the maturity ogives of the three spawning components of mackerel (Western component, Southern component, North Sea component) weighted by the relative size of each component as estimated during the triennial egg survey.
+The maturity ogives for the North Sea and Southern components are fixed over time.
+For the Western component, the ogive is updated every year, using maturity data from commercial catch samples from Germany, Ireland, the Netherlands and the UK collected during the first and second quarters.
+Most of the variability in maturity is observed for age 2 and to a lesser extent for age 3.
 
@@
 
-[text here]
+Natural mortality is assumed to be 0.15 for all age groups and constant over time (ICES 2022).This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (Hamre, 1980).
 
@@
 
-[text here]
+The distribution of weight at age over the whole time series is shown in the plot below (left part).
+The linear regression of the weight residual by age (weight/mean(weight)) is shown in the right hand part of the plot below.
+From ages 4-12 there are clear indications of a downward relationship between total biomass and weight at age.
+For the younger ages, such a relationship is not clearly apparent.
 
@@
 
-**GAMM at age**
+**GAMM model of weight at age vs. Total biomass**
 
-[text here]
+A Generalized Additive Mixed Model was developed to assessment the relationship between age, weight at age residual and a Thin place regression spline based on the total biomass residual.
+
+mgcViz::gammV(log(data) ~ age + s(tb/mean(tb), bs="tp"), data=wt2)
+
+Results for the weight at age residual are shown below with summaries of the model fits and model coefficients in the tables below that.
+Overall, this suggests that there is a negative relationship between total biomass and weight at age residual, i.e.
+with higher biomass, weight at age tends to be lower.
 
@@
 
-**GAMM summary**
-
-[text here]
-
 ```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 _`r tab_nums("cond_gammsummary")`_
-
-\newpage
-
-**GAMM summary**
-
-[text here]
 
@@
 
-**LM of density dependence in weight at age**
+**Comparing observations and GAMM model estimates of density dependence in weight at age**
 
-[text here]
+As a final check on the GAMM analysis, we plotted the predictions from the GAMM model over the observed weight at age, demonstrating that the model is capturing the overall trend in the data.
 
@@
 
-**LM summary**
+**Analysis of density dependence in maturity at age**
 
-[text here]
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-tab_nums(
-  name    = "cond_lmsummary", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))
-
-cat(readLines(file.path(tablesdir, paste(mystk, "lm summary.txt", sep="_"))), sep = '\n')
-
-```
-
-_`r tab_nums("cond_gammcoefficients")`_
-
-\newpage
-
-**DD analysis of maturity at age**
-
-[text here]
+The distribution of maturity at age over the whole time series is shown in the plot below (left part).
+Only ages 1-3 have proportions mature that are between 0 and 1.
+The linear regression of the maturity residual by age (maturity/mean(maturity)) is shown in the right hand part of the plot below.
+For ages 2 and 3 there are weak indications of a downward relationship between total biomass and maturity.
 
@@
 
-$f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
+Maturity at weight was estimated using the following model:
 
-Estimated parameters: k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`
+$f(wt) = \frac{1}{1+e^{-k(wt-W_{50})}}$
+
+With estimated parameters: k=`r format(matPar$par[1], digits=3, nsmall=2)` $W_{50}$=`r format(matPar$par[2], digits=4)`.
+In the graph below, the blue line represents a simple LOESS smoother while the red line represents the model mentioned above.
 
@@
 
-Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
-The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
+In this section, density dependent effects are assessment based on theoretical relationships established in the scientific literature and data that is derived from FishBase.
+Life history parameters are used to develop an example of modelling density dependence for weight, maturity and natural mortality-at-age.
+Natural mortality parameters m1 and m2 are derived from Lorenzen (1997) where the natural mortality at weight was expressed as $M_{W} = M_{u}W^{b}$ with $M_{u}$ is equal to natural mortality at unit weight.
+Given that the assumed natural mortality of `r mystkname` is equal to 0.15 for all ages, and that the natural mortality at 300 gram bodyweight could be equivalent to the general natural mortality, the formula to derive $M_{u}$ has been implemented as: `r myparams[myparams$stock==mystk,"m1txt"]`
+
+The life-history parameters for the equilibrium analyses can then be summarised as follows:
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "dd_pars", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Life-history parameters derived from FishBase"))
+
+cat(readLines(file.path(tablesdir, paste(mystk, "par.txt", sep="_"))), sep = '\n')
+
+```
+
+_`r tab_nums("dd_pars")`_
+
+The life history parameters are then used to construct a `FLBRP` object representing equilibrium values, and are then coerced into a `FLStock` object to model the time series dynamics.
 
 ## Mass-at-age [SHOULD WE CALL IT WEIGHT AT AGE, SINCE WE ARE USING Wa? ]
+
+**DO WE STILL NEED THIS SECTION?**
 
@@
   name    = "sim_dd_mass", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in mass-at-age for different levels of biomass relative to Bmsy")
+  caption = "Simulation of density dependence in mass-at-age for different levels of biomass relative to Bmsy.
+Each line represents the equilibrium values by age for different levels of fishing mortality (from F=0 to Fcrash).
+The red line corresponds to Fmsy")
 
@@
 
-Maturity is then  modelled as a logistic function of mass-at-age i.e.
+**DO WE STILL NEED THIS SECTION?**
+
+Maturity is modelled as a logistic function of mass-at-age i.e.
 
@@
   name    = "sim_dd_mat", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in maturity-at-age for different levels of biomass relative to Bmsy")
+  caption = "Simulation of density dependence in maturity-at-age for different levels of biomass relative to Bmsy.
+Each line represents the equilibrium values by age for different levels of fishing mortality (from F=0 to Fcrash).
+The red line corresponds to Fmsy")
 
@@
 ## Natural mortality
+
+**DO WE STILL NEED THIS SECTION?**
 
@@
 
-$M = 0.0232B + 0.0524$
+$M = 0.0232B + 0.0524$  **HOW DOES THIS RELATE TO THE VALUES USED IN THE REPORT?**
 
@@
 _`r fig_nums("sim_dd_m")`_
+
+\newpage
 
@@
 
-Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA.
+The initial stock assessments was carried out with SAM and a fixed M by years and ages.
+Because SAM allows for deviation from the catch equation, there can be strong discrepancies between the estimated recruitment by year and the materialized recruitment as evident from the appearance in the catches.
+This could lead to spurious recruitment patterns.
+To circumvent that potential problem in the simulations, we estimated a VPA model based on SAM assessment, using the terminal Fs or Ns from the SAM assessment as inputs.
+We then ran that VPA model also with age-varying M to assessment whether we could reproduce the overall equilibrium curves for total biomass and yield.
+
 The VPA reconstructions of the SAM assessment, shows a strongly divergent pattern in recruitment and also some differences in the trend of SSB.
 Mean fishing mortality is somewhat more variable in the VPA reconstruction compared to the SAM assessment.
-Density-dependent processes (on weight, maturity and M) tend to lead to lower SSB in the latter part of the time series.
 
@@
 
+We then explored the impacts of density dependence in mass at age on the equilibrium values of total biomass and yield.
+In the plot below, there are multiple lines for mass at age (one line for each value of F).
+Maturity and M are fixed values by age.
+The equilibrium curves are on the bottom right: the red line is the equilibrium curve for mass at age, while the black line is with age-varying M but without density dependence.
+
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 **Density dependence in mass-at-age and maturity at age**
+
+Next, we explored the impacts of density dependence in mass at age and maturity at age on the equilibrium values of total biomass and yield.
+In the plot below, there are multiple lines for mass at age and maturity at age (one line for each value of F).
+Natural mortality are fixed values by age.
+The equilibrium curves are on the bottom right: the red line is the equilibrium curve for mass at age and maturity at age, while the black line is with age-varying M but without density dependence
 
@@
 
+Finslly, we explored the impacts of density dependence in mass at age and maturity at age and Natural mortality at age on the equilibrium values of total biomass and yield.
+In the plot below, there are multiple lines for mass at age, maturity at age and natural mortality at age (one line for each value of F).
+The equilibrium curves are on the bottom right: the red line is the equilibrium curve for mass at age and maturity at age and natural mortality at age, while the black line is with age-varying M but without density dependence.
+
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 
+Introducing density-dependent process has an impact on the estimate Bmsy (calculated from Total biomass) which is shifted to the left. The expected impact on the equilibrium yield is relatively small.
+
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 \newpage
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "om_eqF", level = 1, display = FALSE,
+  caption = "Equilibrium curves for density dependence analyses: biomass against F")
+
+knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd_F.jpg")))
+
+```
+
+_`r fig_nums("om_eqF")`_
+
+\newpage
+
+**Overview of MSY values from different scenarios**
+
+```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "msy", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". MSY values from different scenarios"))
+
+cat(readLines(file.path(tablesdir, paste(mystk, "msy.txt", sep="_"))), sep = '\n')
+
+```
+
+_`r tab_nums("dd_pars")`_
 
@@
 
+Hamre, J. 1980. Biology, exploitation and management of the NorthEast Atlantic mackerel. Rapp P-V Reun (CIEM), 177, 212-242.
+
+ICES (2019). Interbenchmark workshop on the Assessment of northeast atlantic Mackerel (IBPNEAMAC), 4-7 March 2019. ICES Volume  1, Issue 5: 71 pp.
+
+ICES (2022). Working Group on Widely Distributed Stocks (WGWIDE 2022) Volume 4, Issue 73.
+
+Lorenzen, K.
+(1997).
+"The relationship between body weight and natural mortality in juvenile and adult fish: a comparison of natural ecosystems and aquaculture." Oceanographic Literature Review 4(44): 339.
+
+Nielsen, A. and C. W. Berg (2014). "Estimation of time-varying selectivity in stock assessments using state-space models." Fisheries Research 158(0): 96-101.
+
 <!-- =============================================================================================================== -->

### 2023-02-16 Rmd/report.R

@@ -11,9 +11,6 @@
 
 rm(list=ls())
 
-iFig=0
-iTab=0
-
 library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
 library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
 library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
@@ -56,8 +53,9 @@ myparams <- data.frame(
   stock    = c("mac"                       , "whb"            ),
   mystkname=c("Northeast Atlantic mackerel", "Blue whiting"   ),
   mylatin = c("Scomber scombrus"           ,"Micromesistius poutassou"),
-  m1      = c(0.15/(300^-0.288)            , 0.2/(150^-0.288) ), # 300 g MAC, 150 g WHB 
-  m2      = c(-0.288                        , -0.288          ),
+  m1txt   = c("0.15/(300^-0.288)"          , "0.2/(150^-0.288)"),
+  m1      = c( 0.15/(300^-0.288)           ,  0.2/(150^-0.288) ), # 300 g MAC, 150 g WHB 
+  m2      = c(-0.288                       , -0.288          ),
   m3      = c(NA                           , NA               ),
   m       = c("lorenzen"                   , "lorenzen"       ),
   blim    = c(2000000                      , 1500000          ),
@@ -76,7 +74,7 @@ myparams <- data.frame(
 
 save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-mystk     <- "whb";  
+mystk     <- "mac";  
 # mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   
@@ -104,9 +102,9 @@ mystk     <- "whb";
   stk   = get(mystk)
   stkR  = get(paste0(mystk,"R"))
   
-
   if (mystk == "mac") {
-    ts    = as.data.frame(readxl::read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),sheet="ts"))
+    # ts    = as.data.frame(readxl::read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),sheet="ts"))
+    ts=model.frame(FLQuants(stk,tb=stock,ssb=ssb),drop=T)
     
[…]

### 2023-02-18 Rmd/report mac.Rmd
with ref pts

--- 2023-02-15 Rmd/report mac.Rmd
@@
 source("../R/pm.R")
+```
 
-source("../R/get_dropbox.r")
+```{r}
+source("../R/get_dropbox.R")
+
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
 fig_nums(
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
 
@@
 knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd.jpg")))
-
 ```

### 2023-02-18 R/get_dropbox.R

@@ -1,22 +1,0 @@
-# Taken from: https://www.r-bloggers.com/finding-my-dropbox-in-r/
-
-get_dropbox <- function () {
-  
-  # install.packages("rjson")
-  if (Sys.info()['sysname'] == 'Darwin') {
-    info <- RJSONIO::fromJSON(
-      file.path(path.expand("~"),'.dropbox','info.json'))
-  }
-  if (Sys.info()['sysname'] == 'Windows') {
-    info <- RJSONIO::fromJSON(
-      if (file.exists(file.path(Sys.getenv('APPDATA'), 'Dropbox','info.json'))) {
-        file.path(Sys.getenv('APPDATA'), 'Dropbox', 'info.json')
-      } else {
-        file.path(Sys.getenv('LOCALAPPDATA'),'Dropbox','info.json')
-      }
-    )
-  }
-  
-  dir <- info$personal$path
-  return(dir)
-}

### 2023-02-18 Rmd/report.R

@@ -493,10 +493,18 @@ mystk     <- "whb";
   
   df <- 
     bind_rows(
-      model.frame(FLQuants(eq,biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T) %>% mutate(scen="no-DD"),
-      model.frame(FLQuants(x, biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T) %>% mutate(scen="DD mass")
+      model.frame(FLQuants(eq,ssb    =function(x) ssb(x),
+                              biomass=function(x) biomass(x), 
+                              catch  =function(x) catch(x),
+                              f      =function(x) fbar(x)),
+                           drop=T) %>% mutate(scen="no-DD"),
+      model.frame(FLQuants(x, ssb    =function(x) ssb(x),
+                              biomass=function(x) biomass(x), 
+                              catch  =function(x) catch(x),
+                              f      =function(x) fbar(x)),
+                  drop=T) %>% mutate(scen="DD mass")
     )
-  
+
   # Comparison of equilibrium yields curves with DD mass+maturity
   
   x=propagate(eq,length(dimnames(eq)$year))
@@ -521,9 +529,13 @@ mystk     <- "whb";
   df <- 
     bind_rows(
       df,
-      model.frame(FLQuants(x, biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T) %>% mutate(scen="DD mass+mat")
+      model.frame(FLQuants(x,  ssb    =function(x) ssb(x),
+                               biomass=function(x) biomass(x), 
+                               catch  =function(x) catch(x),
+                               f      =function(x) fbar(x)),
+                  drop=T) %>% mutate(scen="DD mass+mat")
     )
-  
+ 
   # Comparison of equilibrium yields curves with DD mass+mat+M 
   
   x=propagate(eq,length(dimnames(eq)$year))
@@ -546,8 +558,13 @@ mystk     <- "whb";
   df <- 
     bind_rows(
       df,
-      model.frame(FLQuants(x, biomass=function(x) biomass(x), catch=function(x) catch(x)),drop=T) %>% mutate(scen="DD mass+mat+M")
+      model.frame(FLQuants(x, ssb    =function(x) ssb(x),
+                           biomass=function(x) biomass(x), 
[…]

### 2023-02-20 Rmd/mseMac.Rmd
1st run

--- 2023-01-10 Rmd/mseMac.Rmd
@@
 title: "**Density Dependence**"
-subtitle: "Example of MSE for Mackerel"
+subtitle: "MSE for Mackerel"
 author: "L Kell"
@@
 ---
+
+## Conditioning procedure
+
++  Load ICES assessment
++  Time varying M
++  Conduct VPA
 
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/mseMac/",
-               fig.path  ="../tex/figs/mseMac/",
+               cache.path="../cache/mse/",
+               fig.path  ="../tex/figs/mse/",
                fig.width =8,
@@
 ```{r, pkgs}
-library(FLCore)
-library(FLBRP)
-library(FLasher)
-library(FLAssess)
-library(FLife)
+library(FLCore)    # install.packages("FLCore", repos="http://flr-project.org/R")
+library(FLBRP)     # install.packages("FLBRP", repos="http://flr-project.org/R")
+library(FLasher)   # install.packages("FLasher", repos="http://flr-project.org/R")
+library(FLAssess)  # install.packages("FLAssess", repos="http://flr-project.org/R")
+library(FLife)     # install.packages("FLife", repos="http://flr-project.org/R")
 library(ggplotFL)
@@
 
-```{r, mac.data}
-load(file.path(dropboxdir,"data/om/mac.RData"))
+```{r, ices}
+stock="Mac"
+load(paste("../data/om/ices",stock,".RData",sep=""))
+stk=get(tolower(stock))
 
-par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bmsy=3500000,b=0.3,k=30.1, w50=0.2))
+stock.wt(   stk)[1]=0.01
+catch.wt(   stk)[1]=0.01
+landings.wt(stk)[1]=0.01
+discards.wt(stk)[1]=0.01
+```
 
-stock.wt(   mac)[1]=0.01
-catch.wt(   mac)[1]=0.01
-landings.wt(mac)[1]=0.01
-discards.wt(mac)[1]=0.01
+```{r, pars}
+par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bref=NA,b=0.3,matk=30.1, w50=0.2))
+```
 
-om=ddFn2(mac,par)
-om=om+VPA(om)
-om=ddFn2(om,par)
+```{r, m}
+m(stk)=par["m1"]%*%(stock.wt(stk)%^%par["m2"])
+vpa=stk+VPA(stk)
+```
 
-sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
-eq=FLBRP(om,sr=sr)
+```{r, vpa}
+plot(FLStocks("ICES"=stk,"VPA"=vpa))
+```
 
-om=fwdWindow(om,end=2080,eq)
-f =fbar(om)[,ac(2021:2080)]%=%refpts(eq)["msy","harvest"]
-om=fwd(om,fbar=f,sr=eq)
+```{r,eq}
+sr=fmle(as.FLSR(iter(vpa,1),model="bevholt"),control=list(silent=TRUE))
+eq=FLBRP(iter(vpa,1),sr=sr)
+```
 
-om2=om
-om3=om
-om4=om
-for (iYear in ac(2021:2080)){
-  om2=ddFn(iYear,om2,par)
-  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)
+```{r, ddM}
+par["bref"]=refpts(eq)["msy","biomass"]
 
-  om3=ddFn(iYear,om3,par)
-  om3=fwd(om3,fbar=f[,iYear],sr=eq)
-
-  om4=ddFn(iYear,om4,par)
-  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
-  }
-
-plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
+for (i in dimnames(vpa)$year)
+  om=ddFn(i,stk,par,FALSE,FALSE)
 ```
@@
 ```{r}
-ggplot(FLQuants("ICES"=apply(stock.wt(mac[,"2021"]),1,mean),
-                "OM"  =stock.wt(om[ ,"2080"]),
-                "0.8" =stock.wt(om2[,"2080"]),
-                "1.0" =stock.wt(om3[,"2080"]),
-                "1.2" =stock.wt(om4[,"2080"])))+
-  geom_line(aes(age,data,col=qname))+
-  xlab("Age")+ylab("Mass-at-age")
+save(om,eq,file="~/Desktop/tmp/t.RData")
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age
+```{r}
+load("~/Desktop/tmp/t.RData")
+```
+
+```{r, omCheck}
+om=fwdWindow(window(om,end=2020),end=2080,eq)
+f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.5,length.out=100),each=61),dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
+om=fwd(om,fbar=f,sr=eq)
+
+oms=FLStocks("Base"     =om)
+oms[["DD Mass"]]        =om
+oms[["DD Mass, Mat"]]   =om
+oms[["DD Mass, Mat, M"]]=om
+for (iYear in ac(2021:2080)){
+  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
+  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eq)
+
+  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],par,TRUE,TRUE,FALSE)
+  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eq)
+
+  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],par,TRUE,TRUE,TRUE)
+  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eq)
+  }
+```
+
+```{r, check2}
+p1=ggplot(stock.wt(oms[[2]][,ac(2080)]))+ 
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mass-at-age") 
+p2=ggplot(mat(oms[[2]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mat-at-age")
+p3=ggplot(m(oms[[2]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("M-at-age")
+
+d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+d2=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+p4=ggplot(d2)+
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+
+ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence
+
+```{r, check3}
+p1=ggplot(stock.wt(oms[[3]][,ac(2080)]))+ 
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mass-at-age") 
+p2=ggplot(mat(oms[[3]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mat-at-age")
+p3=ggplot(m(oms[[3]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("M-at-age")
+
+d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+d3=model.frame(FLQuants(oms[[3]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+p4=ggplot(d3)+
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+
+ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence
+
+```{r, check4}
+p1=ggplot(stock.wt(oms[[4]][,ac(2080)]))+ 
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mass-at-age") 
+p2=ggplot(mat(oms[[4]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mat-at-age")
+p3=ggplot(m(oms[[4]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("M-at-age")
+
+d1=model.frame(FLQuants(oms[[4]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+d4=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+p4=ggplot(d4)+
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+
+ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence
+
+```{r, eqlCurves}
+rfs=ldply(oms, function(x) { 
+  model.frame(FLQuants(x[,"2080"],
+             biomass=function(x) stock(x),
+             ssb    =function(x) ssb(  x),
+             catch  =function(x) catch(x),
+             f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+
+ggplot(rfs)+
+  geom_line(aes(ssb,catch,col=.id))+
+  xlab("SSB")+ylab("Catch")+
+  theme(legend.position="bottom")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for assumptions about density dependence
+
+**Table `r iTab=iTab+1; iTab`** Reference points
+
+```{r, refptsTable}
+rfpts=ddply(rfs,.(.id), with, {  
+  flag=catch==max(catch)
+  data.frame(Bmsy  =biomass[flag],
+             SSBmsy=ssb[flag],
+             MSY   =catch[flag],
+             Fmsy  =f[flag],
+             Virgin=max(ssb),
+             B0    =max(biomass))})
+
+rfpts
+```
+
+```{r, projectons}
+prj=list("1"=iter(oms[[1]],1:4),
+         "2"=iter(oms[[2]],1:4),
+         "3"=iter(oms[[3]],1:4),
+         "4"=iter(oms[[4]],1:4))
+
+F=fbar(prj[[1]][,ac(2021:2080)])%=%rep(rfpts$Fmsy,each=60)
+
+control=as(FLQuants("f"=F),"fwdControl")
+prj[[1]] =fwd(prj[[1]],control=control,sr=eq)
+
+for (i in ac(2021:2080)) {
+    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    control=as(FLQuants("f"=F[,i]),"fwdControl")
+    prj[[2]] =fwd(prj[[2]],control=control,sr=eq)
+
+    prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    control=as(FLQuants("f"=F[,i]),"fwdControl")
+    prj[[3]] =fwd(prj[[3]],control=control,sr=eq)
+
+    prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+    control=as(FLQuants("f"=F[,i]),"fwdControl")
+    prj[[4]] =fwd(prj[[4]],control=control,sr=eq)
+    }
+```
 
 ```{r}
-ggplot(FLQuants("ICES"=m(mac[,"2021"]), 
-                "OM"  =m(om[ ,"2080"]),
-                "0.8" =m(om2[,"2080"]),
-                "1.0" =m(om3[,"2080"]),
-                "1.2" =m(om4[,"2080"])))+
-  geom_line(aes(age,data,col=qname))+
-  xlab("Age")+ylab("M-at-age")
+plot(FLStocks(prj[[1]]),iter=1:4)
 ```
 
-**Figure `r iFig=iFig+1; iFig`** M-at-age
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming no density dependence
 
 ```{r}
-ggplot(FLQuants("ICES"=apply(mat(mac[,"2021"]),1,mean),
-                "OM"  =mat(om[ ,"2080"]),
-                "0.8" =mat(om2[,"2080"]),
-                "1.0" =mat(om3[,"2080"]),
-                "1.2" =mat(om4[,"2080"])))+
-  geom_line(aes(age,data,col=qname))+
-  xlab("Age")+ylab("Maturity-at-age")
+names(prj)=names(oms)
+plot(FLStocks(prj[[2]]),iter=1:4)
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass-at-age only
+
+```{r}
+plot(FLStocks(prj[[3]]),iter=1:4)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass and maturity-at-age only
+
+```{r}
+plot(FLStocks(prj[[4]]),iter=1:4)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass, maturity and natural mortality-at-age
+
+```{r, eval=FALSE}
+ldply(prjs,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
+                                                     ssb    =function(x) ssb(    x)[,"2080"],
+                                                     catch  =function(x) catch(  x)[,"2080"],
+                                                     f      =function(x) fbar(   x)[,"2080"]),drop=T)) 
+```
+
+```{r, MSE, eval=FALSE}
+mse=hcrICES(fmsy,eql,rec(fmsy),
+                      par,
+                      start,   end,
+                      interval,lag=lag,
+                      err=err,
+                      bndTac=c(0,Inf))
+```

### 2023-02-20 Rmd/mseMac.Rmd
1st

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: html_document
+output: pdf_document
 mathjax: TRUE

### 2023-02-20 Rmd/mseMac.Rmd
with SRR

--- 2023-02-20 Rmd/mseMac.Rmd
@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/mse/",
-               fig.path  ="../tex/figs/mse/",
+               cache.path="../cache/mseMac/",
+               fig.path  ="../tex/figs/mseMac/",
                fig.width =8,
@@
 ```{r, pars}
-par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bref=NA,b=0.3,matk=30.1, w50=0.2))
+par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,b=0.3,matk=30.1,w50=0.2))
 ```
@@
 ```{r,eq}
-sr=fmle(as.FLSR(iter(vpa,1),model="bevholt"),control=list(silent=TRUE))
+sr=fmle(as.FLSR(iter(stk,1),model="bevholt"),control=list(silent=TRUE))
 eq=FLBRP(iter(vpa,1),sr=sr)
@@
   }
+```
+
+```{r}
+save(oms,file="~/Desktop/tmp/oms.RData")
 ```
@@
 ```
+
+```{r}
+load("/home/laurie/Dropbox/pelagics/results/mac/figures/mac_objects.RData")
+
+dt1=model.frame(FLQuants("Rec"=rec(oms[[1]][,"2050"]),"SSB"=ssb(oms[[1]][,"2050"])),drop=T)
+dt2=model.frame(FLQuants("Rec"=rec(vpa[,"2050"]),"SSB"=ssb(vpa[,"2050"])),drop=T)
+
+ggplot()+
+  geom_line(aes(SSB,Rec),data=dt1)+
+  geom_line(aes(SSB,Rec),data=dt2)
+```
+
+```{r}
+stk=window(iter(vpa,1),end=2019)
+sr=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eq=FLBRP(stk,sr=ab(sr))
+
+stk=window(iter(vpaM,1),end=2019)
+srM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqM=FLBRP(stk,sr=ab(srM))
+
+stk=window(iter(vpaDDM,1),end=2019)
+srDDM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDM=FLBRP(stk,sr=ab(srDDM))
+
+stk=window(iter(vpaDDMM,1),end=2019)
+srDDMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDMM=FLBRP(stk,sr=ab(srDDMM))
+
+stk=window(iter(vpaDDMMM,1),end=2019)
+srDDMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDMMM=FLBRP(stk,sr=ab(srDDMMM))
+
+dat=rbind(
+  cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA",     model.frame(FLQuants(eqM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA M",   model.frame(FLQuants(eqDDM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA MM",  model.frame(FLQuants(eqDDMM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA MMM", model.frame(FLQuants(eqDDMMM,rec=rec,ssb=ssb,catch=catch),drop=T)))
+
+ggplot(dat)+
+  geom_line(aes(ssb,catch,col=What))
+
+ggplot(dat)+
+  geom_line(aes(ssb,rec,col=What))
+
+```

### 2023-02-21 Rmd/report.R
Updated report.R

@@ -367,7 +367,10 @@ mystk     <- "mac";
   
   # derive a FLBRP object with Lorenzen natural mortality
   eq =FLife::lhEql(par,m=myparams[myparams$stock==mystk,"m"])
-  
+
+  # sr  = fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(ices))),control=list(silent=TRUE))
+  # eq2  = FLBRP(ices,ab(sr))
+  # plot(eq, eq2)
   
   # plot the BRP object --------------------------------------------------------
   p <-ggplotFL::plot(eq) + 
@@ -647,10 +650,6 @@ mystk     <- "mac";
   # 4. OM
   # ------------------------------------------------------------------------------
 
-  # load(file.path(dropboxdir,"data/inputs/ices.RData"))
-  # ices=ices[["mac.27.nea"]]
-  # ices=window(ices,start=1991)
-  
   ices=window(stk,start=1991)
   stock(ices) =computeStock(ices)
   name(ices) <- mystkname
@@ -663,8 +662,9 @@ mystk     <- "mac";
                                "Rec"=function(x) rec(x)),drop=T) %>% mutate(scen="sam")
   
   # Start with the SAM assessment in forward projection
-  sr  = as.FLSR(ices,model=bevholtSV)
-  sr  = fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(ices))),control=list(silent=TRUE))
+  sr  = fmle(as.FLSR(window(ices, start=1999, end=2020),model="bevholtSV"),
+             fixed=list(s=0.8, spr0=mean(spr0(ices))), 
+             control=list(silent=TRUE))
   eq  = FLBRP(ices,ab(sr))
   sam = fwdWindow(ices,eq,end=2050) 
   
@@ -717,9 +717,10 @@ mystk     <- "mac";
                                 "Yield"=function(x) catch(x),
                                 "Rec"=function(x) rec(x)),drop=T) %>% mutate(scen="vpa") )
     
-  
-  sr  =as.FLSR(vpa,model=bevholtSV)
-  sr  =fmle(sr,fixed=list(s=0.8,spr0=mean(spr0(vpa))),control=list(silent=TRUE))
[…]

### 2023-02-21 Rmd/report.R
started on forward simulations

@@ -682,25 +682,6 @@ mystk     <- "mac";
                                     "Yield"=function(x) catch(x)),drop=T) %>% mutate(scen="sam")
                 
   
-  # t <-
-  #   model.frame(FLQuants(sam,"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
-  #                      "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T) 
-  # tt <- t %>% filter(year==2050)
-  # t %>%   
-  #   ggplot(aes(x=year, y=Biomass, group=iter)) +
-  #   theme_bw() +
-  #   geom_line() +
-  #   geom_text(data=tt, aes(label=iter), hjust=0)
-  
-  # t  <- as.data.frame(F) %>% filter(data <= 0.8)
-  # tt <- t %>% filter(year==2050)
-  # 
-  # t %>%   
-  #   ggplot(aes(x=year, y=data, group=iter)) + 
-  #   theme_bw() + 
-  #   geom_line() +
-  #   geom_text(data=tt, aes(label=iter), hjust=0)
-  
   # VPA assessment based on SAM assessment -------------------------------------
   
   # run VPA
@@ -739,22 +720,6 @@ mystk     <- "mac";
                                        "F"=function(x) fbar(x),
                                        "Yield"=function(x) catch(x)),drop=T) %>% mutate(scen="vpa"))
 
-  # model.frame(FLQuants(vpa,"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
-  #                      "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T) %>%
-  #   ggplot(aes(x=year, y=Biomass, group=iter)) +
-  #   theme_bw() +
-  #   geom_line()
-
-  # t <-
-  #   model.frame(FLQuants(vpa,"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
-  #                        "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T) 
-  # tt <- t %>% filter(year==2050)
-  # t %>%   
-  #   ggplot(aes(x=year, y=Biomass, group=iter)) +
-  #   theme_bw() +
[…]

### 2023-02-21 Rmd/report.R
Almost done with mse

@@ -1097,7 +1097,7 @@ mystk     <- "mac";
                Virgin=max(ssb),
                B0    =max(biomass))})
   
-  rfpts  
+  rfpts$Fmsy  
   
 
   # forward simulations
@@ -1132,10 +1132,10 @@ mystk     <- "mac";
   plot(FLStocks(prj[[1]]),iter=1:4)
   plot(FLStocks(prj[[2]]),iter=1:4)
   plot(FLStocks(prj[[3]]),iter=1:4)
-  plot(FLStocks(prj[[4]]),iter=1:4)
+  plot(FLStocks(prj[[4]]),iter=1:4) 
 
   plot(FLStocks(prj),iter=1:4)
-  plot(FLStocks(prj),iter=1)
+  plot(FLStocks(prj),iter=1:4) +scale_fill_manual(values=rep("white",4))
   
   # mse=hcrICES(fmsy,eql,rec(fmsy),
   #             par,
@@ -1143,7 +1143,31 @@ mystk     <- "mac";
   #             interval,lag=lag,
   #             err=err,
   #             bndTac=c(0,Inf))
+
+  as.data.frame(FLStocks(prj)) %>% 
+    filter(slot %in% c("catch", "stock")) %>% 
+    ggplot(aes(x=year, y=data)) +
+    theme_bw() +
+    geom_line(aes(colour=iter)) +
+    facet_grid(slot ~cname , scales="free_y")
+  
+  plot(propagate(fbar(prj[[1]]),4)%*%FLQuant(seq(0.5,1.5,length.out=4),
+                                             dimnames=list(iter=1:4)),
+       probs=rep(0,5), 
+       worm=1:4)
   
 # } # end of stk loop
 
+df=ldply(prj, function(x) { 
+  model.frame(FLQuants(x,
+                       biomass=function(x) stock(x),
+                       ssb    =function(x) ssb(  x),
+                       catch  =function(x) catch(x),
+                       f      =function(x) fbar( x)),drop=TRUE)})
 
+df %>% 
+  tidyr::pivot_longer(names_to = "variable", values_to = "data", biomass:f) %>% 
+  ggplot(aes(x=year, y=data)) +
+  theme_bw() +
+  geom_line(aes(colour=iter)) +
+  facet_grid(variable ~.id , scales="free_y")

### 2023-02-22 Rmd/mseMac.Rmd
with alpha set by data

--- 2023-02-20 Rmd/mseMac.Rmd
@@
 ```{r, pars}
-par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.28,bref=NA,b=0.3,matk=30.1, w50=0.2))
+par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,b=-0.3,matk=30.1,w50=0.2))
 ```
@@
 
-```{r,eq}
-sr=fmle(as.FLSR(iter(vpa,1),model="bevholt"),control=list(silent=TRUE))
-eq=FLBRP(iter(vpa,1),sr=sr)
+```{r, sr}
+load(paste("/home/laurie/Dropbox/pelagics/results/mac/figures/",tolower(stock),"_objects.RData",sep=""))
+
+stk=window(iter(vpa,1),start=2000,end=2019)
+sr=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eq=FLBRP(stk,sr=ab(sr))
+
+stk=window(iter(vpaM,1),start=2000,end=2019)
+srM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqM=FLBRP(stk,sr=ab(srM))
+
+stk=window(iter(vpaDDM,1),start=2000,end=2019)
+srDDM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDM=FLBRP(stk,sr=ab(srDDM))
+
+stk=window(iter(vpaDDMM,1),start=2000,end=2019)
+srDDMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDMM=FLBRP(stk,sr=ab(srDDMM))
+
+stk=window(iter(vpaDDMMM,1),start=2000,end=2019)
+srDDMMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDMMM=FLBRP(stk,sr=ab(srDDMMM))
+
+dat=rbind(
+  cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA",     model.frame(FLQuants(eqM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA M",   model.frame(FLQuants(eqDDM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA MM",  model.frame(FLQuants(eqDDMM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA MMM", model.frame(FLQuants(eqDDMMM,rec=rec,ssb=ssb,catch=catch),drop=T)))
 ```
 
-```{r, ddM}
-par["bref"]=refpts(eq)["msy","biomass"]
-
-for (i in dimnames(vpa)$year)
-  om=ddFn(i,stk,par,FALSE,FALSE)
+```{r, eqs}
+ggplot(dat)+
+  geom_line(aes(ssb,catch,col=What))
 ```
 
-```{r}
-save(om,eq,file="~/Desktop/tmp/t.RData")
+**Figure `r iFig=iFig+1; iFig`** Stock recruitment relationships
+
+```{r, srs}
+ggplot(dat)+
+  geom_line(aes(ssb,rec,col=What))
 ```
 
-```{r}
-load("~/Desktop/tmp/t.RData")
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for assumptions about density dependence
+
+```{r, ddM}
+par["bref"]=refpts(eqM)["msy","biomass"]
+
+for (i in dimnames(vpaM)$year)
+  om=ddFn(i,vpaM,par,FALSE,FALSE,TRUE)
 ```
@@
 oms[["DD Mass, Mat, M"]]=om
-for (iYear in ac(2021:2080)){
+for (iYear in ac(2020:2080)){
   oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
@@
 
-```{r, projectons}
-prj=list("1"=iter(oms[[1]],1:4),
+```{r, projectons-1}
+prj=FLStocks("1"=iter(oms[[1]],1:4),
              "2"=iter(oms[[2]],1:4),
@@
 
-F=fbar(prj[[1]][,ac(2021:2080)])%=%rep(rfpts$Fmsy,each=60)
+F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)
 
@@
 
-for (i in ac(2021:2080)) {
+for (i in ac(2020:2080)) {
     prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
@@
 ```{r, eval=FALSE}
-ldply(prjs,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
+ldply(prj,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
                                                     ssb    =function(x) ssb(    x)[,"2080"],
@@
 ```
+
+```{r, projectons-2}
+devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)
+
+base=propagate(iter(prj[[1]],1),100)
+
+prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {
+
+            ## FMSY estimate by OM ######################################
+            F=FLQuant(F,dimnames=list(year=2020:2080))
+
+            ## No DD ####################################################
+            control=as(FLQuants("f"=F),"fwdControl")
+            x1     =fwd(base,control=control,sr=eq,residuals=devRec)
+
+            ## With DD ####################################################
+            x2=x1
+            x3=x1
+            x4=x1
+            for (iYr in ac(2020:2080)) {
+              ## DD M ##################################################         
+              x2 =ddFn(iYr,x2,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+              x2 =fwd(x2,control=control,sr=eq,residuals=devRec)
+
+              ## DD MM##################################################
+              x3 =ddFn(iYr,x3,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+              x3 =fwd(x3,control=control,sr=eq,residuals=devRec)
+
+              ## DD MMM ################################################
+              x4 =ddFn(iYr,x4,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+              x4 =fwd(x4,control=control,sr=eq,residuals=devRec)
+              }  ## year loop
+
+            FLStocks(list("Base"=x1,"M"=x2,"MM"=x3,"MMM"=x4))        
+    }) ## F loop
+
+save(prjs,file=paste(dirRuns,"/",stock,"Prjs.RData",sep=""))
+```
+
+```{r}
+plot(FLStocks(prjs[[1]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at Base $F_{MSY}$
+
+```{r}
+plot(FLStocks(prjs[[2]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, with  $F_{MSY}$ for density dependence in mass-at-age
+
+```{r}
+plot(FLStocks(prjs[[3]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity-at-age
+
+```{r}
+plot(FLStocks(prjs[[4]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity, natural mortality-at-age
 
@@
 ```
+
+```{r, chk}
+dt1=model.frame(FLQuants("Rec"=rec(oms[[1]][,"2050"]),"SSB"=ssb(oms[[1]][,"2050"])),drop=T)
+dt2=model.frame(FLQuants("Rec"=rec(vpa[,"2050"]),"SSB"=ssb(vpa[,"2050"])),drop=T)
+
+ggplot()+
+  geom_line(aes(SSB,Rec),data=dt1)+
+  geom_line(aes(SSB,Rec),data=dt2)
+```

### 2023-02-22 R/dd.R

@@ -3,7 +3,6 @@ ddWt<-function(wt,biomass,bref,alpha=-0.2,min=0.5,max=1.5){
   
   wt%*%qmin(qmax(adj,min),max)}
 
-
 ddMat<-function(wt,x) 
   1/(1+exp(-x[1]%*%(wt%-%x[2])))
 
@@ -17,7 +16,7 @@ ddFn<-function(year,x,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE){
     dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
     dsr[!is.finite(dsr)]=0
     
-    stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bref"])
+    stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bref"],par["b"])
     landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
     discards.wt(x)[,year]=stock.wt(x)[,year]*dsr}
     

### 2023-02-25 Rmd/report.R
report.R development; not working?

@@ -87,6 +87,9 @@ mystk     <- "mac";
   
   figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
   tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
+
+  save(.,
+       file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
   
   # ------------------------------------------------------------------------------
   # 1. Introduction
@@ -95,6 +98,10 @@ mystk     <- "mac";
   # ------------------------------------------------------------------------------
   # 2. Conditioning
   # ------------------------------------------------------------------------------
+  
+  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
+  
+  section <- "02"
   
   # Get FLStock and FLBRP object  ----------------------------------------------
   
@@ -152,7 +159,7 @@ mystk     <- "mac";
     # guides(colour = guide_legend(nrow = 1)) +
     scale_color_manual("Age",values=rainbow(length(unique(wt$age))))
   
-  jpeg(filename=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")),
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "waa.jpg", sep="_")),
        width=10, height=6, units="in", res=300)
   print(p) 
   dev.off()
@@ -172,7 +179,7 @@ mystk     <- "mac";
     scale_color_manual("Age",values=rainbow(length(unique(mat$age))))
   
 
-  jpeg(filename=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")),
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "mat.jpg", sep="_")),
        width=10, height=6, units="in", res=300)
   print(p) 
   dev.off()
@@ -192,22 +199,17 @@ mystk     <- "mac";
     # guides(colour = guide_legend(nrow = 1)) +
     scale_color_manual("Age",values=rainbow(length(unique(m$age))))
   
-  jpeg(filename=file.path(figuresdir, paste(mystk, "m.jpg", sep="_")),
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "m.jpg", sep="_")),
        width=10, height=6, units="in", res=300)
   print(p) 
   dev.off()
[…]

### 2023-02-27 Rmd/report.R
report.R

@@ -95,6 +95,22 @@ mystk     <- "mac";
   # 1. Introduction
   # ------------------------------------------------------------------------------
   
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
   # ------------------------------------------------------------------------------
   # 2. Conditioning
   # ------------------------------------------------------------------------------
@@ -340,6 +356,23 @@ mystk     <- "mac";
   save(.,
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section2.RData", sep="_")))
 
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
   # ----------------------------------------------------------------------------
   # 3. Density dependence and theoretical OM
   # ----------------------------------------------------------------------------
@@ -625,18 +658,33 @@ mystk     <- "mac";
   #      file = file.path(dropboxdir, "results", mystk, paste(mystk,"section3.RData", sep="_")))
 
 
-  # ------------------------------------------------------------------------------
+
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+    # ------------------------------------------------------------------------------
   # 4. OM based on ICES assessment
   # ------------------------------------------------------------------------------
-  load(file.path(dropboxdir, "data/om/parMac.RData"))
-  par_laurie <- par
+  # load(file.path(dropboxdir, "data/om/parMac.RData"))
+  # par_laurie <- par
   
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section3.RData", sep="_")))
 
-  bind_rows(
-    as.data.frame(par_laurie) %>% mutate(par="par_laurie"),
-    as.data.frame(par) %>% mutate(par="par")
-  )  
+  # bind_rows(
+  #   as.data.frame(par_laurie) %>% mutate(par="par_laurie"),
+  #   as.data.frame(par) %>% mutate(par="par")
+  # )  
+  
   section <- "04"
   
   ices=window(stk,start=1991)
@@ -726,8 +774,7 @@ mystk     <- "mac";
[…]

### 2023-02-27 R/dd.R
delta

@@ -1,5 +1,5 @@
-ddWt<-function(wt,biomass,bref,alpha=-0.2,min=0.5,max=1.5){
-  adj=(biomass%/%bref)^alpha
+ddWt<-function(wt,biomass,bref,delta=-0.2,min=0.5,max=1.5){
+  adj=(biomass%/%bref)^delta
   
   wt%*%qmin(qmax(adj,min),max)}
 
@@ -16,7 +16,7 @@ ddFn<-function(year,x,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE){
     dsr=discards.wt(x)[,year]/stock.wt(x)[,year]
     dsr[!is.finite(dsr)]=0
     
-    stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bref"],par["b"])
+    stock.wt(   x)[,year]=ddWt(stock.wt(x)[,year],biomass(x)[,ac(an(year)-1)],par["bref"],par["delta"])
     landings.wt(x)[,year]=stock.wt(x)[,year]*lsr
     discards.wt(x)[,year]=stock.wt(x)[,year]*dsr}
     

### 2023-02-27 Rmd/mseMac.Rmd
delta bug fixed

@@
 ```{r}
+dirRuns="/home/laurie/Dropbox/ddmse/data/runs"
+```
+
+```{r, eval=FALSE}
+setwd("/home/laurie/Desktop/projects/ddmse/Rmd")
+```
+
+```{r}
 source("../R/dd.R")
@@
 ```{r, sr}
-load(paste("/home/laurie/Dropbox/pelagics/results/mac/figures/",tolower(stock),"_objects.RData",sep=""))
+load(paste("/home/laurie/Dropbox/pelagics/results/",tolower(stock),"/figures/",tolower(stock),"_objects.RData",sep=""))
 
@@
 
-```{r, eqs}
+```{r, srs}
 ggplot(dat)+
-  geom_line(aes(ssb,catch,col=What))
+  geom_line(aes(ssb,rec,col=What))
 ```
@@
 
-```{r, srs}
+```{r, eqs}
 ggplot(dat)+
-  geom_line(aes(ssb,rec,col=What))
+  geom_line(aes(ssb,catch,col=What))
 ```
@@
 ```{r, omCheck}
-om=fwdWindow(window(om,end=2020),end=2080,eq)
-f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.5,length.out=100),each=61),dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
-om=fwd(om,fbar=f,sr=eq)
+om=fwdWindow(window(om,end=2020),end=2080,eqM)
+f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.5,length.out=100),each=61),
+           dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
+om=fwd(om,fbar=f,sr=eqM)
 
@@
   oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
-  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eq)
+  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eqM)
 
   oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],par,TRUE,TRUE,FALSE)
-  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eq)
+  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eqM)
 
   oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],par,TRUE,TRUE,TRUE)
-  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eq)
+  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eqM)
   } 
@@
     }) ## F loop
+```
 
+```{r, projectons-save}
 save(prjs,file=paste(dirRuns,"/",stock,"Prjs.RData",sep="")) 
@@
 ```
-
-```{r, chk}
-dt1=model.frame(FLQuants("Rec"=rec(oms[[1]][,"2050"]),"SSB"=ssb(oms[[1]][,"2050"])),drop=T)
-dt2=model.frame(FLQuants("Rec"=rec(vpa[,"2050"]),"SSB"=ssb(vpa[,"2050"])),drop=T)
-
-ggplot()+
-  geom_line(aes(SSB,Rec),data=dt1)+
-  geom_line(aes(SSB,Rec),data=dt2)
-```

### 2023-02-27 Rmd/mseWhb.Rmd
delta bug fixed

@@
 title: "**Density Dependence**"
-subtitle: "Example of MSE for Blue whiting"
+subtitle: "MSE for Blue Whiting"
 author: "L Kell"
@@
 ```{r}
+dirRuns="/home/laurie/Dropbox/ddmse/data/runs/"
+```
+
+```{r}
 source("../R/dd.R")
+source("../R/get_dropbox.r")
 ```
@@
 
-```{r, whb.data}
-load(file.path(dropboxdir,"data/om/whb.RData"))
+```{r, ices}
+stock="Whb"
+load(paste("../data/om/ices",stock,".RData",sep=""))
+stk=get(tolower(stock))
 
-par=FLPar(c(m1=0.08/(0.2^-0.288),m2=-0.28,bmsy=6e6,k=63,w50=0.08))
+stock.wt(   stk)[1]=0.01
+catch.wt(   stk)[1]=0.01
+landings.wt(stk)[1]=0.01
+discards.wt(stk)[1]=0.01
+```
 
-stock.wt(   whb)[1]=0.01
-catch.wt(   whb)[1]=0.01
-landings.wt(whb)[1]=0.01
-discards.wt(whb)[1]=0.01
+```{r, pars}
+par=FLPar(c(m1=0.2/(0.15^-0.288),m2=-0.28,bref=NA,b=-0.2,matk=30.1, w50=0.1))  
+```
 
-om=ddFn2(whb,par)
-om=om+VPA(om)
-om=ddFn2(om,par)
+```{r, m}
+m(stk)=par["m1"]%*%(stock.wt(stk)%^%par["m2"])
+vpa=stk+VPA(stk)  
+```
 
-sr=fmle(as.FLSR(om,model="bevholt"),control=list(silent=TRUE))
-eq=FLBRP(om,sr=sr)
+```{r, vpa}
+plot(FLStocks("ICES"=stk,"VPA"=vpa)) 
+```
 
-om=fwdWindow(om,end=2080,eq)
-f =fbar(om)[,ac(2021:2080)]%=%refpts(eq)["msy","harvest"]
-om=fwd(om,fbar=f,sr=eq)
+```{r, sr}
+stk=window(iter(vpa,1),start=2000,end=2019)
+sr=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eq=FLBRP(stk,sr=ab(sr))
 
-om2=om
-om3=om
-om4=om
-for (iYear in ac(2021:2080)){
-  om2=ddFn(iYear,om2,par)
-  om2=fwd(om2,fbar=f[,iYear]*0.8,sr=eq)
+stk=window(iter(vpa,1),start=2000,end=2019)
+srM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqM=FLBRP(stk,sr=ab(srM))
 
-  om3=ddFn(iYear,om3,par)
-  om3=fwd(om3,fbar=f[,iYear],sr=eq)
+stk=window(iter(vpa,1),start=2000,end=2019)
+srDDM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDM=FLBRP(stk,sr=ab(srDDM))
 
-  om4=ddFn(iYear,om4,par)
-  om4=fwd(om4,fbar=f[,iYear]*1.2,sr=eq)
+stk=window(iter(vpa,1),start=2000,end=2019)
+srDDMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDMM=FLBRP(stk,sr=ab(srDDMM))
+
+stk=window(iter(vpa,1),start=2000,end=2019)
+srDDMMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
+                control=list(silent=TRUE))
+eqDDMMM=FLBRP(stk,sr=ab(srDDMMM))
+
+dat=rbind(
+  cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA",     model.frame(FLQuants(eqM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA M",   model.frame(FLQuants(eqDDM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA MM",  model.frame(FLQuants(eqDDMM,rec=rec,ssb=ssb,catch=catch),drop=T)),
+  cbind(What="VPA MMM", model.frame(FLQuants(eqDDMMM,rec=rec,ssb=ssb,catch=catch),drop=T)))
+```
+
+```{r, eqs}
+ggplot(dat)+
+  geom_line(aes(ssb,catch,col=What))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Stock recruitment relationships
+
+```{r, srs}
+ggplot(dat)+
+  geom_line(aes(ssb,rec,col=What))
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for assumptions about density dependence
+
+```{r, ddM}
+par["bref"]=refpts(eqM)["msy","biomass"]
+
+for (i in dimnames(vpa)$year)
+  om=ddFn(i,vpa,par,FALSE,FALSE,TRUE)
+```
+
+```{r, omCheck}
+om=fwdWindow(window(vpa,end=2020),end=2080,eqM)
+f =FLQuant(rep(seq(0,c(refpts(eqM)["crash","harvest"])*1.5,length.out=100),each=61),
+           dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
+om=fwd(om,fbar=f,sr=eqM)
+
+oms=FLStocks("Base"     =om)
+oms[["DD Mass"]]        =om
+oms[["DD Mass, Mat"]]   =om
+oms[["DD Mass, Mat, M"]]=om
+for (iYear in ac(2020:2080)){
+  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
+  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eqM)
+
+  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],par,TRUE,TRUE,FALSE)
+  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eqM)
+
+  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],par,TRUE,TRUE,TRUE)
+  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eqM)
   }
+```
 
-plot(FLStocks("fwd"=om,"0.8"=om2,"1.0"=om3,"1.2"=om4))
+```{r, check2}
+p1=ggplot(stock.wt(oms[[2]][,ac(2080)]))+  
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mass-at-age") 
+p2=ggplot(mat(oms[[2]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mat-at-age")
+p3=ggplot(m(oms[[2]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("M-at-age")
+
+d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+d2=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+p4=ggplot(d2)+
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence
+
+```{r, check3}
+p1=ggplot(stock.wt(oms[[3]][,ac(2080)]))+ 
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mass-at-age") 
+p2=ggplot(mat(oms[[3]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mat-at-age")
+p3=ggplot(m(oms[[3]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("M-at-age")
+
+d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+d3=model.frame(FLQuants(oms[[3]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+p4=ggplot(d3)+
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+
+ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence
+
+```{r, check4}
+p1=ggplot(stock.wt(oms[[4]][,ac(2080)]))+ 
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(stock.wt(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mass-at-age") 
+p2=ggplot(mat(oms[[4]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(mat(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("Mat-at-age")
+p3=ggplot(m(oms[[4]][,ac(2080)]))+
+  geom_line(aes(age,data,group=iter))+
+  geom_line(aes(age,data),data=as.data.frame(m(oms[[1]][,ac("2080")])),col="red")+
+  xlab("Age")+ylab("M-at-age")
+
+d1=model.frame(FLQuants(oms[[4]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+d4=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+                                           "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
+p4=ggplot(d4)+
+  geom_line(aes(SSB,Yield))+
+  geom_line(aes(SSB,Yield),col="red",data=d1)
+
+ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves and biological parameters for assumptions about density dependence
+
+```{r, eqlCurves}
+rfs=ldply(oms, function(x) { 
+  model.frame(FLQuants(x[,"2080"],
+             biomass=function(x) stock(x),
+             ssb    =function(x) ssb(  x),
+             catch  =function(x) catch(x),
+             f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+
+ggplot(rfs)+
+  geom_line(aes(ssb,catch,col=.id))+
+  xlab("SSB")+ylab("Catch")+
+  theme(legend.position="bottom")
+```
+
+**Figure `r iFig=iFig+1; iFig`** Equilibrium curves for assumptions about density dependence
+
+**Table `r iTab=iTab+1; iTab`** Reference points
+
+```{r, refptsTable}
+rfpts=ddply(rfs,.(.id), with, {  
+  flag=catch==max(catch)
+  data.frame(Bmsy  =biomass[flag],
+             SSBmsy=ssb[flag],
+             MSY   =catch[flag],
+             Fmsy  =f[flag],
+             Virgin=max(ssb),
+             B0    =max(biomass))})
+
+rfpts
+```
+
+```{r, projectons-1}
+prj=FLStocks("1"=iter(oms[[1]],1:4),
+             "2"=iter(oms[[2]],1:4),
+             "3"=iter(oms[[3]],1:4),
+             "4"=iter(oms[[4]],1:4))
+
+F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)
+
+control=as(FLQuants("f"=F),"fwdControl")
+prj[[1]] =fwd(prj[[1]],control=control,sr=eqM)
+
+for (i in ac(2020:2080)) {
+    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    control=as(FLQuants("f"=F[,i]),"fwdControl")
+    prj[[2]] =fwd(prj[[2]],control=control,sr=eqM)
+
+    prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    control=as(FLQuants("f"=F[,i]),"fwdControl")
+    prj[[3]] =fwd(prj[[3]],control=control,sr=eqM)
+
+    prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+    control=as(FLQuants("f"=F[,i]),"fwdControl")
+    prj[[4]] =fwd(prj[[4]],control=control,sr=eqM)
+    }
 ```
@@
 ```{r}
-ggplot(FLQuants("ICES"=stock.wt(om[ ,"2080"]), 
-                "0.8" =stock.wt(om2[,"2080"]), 
-                "1.0" =stock.wt(om3[,"2080"]),
-                "1.2" =stock.wt(om4[,"2080"])))+
-  geom_line(aes(age,data,col=qname))+
-  xlab("Age")+ylab("Mass-at-age")
+plot(FLStocks(prj[[1]]),iter=1:4) 
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Mass-at-age
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming no density dependence
 
 ```{r}
-ggplot(FLQuants("ICES"=m(om[ ,"2080"]),  
-                "0.8" =m(om2[,"2080"]), 
-                "1.0" =m(om3[,"2080"]),
-                "1.2" =m(om4[,"2080"])))+
-  geom_line(aes(age,data,col=qname))+
-  xlab("Age")+ylab("M") 
+names(prj)=names(oms)
+plot(FLStocks(prj[[2]]),iter=1:4)
 ```
 
-**Figure `r iFig=iFig+1; iFig`** M-at-mass
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass-at-age only
 
 ```{r}
-ggplot(FLQuants("ICES"=mat(om[ ,"2080"]),  
-               "0.8"  =mat(om2[,"2080"]),
-               "1.0"  =mat(om3[,"2080"]), 
-               "1.2"  =mat(om4[,"2080"])))+
-  geom_line(aes(age,data,col=qname))+
-  xlab("Age")+ylab("Maturity-a-age")
+plot(FLStocks(prj[[3]]),iter=1:4)
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Maturity-at-age
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass and maturity-at-age only
 
 ```{r}
-ggplot(rbind(cbind(What="ICES",model.frame(FLQuants(Mat=mat(om[  ,"2080"]),Mass=stock.wt(om[ ,"2080"])))),  
-             cbind(What= "0.8",model.frame(FLQuants(Mat=mat(om2[ ,"2080"]),Mass=stock.wt(om2[,"2080"])))),
-             cbind(What= "1.0",model.frame(FLQuants(Mat=mat(om3[ ,"2080"]),Mass=stock.wt(om3[,"2080"])))),
-             cbind(What= "1.2",model.frame(FLQuants(Mat=mat(om4[ ,"2080"]),Mass=stock.wt(om4[,"2080"]))))))+
-  geom_line(aes(Mass,Mat,col=What))+
-  geom_point(aes(Mass,Mat,col=What))
+plot(FLStocks(prj[[4]]),iter=1:4)
 ```
 
-**Figure `r iFig=iFig+1; iFig`** M-at-mass
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass, maturity and natural mortality-at-age
+
+```{r, eval=FALSE}
+ldply(prj,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
+                                                    ssb    =function(x) ssb(    x)[,"2080"],
+                                                    catch  =function(x) catch(  x)[,"2080"],
+                                                    f      =function(x) fbar(   x)[,"2080"]),drop=T)) 
+```
+
+```{r, projectons-2}
+devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)
+
+base=propagate(iter(prj[[1]],1),100)
+
+prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {
+
+            ## FMSY estimate by OM ######################################
+            F=FLQuant(F,dimnames=list(year=2020:2080))
+
+            ## No DD ####################################################
+            control=as(FLQuants("f"=F),"fwdControl")
+            x1     =fwd(base,control=control,sr=eqM,residuals=devRec)
+
+            ## With DD ####################################################
+            x2=x1
+            x3=x1
+            x4=x1
+            for (iYr in ac(2020:2080)) {
+              ## DD M ##################################################         
+              x2 =ddFn(iYr,x2,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+              x2 =fwd(x2,control=control,sr=eqM,residuals=devRec)
+
+              ## DD MM##################################################
+              x3 =ddFn(iYr,x3,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+              x3 =fwd(x3,control=control,sr=eqM,residuals=devRec)
+
+              ## DD MMM ################################################
+              x4 =ddFn(iYr,x4,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+              x4 =fwd(x4,control=control,sr=eqM,residuals=devRec)
+              }  ## year loop
+
+            FLStocks(list("Base"=x1,"M"=x2,"MM"=x3,"MMM"=x4))        
+    }) ## F loop
+```
+
+```{r, projectons-save}
+save(prjs,file=paste(dirRuns,"/",stock,"Prjs.RData",sep="")) 
+```
+
+```{r}
+plot(FLStocks(prjs[[1]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at Base $F_{MSY}$
+
+```{r}
+plot(FLStocks(prjs[[2]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, with  $F_{MSY}$ for density dependence in mass-at-age
+
+```{r}
+plot(FLStocks(prjs[[3]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity-at-age
+
+```{r}
+plot(FLStocks(prjs[[4]])) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity, natural mortality-at-age
+
+```{r, MSE, eval=FALSE}
+mse=hcrICES(fmsy,eql,rec(fmsy),
+                      par,
+                      start,   end,
+                      interval,lag=lag,
+                      err=err,
+                      bndTac=c(0,Inf))
+```

### 2023-02-27 Rmd/report.R
report.R

@@ -46,7 +46,7 @@ source("R/pm.R")
 
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
 
-if ("try-error" %in% is(dropboxdir))
+if ("try-error" %in% is(dropboxdir)) 
   dropboxdir="~/Dropbox/pelagics"
 
 myparams <- data.frame(
@@ -88,6 +88,30 @@ mystk     <- "mac";
   figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
   tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
 
+  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
+  
+  
+  # Get FLStock and FLBRP object  ----------------------------------------------
+  
+  load(file.path(dropboxdir,"data","om",paste0(mystk,".RData"))) # only used for FLBRP part; comes from 
+  stk   = get(mystk)
+  stkR  = get(paste0(mystk,"R"))
+  
+  if (mystk == "mac") {
+    # ts    = as.data.frame(readxl::read_excel(file.path(dropboxdir,"data/inputs/dat.xlsx"),sheet="ts"))
+    ts=model.frame(FLQuants(stk,tb=stock,ssb=ssb),drop=T)
+    
+    stock.wt(   stk)[1] = myparams[myparams$stock==mystk,"wt1"]
+    catch.wt(   stk)[1] = myparams[myparams$stock==mystk,"wt1"]
+    landings.wt(stk)[1] = myparams[myparams$stock==mystk,"wt1"]
+    discards.wt(stk)[1] = myparams[myparams$stock==mystk,"wt1"]
+    mat(        stk)[1] = 0
+  } else if (mystk == "whb") {
+    
+    ts=model.frame(FLQuants(stk,tb=stock,ssb=ssb),drop=T)
+    ts$tb=ts$tb/1e6
+  }
+  
   save(.,
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
   
@@ -114,31 +138,9 @@ mystk     <- "mac";
   # ------------------------------------------------------------------------------
   # 2. Conditioning
[…]

### 2023-02-27 Rmd/report.R
report.R

@@ -69,8 +69,8 @@ myparams <- data.frame(
   maxage  = c(12                           , 10               ),
   maxyear = c(2050                         , 2050             ),
   m1scaler= c(8                            , 10               ),
-  w50scaler= c(1000                        , 500              ),
-  matk2   = c(30                           , 100               ) )
+  w50     = c(0.166                         , 0.17              ),
+  matk2   = c(28                           , 63               ) )
 
 save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
@@ -417,7 +417,7 @@ mystk     <- "mac";
             FLPar(c(bref =c(refpts(eq)["msy","biomass"]),
                     delta= myparams[myparams$stock==mystk,"delta"],
                     matk = myparams[myparams$stock==mystk,"matk"],
-                    w50  =par["a"]*par["l50"]^par["b"],
+                    w50  = par["a"]*par["l50"]^par["b"],
                     alpha=myparams[myparams$stock==mystk,"alpha"])))
 
   fileConn <-file(file.path(tablesdir, paste(section,mystk, "par.txt", sep="_")))
@@ -507,7 +507,7 @@ mystk     <- "mac";
     stock.wt(   x)=ddWt(stock.wt(eq),
                         biomass(x),
                         refpts(eq)["msy","biomass"],
-                        alpha=-0.2)
+                        delta=myparams[myparams$stock==mystk,"alpha"])
     landings.wt(x)=stock.wt(x)*lsr
     discards.wt(x)=stock.wt(x)*dsr
     x=brp(x)}
@@ -539,7 +539,7 @@ mystk     <- "mac";
     stock.wt(   x)=ddWt(stock.wt(eq),
                         biomass(x),
                         refpts(eq)["msy","biomass"],
-                        alpha=-0.2)
+                        delta=myparams[myparams$stock==mystk,"alpha"])
     landings.wt(x)=stock.wt(x)*lsr
     discards.wt(x)=stock.wt(x)*dsr
     
[…]

### 2023-02-28 Rmd/report.R
Fixed plotting in projections

@@ -1203,11 +1203,7 @@ mystk     <- "mac";
                Virgin=max(ssb),
                B0    =max(biomass))})
   
-  # projection
-  # prj=FLStocks("1"=iter(oms[[1]],1:4),
-  #              "2"=iter(oms[[2]],1:4),
-  #              "3"=iter(oms[[3]],1:4),
-  #              "4"=iter(oms[[4]],1:4))
+  # projection with different Fs
   prj=FLStocks("Base"           =iter(oms[[1]],1:4),
                "DD Mass"        =iter(oms[[2]],1:4),
                "DD Mass, Mat"   =iter(oms[[3]],1:4),
@@ -1216,20 +1212,27 @@ mystk     <- "mac";
   F=fbar(prj[[1]][,ac(2020:2050)])%=%rep(rfpts$Fmsy,each=31)
   
   control=as(FLQuants("f"=F),"fwdControl")
-  prj[[1]] =fwd(prj[[1]],control=control,sr=vpaM_eq)
+  prj[[1]] =fwd(prj[[1]],
+                control=control,
+                sr=vpaM_eq)
   
   for (i in ac(2020:2050)) {
-    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[2]] =fwd(prj[[2]],control=control,sr=vpaM_eq)
+    
+    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[2]] =fwd(prj[[2]],
+                  control=control,
+                  sr=vpaM_eq)
     
     prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
-    control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[3]] =fwd(prj[[3]],control=control,sr=vpaM_eq)
+    prj[[3]] =fwd(prj[[3]],
+                  control=control,
+                  sr=vpaM_eq)
     
     prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-    control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[4]] =fwd(prj[[4]],control=control,sr=vpaM_eq)
[…]

### 2023-02-28 Rmd/report mac.Rmd
report.R & report Mac.Rmd

@@
 # tab_nums(name    = "section2_tab", caption = "section2_tab", level = 1, display = FALSE)
+section <- "02"
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "waa.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section, mystk, "waa.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "mat.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "m.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "m.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_waa.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "dd_waa.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "gamm.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "gamm.jpg", sep="_")))
 
@@
 
-cat(readLines(file.path(tablesdir, paste(mystk, "gamm summary.txt", sep="_"))), sep = '\n')
+cat(readLines(file.path(tablesdir, paste(section,mystk, "gamm summary.txt", sep="_"))), sep = '\n')
 
@@
 
-cat(readLines(file.path(tablesdir, paste(mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')
+cat(readLines(file.path(tablesdir, paste(section,mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "lm_waa.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "lm_waa.jpg", sep="_")))
 
@@
 
-cat(readLines(file.path(tablesdir, paste(mystk, "lm summary.txt", sep="_"))), sep = '\n')
+cat(readLines(file.path(tablesdir, paste(section,mystk, "lm summary.txt", sep="_"))), sep = '\n')
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "dd_mat.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "dd_mat.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste(mystk, "mat_ogive.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "mat_ogive.jpg", sep="_")))
 
@@
 
+section <- "03"
+
 fig_nums(
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_mass.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd_mass.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_maturity.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd_maturity.jpg",sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_m.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd_m.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_eq_yield.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"eq_yield.jpg",sep="_")))
 
@@
 
+section <- "04"
+
 fig_nums(
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_assessments.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"comparing_assessments.jpg",sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDM_4panels.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, 
+                                       paste(section,mystk,"VPADDM_4panels.jpg", sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMM_4panels.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"VPADDMM_4panels.jpg",sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMMM_4panels.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"VPADDMMM_4panels.jpg",sep="_")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"comparing_dd.jpg",sep="_")))
 ```
@@
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+section <- "05"
 

### 2023-02-28 Rmd/report.R

@@ -69,10 +69,10 @@ myparams <- data.frame(
   maxage  = c(12                           , 10               ),
   maxyear = c(2050                         , 2050             ),
   m1scaler= c(8                            , 10               ),
-  w50     = c(0.166                         , 0.17              ),
+  w50     = c(0.166                        , 0.08             ),
   matk2   = c(28                           , 63               ) )
 
-save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
+# save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
 mystk     <- "mac";  
 # mystk     <- "whb";  
@@ -88,7 +88,7 @@ mystk     <- "mac";
   figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
   tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
 
-  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
+  # load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
   
   
   # Get FLStock and FLBRP object  ----------------------------------------------
@@ -655,9 +655,9 @@ mystk     <- "mac";
 
 
   # clean up and save
-  # rm(dat, df, p, p1, p2, p3, t, x)
-  # save(.,
-  #      file = file.path(dropboxdir, "results", mystk, paste(mystk,"section3.RData", sep="_")))
+  rm(dat, df, p, p1, p2, p3, t, x)
+  save(.,
+       file = file.path(dropboxdir, "results", mystk, paste(mystk,"section3.RData", sep="_")))
 
 
 
@@ -674,7 +674,7 @@ mystk     <- "mac";
   
   
   
-    # ------------------------------------------------------------------------------
+  # ------------------------------------------------------------------------------
   # 4. OM based on ICES assessment
   # ------------------------------------------------------------------------------
[…]

### 2023-03-01 Rmd/report mac.Rmd
REPORT.r AND rmdS

@@
 source("../R/pm.R")
-```
-
-```{r}
 source("../R/get_dropbox.R")
@@
 
+All conditioning of the operating models and estimation of potential density dependent processes have been based on the ICES 2022 SAM assessment of `r mystkname`.
+
 ```{r}
@@
 
-[text here]
+Trends in (stock) weight at age are shown in figure 1.
+Weight-at-age in the stock is calculated as the average of the weight-at-age in the three spawning components, weighted by the relative size of each component (as estimated in the egg surveys).
+The decreasing trend in weight-at-age observed since 2005 for fish of age 3 and older seems to have stopped in 2013 and values in the last 8 years show an increasing trend.
 
@@
 
-[text here]
+The annual maturity ogive for mackerel is calculated as the average of the maturity ogives of the three spawning components weighted by the relative size of each component calculated as described above for
+the stock weights. The ogives for the North Sea and Southern components are fixed over time.
+For the Western component the maturity ogive is updated every year, using maturity data from commercial
+catch samples from Germany, Ireland, the Netherlands and the UK collected during the first
+and second quarters. Most of the variability in maturity at age is on ages 2 and 3. Nearly all fish of age 4 and older are estimated to be mature.
 
@@
 
-[text here]
+Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (Hamre, 1980; ICES 2022; ICES 2020).
+
+In order to assess the impacts of density dependent processes on natural mortality, additional assumptions need to be made about the distribution of natural mortality across ages.
+This will be done in section 3 of this report.
+
+Hamre, J. (1980). "Biology, exploitation, and management of the northeast atlantic mackrel." Conseil International pour l'exploration de la Mer, Rapports et Proces-verbaux des Reunions 177: 212-242.
+
+ICES (2022). Working Group on Widely Distributed Stocks (WGWIDE 2022) Volume 4, Issue 73.
+
+ICES (2020). Stock Annex: Mackerel (Scomber scombrus) in subareas 1-7 and 14 and divi-sions 8.a-e, 9.a (the Northeast Atlantic and adjacent waters), ICES.
 
@@
 
-[text here]
+The distribution of weight at age is shown in the figure below (left) and the relationship between weight at age and total biomass of mackerel is shown to the right.
+Consistent negative slopes on the biomass-weight at age plots can be observed for ages 4 and beyond.
+The patterns for the younger ages are less clear.
+It should be remembered that the younger ages do not appear strongly in the catches.
 
@@
 
-[text here]
+A Generalized Additive Mixed Model (GAMM) was estimated to assess the relationship between weight at age and total biomass.
+
+weights are predicted from two variables: a smoother of total biomass/mean(total biomass) and age.
+The resulting factor is shown below, indicating a nearly linear decline in weight at age as total biomass increases.
 
@@
 
-\newpage
-
-**GAMM summary**
-
-[text here]
+A summary of the GAMM model is shown in the table below.
 
@@
 
-**GAMM summary**
+**Linear model of density dependence in weight at age**
 
-[text here]
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-tab_nums(
-  name    = "cond_gammcoefficients", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))
-
-cat(readLines(file.path(tablesdir, paste(section,mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')
-
-```
-
-_`r tab_nums("cond_gammcoefficients")`_
-
-\newpage
-
-**LM of density dependence in weight at age**
-
-[text here]
+NEEDS BETTER EXPLANATION [ The estimated linear model from the GAMM model relative to the observations ]
 
@@
 
-\newpage
+[NEEDS MORE TEXT]
 
-**LM summary**
-
-[text here]
+A summary of the linear model is show below
 
@@
 
-**DD analysis of maturity at age**
+**Density-dependence analysis of maturity at age**
 
-[text here]
+The distribution of maturity at age is shown in the figure below (left) and the relationship between maturity at age and total biomass of mackerel is shown to the right.
+Negative slopes on the biomass-maturity at age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.
 
@@
 
+A maturity at weight ogive was generated from this formula:
+
 $f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
 
-Estimated parameters: k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`
+The resulted in the parameters k=`r matPar$par[1]` and $W_{50}$=`r matPar$par[2]` which have been used in subsequent simulations.
 
@@
 
-# Estimating Density Dependent effects based on equilibrium analyses
+# Estimating Density Dependent effects based on life-history parameters
 
 Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
-The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
+The parameters are first used to construct an `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
 
-## Mass-at-age [SHOULD WE CALL IT WEIGHT AT AGE, SINCE WE ARE USING Wa? ]
+## Mass-at-age
 
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-section <- "03"
-
-fig_nums(
-  name    = "sim_dd_mass", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in mass-at-age for different levels of biomass relative to Bmsy")
-
-knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd_mass.jpg", sep="_")))
-
-```
-
-_`r fig_nums("sim_dd_mass")`_
-
 ## Maturity-at-weight
@@
 $O(W_a) = 1/(1+e^{k(W_a-W_{50})})$
-
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "sim_dd_mat", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in maturity-at-age for different levels of biomass relative to Bmsy")
-
-knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd_maturity.jpg",sep="_")))
-
-```
-
-_`r fig_nums("sim_dd_mat")`_
 
@@
 
+[WHERE DO THE PARAMETERS COME FROM AND HOW TO THEY RELATE TO SUBSEQUENT ANALYSIS ??? ]
+
+[IT IS VERY CONFUSING TO USE M OF NATURAL MORTALITY AND FOR MASS !!]
+
+The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to Bmsy.
+The red line indicates the values from the current assessment (for natural mortality with an assumed distribution over ages).
+[ CHECK !]
+
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
+section <- "03"
+
 fig_nums(
-  name    = "sim_dd_m", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")
+  name    = "sim_dd3", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in mass-at-age, maturity at age and Natural mortality at age for different levels of biomass relative to Bmsy")
 
-knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd_m.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd3.jpg", sep="_")))
 
@@
 
-_`r fig_nums("sim_dd_m")`_
+_`r fig_nums("sim_dd3")`_
 
-Comparison of equilibrium yield plots for simulation with and without density-dependent processes
+From the theoretical life-history parameters and simulations, we can compare the equilibrium total biomass and yield from an assessment without density dependence, and then adding, step-by-step, the density dependent processes for weight at age, maturity at age and natural mortality at age.
+Comparison of equilibrium yield plots for simulation with and without density-dependent processes indicates that the expected values for Bmsy tend to become lower when density dependence is included and that MSY becomes somewhat higher.
 
@@
 
-_`r fig_nums("sim_dd_m")`_
+_`r fig_nums("sim_yield")`_
 
@@
 
-Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA.
-The VPA reconstructions of the SAM assessment, shows a strongly divergent pattern in recruitment and also some differences in the trend of SSB.
-Mean fishing mortality is somewhat more variable in the VPA reconstruction compared to the SAM assessment.
-Density-dependent processes (on weight, maturity and M) tend to lead to lower SSB in the latter part of the time series.
+The operating models for this study are derived from the 2022 stock assessment.
+SAM is a statistical catch-at-age model that estimates error in the catch-at-age.
+In the case of mackerel, the assessment gives a higher weight to the recruitment index than to the catch data for ages 0 and 1.
+This means that the estimates of recruitment in the assessment can be very different from the actual number of fish that are recruited to the fishery.
+To ensure the MSE projections and the historical estimates are comparable, a Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
+In the next step, we applied a natural mortality at age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M at age.
+The figure below  compares the SAM assessment with the VPA and the VPAM assessment.
+A main difference is in the recruitment estimates.
+Using age-varying M tends to flatten the recruitment pattern.
+The vpaM model was used as the base case to be compared against the density dependent simulations.
 
@@
 
-_`r fig_nums("om_VPADDM")`_
+_`r fig_nums("om_assessments")`_
 
@@
 **Density dependence in mass-at-age**
+
+When applying density dependent processes to weight at age (i.e. lower weight at age for higher biomass), Bmsy is expected to be somewhat lower. The expected increase in MSY is small.
 
@@
 
+When applying density dependent processes to both weight at age and maturity at age, Bmsy is expected to be lower again and MSY is expected to increase.
+
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 
+When applying density dependent processes to all three variables, Bmsy is expected to be lower again and MSY is expected to increase a bit more.
+
 ```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
 fig_nums(
@@
 
-**Comparison of equilibrium curves for density dependence analyses: biomass against yield**
+**Comparison of equilibrium curves for density dependence analyses: biomass against yield and F**
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
+The comparison of the equilibrium curves, indicate that Bmsy is expected to decrease by 30%, Fmsy is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
+So while, inclusion of density dependent processes, has a substantial impact on biological reference points, the results change in equilibrium catch is expected to be small.
+
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
   name    = "om_eqyield", level = 1, display = FALSE,
-  caption = "Equilibrium curves for density dependence analyses: biomass against yield")
+  caption = "Equilibrium curves for density dependence analyses: biomass against yield and biomass against F")
 
-knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"comparing_dd.jpg",sep="_")))
+# print("test")
+# print(figuresdir)
+# print(paste(section,mystk,"om_biomass_yield_F.jpg",sep="_"))
+# print(file.path(figuresdir,paste(section,mystk,"om_biomass_yield_F.jpg",sep="_")))
+
+knitr::include_graphics(
+  path=file.path(figuresdir,paste(section,mystk,"om_biomass_yield_F.jpg",sep="_")))
+
 ```
@@
 
+Forward projections were run in two modes.
+Firstly in deterministic mode and secondly in probabilistic mode.
+We used the Fmsy to project forward from 2020 until 2050.
+We used both the Fmsy that is applicable to the density dependent scenario, but also evaluated what would happen if the 'wrong Fmsy' would be used with a particular scenario.
+
+The plot below shows the deterministic runs with the appropriate Fmsy value being used to project forward.
+Fmsy is highest when density dependence is taken into account for all three processes.
+Nevertheless, expected catches are somewhat lower when density dependence is taken into account.
+
+**[ THIS IS CONTRADICTING THE RESULTS FROM THE EQUILIBRIUM ANALYSIS ?? ]**
+
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 
+fig_nums(
+  name    = "mse_determ1", level = 1, display = FALSE,
+  caption = "Deterministic projections at Fmsy for different scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"deterministic_projections_at_Fmsy.jpg",sep="_")))
 ```
+
+_`r fig_nums("mse_determ1")`_
+
+\newpage
+
+**Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age**
+
+How can we understand the results of the deterministic runs?
+Below is a plot of the trends in stock weight at age, maturity at age and natural mortality at age, for the most relevant ages (1-6) and for the four different scenarios (Base, DDM, DDMM and DDMMM).
+The density dependent effect of maturity is strongest when both stock weight and maturity are density dependent.
+When natural mortality is also density dependent, this affects the stock size and hence also the density dependence for stock weight and maturity, leading to a smaller density dependent effects for those variables.
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_determ2", level = 1, display = FALSE,
+  caption = "Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"deterministic_projections_of_MMM.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_determ2")`_
+
+\newpage
+
+**Stochastic projections at Fmsy for different scenarios**
+
+Using stochastic projections by including random processes for recruitment deviations and the Fmsy values that are consistent with the different scenarios, leads to the overview below.
+SSB and total biomass are highest in the case of no density dependence.
+Catches are different in the first years of 'switching' to the new reference points, but afterwards, the catches for the different scenarios are highly comparable.
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_sim1", level = 1, display = FALSE,
+  caption = "Stochastic projections at Fmsy for different scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"stochastic_projections_at_Fmsy.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_sim1")`_
+
+\newpage
+
+**Histograms of probability distributions in 2050 for different scenarios**
+
+The histograms of the probability distributions in 2050 for different scenarios, again demonstrate the expected larger stock size without taking density dependence into account, and only marginal differences in catch.
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_eq2050", level = 1, display = FALSE,
+  caption = "Histograms of probability distributions in 2050 for different scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"histograms-2050.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_eq2050")`_
+
+\newpage
+
+**Stochastic projections at all Fmsy's from all scenarios**
+
+[ NEED TO CHECK THIS FIGURE; IS THIS A PLAUSIBLE RESULT? ]
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_sim2", level = 1, display = FALSE,
+  caption = "Stochastic projections at all Fmsy's from all scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"stochastic_projections_at_different_Fmsys.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_eq2050")`_
 

### 2023-03-01 Rmd/report whb.Rmd
with ref pts
REPORT.r AND rmdS

@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "sim_dd_mass", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in mass-at-age for different levels of biomass relative to Bmsy")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_mass.jpg")))
-
-```
-
-_`r fig_nums("sim_dd_mass")`_
-
 ## Maturity-at-weight
@@
 $O(W_a) = 1/(1+e^{k(W_a-W_{50})})$
-
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "sim_dd_mat", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in maturity-at-age for different levels of biomass relative to Bmsy")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_maturity.jpg")))
-
-```
-
-_`r fig_nums("sim_dd_mat")`_
 
@@
 
-Mackerel:
+<!-- Mackerel: -->
 
-$M = 0.0232B + 0.0524$
+<!-- $M = 0.0232B + 0.0524$ -->
 
@@
 
+section <- "03"
+
 fig_nums(
-  name    = "sim_dd_m", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")
+  name    = "sim_dd3", level = 1, display = FALSE,
+  caption = "Simulation of density dependence in mass-at-age, maturity at age and Natural mortality at age for different levels of biomass relative to Bmsy")
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_m.jpg")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"sim_dd3.jpg", sep="_")))
 
@@
 
-_`r fig_nums("sim_dd_m")`_
+_`r fig_nums("sim_dd3")`_
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_eq_yield.jpg")))
+knitr::include_graphics(
+  path=file.path(figuresdir, paste(section,mystk,"eq_yield.jpg",sep="_")))
 
@@
 
-**Comparison of equilibrium curves for density dependence analyses: biomass against yield**
+**Comparison of equilibrium curves for density dependence analyses: biomass against yield and F**
 
@@
   name    = "om_eqyield", level = 1, display = FALSE,
-  caption = "Equilibrium curves for density dependence analyses: biomass against yield")
+  caption = "Equilibrium curves for density dependence analyses: biomass against yield and biomass against F")
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd.jpg")))
+# print("test")
+# print(figuresdir)
+# print(paste(section,mystk,"om_biomass_yield_F.jpg",sep="_"))
+# print(file.path(figuresdir,paste(section,mystk,"om_biomass_yield_F.jpg",sep="_")))
+
+knitr::include_graphics(
+  path=file.path(figuresdir,paste(section,mystk,"om_biomass_yield_F.jpg",sep="_")))
 
@@
 
+[explanations here]
+
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
+section <- "05"
+
+fig_nums(
+  name    = "mse_determ1", level = 1, display = FALSE,
+  caption = "Deterministic projections at Fmsy for different scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"deterministic_projections_at_Fmsy.jpg",sep="_")))
 ```
+
+_`r fig_nums("mse_determ1")`_
+
+\newpage
+
+**Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age**
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_determ2", level = 1, display = FALSE,
+  caption = "Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"deterministic_projections_of_MMM.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_determ2")`_
+
+\newpage
+
+**Stochastic projections at Fmsy for different scenarios**
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_sim1", level = 1, display = FALSE,
+  caption = "Stochastic projections at Fmsy for different scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"stochastic_projections_at_Fmsy.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_sim1")`_
+
+\newpage
+
+**Histograms of probability distributions in 2050 for different scenarios**
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_eq2050", level = 1, display = FALSE,
+  caption = "Histograms of probability distributions in 2050 for different scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"histograms-2050.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_eq2050")`_
+
+\newpage
+
+**Stochastic projections at all Fmsy's from all scenarios**
+
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "mse_sim2", level = 1, display = FALSE,
+  caption = "Stochastic projections at all Fmsy's from all scenarios")
+
+knitr::include_graphics(path=
+                    file.path(figuresdir,paste(section,mystk,"stochastic_projections_at_different_Fmsys.jpg",sep="_")))
+
+```
+
+_`r fig_nums("mse_eq2050")`_
 

### 2023-03-01 Rmd/report.R

@@ -74,8 +74,8 @@ myparams <- data.frame(
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-mystk     <- "mac";  
-# mystk     <- "whb";  
+# mystk     <- "mac";  
+mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   
   mystkname  <- myparams[myparams$stock==mystk,"mystkname"]
@@ -250,14 +250,14 @@ mystk     <- "mac";
   dev.off()
   
   
-  # GAM model and diagnostics of weight vs Total biomass -----------------------
+  # GAMM model and diagnostics of weight vs Total biomass -----------------------
   
   # drop age 0 if existing
   wt2 = subset(wt,age!=0)
   nn  = length(unique(wt2$age))
   
   # run gammV  
-  gmr=mgcViz::gammV(log(data) ~ s(tb/mean(tb), bs="tp")+age, 
+  gmr=mgcViz::gammV(log(data) ~ s(tb/mean(tb), bs="tp") + age, 
                     data=wt2)
   
   # plot gammV
@@ -586,6 +586,25 @@ mystk     <- "mac";
                   drop=T) %>% mutate(scen="DD mass+mat+M")
     )
   save(df,file=paste(dropboxdir,"/data/results/",mystk,"Eq.RData",sep=""))
+  # load(   file=paste(dropboxdir,"/data/results/",mystk,"Eq.RData",sep=""))
+  
+  p <-
+    df %>% 
+    mutate(scen = factor(scen, levels=c("no-DD", "DD mass", "DD mass+mat", "DD mass+mat+M"))) %>% 
+    ggplot(aes(biomass, catch))+
+    theme_bw() +
+    geom_line(aes(colour=scen)) +
+    labs(x="Total biomass", y="Catch", colour="") +
+    scale_colour_manual(values =c("no-DD"         = "black",
+                                  "DD mass"       = "red", 
+                                  "DD mass+mat"   = "blue",
+                                  "DD mass+mat+M" = "darkgreen")) 
+  
+  
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "eq_yield.jpg", sep="_")),
+       width=10, height=10, units="in", res=300)
+  print(p) 
+  dev.off()
   
   p <-
     df %>% 
@@ -1080,15 +1099,15 @@ mystk     <- "mac";
     theme_bw() +
     theme(legend.position="none") +
[…]

### 2023-03-01 Rmd/backUp/mseWhb.Rmd
all runnuning

--- 2023-02-27 Rmd/mseWhb.Rmd
@@
 ```{r, sr}
+load(paste("/home/laurie/Dropbox/pelagics/results/",tolower(stock),"/figures/",tolower(stock),"_objects.RData",sep=""))
+
 stk=window(iter(vpa,1),start=2000,end=2019)
@@
 
-stk=window(iter(vpa,1),start=2000,end=2019)
+stk=window(iter(vpaM,1),start=2000,end=2019)
 srM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
@@
 
-stk=window(iter(vpa,1),start=2000,end=2019)
+stk=window(iter(vpaDDM,1),start=2000,end=2019)
 srDDM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
@@
 
-stk=window(iter(vpa,1),start=2000,end=2019)
+stk=window(iter(vpaDDMM,1),start=2000,end=2019)
 srDDMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
@@
 
-stk=window(iter(vpa,1),start=2000,end=2019)
+stk=window(iter(vpaDDMMM,1),start=2000,end=2019)
 srDDMMM=fmle(as.FLSR(stk,model="bevholtSV"),fixed=list(s=0.8,spr0=mean(spr0(stk))),
@@
 ```{r}
-plot(FLStocks(prj[[1]]),iter=1:4) 
+plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4) 
 ```
@@
 names(prj)=names(oms)
-plot(FLStocks(prj[[2]]),iter=1:4)
+plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)
 ```
@@
 ```{r}
-plot(FLStocks(prj[[3]]),iter=1:4)
+plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)
 ```
@@
 ```{r}
-plot(FLStocks(prj[[4]]),iter=1:4)
+plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)
 ```

### 2023-03-01 Rmd/backUp/report mac.Rmd
all runnuning

--- 2023-02-15 Rmd/report whb.Rmd
@@
                crop      =TRUE,
-               cache.path="../cache/figs/cond/",
-               fig.path  ="../tex/figs",
+               cache.path="../cache/report/",
+               fig.path  ="../tex/report",
                fig.width =10,
@@
 library(captioner)     # captions
-tab_nums <- captioner::captioner(prefix = "Table", levels=1, type=c("n"), infix=".")
-fig_nums <- captioner::captioner(prefix = "Figure", levels=1, type=c("n"), infix=".")
+tab_nums <- captioner::captioner(prefix = "Table", levels=2, type=c("n","n"), infix=".")
+fig_nums <- captioner::captioner(prefix = "Figure", levels=2, type=c("n","n"), infix=".")
 
@@
 
-source("../R/get_dropbox.r")
+source("../R/get_dropbox.R")
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
 
-# mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
-mystk     <- "whb"; mystkname <- "Blue whiting";                mylatin <- "Micromesistius poutassou" 
+mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
+# mystk     <- "whb"; mystkname <- "Blue whiting";                mylatin <- "Micromesistius poutassou" 
 maxyear <- 2050
@@
 load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
+```
 
+```{r, message=FALSE, warning=FALSE, cache=FALSE}
 # ------------------------------------------------------------------------------
@@
 
-**Executive summary**
+# Executive summary
 
@@
 
-# captioner::bump(fig_nums, level=1)
-# fig_nums(name    = "section1_fig", caption = "section1_fig", level = 1, display = FALSE)
-# tab_nums(name    = "section1_tab", caption = "section1_tab", level = 1, display = FALSE)
+tab_nums(name    = "section1_tab", level = 1, display = FALSE)
+fig_nums(name    = "section1_fig", level = 1, display = FALSE)
 
@@
 
-```{r}
-
-# fig_nums(name    = "section2_fig", caption = "section2_fig", level = 1, display = FALSE)
-# tab_nums(name    = "section2_tab", caption = "section2_tab", level = 1, display = FALSE)
-
-```
-
 **Weight at age**
@@
 fig_nums(
-  name    = "cond_mat", level = 1, display = FALSE,
+  name    = "cond_mat", level = 2, display = FALSE,
   caption = paste0(mystkname, ". Maturity at age by year and cohort"))
@@
 fig_nums(
-  name    = "cond_natmor", level = 1, display = FALSE,
+  name    = "cond_natmor", level = 2, display = FALSE,
   caption = paste0(mystkname, ". Natural mortality at age by year and cohort (all ages and years the same M)"))
@@
 fig_nums(
-  name    = "cond_ddwaa", level = 1, display = FALSE,
+  name    = "cond_ddwaa", level = 2, display = FALSE,
   caption = paste0(mystkname, ". Density dependence and weight at age."))
@@
 fig_nums(
-  name    = "cond_gammwaa", level = 1, display = FALSE,
+  name    = "cond_gammwaa", level = 2, display = FALSE,
   caption = paste0(mystkname, ". GAMM analysis of density dependence and weight at age."))
@@
 tab_nums(
-  name    = "cond_gammcoefficients", level = 1, display = FALSE,
+  name    = "cond_gammcoefficients", level = 2, display = FALSE,
   caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))
@@
 fig_nums(
-  name    = "cond_lmwaa", level = 1, display = FALSE,
+  name    = "cond_lmwaa", level = 2, display = FALSE,
   caption = paste0(mystkname, ". LM analysis of density dependence and weight at age."))
@@
 tab_nums(
-  name    = "cond_lmsummary", level = 1, display = FALSE,
+  name    = "cond_lmsummary", level = 2, display = FALSE,
   caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))
@@
 fig_nums(
-  name    = "cond_ddmat", level = 1, display = FALSE,
+  name    = "cond_ddmat", level = 2, display = FALSE,
   caption = paste0(mystkname, ". Analysis of density dependence and maturity at age."))
@@
 fig_nums(
-  name    = "cond_matogive", level = 1, display = FALSE,
+  name    = "cond_matogive", level = 2, display = FALSE,
   caption = paste0(mystkname, ". Estimated maturity ogive at weight"))
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_mass.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_mass.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_maturity.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_maturity.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_sim_dd_m.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_sim_dd_m.jpg")))
 
@@
 
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_eq_yield.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_eq_yield.jpg")))
 
@@
 
-**Stock assessments**
-
 Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA.
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
-fig_nums(
-  name    = "om_assessments", level = 1, display = FALSE,
-  caption = "Comparison of different assessments. VPA-M is used going forward")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_assessments.jpg")))
+# knitr::include_graphics(path=file.path(resultsdir, paste0(mystk,"_om.jpg")))
 
@@
 
-_`r fig_nums("om_VPADDM")`_
-
 \newpage
 
-**Density dependence in mass-at-age**
+Forward projections with F=fmsy (fwd), 0.8xfmsy?, 1.0xfmsy, 1.2xfmsy?
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "om_VPADDM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDM_4panels.jpg")))
+```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("om_VPADDM")`_
-
-\newpage
-
-**Density dependence in mass-at-age and maturity at age**
-
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "om_VPADDMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age and Maturity at age")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMM_4panels.jpg")))
-
-```
-
-_`r fig_nums("om_VPADDMM")`_
-
-\newpage
-
-**Density dependence in mass-at-age, maturity at age and M at age**
-
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "om_VPADDMMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_VPADDMMM_4panels.jpg")))
-
-```
-
-_`r fig_nums("om_VPADDMMM")`_
-
-\newpage
-
-**Comparison of equilibrium curves for density dependence analyses: biomass against yield**
-
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-fig_nums(
-  name    = "om_eqyield", level = 1, display = FALSE,
-  caption = "Equilibrium curves for density dependence analyses: biomass against yield")
-
-knitr::include_graphics(path=file.path(figuresdir, paste0(mystk,"_comparing_dd.jpg")))
-
-```
-
-_`r fig_nums("om_eqyield")`_
-
-\newpage
-
 <!-- =============================================================================================================== -->
-<!-- Section 5: Running forward projections                                                                          -->
+<!-- Section 5: Running MSE's                                                                                        -->
 <!-- =============================================================================================================== -->
 
-# Forward projections without feedback
+# Running MSE's
 

### 2023-03-01 Rmd/condMac.Rmd
m1 corrected
all runnuning

@@
 date: "`r format(Sys.time(), '%d %B, %Y')`"
-output: word_document
+output: pdf_document
 mathjax: TRUE
@@
 ```{r, dropbox}
-source("../R/get_dropbox.R")
+source("../R/get_dropbox.r")
 
@@
 ```{r}
-matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))
-
 dat=model.frame(FLQuants(mac,"wt"=stock.wt,"mat"=function(x) mat(x)))
 dat$age=factor(dat$age)
+
+matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))
 

### 2023-03-01 Rmd/ddMac.Rmd
all runnuning

@@
 
-source("../R/get_dropbox.R")
+source("../R/get_dropbox.r")
 
@@
 for (i in seq(10)){
-  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"],alpha=-0.2)
+  stock.wt(   x)=ddWt(stock.wt(eq),biomass(x),refpts(eq)["msy","biomass"],delta=-0.2)
   landings.wt(x)=stock.wt(x)*lsr

### 2023-03-01 Rmd/mseMac.Rmd
all runnuning

@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/mseMac/",
-               fig.path  ="../tex/figs/mseMac/",
+               cache.path="../cache/mac/mse/",
+               fig.path  ="../tex/figs/mac/mse/",
                fig.width =8,
@@
 ```{r, pars}
-par=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,b=-0.3,matk=30.1,w50=0.2))
+params=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,delta=-0.3,matk=30,w50=0.17))
 ```
@@
 ```{r, m}
-m(stk)=par["m1"]%*%(stock.wt(stk)%^%par["m2"])
+m(stk)=params["m1"]%*%(stock.wt(stk)%^%params["m2"])
 vpa=stk+VPA(stk)
@@
 ```{r, ddM}
-par["bref"]=refpts(eqM)["msy","biomass"]
+params["bref"]=refpts(eqM)["msy","biomass"]
 
 for (i in dimnames(vpaM)$year)
-  om=ddFn(i,vpaM,par,FALSE,FALSE,TRUE)
+  om=ddFn(i,vpaM,params,FALSE,FALSE,TRUE) 
+```
+
+```{r}
+save(params,file="~/Desktop/tmp/parMac.RData")  
 ```
@@
 for (iYear in ac(2020:2080)){
-  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],par,TRUE,FALSE,FALSE)
+  oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],params,TRUE,FALSE,FALSE)
   oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eqM)
 
-  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],par,TRUE,TRUE,FALSE)
+  oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],params,TRUE,TRUE,FALSE)
   oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eqM)
 
-  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],par,TRUE,TRUE,TRUE)
+  oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],params,TRUE,TRUE,TRUE)
   oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eqM)
@@
 
-```{r, projectons-1}
+```{r, projections-1}
 prj=FLStocks("1"=iter(oms[[1]],1:4),
@@
 control=as(FLQuants("f"=F),"fwdControl")
-prj[[1]] =fwd(prj[[1]],control=control,sr=eq)
+prj[[1]] =fwd(prj[[1]],control=control,sr=eqM)
 
 for (i in ac(2020:2080)) {
-    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[2]] =ddFn(i,prj[[2]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[2]] =fwd(prj[[2]],control=control,sr=eq)
+    prj[[2]] =fwd(prj[[2]],control=control,sr=eqM)
 
-    prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[3]] =ddFn(i,prj[[3]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[3]] =fwd(prj[[3]],control=control,sr=eq)
+    prj[[3]] =fwd(prj[[3]],control=control,sr=eqM)
 
-    prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+    prj[[4]] =ddFn(i,prj[[4]],params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[4]] =fwd(prj[[4]],control=control,sr=eq)
+    prj[[4]] =fwd(prj[[4]],control=control,sr=eqM)
     }
@@
 
-```{r, projectons-2}
+```{r, projections-2}
 devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)
@@
             for (iYr in ac(2020:2080)) {
+              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+
               ## DD M ##################################################         
-              x2 =ddFn(iYr,x2,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
-              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
-              x2 =fwd(x2,control=control,sr=eq,residuals=devRec)
+              x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+              x2 =fwd(x2,control=control,sr=eqM,residuals=devRec)
 
               ## DD MM##################################################
-              x3 =ddFn(iYr,x3,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
-              x3 =fwd(x3,control=control,sr=eq,residuals=devRec)
+              x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+              x3 =fwd(x3,control=control,sr=eqM,residuals=devRec)
 
               ## DD MMM ################################################
-              x4 =ddFn(iYr,x4,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
-              x4 =fwd(x4,control=control,sr=eq,residuals=devRec)
+              x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+              x4 =fwd(x4,control=control,sr=eqM,residuals=devRec)
               }  ## year loop
@@
 mse=hcrICES(fmsy,eql,rec(fmsy),
-                      par,
+                      params,
                       start,   end,
@@
 ```
+
+```{r}
+ggarrange(plot(prjs[[1]]) +
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[2]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[3]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(prjs[[4]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1))
+```
+
+```{r}
+save(params,file="~/Desktop/tmp/params.RData")
+```

### 2023-03-01 Rmd/omMac.Rmd
all runnuning

@@
 ```{r warning=FALSE, cache=FALSE, include=FALSE, knitr, eval=TRUE}
-
 knitr::opts_chunk$set(echo = FALSE)
@@
 iTab=0
+```
 
+```{r, pkgs}
 library(FLCore)
@@
 theme_set(theme_bw(16))
+```
 
-source("../R/get_dropbox.R")
+```{r, dropbox}
+source("../R/get_dropbox.r")
 
@@
 source("../R/dd.R")
-
 ```
@@
 ```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
-
 load(file.path(dropboxdir,"data/om/mac.RData"))
-
 ```
@@
 ```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
-
-# Start the clock
-ptm <- proc.time()
-
 load(file.path(dropboxdir,"data/inputs/ices.RData"))
@@
 
-F  =propagate(window(fbar(ices),start=2020),101)
+F  =propagate(window(fbar(ices),end=2050),101)
 F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
@@
 
-control=as(FLQuants("f"=F),"fwdControl")
-ices   =fwd(ices,control=control,sr=eq)
+control=as(FLQuants("f"=F[,-1]),"fwdControl")
+ices   =fwd(propagate(ices,101),control=control,sr=eq)
 
@@
   geom_line(aes(SSB,Yield))
-
-# Stop the clock
-proc.time() - ptm
-
 ```
@@
 ```{r, vpa}
-
 # Start the clock
@@
 ```{r, vpaM}
-
 # Start the clock
@@
 
-load(file.path(dropboxdir, "data/om/parMac.RData"))
+#load(file.path(dropboxdir, "data/om/parMac.RData"))
+load("/home/laurie/Desktop/projects/ddmse/data/om/parMac.RData")
 par["m1"]  =par["m1"]/10
@@
 vpaM=fwdWindow(vpaM,eq,end=2050)
+
+# plot(sr)
+# plot(vpaM)
 

### 2023-03-01 Rmd/report mac.Rmd
all runnuning

--- 2023-02-18 Rmd/report mac.Rmd
@@
 ```{r}
-source("../R/get_dropbox.R")
+source("../R/get_dropbox.r")
 
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE, eval=FALSE}
+Introducing density-dependent process has an impact on the estimate Bmsy (calculated from Total biomass) which is shifted to the left. The expected impact on the equilibrium yield is relatively small.
 
+```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 fig_nums(

### 2023-03-02 Rmd/report mac.Rmd

--- 2023-03-01 Rmd/report mac.Rmd
@@
 mathjax: TRUE
+bibliography: refs.bib
 ---
+
+@lorenzen2000allometry
 
@@
 source("../R/pm.R")
-source("../R/get_dropbox.R")
+source("../R/get_dropbox.r")
 
@@
 <!-- =============================================================================================================== -->
+
 <!-- ## Mackerel report                                                                                              -->
+
 <!-- =============================================================================================================== -->
@@
 <!-- =============================================================================================================== -->
+
 <!-- Section 1: introduction                                                                                         -->
+
 <!-- =============================================================================================================== -->
@@
 
-The purpose of this work is to support the project “Fmsy-project for six high profile fish stocks”.
+The purpose of this work is to support the project "Fmsy-project for six high profile fish stocks".
 The project focuses on the current management approach for fish stocks based on single-species Fmsy Values where the only density-dependent process that is included is recruitment.
@@
 
-All MSE's have been developed within the FLR framework (https://flr-project.org/) and all the code is stored on github (https://github.com/martinpastoors/ddmse).
+All MSE's have been developed within the FLR framework (<https://flr-project.org/>) and all the code is stored on github (<https://github.com/martinpastoors/ddmse>).
 
-All data and results are available on DropBox (https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0)
+All data and results are available on DropBox (<https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0>)
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 2: conditioning and summaries of ICES assessment                                                        -->
+
 <!-- =============================================================================================================== -->
@@
 
-_`r fig_nums("cond_waa")`_
+*`r fig_nums("cond_waa")`*
 
@@
 
-The annual maturity ogive for mackerel is calculated as the average of the maturity ogives of the three spawning components weighted by the relative size of each component calculated as described above for
-the stock weights. The ogives for the North Sea and Southern components are fixed over time.
-For the Western component the maturity ogive is updated every year, using maturity data from commercial
-catch samples from Germany, Ireland, the Netherlands and the UK collected during the first
-and second quarters. Most of the variability in maturity at age is on ages 2 and 3. Nearly all fish of age 4 and older are estimated to be mature.
+The annual maturity ogive for mackerel is calculated as the average of the maturity ogives of the three spawning components weighted by the relative size of each component calculated as described above for the stock weights.
+The ogives for the North Sea and Southern components are fixed over time.
+For the Western component the maturity ogive is updated every year, using maturity data from commercial catch samples from Germany, Ireland, the Netherlands and the UK collected during the first and second quarters.
+Most of the variability in maturity at age is on ages 2 and 3.
+Nearly all fish of age 4 and older are estimated to be mature.
 
@@
 
-_`r fig_nums("cond_mat")`_
+*`r fig_nums("cond_mat")`*
 
@@
 
-_`r fig_nums("cond_natmor")`_
+*`r fig_nums("cond_natmor")`*
 
@@
 
-_`r fig_nums("cond_ddwaa")`_
+*`r fig_nums("cond_ddwaa")`*
 
@@
 
-_`r fig_nums("cond_gammwaa")`_
+*`r fig_nums("cond_gammwaa")`*
 
@@
 
-_`r tab_nums("cond_gammsummary")`_
+*`r tab_nums("cond_gammsummary")`*
 
@@
 
-_`r fig_nums("cond_lmwaa")`_
+*`r fig_nums("cond_lmwaa")`*
 
@@
 
-_`r tab_nums("cond_gammcoefficients")`_
+*`r tab_nums("cond_gammcoefficients")`*
 
@@
 
-_`r fig_nums("cond_ddmat")`_
+*`r fig_nums("cond_ddmat")`*
 
@@
 
-_`r fig_nums("cond_matogive")` _
+*`r fig_nums("cond_matogive")`*
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 3: Estimating Density Dependent effects based on equilibrium analyses                                   -->
+
 <!-- =============================================================================================================== -->
@@
 
-_`r fig_nums("sim_dd3")`_
+*`r fig_nums("sim_dd3")`*
 
@@
 
-_`r fig_nums("sim_yield")`_
+*`r fig_nums("sim_yield")`*
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 4: Setting up the Operating Models (OM)                                                                 -->
+
 <!-- =============================================================================================================== -->
@@
 
-_`r fig_nums("om_assessments")`_
+*`r fig_nums("om_assessments")`*
 
@@
 
-_`r fig_nums("om_VPADDM")`_
+*`r fig_nums("om_VPADDM")`*
 
@@
 
-_`r fig_nums("om_VPADDMM")`_
+*`r fig_nums("om_VPADDMM")`*
 
@@
 
-_`r fig_nums("om_VPADDMMM")`_
+*`r fig_nums("om_VPADDMMM")`*
 
@@
 
-_`r fig_nums("om_eqyield")`_
+*`r fig_nums("om_eqyield")`*
 
@@
 <!-- =============================================================================================================== -->
+
 <!-- Section 5: Running forward projections                                                                          -->
+
 <!-- =============================================================================================================== -->
@@
 
-_`r fig_nums("mse_determ1")`_
+*`r fig_nums("mse_determ1")`*
 
@@
 
-_`r fig_nums("mse_determ2")`_
+*`r fig_nums("mse_determ2")`*
 
@@
 
-_`r fig_nums("mse_sim1")`_
+*`r fig_nums("mse_sim1")`*
 
@@
 
-_`r fig_nums("mse_eq2050")`_
+*`r fig_nums("mse_eq2050")`*
 
@@
 
-_`r fig_nums("mse_eq2050")`_
+*`r fig_nums("mse_eq2050")`*
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 6: Discussion and conclusions                                                                           -->
+
 <!-- =============================================================================================================== -->
@@
 <!-- =============================================================================================================== -->
+
 <!-- Funding                                                                                                         -->
+
 <!-- =============================================================================================================== -->
@@
 <!-- =============================================================================================================== -->
+
 <!-- References                                                                                                      -->
+
 <!-- =============================================================================================================== -->
@@
 <!-- =============================================================================================================== -->
+
 <!-- Annex 1: OM conditioning                                                                           -->
+
 <!-- =============================================================================================================== -->
@@
 <!-- =============================================================================================================== -->
+
 <!-- Annex 2: Density dependent processes                                                                           -->
+
 <!-- =============================================================================================================== -->

### 2023-03-04 Rmd/mseMac.Rmd
fixed initialisation

@@
 
-dat=rbind(
+eqCs=rbind(
   cbind(What="ICES",    model.frame(FLQuants(eq,rec=rec,ssb=ssb,catch=catch),drop=T)),
@@
 ```{r, srs}
-ggplot(dat)+
+ggplot(eqCs)+
   geom_line(aes(ssb,rec,col=What))
@@
 ```{r, eqs}
-ggplot(dat)+
+ggplot(eqCs)+
   geom_line(aes(ssb,catch,col=What))
@@
   om=ddFn(i,vpaM,params,FALSE,FALSE,TRUE) 
-```
-
-```{r}
-save(params,file="~/Desktop/tmp/parMac.RData")  
 ```
@@
 
-d1=model.frame(FLQuants(oms[[4]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+d1=model.frame(FLQuants(oms[[1]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                            "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
-d4=model.frame(FLQuants(oms[[2]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
+d4=model.frame(FLQuants(oms[[4]][,ac(2080)],"Biomass"=function(x) biomass(x),"SSB"=function(x) ssb(x),
                                            "F"=function(x) fbar(x),"Yield"=function(x) catch(x)),drop=T)
@@
 ```{r, eqlCurves}
-rfs=ldply(oms, function(x) { 
+eqCurves=ldply(oms, function(x) { 
   model.frame(FLQuants(x[,"2080"],
@@
 
-ggplot(rfs)+
+ggplot(eqCurves)+
   geom_line(aes(ssb,catch,col=.id))+
@@
 ```{r, refptsTable}
-rfpts=ddply(rfs,.(.id), with, {  
+rfpts=ddply(eqCurves,.(.id), with, {  
   flag=catch==max(catch)
@@
 
+```{r}
+save(om,eqM,rfpts,params,file="~/Desktop/tmp/t.RData")
+```
+
 ```{r, projections-1}
-prj=FLStocks("1"=iter(oms[[1]],1:4),
-             "2"=iter(oms[[2]],1:4),
-             "3"=iter(oms[[3]],1:4),
-             "4"=iter(oms[[4]],1:4))
+load("~/Desktop/tmp/t.RData")
+source("/home/laurie/Desktop/projects/ddmse/R/dd.R")
 
-F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)
+om=fwdWindow(window(iter(om,51),end=2020),end=2080,eqM)
+F =propagate(fbar(om[,ac(2020:2080)]),4)
+F[]=rep(rfpts$Fmsy,each=61)
+control=as(FLQuants("f"=F),"fwdControl")
 
-control=as(FLQuants("f"=F),"fwdControl")
-prj[[1]] =fwd(prj[[1]],control=control,sr=eqM)
+om =fwd(om,control=control,sr=eqM)
 
+prj=FLStocks("Base"     =om)
+prj[["DD Mass"]]        =om
+prj[["DD Mass, Mat"]]   =om
+prj[["DD Mass, Mat, M"]]=om
 for (i in ac(2020:2080)) {
@@
 
-    prj[[3]] =ddFn(i,prj[[3]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[3]] =ddFn(i,prj[[3]],params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
     }
+```
+
+```{r}
+ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
+                                                  f    =function(x) fbar(x), 
+                                                  catch=function(x) catch(x)),drop=T))
+ts=transform(ts,What=.id)
+ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
+
+#plot(prj,probs=0,worm=seq(4))+facet_grid(qname~stock,scale="free")
+
+ggplot()+
+  geom_line( aes(ssb,catch,col=.id),   data=eqCurves)+
+  geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))
+
+cast(subset(ts,year==2080),iter~What,value="catch")
 ```
@@
 ```{r, projections-2}
-devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)
-
-base=propagate(iter(prj[[1]],1),100)
+base=propagate(iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1),100)
+devRec=rlnorm(100,rec(base)[,ac(2020:2080),,,,1]%=%0,0.3)
 
@@
     control=as(FLQuants("f"=F),"fwdControl")
-            x1     =fwd(base,control=control,sr=eq,residuals=devRec)
+    x1     =fwd(base,control=control,sr=eqM,residuals=devRec)
 
@@
           ## DD MM##################################################
-              x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+          x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
           x3 =fwd(x3,control=control,sr=eqM,residuals=devRec)
@@
           x4 =fwd(x4,control=control,sr=eqM,residuals=devRec)
-              }  ## year loop
+          }
 
@@
 ```{r}
-save(params,file="~/Desktop/tmp/params.RData")
+save(prjs,file="~/Desktop/tmp/t.RData")
+
+load("~/Desktop/tmp/t.RData")
+
+smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),
+            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="F")
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
 ```

### 2023-03-04 Rmd/mseWhb.Rmd
fixed initialisation

@@
                eval      =TRUE,cache   =TRUE,
-               cache.path="../cache/mseWhb/",
-               fig.path  ="../tex/figs/mseWhb/",
+               cache.path="../cache/whb/mse/",
+               fig.path  ="../tex/figs/whb/mse/",
                fig.width =8,
@@
 ```{r, pars}
-par=FLPar(c(m1=0.2/(0.15^-0.288),m2=-0.28,bref=NA,b=-0.2,matk=30.1, w50=0.1))  
+params=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,delta=-0.2,matk=62.1, w50=0.08))  
 ```
@@
 
-```{r, eqs}
-ggplot(dat)+
-  geom_line(aes(ssb,catch,col=What))
+```{r, srs}
+ggplot(eqCs)+
+  geom_line(aes(ssb,rec,col=What))
 ```
@@
 
-```{r, srs}
-ggplot(dat)+
-  geom_line(aes(ssb,rec,col=What))
+```{r, eqs}
+ggplot(eqCs)+
+  geom_line(aes(ssb,catch,col=What))
 ```
@@
 ```{r, ddM}
-par["bref"]=refpts(eqM)["msy","biomass"]
+params["bref"]=refpts(eqM)["msy","biomass"]
 
-for (i in dimnames(vpa)$year)
-  om=ddFn(i,vpa,par,FALSE,FALSE,TRUE)
+for (i in dimnames(vpaM)$year)
+  om=ddFn(i,vpaM,params,FALSE,FALSE,TRUE) 
 ```
@@
 ```{r, omCheck}
-om=fwdWindow(window(vpa,end=2020),end=2080,eqM)
-f =FLQuant(rep(seq(0,c(refpts(eqM)["crash","harvest"])*1.5,length.out=100),each=61),
+om=fwdWindow(window(om,end=2020),end=2080,eqM)
+f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.5,length.out=100),each=61),
            dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
@@
   geom_line(aes(SSB,Yield),col="red",data=d1)
+
 ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
@@
 
-```{r, projectons-1}
-prj=FLStocks("1"=iter(oms[[1]],1:4),
-             "2"=iter(oms[[2]],1:4),
-             "3"=iter(oms[[3]],1:4),
-             "4"=iter(oms[[4]],1:4))
+```{r}
+save(om,eqM,rfpts,params,file="~/Desktop/tmp/t.RData")
+```
 
-F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)
+```{r, projections-1}
+load("~/Desktop/tmp/t.RData")
+source("/home/laurie/Desktop/projects/ddmse/R/dd.R")
 
+om=fwdWindow(window(iter(om,51),end=2020),end=2080,eqM)
+F =propagate(fbar(om[,ac(2020:2080)]),4)
+F[]=rep(rfpts$Fmsy,each=61)
 control=as(FLQuants("f"=F),"fwdControl")
-prj[[1]] =fwd(prj[[1]],control=control,sr=eqM)
 
+om =fwd(om,control=control,sr=eqM)
+
+prj=FLStocks("Base"     =om)
+prj[["DD Mass"]]        =om
+prj[["DD Mass, Mat"]]   =om
+prj[["DD Mass, Mat, M"]]=om
 for (i in ac(2020:2080)) {
-    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[2]] =ddFn(i,prj[[2]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
 
-    prj[[3]] =ddFn(i,prj[[3]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[3]] =ddFn(i,prj[[3]],params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
 
-    prj[[4]] =ddFn(i,prj[[4]],par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+    prj[[4]] =ddFn(i,prj[[4]],params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
 ```{r}
-plot(FLStocks(prj[[1]]),iter=1:4) 
+ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
+                                                  f    =function(x) fbar(x), 
+                                                  catch=function(x) catch(x)),drop=T))
+ts=transform(ts,What=.id)
+ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
+
+#plot(prj,probs=0,worm=seq(4))+facet_grid(qname~stock,scale="free")
+
+ggplot()+
+  geom_line( aes(ssb,catch,col=.id),   data=eqCurves)+
+  geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))
+
+cast(subset(ts,year==2080),iter~What,value="catch")
+```
+
+```{r}
+plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4) 
 ```
@@
 
-```{r, projectons-2}
-devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)
-
-base=propagate(iter(prj[[1]],1),100)
+```{r, projections-2}
+base=propagate(iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1),100)
+devRec=rlnorm(100,rec(base)[,ac(2020:2080),,,,1]%=%0,0.3)
 
@@
     for (iYr in ac(2020:2080)) {
+        control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+
           ## DD M ##################################################         
-              x2 =ddFn(iYr,x2,par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
-              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+          x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
           x2 =fwd(x2,control=control,sr=eqM,residuals=devRec)
@@
           ## DD MM##################################################
-              x3 =ddFn(iYr,x3,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+          x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
           x3 =fwd(x3,control=control,sr=eqM,residuals=devRec)
@@
           ## DD MMM ################################################
-              x4 =ddFn(iYr,x4,par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-              control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+          x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
           x4 =fwd(x4,control=control,sr=eqM,residuals=devRec)
-              }  ## year loop
+          }
 
@@
 ```
+
+```{r}
+ggarrange(plot(prjs[[1]]) +
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[2]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[3]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(prjs[[4]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1))
+```
+
+```{r}
+save(prjs,file="~/Desktop/tmp/t.RData")
+
+load("~/Desktop/tmp/t.RData")
+
+smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),
+            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="F")
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
+```

### 2023-03-04 Rmd/omMac.Rmd
fixed initialisation

@@
 ices=ices[["mac.27.nea"]]
-ices=window(ices,start=1991)
+#ices=window(ices,start=1991)
 

### 2023-03-04 Rmd/report.R

@@ -1202,10 +1202,10 @@ mystk     <- "whb";
   
   
   # projection with different Fs
-  prj=FLStocks("Base" =iter(oms[[1]],1:4),
-               "M"    =iter(oms[[2]],1:4),
-               "MM"   =iter(oms[[3]],1:4),
-               "MMM"  =iter(oms[[4]],1:4))
+  prj=FLStocks("Base" =fwdWindow(window(iter(oms[[1]],1:4),end=2020),end=2050,eqM),
+               "M"    =fwdWindow(window(iter(oms[[2]],1:4),end=2020),end=2050,eqM),
+               "MM"   =fwdWindow(window(iter(oms[[3]],1:4),end=2020),end=2050,eqM),
+               "MMM"  =fwdWindow(window(iter(oms[[4]],1:4),end=2020),end=2050,eqM))
   
   F=fbar(prj[[1]][,ac(2020:2050)])%=%rep(rfpts$Fmsy,each=31)
   
@@ -1224,7 +1224,7 @@ mystk     <- "whb";
                   control=control,
                   sr=vpaM_eq)
     
-    prj[[3]] =ddFn(i,prj[[3]],vpaM_par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[3]] =ddFn(i,prj[[3]],vpaM_par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
     prj[[3]] =fwd(prj[[3]],
                   control=control,
                   sr=vpaM_eq)
@@ -1315,7 +1315,7 @@ mystk     <- "whb";
   
   # Next running the stochastic loop
   devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2050),,,,1]%=%0,0.3)
-  base=propagate(iter(prj[[1]],1),100)
+  base=fwdWindow(window(propagate(iter(prj[[1]],1),100),end=2020),end=2050,eqM)
   prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {
     
     ## FMSY estimate by OM ######################################
@@ -1346,7 +1346,7 @@ mystk     <- "whb";
               residuals=devRec)
       
       ## DD MM##################################################
-      x3 =ddFn(iYr,x3,vpaM_par,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+      x3 =ddFn(iYr,x3,vpaM_par,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
[…]

### 2023-03-04 Rmd/report.R
uses vpaM_eq

@@ -1200,12 +1200,11 @@ mystk     <- "whb";
   
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
   
-  
   # projection with different Fs
-  prj=FLStocks("Base" =fwdWindow(window(iter(oms[[1]],1:4),end=2020),end=2050,eqM),
-               "M"    =fwdWindow(window(iter(oms[[2]],1:4),end=2020),end=2050,eqM),
-               "MM"   =fwdWindow(window(iter(oms[[3]],1:4),end=2020),end=2050,eqM),
-               "MMM"  =fwdWindow(window(iter(oms[[4]],1:4),end=2020),end=2050,eqM))
+  prj=FLStocks("Base" =fwdWindow(window(iter(oms[[1]],1:4),end=2020),end=2050,vpaM_eq),
+               "M"    =fwdWindow(window(iter(oms[[2]],1:4),end=2020),end=2050,vpaM_eq),
+               "MM"   =fwdWindow(window(iter(oms[[3]],1:4),end=2020),end=2050,vpaM_eq),
+               "MMM"  =fwdWindow(window(iter(oms[[4]],1:4),end=2020),end=2050,vpaM_eq))
   
   F=fbar(prj[[1]][,ac(2020:2050)])%=%rep(rfpts$Fmsy,each=31)
   
@@ -1315,7 +1314,7 @@ mystk     <- "whb";
   
   # Next running the stochastic loop
   devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2050),,,,1]%=%0,0.3)
-  base=fwdWindow(window(propagate(iter(prj[[1]],1),100),end=2020),end=2050,eqM)
+  base=fwdWindow(window(propagate(iter(prj[[1]],1),100),end=2020),end=2050,vpaM_eq)
   prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {
     
     ## FMSY estimate by OM ######################################
@@ -1333,8 +1332,6 @@ mystk     <- "whb";
     x3=x1
     x4=x1
     for (iYr in ac(2020:2050)) {
-      
-      print(iYr)
       
       control=as(FLQuants("f"=F[,iYr]),"fwdControl")
       

### 2023-03-04 Rmd/mseMac.Rmd
removed saves

@@
 
-```{r}
-save(om,eqM,rfpts,params,file="~/Desktop/tmp/t.RData")
-```
-
 ```{r, projections-1}
-load("~/Desktop/tmp/t.RData")
 source("/home/laurie/Desktop/projects/ddmse/R/dd.R")
@@
 ```{r}
-save(prjs,file="~/Desktop/tmp/t.RData")
-
-load("~/Desktop/tmp/t.RData")
-
 smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),

### 2023-03-06 Rmd/mseMac.Rmd
checked for double DD

@@
 library(FLife)     # install.packages("FLife",    repos="http://flr-project.org/R")
-library(ggplotFL)
+library(ggplotFL)  # install.packages("ggplotFL", repos="http://flr-project.org/R")
 
@@
   geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))
+```
 
+```{r}
 cast(subset(ts,year==2080),iter~What,value="catch")
@@
 ```{r}
-plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4) 
+ggarrange(plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4)+ 
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)+ 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)+ 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)+ 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming no density dependence
-
-```{r}
-names(prj)=names(oms)
-plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass-at-age only
-
-```{r}
-plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass and maturity-at-age only
-
-```{r}
-plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass, maturity and natural mortality-at-age
-
-```{r, eval=FALSE}
-ldply(prj,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
-                                                    ssb    =function(x) ssb(    x)[,"2080"],
-                                                    catch  =function(x) catch(  x)[,"2080"],
-                                                    f      =function(x) fbar(   x)[,"2080"]),drop=T)) 
-```
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$
 
@@
 
+```{r}
+smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),  
+            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
+
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
+```
+
+```{r}
+ggarrange(plot(prjs[[1]])+ 
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[2]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[3]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(prjs[[4]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$
+
 ```{r, projectons-save}
@@
 
-```{r}
-plot(FLStocks(prjs[[1]])) 
+```{r, projections-3}
+base=iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1)
+
+prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {
+
+    ## FMSY estimate by OM ######################################
+    F=FLQuant(F,dimnames=list(year=2020:2080))
+
+    ## No DD ####################################################
+    control=as(FLQuants("f"=F),"fwdControl")
+    x1     =fwd(iter(base,1),control=control,sr=eqM)
+
+    ## With DD ####################################################
+    x2=x1
+    x3=x1
+    x4=x1
+    for (iYr in ac(2020:2080)) {
+        control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+
+          ## DD M ##################################################         
+          x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+          x2 =fwd(x2,control=control,sr=eqM)
+
+          ## DD MM##################################################
+          x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
+          x3 =fwd(x3,control=control,sr=eqM)
+
+          ## DD MMM ################################################
+          x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+          x4 =fwd(x4,control=control,sr=eqM)
+          }
+
+    FLStocks(list("Base"=x1,"M"=x2,"MM"=x3,"MMM"=x4))        
+    }) ## F loop
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at Base $F_{MSY}$
+```{r}
+smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),  
+            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
 
-```{r}
-plot(FLStocks(prjs[[2]])) 
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, with  $F_{MSY}$ for density dependence in mass-at-age
-
-```{r}
-plot(FLStocks(prjs[[3]])) 
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity-at-age
-
-```{r}
-plot(FLStocks(prjs[[4]])) 
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity, natural mortality-at-age
-
-```{r, MSE, eval=FALSE}
-mse=hcrICES(fmsy,eql,rec(fmsy),
-                      params,
-                      start,   end,
-                      interval,lag=lag,
-                      err=err,
-                      bndTac=c(0,Inf))
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
 ```
@@
 
-```{r}
-smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),
-            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
-cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="F")
-cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
+```{r, MSE, eval=FALSE}
+mse=hcrICES(fmsy,eql,rec(fmsy),
+                      params,
+                      start,   end,
+                      interval,lag=lag,
+                      err=err,
+                      bndTac=c(0,Inf))
 ```

### 2023-03-06 Rmd/report mac.Rmd
Updates to report.R (plotting), report mac.Rmd and refs.bib

@@
 ---
-
-@lorenzen2000allometry
 
@@
 
-dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
+dropboxdir<-try(file.path(get_dropbox(), "DDMSE"))
 
 if ("try-error"%in%is(dropboxdir))
-  dropboxdir="~/Dropbox/pelagics"
+  dropboxdir="~/Dropbox/DDMSE"
 
@@
 
-[summary here]
+The current management of fish stocks is based on achieving the fishing mortality ( (FMSY) that achieves the maximum sustainable yield (MSY).
+Estimation of MSY generally only considers the stock recruitment relationship when considering density dependent population regulation mechanisms.
+Density dependence, however, may also operate in  other processes such as growth, sexual maturity, and natural mortality.
+A main aim of this study was to condition  alternatives Operating Models (OMs) for use in Management Strategy Evaluation (MSE), that allows comparison with the  assessments conducted to support the ICES advice framework.
+To do this the OM include density dependence in growth, maturity, natural mortality, individually and combined.
+We use these OMs to evaluate the consequences of not including density dependence management advice.
+
+The Operating Models were
+
+* Base Case, with varying M-at-age
+* Density dependence in mass-at-age,
+* Density dependence in mass, maturity-at-age,
+* Density dependence in mass, maturity and M-at-age,
+
+The MSY reference points BMSY and FMSY are generally based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship ([**Sissenwine and Shepherd, 1986**]) or by conducting long-term projections under constant fishing moralities.
+
+Initial evaluations were conducted based on life history theory, as used by ICES to develop advice conducting MSE for data poor stocks (**Fischer et al., 2021**).
+**Figure 10** shows the  the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
+These allow the relationships between the MSY reference points and virgin biomass to be seen.
+
+Each of the OMs were then projected for the values of FMSY based on the four assumptions about density dependence.
+This allows a comparison between the yield that could potentially be achieved and that forgone by making an incorrect assumption about density dependence.
+The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to Fcrash) for 30 years into the future.
+The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure X2**).
+The points on the curve correspond to the projected outcomes for the values of FMSY from each OM.
+This was run as a check, as the equilibrium and projections should agree, I.e.
+the points at MSY should be the same colour as the curve.
+
+The OMs were also projected for the estimates of different estimates FMSY, with stochasticty in the recruitment deviates.
+This allows the impact of misspecification of density dependence to be evaluated, i.e.
+if FMSY is underestimated then yield will be less than MSY, while if FMSY is  overestimated  yield will be lost due to overfishing.
+Stochastic projections of the four Operating Models for the four estimates of FMSY are shown in **Figures X3a, X3b, X3c and X3d**
+
+**Figure X4** summaries the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence.
+
+Conclusions
+
+Estimating density-dependence from observations only possible for weight at age and maturity age.
+Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
+
+* The impact of including density dependence: major impact on reference points.
+Especially Fmsy is much higher when including DD.
+However, the expected yields in equilibrium conditions are expected to be very similar.
+The main difference is in the initial step from the current Fmsy to any new Fmsy that would include DD, as the higher Fmsy would be applied to a stock that is estimated without DD.
+* The impact of ignoring density dependence
+* The impact of assuming density-dependence incorrectly
+* Robustness of the current ICES framework
 
@@
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-\newpage
-
 **Maturity at age**
@@
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-\newpage
-
 **Natural mortality at age (fixed)**
 
-Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (Hamre, 1980; ICES 2022; ICES 2020).
+Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (@hamre1980biology; @ices2022wgwide; @ices2022macstockannex).
 
@@
 
-Hamre, J. (1980). "Biology, exploitation, and management of the northeast atlantic mackrel." Conseil International pour l'exploration de la Mer, Rapports et Proces-verbaux des Reunions 177: 212-242.
-
-ICES (2022). Working Group on Widely Distributed Stocks (WGWIDE 2022) Volume 4, Issue 73.
-
-ICES (2020). Stock Annex: Mackerel (Scomber scombrus) in subareas 1-7 and 14 and divi-sions 8.a-e, 9.a (the Northeast Atlantic and adjacent waters), ICES.
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-\newpage
-
 **GAMM at age**
@@
 
-\newpage
-
 **Linear model of density dependence in weight at age**
 
-NEEDS BETTER EXPLANATION [ The estimated linear model from the GAMM model relative to the observations ]
+The change in relative mass-at-age (i.e.
+observed - mean values) is plotted against total biomass, this allows the relationship between growth and total biomass to be compared across all ages.
+It is seen that there is a common negative relationship for all ages.
+A summary of the linear model is in the table below.
+The trend of growth with biomass is -0.18, this was then used to model mass-at-age as:
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+[ CHECK THE FORMULA ]
+W^'=¯W^-0.18
+$W_a = {\overline{W_a}} B^{{\beta}_a}$
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-[NEEDS MORE TEXT]
-
-A summary of the linear model is show below
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-*`r tab_nums("cond_gammcoefficients")`*
+*`r tab_nums("cond_lmsummary")`*
 
@@
 
-\newpage
-
 **Estimated maturity at weight**
@@
 
-The resulted in the parameters k=`r matPar$par[1]` and $W_{50}$=`r matPar$par[2]` which have been used in subsequent simulations.
+This resulted in the parameters k=`r format(matPar$par[1], digits=4)` and $W_{50}$=`r format(matPar$par[2], digits=4)` which have been used in subsequent simulations.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
-The parameters are first used to construct an `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
+The parameters are first used to construct an *`**FLBRP** object representing the equilibrium, and are then coerced into an **FLStock** to model the time series dynamics.
 
@@
 Where $\overline{W_a}$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data.
-$W_a$ is constrained so sizes do not become unfeasible large or small.
+$W_a$ is constrained so that sizes do not become unfeasible large or small.
 
@@
 
-Is assumed to vary at unit length ($L_1$) e.g.
+Natural mortality is assumed to vary at unit length ($L_1$).
 
-$ln(M_1) =\alpha + \beta ln(L_{\infty}) + \gamma ln(k)$
+Empirical and theoretical studies (@lorenzen2000allometry) have shown that natural mortality (M) decrease with fish length, but increase with the von Bertalanffy parameters of curvature k and asymptotic length L_∞.
+Therefore, M can be derived from life history parameters, e.g.
+M=1.5k(L_∞/L) (@POPE2021105952) or M=1.5k @jensen1985comparison.
+@lorenzen2008fish @lorenzen2022natural   showed that natural mortality rates vary with body size and age, often by orders of magnitude over the life cycle and support the existence of an allometric relationship between M and body mass (W) at-age of the form:
 
-where $\alpha=0.65$, $\beta=0.91$, $\gamma=0.87$
+[Formula here]
 
-Length-at-age ($L_a$) can be obtained from the length-weight relationship
+Where M_u is M at unit weight, and d is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
+d is approximately equal to –1/3, which implies that M is inversely proportional to body length, since mass is approximately proportional to the third power of length.
+This is consistent with [**Gislason et al., (2008)**], who proposed a relationship for natural mortality based on length
 
-$L_a=(W/a)^{1/b}$
+$M=\alphak(L_{\infty}χ)L\beta$
 
-giving
-
-$ln(M_1) =(W/a)^{1/b}(\alpha + \beta ln(L_{\infty}) + \gamma ln(k))$
-
-For DD there is a relationship between Mass at age $M_a$ and total biomass ($B$)
-
-Mackerel:
-
-$M = 0.0232B + 0.0524$
-
-[WHERE DO THE PARAMETERS COME FROM AND HOW TO THEY RELATE TO SUBSEQUENT ANALYSIS ??? ]
-
-[IT IS VERY CONFUSING TO USE M OF NATURAL MORTALITY AND FOR MASS !!]
-
-The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to Bmsy.
+The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to B_MSY.
 The red line indicates the values from the current assessment (for natural mortality with an assumed distribution over ages).
@@
   name    = "sim_dd3", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in mass-at-age, maturity at age and Natural mortality at age for different levels of biomass relative to Bmsy")
+  caption = "Simulation of density dependence in mass-at-age, maturity at age and Natural mortality at age for different levels of biomass relative to Bmsy. Red line is without density dependence.")
 
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "500px",  fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 Using age-varying M tends to flatten the recruitment pattern.
-The vpaM model was used as the base case to be compared against the density dependent simulations.
+The vpaM model was used as the base case to be compared against the density-dependent simulations.
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "600px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
   name    = "om_VPADDM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age")
+  caption = "Simulation of density dependence in Mass-at-age. Red lines are the non-density dependent values.")
 
@@
   name    = "om_VPADDMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age and Maturity at age")
+  caption = "Simulation of density dependence in Mass-at-age and Maturity at age. Red lines are the non-density dependent values.")
 
@@
   name    = "om_VPADDMMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age")
+  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age. Red lines are the non-density dependent values.")
 
@@
 
-The comparison of the equilibrium curves, indicate that Bmsy is expected to decrease by 30%, Fmsy is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
-So while, inclusion of density dependent processes, has a substantial impact on biological reference points, the results change in equilibrium catch is expected to be small.
+The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 30%, Fmsy is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
+So while, inclusion of density-dependent processes, has a substantial impact on biological reference points, the resulting change in equilibrium catch is expected to be small.
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "500px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 Nevertheless, expected catches are somewhat lower when density dependence is taken into account.
-
-**[ THIS IS CONTRADICTING THE RESULTS FROM THE EQUILIBRIUM ANALYSIS ?? ]**
 

### 2023-03-06 Rmd/report.R

@@ -44,10 +44,10 @@ source("R/dd.R")
 source("R/get_dropbox.r")
 source("R/pm.R")
 
-dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
+dropboxdir<-try(file.path(get_dropbox(), "DDMSE"))
 
 if ("try-error" %in% is(dropboxdir)) 
-  dropboxdir="~/Dropbox/pelagics"
+  dropboxdir="~/Dropbox/DDMSE"
 
 myparams <- data.frame(
   stock    = c("mac"                       , "whb"            ),
@@ -74,8 +74,8 @@ myparams <- data.frame(
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-# mystk     <- "mac";  
-mystk     <- "whb";  
+mystk     <- "mac";  
+# mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   
   mystkname  <- myparams[myparams$stock==mystk,"mystkname"]
@@ -355,10 +355,10 @@ mystk     <- "whb";
 
   # clean up and save
   rm(dat, gmr, m, mat, p, p1, p2, ts, wt, wt2)
-  save(.,
+  save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section2.RData", sep="_")))
 
-  
+
   
   
   
@@ -585,9 +585,8 @@ mystk     <- "whb";
                            f      =function(x) fbar(x)),
                   drop=T) %>% mutate(scen="DD mass+mat+M")
     )
-  save(df,file=paste(dropboxdir,"/data/results/",mystk,"Eq.RData",sep=""))
-  # load(   file=paste(dropboxdir,"/data/results/",mystk,"Eq.RData",sep=""))
-  
+  save(df,file=file.path(dropboxdir,"data", "om", paste(mystk, "Eq.RData", sep="_")))
+
   p <-
     df %>% 
     mutate(scen = factor(scen, levels=c("no-DD", "DD mass", "DD mass+mat", "DD mass+mat+M"))) %>% 
@@ -602,7 +601,7 @@ mystk     <- "whb";
   
   
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "eq_yield.jpg", sep="_")),
-       width=10, height=10, units="in", res=300)
+       width=10, height=6, units="in", res=300)
[…]

### 2023-03-06 Rmd/mseMac.Rmd
with same code

@@
 #plot(prj,probs=0,worm=seq(4))+facet_grid(qname~stock,scale="free")
+```
 
+```{r, check, fig.height=6,fig.width=8}
 ggplot()+
   geom_line( aes(ssb,catch,col=.id),   data=eqCurves)+
-  geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))
+  geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))+
+  theme_bw()+theme(legend.position="bottom")+
+  xlab("SSB")+ylab("Yield")
 ```

### 2023-03-06 Rmd/omMac.Rmd
with same code

--- 2023-03-01 Rmd/omMac.Rmd
@@
 ```{r echo=FALSE, warning=FALSE, cache=FALSE, eval=TRUE}
+library(FLCore)
+library(FLBRP)
+library(FLasher)
+
 load(file.path(dropboxdir,"data/inputs/ices.RData"))
 ices=ices[["mac.27.nea"]]
-ices=window(ices,start=1991)
+#ices=window(ices,start=1991)
 

### 2023-03-06 Rmd/omMac.Rmd
draft for mac report

@@
 ---
+
+3.3 Natural mortality
+Is assumed to vary at unit length ()
+Empirical and theoretical studies (lorenzen2002density?) have shown that natural mortality (M) decrease with fish length, but increase with the von Bertalanffy parameters of curvature  and asymptotic length .
+Therefore, M can be derived from life history parameters, e.g.  (Pope et al., 2021) or  (Jensen, 1985).
+Lorenzen (2008, 2022) showed that natural mortality rates vary with body size and age, often by orders of magnitude over the life cycle and support the existence of an allometric relationship between M and body mass () at-age of the form:
+
+Where  is M at unit weight, and d is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
+d is approximately equal to –1/3, which implies that M is inversely proportional to body length, since mass is approximately proportional to the third power of length.
+This is consistent with Gislason et al., (2008), who proposed a relationship for natural mortality based on length
+  $M=\alphak(L_{\infty}χ)L\beta$
+The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to .
+The red line indicates the values from the current assessment (for natural mortality with an assumed distribution over ages).
+[ CHECK !]
 

### 2023-03-06 Rmd/report mac.Rmd
draft for mac report

@@
 output:
-  word_document:
+  html_document:
     reference_docx: ../report_template.dotx
@@
 source("../R/pm.R")
-source("../R/get_dropbox.r")
+source("../R/get_dropbox.R")
 
@@
   dropboxdir="~/Dropbox/DDMSE"
+```
 
+```{r}
 mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
@@
 
-load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
+load(file=file.path(dropboxdir,paste0("data/om/",mystk,"/parMat.RData")))
+```
 
+```{r}
 # ------------------------------------------------------------------------------
@@
 tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
-
 ```
@@
 
-The current management of fish stocks is based on achieving the fishing mortality ( (FMSY) that achieves the maximum sustainable yield (MSY).
-Estimation of MSY generally only considers the stock recruitment relationship when considering density dependent population regulation mechanisms.
-Density dependence, however, may also operate in  other processes such as growth, sexual maturity, and natural mortality.
-A main aim of this study was to condition  alternatives Operating Models (OMs) for use in Management Strategy Evaluation (MSE), that allows comparison with the  assessments conducted to support the ICES advice framework.
-To do this the OM include density dependence in growth, maturity, natural mortality, individually and combined.
-We use these OMs to evaluate the consequences of not including density dependence management advice.
+The current management of fish stocks is based on achieving the fishing mortality ($F_{MSY}$) that achieves the maximum sustainable yield (MSY).
+Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms.
+Density dependence, however, may also operate in processes such as growth, sexual maturity, and natural mortality.
+A main aim of this study was to condition Operating Models (OMs) on alternative hypotheses about density dependence for use in Management Strategy Evaluation (MSE).
+The objective is to evaluate the robustness of the assessments conducted to support the ICES advice framework.
+To do this OMs were developed that include density dependence in growth, maturity, natural mortality, individually and combined.
+These OMs were then used to evaluate the consequences of not including density dependence management advice.
 
@@
 
-* Base Case, with varying M-at-age
-* Density dependence in mass-at-age,
-* Density dependence in mass, maturity-at-age,
-* Density dependence in mass, maturity and M-at-age,
+-   Base Case, with varying M-at-age
+-   Density dependence in mass-at-age,
+-   Density dependence in mass, maturity-at-age,
+-   Density dependence in mass, maturity and M-at-age,
 
-The MSY reference points BMSY and FMSY are generally based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship ([**Sissenwine and Shepherd, 1986**]) or by conducting long-term projections under constant fishing moralities.
+The MSY reference points $B_{MSY}$ and $F_{MSY}$ are generally based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship [@sissenwine1987alternative] or by conducting long-term projections under constant fishing moralities.
 
-Initial evaluations were conducted based on life history theory, as used by ICES to develop advice conducting MSE for data poor stocks (**Fischer et al., 2021**).
+Initial evaluations were conducted based on life history theory, as used by the International Council for the Exploration of the Sea (ICES) to develop advice conducting MSE for data poor stocks [@fischer2020ga].
 **Figure 10** shows the the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
@@
 
-Each of the OMs were then projected for the values of FMSY based on the four assumptions about density dependence.
+Each of the OMs were then projected for the values of $F_{MSY}$ based on the four assumptions about density dependence.
 This allows a comparison between the yield that could potentially be achieved and that forgone by making an incorrect assumption about density dependence.
-The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to Fcrash) for 30 years into the future.
+The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to $F_{crash}$) for 30 years into the future.
 The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure X2**).
-The points on the curve correspond to the projected outcomes for the values of FMSY from each OM.
+The points on the curve correspond to the projected outcomes for the values of $F_{MSY}$ from each OM.
 This was run as a check, as the equilibrium and projections should agree, I.e.
@@
 
-The OMs were also projected for the estimates of different estimates FMSY, with stochasticty in the recruitment deviates.
+The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticty in the recruitment deviates.
 This allows the impact of misspecification of density dependence to be evaluated, i.e.
-if FMSY is underestimated then yield will be less than MSY, while if FMSY is  overestimated  yield will be lost due to overfishing.
-Stochastic projections of the four Operating Models for the four estimates of FMSY are shown in **Figures X3a, X3b, X3c and X3d**
+if $F_{MSY}$ is underestimated then yield will be less than MSY, while if $F_{MSY}$ is overestimated yield will be lost due to overfishing.
+Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures X3a, X3b, X3c and X3d**
 
@@
 
-Estimating density-dependence from observations only possible for weight at age and maturity age.
+Estimating density-dependence from observations only possible for mass-at-age and maturity age.
 Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
 
-* The impact of including density dependence: major impact on reference points.
-Especially Fmsy is much higher when including DD.
+-   The impact of including density dependence: major impact on reference points.
+Especially $F_{MSY}$ is much higher when including DD.
 However, the expected yields in equilibrium conditions are expected to be very similar.
-The main difference is in the initial step from the current Fmsy to any new Fmsy that would include DD, as the higher Fmsy would be applied to a stock that is estimated without DD.
-* The impact of ignoring density dependence
-* The impact of assuming density-dependence incorrectly
-* Robustness of the current ICES framework
+The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
+-   The impact of ignoring density dependence
+-   The impact of assuming density-dependence incorrectly
+-   Robustness of the current ICES framework
 
@@
 
-The purpose of this work is to support the project "Fmsy-project for six high profile fish stocks".
-The project focuses on the current management approach for fish stocks based on single-species Fmsy Values where the only density-dependent process that is included is recruitment.
-The project includes three other density-dependent population dynamics: growth (weight at age), sexual maturity and natural mortality.
+The purpose of this work is to support the project "$F_{MSY}$-project for six high profile fish stocks".
+The project focuses on the current management approach for fish stocks based on single-species $F_{MSY}$ Values where the only density-dependent process that is included is recruitment.
+The project includes three other density-dependent population dynamics: growth (mass-at-age), sexual maturity and natural mortality.
 
@@
 
-**Weight at age**
+**Mass-at-age**
 
-Trends in (stock) weight at age are shown in figure 1.
+Trends in (stock) mass-at-age are shown in figure 1.
 Weight-at-age in the stock is calculated as the average of the weight-at-age in the three spawning components, weighted by the relative size of each component (as estimated in the egg surveys).
@@
   name    = "cond_waa", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Weight at age by year and cohort"))
+  caption = paste0(mystkname, ". mass-at-age by year and cohort"))
 
@@
 
-**Maturity at age**
+**Maturity-at-age**
 
@@
 For the Western component the maturity ogive is updated every year, using maturity data from commercial catch samples from Germany, Ireland, the Netherlands and the UK collected during the first and second quarters.
-Most of the variability in maturity at age is on ages 2 and 3.
+Most of the variability in maturity-at-age is on ages 2 and 3.
 Nearly all fish of age 4 and older are estimated to be mature.
@@
   name    = "cond_mat", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Maturity at age by year and cohort"))
+  caption = paste0(mystkname, ". Maturity-at-age by year and cohort"))
 
@@
 
-**Natural mortality at age (fixed)**
+**Natural mortality-at-age**
 
-Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (@hamre1980biology; @ices2022wgwide; @ices2022macstockannex).
+Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (@hamre1980biology; @ICES2022WGWIDE; @ICES2022MAC).
 
@@
   name    = "cond_natmor", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Natural mortality at age by year and cohort (all ages and years the same M)"))
+  caption = paste0(mystkname, ". Natural mortality-at-age by year and cohort (all ages and years the same M)"))
 
@@
 
-**Density dependence in Weight at age vs. Total biomass**
+**Density dependence in mass-at-age vs. Total biomass**
 
-The distribution of weight at age is shown in the figure below (left) and the relationship between weight at age and total biomass of mackerel is shown to the right.
-Consistent negative slopes on the biomass-weight at age plots can be observed for ages 4 and beyond.
+The distribution of mass-at-age is shown in the figure below (left) and the relationship between mass-at-age and total biomass of mackerel is shown to the right.
+Consistent negative slopes on the biomass-mass-at-age plots can be observed for ages 4 and beyond.
 The patterns for the younger ages are less clear.
@@
   name    = "cond_ddwaa", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Density dependence and weight at age."))
+  caption = paste0(mystkname, ". Density dependence and mass-at-age."))
 
@@
 
-**GAMM at age**
+**GAMM-at-age**
 
-A Generalized Additive Mixed Model (GAMM) was estimated to assess the relationship between weight at age and total biomass.
+A Generalized Additive Mixed Model (GAMM) was estimated to assess the relationship between mass-at-age and total biomass.
 
 weights are predicted from two variables: a smoother of total biomass/mean(total biomass) and age.
-The resulting factor is shown below, indicating a nearly linear decline in weight at age as total biomass increases.
+The resulting factor is shown below, indicating a nearly linear decline in mass-at-age as total biomass increases.
 
@@
   name    = "cond_gammwaa", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". GAMM analysis of density dependence and weight at age."))
+  caption = paste0(mystkname, ". GAMM analysis of density dependence and mass-at-age."))
 
@@
   name    = "cond_gammsummary", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". GAMM summary of density dependence and weight at age."))
+  caption = paste0(mystkname, ". GAMM summary of density dependence and mass-at-age."))
 
@@
 
-**Linear model of density dependence in weight at age**
+**Linear model of density dependence in mass-at-age**
 
@@
 
-[ CHECK THE FORMULA ]
-W^'=¯W^-0.18
-$W_a = {\overline{W_a}} B^{{\beta}_a}$
+[ CHECK THE FORMULA ] W^'=¯W^-0.18 $W_a = {\overline{W_a}} B^{{\beta}_a}$
 
@@
   name    = "cond_lmwaa", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". LM analysis of density dependence and weight at age."))
+  caption = paste0(mystkname, ". LM analysis of density dependence and mass-at-age."))
 
@@
   name    = "cond_lmsummary", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". LM summary of density dependence and weight at age."))
+  caption = paste0(mystkname, ". LM summary of density dependence and mass-at-age."))
 
@@
 
-**Density-dependence analysis of maturity at age**
+**Density-dependence analysis of maturity-at-age**
 
-The distribution of maturity at age is shown in the figure below (left) and the relationship between maturity at age and total biomass of mackerel is shown to the right.
-Negative slopes on the biomass-maturity at age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.
+The distribution of maturity-at-age is shown in the figure below (left) and the relationship between maturity-at-age and total biomass of mackerel is shown to the right.
+Negative slopes on the biomass-maturity-at-age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.
 
@@
   name    = "cond_ddmat", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Analysis of density dependence and maturity at age."))
+  caption = paste0(mystkname, ". Analysis of density dependence and maturity-at-age."))
 
@@
 $f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
+
+```{r}
+load("~/Dropbox/DDMSE/data/om/mac/parMat.RData") 
+```
 
@@
 Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
-The parameters are first used to construct an *`**FLBRP** object representing the equilibrium, and are then coerced into an **FLStock** to model the time series dynamics.
+The parameters are first used to construct an \*\`**FLBRP** object representing the equilibrium, and are then coerced into an **FLStock** to model the time series dynamics.
 
@@
 
-Empirical and theoretical studies (@lorenzen2000allometry) have shown that natural mortality (M) decrease with fish length, but increase with the von Bertalanffy parameters of curvature k and asymptotic length L_∞.
+Empirical and theoretical studies (@lorenzen2000allometry) have shown that natural mortality (M) decrease with fish length, but increase with the von Bertalanffy parameters of curvature $k$ and asymptotic length $L_{\infty}$.
 Therefore, M can be derived from life history parameters, e.g.
-M=1.5k(L_∞/L) (@POPE2021105952) or M=1.5k @jensen1985comparison.
+$M=1.5kL{\infty}/L$) (@POPE2021105952) or M=1.5k @jensen1985comparison.
 @lorenzen2008fish @lorenzen2022natural showed that natural mortality rates vary with body size and age, often by orders of magnitude over the life cycle and support the existence of an allometric relationship between M and body mass (W) at-age of the form:
 
-[Formula here]
+$M=M_uW^d$
 
-Where M_u is M at unit weight, and d is the allometric exponent from empirical relationships shown to range from –0.3 to –0.37.
-d is approximately equal to –1/3, which implies that M is inversely proportional to body length, since mass is approximately proportional to the third power of length.
-This is consistent with [**Gislason et al., (2008)**], who proposed a relationship for natural mortality based on length
+Where $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from -0.3 to -0.37.
+$d$ is approximately equal to -1/3, which implies that $M$ is inversely proportional to body length, since mass is approximately proportional to the third power of length.
+This is consistent with [@gislason2008coexistence], who proposed a relationship for natural mortality based on length
 
-$M=\alphak(L_{\infty}χ)L\beta$
+$M={\alpha}k(L_{\infty}χ)L\beta$
 
-The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to B_MSY.
-The red line indicates the values from the current assessment (for natural mortality with an assumed distribution over ages).
-[ CHECK !]
+The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to $B_{MSY}$.
+The red line indicates the values from the current assessment (for natural mortality that varies by mass and hence age).
 
@@
   name    = "sim_dd3", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in mass-at-age, maturity at age and Natural mortality at age for different levels of biomass relative to Bmsy. Red line is without density dependence.")
+  caption = "Simulation of density dependence in mass-at-age, maturity-at-age and Natural mortality-at-age for different levels of biomass relative to $B_{MSY}$.
+Red line is without density dependence.")
 
@@
 
-From the theoretical life-history parameters and simulations, we can compare the equilibrium total biomass and yield from an assessment without density dependence, and then adding, step-by-step, the density dependent processes for weight at age, maturity at age and natural mortality at age.
-Comparison of equilibrium yield plots for simulation with and without density-dependent processes indicates that the expected values for Bmsy tend to become lower when density dependence is included and that MSY becomes somewhat higher.
+From the theoretical life-history parameters and simulations, we can compare the equilibrium total biomass and yield from an assessment without density dependence, and then adding, step-by-step, the density dependent processes for mass-at-age, maturity-at-age and natural mortality-at-age.
+Comparison of equilibrium yield plots for simulation with and without density-dependent processes indicates that the expected values for $B_{MSY}$ tend to become lower when density dependence is included and that MSY becomes somewhat higher.
 
@@
   name    = "sim_yield", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")
+  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to $B_{MSY}$")
 
@@
 To ensure the MSE projections and the historical estimates are comparable, a Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
-In the next step, we applied a natural mortality at age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M at age.
+In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M-at-age.
 The figure below compares the SAM assessment with the VPA and the VPAM assessment.
@@
 
-When applying density dependent processes to weight at age (i.e. lower weight at age for higher biomass), Bmsy is expected to be somewhat lower. The expected increase in MSY is small.
+When applying density dependent processes to mass-at-age (i.e. lower mass-at-age for higher biomass), $B_{MSY}$ is expected to be somewhat lower. The expected increase in MSY is small.
 
@@
 
-**Density dependence in mass-at-age and maturity at age**
+**Density dependence in mass-at-age and maturity-at-age**
 
-When applying density dependent processes to both weight at age and maturity at age, Bmsy is expected to be lower again and MSY is expected to increase.
+When applying density dependent processes to both mass-at-age and maturity-at-age, $B_{MSY}$ is expected to be lower again and MSY is expected to increase.
 
@@
   name    = "om_VPADDMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age and Maturity at age. Red lines are the non-density dependent values.")
+  caption = "Simulation of density dependence in Mass-at-age and Maturity-at-age. Red lines are the non-density dependent values.")
 
@@
 
-**Density dependence in mass-at-age, maturity at age and M at age**
+**Density dependence in mass-at-age, maturity-at-age and M-at-age**
 
-When applying density dependent processes to all three variables, Bmsy is expected to be lower again and MSY is expected to increase a bit more.
+When applying density dependent processes to all three variables, $B_{MSY}$ is expected to be lower again and MSY is expected to increase a bit more.
 
@@
   name    = "om_VPADDMMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age. Red lines are the non-density dependent values.")
+  caption = "Simulation of density dependence in Mass-at-age, maturity-at-age and M-at-age. Red lines are the non-density dependent values.")
 
@@
 
-The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 30%, Fmsy is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
+The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 30%, $F_{MSY}$ is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
 So while, inclusion of density-dependent processes, has a substantial impact on biological reference points, the resulting change in equilibrium catch is expected to be small.
@@
 Firstly in deterministic mode and secondly in probabilistic mode.
-We used the Fmsy to project forward from 2020 until 2050.
-We used both the Fmsy that is applicable to the density dependent scenario, but also evaluated what would happen if the 'wrong Fmsy' would be used with a particular scenario.
+We used the $F_{MSY}$ to project forward from 2020 until 2050.
+We used both the $F_{MSY}$ that is applicable to the density dependent scenario, but also evaluated what would happen if the 'wrong $F_{MSY}$' would be used with a particular scenario.
 
-The plot below shows the deterministic runs with the appropriate Fmsy value being used to project forward.
-Fmsy is highest when density dependence is taken into account for all three processes.
+The plot below shows the deterministic runs with the appropriate $F_{MSY}$ value being used to project forward.
+$F_{MSY}$ is highest when density dependence is taken into account for all three processes.
 Nevertheless, expected catches are somewhat lower when density dependence is taken into account.
@@
   name    = "mse_determ1", level = 1, display = FALSE,
-  caption = "Deterministic projections at Fmsy for different scenarios")
+  caption = "Deterministic projections at $F_{MSY}$ for different scenarios")
 
@@
 
-**Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age**
+**Deterministic projections at $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age**
 
 How can we understand the results of the deterministic runs?
-Below is a plot of the trends in stock weight at age, maturity at age and natural mortality at age, for the most relevant ages (1-6) and for the four different scenarios (Base, DDM, DDMM and DDMMM).
+Below is a plot of the trends in stock mass-at-age, maturity-at-age and natural mortality-at-age, for the most relevant ages (1-6) and for the four different scenarios (Base, DDM, DDMM and DDMMM).
 The density dependent effect of maturity is strongest when both stock weight and maturity are density dependent.
@@
   name    = "mse_determ2", level = 1, display = FALSE,
-  caption = "Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age")
+  caption = "Deterministic projections at $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age")
 
@@
 
-**Stochastic projections at Fmsy for different scenarios**
+**Stochastic projections at $F_{MSY}$ for different scenarios**
 
-Using stochastic projections by including random processes for recruitment deviations and the Fmsy values that are consistent with the different scenarios, leads to the overview below.
+Using stochastic projections by including random processes for recruitment deviations and the $F_{MSY}$ values that are consistent with the different scenarios, leads to the overview below.
 SSB and total biomass are highest in the case of no density dependence.
@@
   name    = "mse_sim1", level = 1, display = FALSE,
-  caption = "Stochastic projections at Fmsy for different scenarios")
+  caption = "Stochastic projections at $F_{MSY}$ for different scenarios")
 
@@
 
-**Stochastic projections at all Fmsy's from all scenarios**
+**Stochastic projections at all $F_{MSY}$'s from all scenarios**
 
@@
   name    = "mse_sim2", level = 1, display = FALSE,
-  caption = "Stochastic projections at all Fmsy's from all scenarios")
+  caption = "Stochastic projections at all $F_{MSY}$'s from all scenarios")
 
@@
 
-The quality shall live up to that demanded by the International Council for the Exploration of the Sea (ICES) and potentially be used in ICES advice to fisheries managers.
-The Consultant shall be interacting with the leadership of the project Henrik Sparholt, or his representative, in the process of developing the software.
-The software shall be made publicly available via Github or on alternative platform if so directed by Henrik Sparholt or his representative.
+The quality lives up to that required by ICES and so can potentially be used when providing advice to fisheries managers. The software is also publicly available via Github.
+
+Since the early 1980s, natural mortality ($M$) for mackerel has been fixed within the assessment model at 0.15, for all ages and all assessment years.
+This value was calculated based on estimates of total mortality derived from tagging data combined with catch data [@@hamre1980biology].
+M must vary both by age and over time due to changes in abundance of predators, feeding conditions, disease etc.
+\cite{siegfried2009review}.
+
+Changes in mackerel distribution and growth will also affect natural mortality. For example, the recent expansion in distribution area and change in
+mass-at-age could also result in changes in natural mortality.
 

### 2023-03-06 Rmd/report mac.Rmd
update critique of M=0.15

@@
 Since the early 1980s, natural mortality ($M$) for mackerel has been fixed within the assessment model at 0.15, for all ages and all assessment years.
-This value was calculated based on estimates of total mortality derived from tagging data combined with catch data [@@hamre1980biology].
-M must vary both by age and over time due to changes in abundance of predators, feeding conditions, disease etc.
-\cite{siegfried2009review}.
-
-Changes in mackerel distribution and growth will also affect natural mortality. For example, the recent expansion in distribution area and change in
-mass-at-age could also result in changes in natural mortality.
+This value was calculated based on estimates of total mortality derived from tagging data combined with catch data [@hamre1980biology].
+M must, however, vary both by age and over time due to changes in abundance of predators, feeding conditions, disease etc.
+[@siegfried2009review].
+Particularly any changes in mackerel distribution and growth will affect natural mortality.
 

### 2023-03-06 Rmd/report.R
report.R

@@ -706,6 +706,8 @@ mystk     <- "mac";
   #   as.data.frame(par_laurie) %>% mutate(par="par_laurie"),
   #   as.data.frame(par) %>% mutate(par="par")
   # )  
+
+  rm(eq, control)
   
   section <- "04"
   
@@ -721,24 +723,23 @@ mystk     <- "mac";
                                "Rec"=function(x) rec(x)),drop=T) %>% mutate(scen="sam")
   
   # Start with the SAM assessment in forward projection
-  sr  = fmle(as.FLSR(window(ices, start=1999, end=2020),model="bevholtSV"),
-             fixed=list(s=0.8, spr0=mean(spr0(ices))), 
-             control=list(silent=TRUE))
-  eq  = FLBRP(ices,ab(sr))
-  sam = fwdWindow(ices,eq,end=2050) 
+  sam_sr  = fmle(as.FLSR(window(ices, start=1999, end=2020),model="bevholtSV"),
+                 fixed=list(s=0.8, spr0=mean(spr0(ices))), 
+                 control=list(silent=TRUE))
+  sam_eq  = FLBRP(ices,ab(sam_sr))
+  sam     = fwdWindow(ices,sam_eq,end=2050) 
 
-  sam_sr = sr
-  sam_eq = eq  
-  
   # plot(sam_sr)
   
   F  =propagate(window(fbar(sam),start=2020),101)
-  F[]=rep(c(seq(0,                               c(refpts(eq)["msy",  "harvest"]),length.out=51),
-            seq(c(refpts(eq)["msy",  "harvest"]),c(refpts(eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])
+  F[]=rep(c(seq(0,                                   c(refpts(sam_eq)["msy",  "harvest"]),length.out=51),
+            seq(c(refpts(sam_eq)["msy",  "harvest"]),c(refpts(sam_eq)["crash","harvest"])*1.2,length.out=51)[-1]),each=dim(F)[2])
   # F[]=rep(c(seq(0,                               c(refpts(eq)["crash","harvest"])*1.2,length.out=101)),each=dim(F)[2])
   
-  control=as(FLQuants("f"=F),"fwdControl")
-  sam   =fwd(sam,control=control,sr=eq)
+  sam_control=as(FLQuants("f"=F),"fwdControl")
+  sam        =fwd(sam,
[…]

### 2023-03-06 Rmd/report mac.Rmd
report mac

@@
 output:
-  html_document:
+  word_document:
     reference_docx: ../report_template.dotx
@@
 
-load(file=file.path(dropboxdir,paste0("data/om/",mystk,"/parMat.RData")))
+load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
 ```
@@
 
-```{r, echo=FALSE, out.width = "400px", fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 $f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
-
-```{r}
-load("~/Dropbox/DDMSE/data/om/mac/parMat.RData") 
-```
 

### 2023-03-06 Rmd/report mac.Rmd
mac report + refs bib

@@
   name    = "om_assessments", level = 1, display = FALSE,
-  caption = "Comparison of different assessments. VPA-M is used going forward")
+  caption = "Comparison of different assessments. VPA-M is used as. the base case going forward")
 
@@
 
-The quality lives up to that required by ICES and so can potentially be used when providing advice to fisheries managers. The software is also publicly available via Github.
+The current management of fish stocks is based on achieving the fishing mortality ($F_{MSY}$) that is expected to achieve the maximum sustainable yield (MSY).
+Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms.
+Density dependence, however, may also operate in processes such as growth, sexual maturity, and natural mortality.
+A main aim of this study was to condition Operating Models (OMs) on alternative hypotheses about density dependence for use in Management Strategy Evaluation (MSE).
+The objective is to evaluate the robustness of the assessments conducted to support the ICES advice framework.
+To do this, OMs were developed that include density dependence in growth, maturity, natural mortality, individually and combined.
+These OMs were then used to evaluate the consequences of including or not including density dependence in management advice.
+
+Initial explorations of density dependent effect were conducted based on life history theory, as used by the International Council for the Exploration of the Sea (ICES) to develop advice conducting MSE for data poor stocks [@fischer2020ga].
+**Figure 10** shows the the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
+These allow the relationships between the MSY reference points and virgin biomass to be seen.
+
+Estimating density-dependence from observations was only possible for mass-at-age and maturity age.
+Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
+
+In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M-at-age.
+The figure below compares the SAM assessment with the VPA and the VPAM assessment.
+A main difference is in the recruitment estimates.
+Using age-varying M tends to flatten the recruitment pattern.
+The vpaM model was used as the base case to be compared against the density-dependent simulations.
+
+Next, the operating models were constructed. This was done in a number of steps.
+
+1.
+The ICES (SAM) assessment was taken from the ICES WGWIDE files (2022).
+SAM is a statistical catch-at-age model that estimates error in the catch-at-age.
+In the case of `r mystkname`, the assessment gives a higher weight to the recruitment index than to the catch data for ages 0 and 1.
+This means that the estimates of recruitment in the assessment can be very different from the actual number of fish that are recruited to the fishery.
+
+2.
+Next, a VPA model model was estimated.
+This was done to ensure the MSE projections and the historical estimates are comparable.
+The Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
+
+3.
+The third step was to convert the VPA model with fixed natural mortality at age and by year into an age-varying natural mortality vector.
+We applied a natural mortality-at-age vector derived from the Lorenzen approach (@lorenzen2000allometry) which would mimick the VPA assessment with fixed M-at-age.
+This assessment was then treated as the base case against which the density dependent operating models were compared.
+
+4.
+The density-dependent scenarios were constructed cumulative.
+First by adding density dependence in mass-at-age (M), then in mass and maturity-at-age (MM) and then in mass, maturity and M-at-age (MMM),
+
+Therefore, the four Operating Models were
+
+* VPA Base Case, with varying M-at-age
+* Density dependence in mass-at-age,
+* Density dependence in mass, maturity-at-age,
+* Density dependence in mass, maturity and M-at-age,
+
+The MSY reference points $B_{MSY}$ and $F_{MSY}$ were calculated based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship [@sissenwine1987alternative] or by conducting long-term projections under constant fishing moralities.
+
+Each of the four OMs were then projected for the values of $F_{MSY}$ based on the four assumptions about density dependence.
+This allows a comparison between the yield that could potentially be achieved and that forgone by making an incorrect assumption about density dependence.
+
+The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to $F_{crash}$) for 30 years into the future.
+The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure X2**).
+The points on the curve correspond to the projected outcomes for the values of $F_{MSY}$ from each OM.
+This was run as a check, as the equilibrium and projections should agree, I.e.
+the points at MSY should be the same colour as the curve.
+
+The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticty in the recruitment deviates.
+This allows the impact of misspecification of density dependence to be evaluated, i.e.
+if $F_{MSY}$ is underestimated then yield will be less than MSY, while if $F_{MSY}$ is overestimated yield will be lost due to overfishing.
+Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures X3a, X3b, X3c and X3d**
+
+**Figure X4** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points.
+
+The impact of including density dependence was primarily on the biological reference points.
+Especially $F_{MSY}$ is much higher when including density dependence.
+However, the expected yields in equilibrium conditions are  very similar.
+The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
+
+Ignoring density dependent processes when they would be happening is expected to lead to ...
+
+The impact of assuming the wrong type of density-dependence processes, i.e.
+applying the reference points from the one density dependent operating model to a simulation with other density dependent processes, could lead to ....
 
@@
 Particularly any changes in mackerel distribution and growth will affect natural mortality.
+The impact of including age-varying natural mortality is expected to ....
+
+The quality of this analyses lives up to the quality standards that are required by ICES and so can potentially be used when providing advice to fisheries managers.
+The software is publicly available via Github (https://github.com/martinpastoors/ddmse).
+All data is derived from WGWIDE 2022 @ICES2022WGWIDE.
 

### 2023-03-07 Rmd/mseMac.Rmd
both aligned

@@
 om=fwdWindow(window(om,end=2020),end=2080,eqM)
-f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.5,length.out=100),each=61),
-           dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
-om=fwd(om,fbar=f,sr=eqM)
+F    =rep(c(seq(0,                                  c(refpts(eqM)["msy",  "harvest"]),  length.out=51),
+            seq(c(refpts(eqM)["msy",  "harvest"]),c(refpts(eqM)["crash","harvest"])*1.5,length.out=51)[-1]))
+F    =FLQuant(rep(F,each=61),dimnames=list(year=2020:2080,iter=seq(101)))
+om=fwd(om,fbar=F,sr=eqM)
 
@@
   oms[["DD Mass"]]=ddFn(iYear,oms[["DD Mass"]],params,TRUE,FALSE,FALSE)
-  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=f[,iYear],sr=eqM)
+  oms[["DD Mass"]]=fwd(oms[["DD Mass"]],fbar=F[,iYear],sr=eqM)
 
   oms[["DD Mass, Mat"]]=ddFn(iYear,oms[["DD Mass, Mat"]],params,TRUE,TRUE,FALSE)
-  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=f[,iYear],sr=eqM)
+  oms[["DD Mass, Mat"]]=fwd(oms[["DD Mass, Mat"]],fbar=F[,iYear],sr=eqM)
 
   oms[["DD Mass, Mat, M"]]=ddFn(iYear,oms[["DD Mass, Mat, M"]],params,TRUE,TRUE,TRUE)
-  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=f[,iYear],sr=eqM)
+  oms[["DD Mass, Mat, M"]]=fwd(oms[["DD Mass, Mat, M"]],fbar=F[,iYear],sr=eqM)
   } 
@@
 
-om=fwdWindow(window(iter(om,51),end=2020),end=2080,eqM)
-F =propagate(fbar(om[,ac(2020:2080)]),4)
+omFmsy=fwdWindow(window(iter(om,51),end=2020),end=2080,eqM)
+F =propagate(fbar(omFmsy[,ac(2020:2080)]),4)
 F[]=rep(rfpts$Fmsy,each=61)
@@
 
-om =fwd(om,control=control,sr=eqM)
+omFmsy =fwd(omFmsy,control=control,sr=eqM)
 
-prj=FLStocks("Base"     =om)
-prj[["DD Mass"]]        =om
-prj[["DD Mass, Mat"]]   =om
-prj[["DD Mass, Mat, M"]]=om
+prj=FLStocks("Base"     =omFmsy)
+prj[["DD Mass"]]        =omFmsy
+prj[["DD Mass, Mat"]]   =omFmsy
+prj[["DD Mass, Mat, M"]]=omFmsy
 for (i in ac(2020:2080)) {
@@
 ```{r}
+om.  =om
+oms. =oms
+prj. =prj
+prjs.=prjs   
+save(eqM,    file="~/Desktop/tmp/eqM.RData")
+save(om.,    file="~/Desktop/tmp/om..RData")
+save(omFmsy, file="~/Desktop/tmp/omFmsy.RData")
+save(oms.,   file="~/Desktop/tmp/oms..RData")
+save(prj.,   file="~/Desktop/tmp/prj..RData")
+save(prjs.,  file="~/Desktop/tmp/prjs..RData") 
+```
+
+```{r}
 ggarrange(plot(prjs[[1]])+ 
@@
 **Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$
-
-```{r, projectons-save}
-save(prjs,file=paste(dirRuns,"/",stock,"Prjs.RData",sep=""))
-```
 
@@
 ```
+
+```{r, eval=FALSE}
+load("~/Desktop/tmp/eqM.RData")
+load("~/Desktop/tmp/om..RData")
+load("~/Desktop/tmp/oms..RData")
+load("~/Desktop/tmp/prj..RData")
+load("~/Desktop/tmp/prjs..RData")
+
+plot(FLStocks("mse"=om.,"report"=om))
+
+refpts(vpaM_eq)
+refpts(eqM)
+
+plot(eqM)
+plot(vpaM_eq)
+
+plot(vpaM_eq)
+```

### 2023-03-07 Rmd/report.R

@@ -723,7 +723,8 @@ mystk     <- "mac";
                                "Rec"=function(x) rec(x)),drop=T) %>% mutate(scen="sam")
   
   # Start with the SAM assessment in forward projection
-  sam_sr  = fmle(as.FLSR(window(ices, start=1999, end=2020),model="bevholtSV"),
+  #@# years changed and single iter to get rid of noise in 2020 @#@#@#@#@#@#@#@# 
+  sam_sr  = fmle(as.FLSR(window(iter(ices,1), start=2000, end=2019),model="bevholtSV"),
                  fixed=list(s=0.8, spr0=mean(spr0(ices))), 
                  control=list(silent=TRUE))
   sam_eq  = FLBRP(ices,ab(sam_sr))
@@ -740,6 +741,7 @@ mystk     <- "mac";
   sam        =fwd(sam,
                   control=sam_control,
                   sr=sam_eq)
+  #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   
   df_om  =model.frame(FLQuants(sam[,"2050"], "Biomass"=function(x) biomass(x),
                                     "SSB"=function(x) ssb(x),
@@ -763,8 +765,9 @@ mystk     <- "mac";
                                 "F"=function(x) fbar(x),
                                 "Yield"=function(x) catch(x),
                                 "Rec"=function(x) rec(x)),drop=T) %>% mutate(scen="vpa") )
-    
-  vpa_sr  =fmle(as.FLSR(window(vpa, start=1999, end=2020),
+  
+  #@# years changed and single iter to get rid of noise in 2020 @#@#@#@#@#@#@#@# 
+  vpa_sr  =fmle(as.FLSR(window(iter(vpa,1),start=2000, end=2019),
                         model="bevholtSV"),
             fixed=list(s   =0.8, spr0=mean(spr0(vpa))), 
             control=list(silent=TRUE))
@@ -782,6 +785,7 @@ mystk     <- "mac";
   vpa         = fwd(vpa,
[…]

### 2023-03-07 Rmd/mseMac.Rmd
checked!

@@
 
-+  Load ICES assessment
-+  Time varying M
-+  Conduct VPA
+-   Load ICES assessment
+-   Time varying M
+-   Conduct VPA
 
@@
 ```{r, pars}
-params=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,bref=NA,delta=-0.3,matk=30,w50=0.17))
+params=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,   bref=3.69e+06,delta=-0.3,     matk=30,      w50=0.17))
+params=FLPar(c(m1=9.69e-02         ,m2=-2.88e-01,bref=4.11e+06,delta=-2.00e-01,matk=2.80e+01,w50=1.66e-01)) 
 ```
@@
 F[]=rep(rfpts$Fmsy,each=61)
-control=as(FLQuants("f"=F),"fwdControl")
 
-omFmsy =fwd(omFmsy,control=control,sr=eqM)
+omFmsy =fwd(omFmsy,fbar=F,sr=eqM)
 
@@
     prj[[2]] =ddFn(i,prj[[2]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
-    control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[2]] =fwd(prj[[2]],control=control,sr=eqM)
+    prj[[2]] =fwd(prj[[2]],fbar=F[,i],sr=eqM)
 
     prj[[3]] =ddFn(i,prj[[3]],params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
-    control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[3]] =fwd(prj[[3]],control=control,sr=eqM)
+    prj[[3]] =fwd(prj[[3]],fbar=F[,i],sr=eqM)
 
     prj[[4]] =ddFn(i,prj[[4]],params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-    control=as(FLQuants("f"=F[,i]),"fwdControl")
-    prj[[4]] =fwd(prj[[4]],control=control,sr=eqM)
+    prj[[4]] =fwd(prj[[4]],fbar=F[,i],sr=eqM)
     }
@@
 
-```{r}
+```{r, check, fig.height=6,fig.width=8}
+eqCurves=ldply(oms, function(x) { 
+  model.frame(FLQuants(x[,dim(oms[[1]])[2]],
+             biomass=function(x) stock(x),
+             ssb    =function(x) ssb(  x),
+             catch  =function(x) catch(x),
+             f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+
 ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
@@
 
-#plot(prj,probs=0,worm=seq(4))+facet_grid(qname~stock,scale="free")
-```
-
-```{r, check, fig.height=6,fig.width=8}
 ggplot()+
-  geom_line( aes(ssb,catch,col=.id),   data=eqCurves)+
+  geom_line( aes(ssb,catch,col=.id),data=eqCurves)+
   geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))+
@@
     ## No DD ####################################################
-    control=as(FLQuants("f"=F),"fwdControl")
-    x1     =fwd(base,control=control,sr=eqM,residuals=devRec)
+    x1     =fwd(base,fbar=F,sr=eqM,residuals=devRec)
 
@@
     for (iYr in ac(2020:2080)) {
-        control=as(FLQuants("f"=F[,iYr]),"fwdControl")
-
           ## DD M ##################################################         
           x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
-          x2 =fwd(x2,control=control,sr=eqM,residuals=devRec)
+          x2 =fwd(x2,fbar=F[,iYr],sr=eqM,residuals=devRec)
 
@@
           x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
-          x3 =fwd(x3,control=control,sr=eqM,residuals=devRec)
+          x3 =fwd(x3,fbar=F[,iYr],sr=eqM,residuals=devRec)
 
@@
           x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-          x4 =fwd(x4,control=control,sr=eqM,residuals=devRec)
+          x4 =fwd(x4,fbar=F[,iYr],sr=eqM,residuals=devRec)
           }
@@
 save(prjs.,  file="~/Desktop/tmp/prjs..RData")
+save(params, file="~/Desktop/tmp/params.RData") 
 ```
@@
     ## No DD ####################################################
-    control=as(FLQuants("f"=F),"fwdControl")
-    x1     =fwd(iter(base,1),control=control,sr=eqM)
+    x1     =fwd(iter(base,1),fbar=F,sr=eqM)
 
@@
     for (iYr in ac(2020:2080)) {
-        control=as(FLQuants("f"=F[,iYr]),"fwdControl")
-
           ## DD M ##################################################         
           x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
-          x2 =fwd(x2,control=control,sr=eqM)
+          x2 =fwd(x2,fbar=F[,iYr],sr=eqM)
 
@@
           x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
-          x3 =fwd(x3,control=control,sr=eqM)
+          x3 =fwd(x3,fbar=F[,iYr],sr=eqM)
 
@@
           x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
-          x4 =fwd(x4,control=control,sr=eqM)
+          x4 =fwd(x4,fbar=F[,iYr],sr=eqM)
           }
@@
 
-plot(vpaM_eq)
+plot(FLStocks("MSE"=oms.[[2]],"Report"=oms[[2]]))
+
+eqCurves=ldply(oms., function(x) { 
+    model.frame(FLQuants(x[,dim(oms[[1]])[2]],
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+ts=ldply(prj., function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
+                                                  f    =function(x) fbar(x), 
+                                                  catch=function(x) catch(x)),drop=T))
+ts=transform(ts,What=.id)
+ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
+
+ggplot()+
+    geom_line( aes(ssb,catch,col=.id),data=eqCurves)+
+    geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))+
+    theme_bw()+theme(legend.position="bottom")+
+    xlab("SSB")+ylab("Yield")
+
+eqCurves=ldply(oms, function(x) { 
+    model.frame(FLQuants(x[,dim(oms[[1]])[2]],
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
+                                                  f    =function(x) fbar(x), 
+                                                  catch=function(x) catch(x)),drop=T))
+ts=transform(ts,What=.id)
+ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
+
+ggplot()+
+    geom_line( aes(ssb,catch,col=.id),data=eqCurves)+
+    geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))+
+    theme_bw()+theme(legend.position="bottom")+
+    xlab("SSB")+ylab("Yield")
 ```

### 2023-03-07 Rmd/report.R

@@ -806,7 +806,7 @@ mystk     <- "mac";
   # save(par, file=file.path(dropboxdir, paste0("data/om/par",mystk,".RData")))
   # =======================================================================================
   
-   
+
   fileConn <-file(file.path(tablesdir, paste(section,mystk, "par_om.txt", sep="_")))
   as.data.frame(vpaM_par) %>% dplyr::select(-iter) %>% pander::pandoc.table(style="simple") %>% capture.output() %>% writeLines(., con=fileConn)
   close(fileConn)
@@ -1234,26 +1234,28 @@ mystk     <- "mac";
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
   
   om=fwdWindow(window(vpaM,end=2020),end=2050,vpaM_eq)
-  f =FLQuant(rep(seq(0,c(refpts(vpaM_eq)["crash","harvest"])*1.5,length.out=100),each=31),
-             dimnames=dimnames(fbar(om)[,ac(2020:2050)]))
-  om=fwd(om,fbar=f,sr=vpaM_eq)
+  F    =rep(c(seq(0,                                    c(refpts(vpaM_eq)["msy",  "harvest"]),  length.out=51),
+              seq(c(refpts(vpaM_eq)["msy",  "harvest"]),c(refpts(vpaM_eq)["crash","harvest"])*1.5,length.out=51)[-1]))
+  F    =FLQuant(rep(F,each=31),dimnames=list(year=2020:2050,iter=seq(101)))
+  om=fwd(om,fbar=F,sr=vpaM_eq)
+  
   
   #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   #@# Projection for range of F                                              @#@
   #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
-  oms=FLStocks("Base"     =om)
-  oms[["DD Mass"]]        =om
-  oms[["DD Mass, Mat"]]   =om
-  oms[["DD Mass, Mat, M"]]=om
[…]

### 2023-03-07 Rmd/report mac.Rmd
checked

@@
 
-**Deterministic projections at $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age**
+**Deterministic projections at** $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age
 
@@
 
-**Stochastic projections at $F_{MSY}$ for different scenarios**
+**Stochastic projections at** $F_{MSY}$ for different scenarios
 
@@
 
-**Stochastic projections at all $F_{MSY}$'s from all scenarios**
+**Stochastic projections at all** $F_{MSY}$'s from all scenarios
 
@@
 
-The current management of fish stocks is based on achieving the fishing mortality ($F_{MSY}$) that is expected to achieve the maximum sustainable yield (MSY).
+The current management of fish stocks is based on achieving a fishing mortality ($F_{MSY}$) that is expected to achieve the maximum sustainable yield (MSY).
 Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms.
@@
 
-Estimating density-dependence from observations was only possible for mass-at-age and maturity age.
+Estimating density-dependence from observations was only possible for mass-at-age and maturity-at-qge.
 Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
 
-In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M-at-age.
+In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which would mimick the VPA assessment with fixed M-at-age.
 The figure below compares the SAM assessment with the VPA and the VPAM assessment.
@@
 
-* VPA Base Case, with varying M-at-age
-* Density dependence in mass-at-age,
-* Density dependence in mass, maturity-at-age,
-* Density dependence in mass, maturity and M-at-age,
+-   VPA Base Case, with varying M-at-age
+-   Density dependence in mass-at-age,
+-   Density dependence in mass, maturity-at-age,
+-   Density dependence in mass, maturity and M-at-age,
 
@@
 The quality of this analyses lives up to the quality standards that are required by ICES and so can potentially be used when providing advice to fisheries managers.
-The software is publicly available via Github (https://github.com/martinpastoors/ddmse).
+The software is publicly available via Github (<https://github.com/martinpastoors/ddmse>).
 All data is derived from WGWIDE 2022 @ICES2022WGWIDE.

### 2023-03-07 Rmd/report.R

@@ -1259,7 +1259,27 @@ mystk     <- "mac";
   } 
   #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   
+  #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   # reference points data.frame (equilibrium)
+  #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
+  rfs=ldply(oms, function(x) { 
+    model.frame(FLQuants(x[,"2050"],
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+  
+  # reference points (at maximum catch)
+  rfpts=ddply(rfs,.(.id), with, {  
+    flag=catch==max(catch)
+    data.frame(Bmsy  =biomass[flag],
+               SSBmsy=ssb[flag],
+               MSY   =catch[flag],
+               Fmsy  =f[flag],
+               Virgin=max(ssb),
+               B0    =max(biomass))}) %>% 
+    dplyr::rename(scen=.id)
+  #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   
   #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   #@# projection with different Fs                                           @#@
@@ -1649,8 +1669,8 @@ mystk     <- "mac";
                          f      =function(x) fbar( x)),drop=TRUE)[,-1]})
   
   ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
-                                                 f    =function(x) fbar(x), 
-                                                 catch=function(x) catch(x)),drop=T))
+                                                    f    =function(x) fbar(x), 
[…]

### 2023-03-07 Rmd/report mac.Rmd
Updated report.R and report mac

@@
 
-**L. Kell, M.A. Pastoors**
+**`r ifelse(mystk=="mac", "M.A. Pastoors & L.T. Kell", "L.T. Kell & M.A. Pastoors")`**
 
@@
 
-Commissioned by: University of Copenhagen (Henrik Sparholt)
+Version: 1.0 (draft)
+
+&nbsp;
+
+**Commissioned by:**
+
+Henrik Sparholt
+
+University of Copenhagen
+
+Globe Institute
+
+Center for Macroecology Evolution and Climate
+
+Universitetsparken 15, 3rd floor, room 339
+
+2100 Copenhagen Ø
+
+Denmark
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+```{r, echo=FALSE, out.width = "300px", fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
+
+if(mystk=="mac") {
+  knitr::include_graphics("../MPFF Seaplusplus.png")
+} else {
+  knitr::include_graphics("../Seaplusplus MPFF.png")
+}
+
+```
+
+\newpage
 
@@
 
-**Figure X4** summaries the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence.
+**Figure X4** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
 
@@
 
--   The impact of including density dependence: major impact on reference points.
+The impact of including density dependence: major impact on reference points.
 Especially $F_{MSY}$ is much higher when including DD.
@@
 The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
--   The impact of ignoring density dependence
--   The impact of assuming density-dependence incorrectly
--   Robustness of the current ICES framework
+
+REMOVE 
+[ -   The impact of ignoring density dependence ]
+[ -   The impact of assuming density-dependence incorrectly ]
+[ -   Robustness of the current ICES framework ]
 
@@
 ```
+
+\newpage
 
@@
 
-All data and results are available on DropBox (<https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0>)
+All data and results are available on a dedicated DropBox folder.
 
@@
 
-Ignoring density dependent processes when they would be happening is expected to lead to ...
+REMOVE
 
-The impact of assuming the wrong type of density-dependence processes, i.e.
+[ Ignoring density dependent processes when they would be happening is expected to lead to ... ]
+
+[ The impact of assuming the wrong type of density-dependence processes, i.e.
 applying the reference points from the one density dependent operating model to a simulation with other density dependent processes, could lead to ....
+]
 
@@
 
+This work was funded under a consultancy agreement between the University of Copenhagen and `r ifelse(mystk=="mac", "Martin Pastoors F&F", "Sea++")` as part of the project “117957 Fmsy-project for six high profile fish stocks”.
+The current work focused on `r mystkname` and was carried out in the months of January-March 2023.
+A similar project was carried out for `r ifelse(mystk=="mac", "Blue whiting", mystkname)` by `r ifelse(mystk=="mac", "Sea++", "Martin Pastoors F&F")`.
+Both projects were carried in close coordination between Sea++ and Martin Pastoors F&F.
+
 <!-- =============================================================================================================== -->
@@
 # References
-
-<!-- =============================================================================================================== -->
-
-<!-- Annex 1: OM conditioning                                                                           -->
-
-<!-- =============================================================================================================== -->
-
-# Annex 1: OM conditioning
-
-<!-- =============================================================================================================== -->
-
-<!-- Annex 2: Density dependent processes                                                                           -->
-
-<!-- =============================================================================================================== -->
-
-# Annex 2: Density dependent processes

### 2023-03-07 Rmd/report.R

@@ -67,10 +67,12 @@ myparams <- data.frame(
   wt1     = c(0.01                         , NA               ), # for resetting weight at age 1
   minage  = c(0                            , 1                ),
   maxage  = c(12                           , 10               ),
+  minyear = c(1991                         , 2000             ),
   maxyear = c(2050                         , 2050             ),
   m1scaler= c(8                            , 10               ),
   w50     = c(0.166                        , 0.08             ),
-  matk2   = c(28                           , 63               ) )
+  matk2   = c(28                           , 63               ),
+  steepness=c(0.8                          , 0.5               ))
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
@@ -605,24 +607,6 @@ mystk     <- "mac";
   print(p) 
   dev.off()
   
-  p <-
-    df %>% 
-    mutate(scen = factor(scen, levels=c("no-DD", "DD mass", "DD mass+mat", "DD mass+mat+M"))) %>% 
-    ggplot(aes(biomass, catch))+
-    theme_bw() +
-    geom_line(aes(colour=scen)) +
-    labs(x="Total biomass", y="Catch", colour="") +
-    scale_colour_manual(values =c("no-DD"         = "black",
-                                  "DD mass"       = "red", 
-                                  "DD mass+mat"   = "blue",
-                                  "DD mass+mat+M" = "darkgreen")) 
-  
-  
-  jpeg(filename=file.path(figuresdir, paste(section,mystk, "eq_yield.jpg", sep="_")),
-       width=10, height=10, units="in", res=300)
-  print(p) 
-  dev.off()
-  
   # Check 1 --------------------------------------------------------------------
 
   # ptm <- proc.time()
@@ -711,7 +695,7 @@ mystk     <- "mac";
   
   section <- "04"
   
-  ices=window(stk,start=1991)
+  ices=window(stk,start=myparams[myparams$stock==mystk,"minyear"])
   stock(ices) =computeStock(ices)
   name(ices) <- mystkname
 
@@ -725,7 +709,8 @@ mystk     <- "mac";
   # Start with the SAM assessment in forward projection
[…]

### 2023-03-07 Rmd/report.R
added equilibrium figure

@@ -1210,6 +1210,19 @@ mystk     <- "mac";
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
   
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
+  
   # ------------------------------------------------------------------------------
   # 5. MSE
   # ------------------------------------------------------------------------------
@@ -1644,8 +1657,9 @@ mystk     <- "mac";
   dev.off()
 
 #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
-#@#  Check                                                                   @#@
+#@#  Equilibrium Check                                                       @#@
 #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
+  
   eqCurves=ldply(oms, function(x) { 
     model.frame(FLQuants(x[,dim(oms[[1]])[2]],
                          biomass=function(x) stock(x),
@@ -1659,11 +1673,18 @@ mystk     <- "mac";
   ts=transform(ts,What=.id)
   ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
   
-  ggplot()+
+  p <-
+    ggplot()+
     geom_line( aes(ssb,catch,col=.id),data=eqCurves)+
     geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))+
     theme_bw()+theme(legend.position="bottom")+
     xlab("SSB")+ylab("Yield")
+  
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "equilibrium_check.jpg", sep="_")),
+       width=10, height=6, units="in", res=300)
+  print(p) 
+  dev.off()
+  
 #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   
   rm(p, p1, p2, p3, p4)

### 2023-03-07 Rmd/report.R
checked

@@ -1669,8 +1669,8 @@ mystk     <- "mac";
                          f      =function(x) fbar( x)),drop=TRUE)[,-1]})
   
   ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
-                                                    f    =function(x) fbar(x), 
-                                                    catch=function(x) catch(x)),drop=T))
+                                                 f    =function(x) fbar(x), 
+                                                 catch=function(x) catch(x)),drop=T))
   ts=transform(ts,What=.id)
   ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
   
@@ -1681,10 +1681,21 @@ mystk     <- "mac";
     xlab("SSB")+ylab("Yield")
 #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   
-  rm(p, p1, p2, p3, p4)
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   # load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
+  
+  {r, echo=FALSE, out.width = "500px" fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+  
+  fig_nums(
+    name    = "eq_check", level = 1, display = FALSE,
+    caption = "Check of equilibrium curves against end points of simulated time series at $F_{MSY}$'s from all scenarios")
+  
+  knitr::include_graphics(path=
+                            file.path(figuresdir,paste(section,mystk,"equilibrium_check.jpg",sep="_")))
+  
+  
+  rm(p, p1, p2, p3, p4)
   
 # } # end of stk loop
 

### 2023-03-07 Rmd/report mac.Rmd
checked
updates

@@
 The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to $F_{crash}$) for 30 years into the future.
-The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure X2**).
+The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure 15**).
 The points on the curve correspond to the projected outcomes for the values of $F_{MSY}$ from each OM.
@@
 if $F_{MSY}$ is underestimated then yield will be less than MSY, while if $F_{MSY}$ is overestimated yield will be lost due to overfishing.
-Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures X3a, X3b, X3c and X3d**
-
-**Figure X4** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
+Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
 
@@
 The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
-
-REMOVE 
-[ -   The impact of ignoring density dependence ]
-[ -   The impact of assuming density-dependence incorrectly ]
-[ -   Robustness of the current ICES framework ]
 
@@
 
-[ CHECK THE FORMULA ] W^'=¯W^-0.18 $W_a = {\overline{W_a}} B^{{\beta}_a}$
+$W_a = {\overline{W_a}} (B/B_{MSY})^{{\beta}_a}$
 
@@
 
-$W_a = {\overline{W_a}} B^{{\beta}_a}$
+$W_a = {\overline{W_a}} {(B/B_{MSY})^{{\beta}_a}$
 
@@
 
-[ NEED TO CHECK THIS FIGURE; IS THIS A PLAUSIBLE RESULT? ]
-
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 
-*`r fig_nums("mse_eq2050")`*
+*`r fig_nums("mse_sim2")`*
+
+```{r, echo=FALSE, out.width = "500px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "eq_check", level = 1, display = FALSE,
+  caption = "Check of equilibrium curves against end points of simulated time series at $F_{MSY}$'s from all scenarios")
+
+  knitr::include_graphics(path=
+                            file.path(figuresdir,paste(section,mystk,"equilibrium_check.jpg",sep="_")))
+
+```
+
+*`r fig_nums("eq_check")`*
 
@@
 if $F_{MSY}$ is underestimated then yield will be less than MSY, while if $F_{MSY}$ is overestimated yield will be lost due to overfishing.
-Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures X3a, X3b, X3c and X3d**
+Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**
 
-**Figure X4** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points.
 
@@
 The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
-
-REMOVE
-
-[ Ignoring density dependent processes when they would be happening is expected to lead to ... ]
-
-[ The impact of assuming the wrong type of density-dependence processes, i.e.
-applying the reference points from the one density dependent operating model to a simulation with other density dependent processes, could lead to ....
-]
 

### 2023-03-07 Rmd/report.R

@@ -1495,37 +1495,47 @@ mystk     <- "mac";
   }) ## F loop
 
   # generating stochastic projection data frame
-  prjs_df <- tibble()
-  for (i in 1:length(prjs)) {
-    prjs_df <-
-      bind_rows(
-        prjs_df,
-        ldply(prjs[[i]], function(x) {
-          model.frame(FLQuants(x,
-                               rec    =function(x) rec(x),
-                               biomass=function(x) stock(x),
-                               ssb    =function(x) ssb(  x),
-                               catch  =function(x) catch(x),
-                               f      =function(x) fbar( x)),drop=TRUE)})    
-      )
-  }
+  prjs_df=ldply(prjs, function(x) ldply(x, function(x) { 
+    model.frame(FLQuants(x,
+                         rec    =function(x) rec(x),
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)})) %>% 
+    mutate(Fmsy=factor(F, levels=rfpts$Fmsy,labels=c("Base","M","MM","MMM"))) %>% 
+    as_tibble()
 
-  nyears = length(unique(prjs_df$year))
-  niters = length(unique(prjs_df$iter))
-  nscen  = length(unique(prjs_df$.id))
-  
-  prjs_df <- prjs_df %>% 
-    bind_cols(Fmsy    = rep(rfpts$Fmsy, each=(nyears*niters*nscen))) %>% 
-    bind_cols(Fmsy_id = rep(unique(prjs_df$.id), each=(nyears*niters*nscen)))
+  # prjs_df <- tibble()
+  # for (i in 1:length(prjs)) {
+  #   prjs_df <-
+  #     bind_rows(
+  #       prjs_df,
+  #       ldply(prjs[[i]], function(x) {
+  #         model.frame(FLQuants(x,
+  #                              rec    =function(x) rec(x),
+  #                              biomass=function(x) stock(x),
+  #                              ssb    =function(x) ssb(  x),
[…]

### 2023-03-08 Rmd/report mac.Rmd
Minor updates; added some text.

@@
 
-`r format(Sys.time(), '%d/%m/%Y')`
+Version: 1.0 (07 March 2023)
 
-Version: 1.0 (draft)
+Version: 1.1 (`r format(Sys.time(), '%d %B %Y')`)
 
@@
 Denmark
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
 
@@
 
-**Stochastic projections at all** $F_{MSY}$'s from all scenarios
+**Stochastic projections at all $F_{MSY}$'s from all scenarios**
+
+Misspecification of Fmsy, i.e.
+using an Fmsy from one scenario and applying it to a different scenario has only limited impacts on the overall outcomes in terms of recruitment, biomass or yield.
+In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colours.
+For example a valid combination is the Base case with the red curves (Fmsy=0.33) or the M scenario with the green curve.
+Overall the misspecification does not results in noticeable differences.
 
@@
 *`r fig_nums("mse_sim2")`*
+
+As a chech, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colours).
+The points fall on the equilibrium curve, thereby confirming the consistency between the two approaches.
 
@@
 
+Estimating density-dependence from observations was only possible for mass-at-age and maturity age.
+Therefore, density dependence in natural mortality can only be explored from theoretical considerations, because current assessment does not include age or time varying natural mortality.
+M is likely to vary by age and time in relation to processes such as density dependence.
+While the recent expansion in distribution area and change in weight-at-age could be associated with changes in natural mortality.
+Therefore in the OMs we modelled M as a function of mass-at-age.
+
 The impact of including density dependence was primarily on the biological reference points.
@@
 Particularly any changes in mackerel distribution and growth will affect natural mortality.
-The impact of including age-varying natural mortality is expected to ....
+
+Given the uncertainty in the assessment and the estimation of density dependence, it is likely that the Pretty Good Yield approach would ensure sustainable exploitation.
+Also, because we did not yhet consider stochastic variability in future developments in mass, maturity and natural mortality which are likely to obscure the impacts of density dependence.
+Variations in these processes due to changes in climate and range expansion are also likely to occur.
+
+It is suggested that emprirical harvest control rules that respond to changes in productivity could be evaluated in contrast to the highly complex harvest rules based on SAM stock assessment models.
 

### 2023-03-08 Rmd/report.R

@@ -1611,6 +1611,21 @@ mystk     <- "mac";
     labs(y="", x="", colour="Fmsy", fill="Fmsy") +
     expand_limits(y=0) +
     facet_grid(variable ~ .id, scales="free_y")
+
+  # prjs_df_summ %>% filter(year==2050) %>% 
+  #   mutate(specified=ifelse(.id==Fmsy, "correct","misspecified")) %>% 
+  #   ggplot(aes(x=.id, y=median)) +
+  #   theme_bw() + theme(legend.position="bottom") +
+  #   geom_point(aes(colour=specified),
+  #              position=position_dodge(.4)) +
+  #   geom_errorbar(aes(ymin=lower, ymax=upper, colour=specified), 
+  #                 width=0.2, position=position_dodge(.4)) +
+  #   geom_text(aes(label=paste(Fmsy, F), colour=specified), 
+  #             hjust=0, position=position_dodge(.4)) +
+  #   expand_limits(y=0) +
+  #   facet_wrap(~variable, scales="free_y")
+  
+    
   
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "stochastic_projections_at_different_Fmsys.jpg", sep="_")),
        width=10, height=10, units="in", res=300)
@@ -1700,7 +1715,7 @@ mystk     <- "mac";
   rm(p, p1, p2, p3, p4)
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
-  # load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
+  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   
   
   

### 2023-03-09 Rmd/mseMac.Rmd
uppity

@@
 params=FLPar(c(m1=0.15/(0.2^-0.288),m2=-0.288,   bref=3.69e+06,delta=-0.3,     matk=30,      w50=0.17))
-params=FLPar(c(m1=9.69e-02         ,m2=-2.88e-01,bref=4.11e+06,delta=-2.00e-01,matk=2.80e+01,w50=1.66e-01)) 
+#params=FLPar(c(m1=9.69e-02         ,m2=-2.88e-01,bref=4.11e+06,delta=-2.00e-01,matk=2.80e+01,w50=1.66e-01)) 
 ```
@@
 load("~/Desktop/tmp/prjs..RData")
+load("/home/laurie/Dropbox/DDMSE/results/mac/mac_section5.RData")
 
@@
 eqCurves=ldply(oms., function(x) { 
-    model.frame(FLQuants(x[,dim(oms[[1]])[2]],
+    model.frame(FLQuants(x[,dim(oms.[[1]])[2]],
                          biomass=function(x) stock(x),
@@
                          f      =function(x) fbar( x)),drop=TRUE)[,-1]})
-ts=ldply(prj., function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
+ts=ldply(prj., function(x) model.frame(FLQuants(x,ssb  =function(x) ssb(  x), 
                                                   f    =function(x) fbar(x), 
@@
 ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
+
+fmsy=transform(ddply(eqCurves,.(.id), with, 
+                     data.frame(Fmsy=f[catch==max(catch)])),lower=Fmsy*0.8,upper=Fmsy*1.2)
+msy=ddply(merge(eqCurves,fmsy,by=".id"),.(.id), with, 
+      data.frame(lcatch=catch[min((lower-f)^2)==(lower-f)^2],
+                 msy   =catch[catch==max(catch)],
+                 ucatch=catch[min((upper-f)^2)==(upper-f)^2]))
+bmsy=ddply(merge(eqCurves,fmsy,by=".id"),.(.id), with, 
+      data.frame(lssb   =ssb[min((lower-f)^2)==(lower-f)^2],
+                 bmsy   =ssb[catch==max(catch)],
+                 ussb   =ssb[min((upper-f)^2)==(upper-f)^2]))
+
+ggplot(eqCurves)+
+  geom_line(aes(f,catch,col=.id))+scale_x_continuous(limits=c(0,2))+
+  theme(legend.position="bottom")
 
@@
 ```
+
+```{r}
+load("~/Desktop/tmp/prjs..RData")
+load("~/Desktop/tmp/oms..RData")
+load("/home/laurie/Dropbox/DDMSE/results/mac/mac_section5.RData")
+
+eqCurves=ldply(oms, function(x) { 
+    model.frame(FLQuants(x[,dim(oms[[1]])[2]],
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)[,-1]})
+
+fmsy=transform(ddply(eqCurves,.(.id), with, 
+                     data.frame(Fmsy=f[catch==max(catch)])),lower=Fmsy*0.8,upper=Fmsy*1.2)
+msy=ddply(merge(eqCurves,fmsy,by=".id"),.(.id), with, 
+      data.frame(lcatch=catch[min((lower-f)^2)==(lower-f)^2],
+                 msy   =catch[catch==max(catch)],
+                 ucatch=catch[min((upper-f)^2)==(upper-f)^2]))
+bmsy=ddply(merge(eqCurves,fmsy,by=".id"),.(.id), with, 
+      data.frame(lssb   =ssb[min((lower-f)^2)==(lower-f)^2],
+                 bmsy   =ssb[catch==max(catch)],
+                 ussb   =ssb[min((upper-f)^2)==(upper-f)^2]))
+
+sch=ldply(prjs, function(x) ldply(x, function(x) { 
+    model.frame(FLQuants(x,
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)}))
+sch=transform(sch,.id=factor(.id,labels=c("Base","DD Mass","DD Mass, Mat","DD Mass, Mat, M")))
+
+sch=subset(sch,year==2050)
+sch=merge(sch,bmsy,by=".id")
+sch=merge(sch,msy,by=".id") 
+```
+
+```{r, kobe-base, fig.height=8,fig.width=8}
+p1=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey")) 
+```
+
+```{r,kb-m, fig.height=8,fig.width=8}
+p2=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey")) 
+```
+
+```{r,kobe-mm, fig.height=8,fig.width=8}
+p3=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey")) 
+```
+
+```{r,kobe-mmm, fig.height=8,fig.width=8}
+p4=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+                     xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red")) 
+```
+
+```{r, den-ssb}
+ggdensity(transmute(sch,stock=ssb/bmsy,harvest=catch/msy,run=ac(signif(f,3)),.id=.id),x="stock",fill=".id")+
+  geom_vline(aes(xintercept=1),col="red")+
+  facet_grid(run~.) 
+```
+
+```{r, den-yield}
+ggdensity(transmute(sch,stock=ssb/bmsy,catch=catch/msy,run=ac(signif(f,3)),.id=.id),x="catch",fill=".id")+
+  geom_vline(aes(xintercept=1),col="red")+
+  facet_grid(run~.) 
+```

### 2023-03-09 Rmd/report.R

@@ -1714,11 +1714,66 @@ mystk     <- "mac";
   
   rm(p, p1, p2, p3, p4)
   save(list=ls(),
-       file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
+         file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   
   
+  eqCurves=ldply(oms, function(x) { 
+    model.frame(FLQuants(x[,dim(oms[[1]])[2]],
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)[,-1]})
   
+  fmsy=transform(ddply(eqCurves,.(.id), with, 
+                       data.frame(Fmsy=f[catch==max(catch)])),lower=Fmsy*0.8,upper=Fmsy*1.2)
+  msy=ddply(merge(eqCurves,fmsy,by=".id"),.(.id), with, 
+            data.frame(lcatch=catch[min((lower-f)^2)==(lower-f)^2],
+                       msy   =catch[catch==max(catch)],
+                       ucatch=catch[min((upper-f)^2)==(upper-f)^2]))
+  bmsy=ddply(merge(eqCurves,fmsy,by=".id"),.(.id), with, 
+             data.frame(lssb   =ssb[min((lower-f)^2)==(lower-f)^2],
+                        bmsy   =ssb[catch==max(catch)],
+                        ussb   =ssb[min((upper-f)^2)==(upper-f)^2]))
+  
+  sch=ldply(prjs, function(x) ldply(x, function(x) { 
+    model.frame(FLQuants(x,
+                         biomass=function(x) stock(x),
+                         ssb    =function(x) ssb(  x),
+                         catch  =function(x) catch(x),
+                         f      =function(x) fbar( x)),drop=TRUE)}))
+  sch=transform(sch,.id=factor(.id,labels=c("Base","DD Mass","DD Mass, Mat","DD Mass, Mat, M")))
+  
+  sch=subset(sch,year==2050)
[…]

### 2023-03-09 Rmd/report mac.Rmd
Added kobe plots to report.R and report mac.rmd

@@
   dropboxdir="~/Dropbox/DDMSE"
-```
 
-```{r}
 mystk     <- "mac"; mystkname <- "Northeast Atlantic mackerel"; mylatin <- "Scombrus scomber"
@@
 load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
-```
 
-```{r}
 # ------------------------------------------------------------------------------
@@
 Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**.
-**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
+**Figures 20 and 21** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
 
@@
 
-$W_a = {\overline{W_a}} {(B/B_{MSY})^{{\beta}_a}$
+$W_a = {\overline{W_a}} (B/B_{MSY})^{{\beta}_a}$
 
@@
 For example a valid combination is the Base case with the red curves (Fmsy=0.33) or the M scenario with the green curve.
-Overall the misspecification does not results in noticeable differences.
+Overall the misspecification does not results in noticeable differences in catch although they do have difference for biomass.
 
@@
 
-As a chech, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colours).
+\newpage
+
+For the four scenarios, we plotted the Kobe phase plot for 2050 with the associated density functions for B/Bmsy and Catch/MSY.
+The results confirm the results that equilibrium catch is not really affected by the DD assumptions of the Fmsy applied, but that biomass is sensitive to the assumptions.
+
+```{r, echo=FALSE, out.width = "300px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "kobe", level = 1, display = FALSE,
+  caption = "Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario). Topleft: base, topright: M, bottomleft: MM, bottomright: MMM")
+
+knitr::include_graphics(path=
+                    c(file.path(figuresdir,paste("05",mystk,"kobe base.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe M.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe MM.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe MMM.jpg",sep="_"))
+                      )
+                    )
+
+```
+
+*`r fig_nums("kobe")`*
+
+Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colours).
 The points fall on the equilibrium curve, thereby confirming the consistency between the two approaches.
@@
 
-**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, and **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield.
 

### 2023-03-09 Rmd/report.R

@@ -1717,6 +1717,7 @@ mystk     <- "mac";
          file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   
+  #@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@#@
   
   eqCurves=ldply(oms, function(x) { 
     model.frame(FLQuants(x[,dim(oms[[1]])[2]],
@@ -1748,22 +1749,39 @@ mystk     <- "mac";
   sch=merge(sch,bmsy,by=".id")
   sch=merge(sch,msy,by=".id") 
 
-  
-  p1=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
-                          xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey")) 
+  # Base
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe base.jpg", sep="_")),
+       width=10, height=6, units="in", res=300)
+  kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),
+                                        stock=ssb/bmsy,
+                                        harvest=catch/msy,
+                                        run=ac(f))),
+                        xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey")) 
+  dev.off()
 
-  
-  p2=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
-                          xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey")) 
+  # M
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe M.jpg", sep="_")),
+       width=10, height=6, units="in", res=300)
+  kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),
+                                        stock=ssb/bmsy,
+                                        harvest=catch/msy,
[…]

### 2023-03-09 Rmd/report mac.Rmd
spell checked

@@
 This was run as a check, as the equilibrium and projections should agree, I.e.
-the points at MSY should be the same colour as the curve.
+the points at MSY should be the same color as the curve.
 
@@
 
-Mass-at-age ($W_a$) is modelled as
+Mass-at-age ($W_a$) is modeled as
 
@@
 
-Maturity is then modelled as a logistic function of mass-at-age i.e.
+Maturity is then modeled as a logistic function of mass-at-age i.e.
 
@@
 To ensure the MSE projections and the historical estimates are comparable, a Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
-In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M-at-age.
+In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimic the VPA assessment with fixed M-at-age.
 The figure below compares the SAM assessment with the VPA and the VPAM assessment.
@@
 using an Fmsy from one scenario and applying it to a different scenario has only limited impacts on the overall outcomes in terms of recruitment, biomass or yield.
-In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colours.
+In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colors.
 For example a valid combination is the Base case with the red curves (Fmsy=0.33) or the M scenario with the green curve.
@@
 
-Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colours).
+Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colors).
 The points fall on the equilibrium curve, thereby confirming the consistency between the two approaches.
@@
 
-Estimating density-dependence from observations was only possible for mass-at-age and maturity-at-qge.
+Estimating density-dependence from observations was only possible for mass-at-age and maturity-at-age.
 Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
 
-In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which would mimick the VPA assessment with fixed M-at-age.
+In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which would mimic the VPA assessment with fixed M-at-age.
 The figure below compares the SAM assessment with the VPA and the VPAM assessment.
@@
 The third step was to convert the VPA model with fixed natural mortality at age and by year into an age-varying natural mortality vector.
-We applied a natural mortality-at-age vector derived from the Lorenzen approach (@lorenzen2000allometry) which would mimick the VPA assessment with fixed M-at-age.
+We applied a natural mortality-at-age vector derived from the Lorenzen approach (@lorenzen2000allometry) which would mimic the VPA assessment with fixed M-at-age.
 This assessment was then treated as the base case against which the density dependent operating models were compared.
@@
 While the recent expansion in distribution area and change in weight-at-age could be associated with changes in natural mortality.
-Therefore in the OMs we modelled M as a function of mass-at-age.
+Therefore in the OMs we modeled M as a function of mass-at-age.
 
@@
 Given the uncertainty in the assessment and the estimation of density dependence, it is likely that the Pretty Good Yield approach would ensure sustainable exploitation.
-Also, because we did not yhet consider stochastic variability in future developments in mass, maturity and natural mortality which are likely to obscure the impacts of density dependence.
+Also, because we did not yet consider stochastic variability in future developments in mass, maturity and natural mortality which are likely to obscure the impacts of density dependence.
 Variations in these processes due to changes in climate and range expansion are also likely to occur.
 
-It is suggested that emprirical harvest control rules that respond to changes in productivity could be evaluated in contrast to the highly complex harvest rules based on SAM stock assessment models.
+It is suggested that empirical harvest control rules that respond to changes in productivity could be evaluated in contrast to the highly complex harvest rules based on SAM stock assessment models.
 

### 2023-03-09 Rmd/report mac.Rmd
added refpoints

@@
   word_document:
-    reference_docx: ../report_template.dotx
+    reference_docx: ../report_template_v1.1.dotx
 mathjax: TRUE
@@
 
+**Reference points**
+
+Reference points estimated for the different scenarios are in the table below.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "refpoints", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Reference points for the different scenarios"))
+
+cat(readLines(file.path(tablesdir, paste(section,mystk, "refpoints.txt", sep="_"))), sep = '\n')
+
+```
+
+*`r tab_nums("refpoints")`*
+
 <!-- =============================================================================================================== -->

### 2023-03-10 Rmd/report whb.Rmd
Updates to report.R and report whb.Rmd

@@
   word_document:
-    reference_docx: ../report_template.dotx
+    reference_docx: ../report_template_v1.1.dotx
 mathjax: TRUE
+bibliography: refs.bib
 ---
@@
 source("../R/pm.R")
-```
-
-```{r}
 source("../R/get_dropbox.R")
 
-dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
+dropboxdir<-try(file.path(get_dropbox(), "DDMSE"))
 
 if ("try-error"%in%is(dropboxdir))
-  dropboxdir="~/Dropbox/pelagics"
+  dropboxdir="~/Dropbox/DDMSE"
 
@@
 <!-- =============================================================================================================== -->
-<!-- ## Mackerel report                                                                                              -->
+
+<!-- ## Blue whiting report                                                                                              -->
+
 <!-- =============================================================================================================== -->
@@
 
-**L. Kell, M.A. Pastoors**
+**`r ifelse(mystk=="mac", "M.A. Pastoors & L.T. Kell", "L.T. Kell & M.A. Pastoors")`**
 
-`r format(Sys.time(), '%d/%m/%Y')`
+Version: 1.0 (`r format(Sys.time(), '%d %B %Y')`)
 
-Commissioned by: University of Copenhagen (Henrik Sparholt)
+&nbsp;
+
+**Commissioned by:**
+
+Henrik Sparholt
+
+University of Copenhagen
+
+Globe Institute
+
+Center for Macroecology Evolution and Climate
+
+Universitetsparken 15, 3rd floor, room 339
+
+2100 Copenhagen Ø
+
+Denmark
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+```{r, echo=FALSE, out.width = "300px", fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
+
+if(mystk=="mac") {
+  knitr::include_graphics("../MPFF Seaplusplus.png")
+} else {
+  knitr::include_graphics("../Seaplusplus MPFF.png")
+}
+
+```
+
+\newpage
 
@@
 
-[summary here]
+The current management of fish stocks is based on achieving the fishing mortality ($F_{MSY}$) that achieves the maximum sustainable yield (MSY).
+Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms.
+Density dependence, however, may also operate in processes such as growth, sexual maturity, and natural mortality.
+A main aim of this study was to condition Operating Models (OMs) on alternative hypotheses about density dependence for use in Management Strategy Evaluation (MSE).
+The objective is to evaluate the robustness of the assessments conducted to support the ICES advice framework.
+To do this OMs were developed that include density dependence in growth, maturity, natural mortality, individually and combined.
+These OMs were then used to evaluate the consequences of not including density dependence management advice.
+
+The Operating Models were
+
+-   Base Case, with varying M-at-age
+-   Density dependence in mass-at-age,
+-   Density dependence in mass, maturity-at-age,
+-   Density dependence in mass, maturity and M-at-age,
+
+The MSY reference points $B_{MSY}$ and $F_{MSY}$ are generally based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship [@sissenwine1987alternative] or by conducting long-term projections under constant fishing moralities.
+
+Initial evaluations were conducted based on life history theory, as used by the International Council for the Exploration of the Sea (ICES) to develop advice conducting MSE for data poor stocks [@fischer2020ga].
+**Figure 10** shows the the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
+These allow the relationships between the MSY reference points and virgin biomass to be seen.
+
+Each of the OMs were then projected for the values of $F_{MSY}$ based on the four assumptions about density dependence.
+This allows a comparison between the yield that could potentially be achieved and that forgone by making an incorrect assumption about density dependence.
+The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to $F_{crash}$) for 30 years into the future.
+The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure 15**).
+The points on the curve correspond to the projected outcomes for the values of $F_{MSY}$ from each OM.
+This was run as a check, as the equilibrium and projections should agree, I.e.
+the points at MSY should be the same color as the curve.
+
+The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticty in the recruitment deviates.
+This allows the impact of misspecification of density dependence to be evaluated, i.e.
+if $F_{MSY}$ is underestimated then yield will be less than MSY, while if $F_{MSY}$ is overestimated yield will be lost due to overfishing.
+Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**.
+**Figures 20 and 21** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
+
+Conclusions
+
+Estimating density-dependence from observations only possible for mass-at-age and maturity age.
+Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
+
+The impact of including density dependence: major impact on reference points.
+Especially $F_{MSY}$ is much higher when including DD.
+However, the expected yields in equilibrium conditions are expected to be very similar.
+The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 1: introduction                                                                                         -->
+
 <!-- =============================================================================================================== -->
@@
 
+\newpage
+
 # Introduction
 
-The purpose of this work is to support the project “Fmsy-project for six high profile fish stocks”.
-The project focuses on the current management approach for fish stocks based on single-species Fmsy Values where the only density-dependent process that is included is recruitment.
-The project includes three other density-dependent population dynamics: growth (weight at age), sexual maturity and natural mortality.
+The purpose of this work is to support the project "$F_{MSY}$-project for six high profile fish stocks".
+The project focuses on the current management approach for fish stocks based on single-species $F_{MSY}$ Values where the only density-dependent process that is included is recruitment.
+The project includes three other density-dependent population dynamics: growth (mass-at-age), sexual maturity and natural mortality.
 
@@
 
-All MSE's have been developed within the FLR framework (https://flr-project.org/) and all the code is stored on github (https://github.com/martinpastoors/ddmse).
+All MSE's have been developed within the FLR framework (<https://flr-project.org/>) and all the code is stored on github (<https://github.com/martinpastoors/ddmse>).
 
-All data and results are available on DropBox (https://www.dropbox.com/sh/k87x074xqw7sgae/AAAmF980iR2n24wTkqhOhnpfa?dl=0)
+All data and results are available on a dedicated DropBox folder.
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 2: conditioning and summaries of ICES assessment                                                        -->
+
 <!-- =============================================================================================================== -->
@@
 # Conditioning and summaries based on the 2022 ICES assessment
+
+All conditioning of the operating models and estimation of potential density dependent processes have been based on the ICES 2022 SAM assessment of `r mystkname`.
 
@@
 
-**Weight at age**
+**Mass-at-age**
 
-[text here]
+Trends in (stock) mass-at-age are shown in figure 1.
+Weight-at-age in the stock is calculated as the average of the weight-at-age in the three spawning components, weighted by the relative size of each component (as estimated in the egg surveys).
+The decreasing trend in weight-at-age observed since 2005 for fish of age 3 and older seems to have stopped in 2013 and values in the last 8 years show an increasing trend.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_waa")`_
+*`r fig_nums("cond_waa")`*
 
-\newpage
+**Maturity-at-age**
 
-**Maturity at age**
+[@@ The fixed maturity ogive for blue whiting is derived from ...]
 
-[text here]
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_mat")`_
+*`r fig_nums("cond_mat")`*
 
-\newpage
+**Natural mortality-at-age**
 
-**Natural mortality at age (fixed)**
+Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (@hamre1980biology; @ICES2022WGWIDE; @ICES2022MAC).
 
-[text here]
+In order to assess the impacts of density dependent processes on natural mortality, additional assumptions need to be made about the distribution of natural mortality across ages.
+This will be done in section 3 of this report.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-**Density dependence in Weight at age vs. Total biomass**
+**Density dependence in mass-at-age vs. Total biomass**
 
-[text here]
+The distribution of mass-at-age is shown in the figure below (left) and the relationship between mass-at-age and total biomass is shown to the right.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+[@@ Consistent negative slopes on the biomass-mass-at-age plots can be observed for ages 4 and beyond.
+The patterns for the younger ages are less clear.
+It should be remembered that the younger ages do not appear strongly in the catches.]
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_ddwaa")`_
+*`r fig_nums("cond_ddwaa")`*
 
-\newpage
+**GAMM-at-age**
 
-**GAMM at age**
+A Generalized Additive Mixed Model (GAMM) was estimated to assess the relationship between mass-at-age and total biomass.
 
-[text here]
+weights are predicted from two variables: a smoother of total biomass/mean(total biomass) and age.
+The resulting factor is shown below, indicating a nearly linear decline in mass-at-age as total biomass increases.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_gammwaa")`_
+*`r fig_nums("cond_gammwaa")`*
 
-\newpage
+A summary of the GAMM model is shown in the table below.
 
-**GAMM summary**
-
-[text here]
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r tab_nums("cond_gammsummary")`_
+*`r tab_nums("cond_gammsummary")`*
 
-\newpage
+**Linear model of density dependence in mass-at-age**
 
-**GAMM summary**
+The change in relative mass-at-age (i.e.
+observed - mean values) is plotted against total biomass, this allows the relationship between growth and total biomass to be compared across all ages.
+It is seen that there is a common negative relationship for all ages.
+A summary of the linear model is in the table below.
+The trend of growth with biomass is -0.18, this was then used to model mass-at-age as:
 
-[text here]
+$W_a = {\overline{W_a}} (B/B_{MSY})^{{\beta}_a}$
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-tab_nums(
-  name    = "cond_gammcoefficients", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". GAMM coefficients of density dependence and weight at age."))
-
-cat(readLines(file.path(tablesdir, paste(section,mystk, "gamm coefficients.txt", sep="_"))), sep = '\n')
-
-```
-
-_`r tab_nums("cond_gammcoefficients")`_
-
-\newpage
-
-**LM of density dependence in weight at age**
-
-[text here]
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_lmwaa")`_
+*`r fig_nums("cond_lmwaa")`*
 
-\newpage
-
-**LM summary**
-
-[text here]
-
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r tab_nums("cond_gammcoefficients")`_
+*`r tab_nums("cond_lmsummary")`*
 
@@
 
-**DD analysis of maturity at age**
+**Density-dependence analysis of maturity-at-age**
 
-[text here]
+The distribution of maturity-at-age is shown in the figure below (left) and the relationship between maturity-at-age and total biomass is shown to the right.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+[@@Negative slopes on the biomass-maturity-at-age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.]
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_ddmat")`_
-
-\newpage
+*`r fig_nums("cond_ddmat")`*
 
@@
 
+A maturity at weight ogive was generated from this formula:
+
 $f(x) = \frac{1}{1+e^{-k(x-x_0)}}$
 
-Estimated parameters: k=`r matPar$par[1]` $W_{50}$=`r matPar$par[2]`
+This resulted in the parameters k=`r  format(matPar$par[1], digits=4)` and $W_{50}$=`r  format(matPar$par[2], digits=4)` which have been used in subsequent simulations.
 
-```{r, echo=FALSE, fig.width=10, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-_`r fig_nums("cond_matogive")` _
+*`r fig_nums("cond_matogive")`*
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 3: Estimating Density Dependent effects based on equilibrium analyses                                   -->
+
 <!-- =============================================================================================================== -->
 
-# Estimating Density Dependent effects based on equilibrium analyses
+# Estimating Density Dependent effects based on life-history parameters
 
 Life history parameters are used to develop an example of modelling density dependence for mass, maturity and natural mortality-at-age.
-The parameters are first used to construct An `FLBRP` object representing the equilibrium, and are then coerced into an `FLStock` to model the time series dynamics.
+The parameters are first used to construct an \*\`**FLBRP** object representing the equilibrium, and are then coerced into an **FLStock** to model the time series dynamics.
 
-## Mass-at-age [SHOULD WE CALL IT WEIGHT AT AGE, SINCE WE ARE USING Wa? ]
+## Mass-at-age
 
-Mass-at-age ($W_a$) is modelled as
+Mass-at-age ($W_a$) is modeled as
 
-$W_a = {\overline{W_a}} B^{{\beta}_a}$
+$W_a = {\overline{W_a}} (B/B_{MSY})^{{\beta}_a}$
 
 Where $\overline{W_a}$ and $\beta$ are estimated from a regression of mass-at-age ($W_a$) on total biomass ($B_a$), based on empirical data.
-$W_a$ is constrained so sizes do not become unfeasible large or small.
+$W_a$ is constrained so that sizes do not become unfeasible large or small.
 
@@
 
-Maturity is then  modelled as a logistic function of mass-at-age i.e.
+Maturity is then modeled as a logistic function of mass-at-age i.e.
 
@@
 
-Is assumed to vary at unit length ($L_1$) e.g.
+Natural mortality is assumed to vary at unit length ($L_1$).
 
-$ln(M_1) =\alpha + \beta ln(L_{\infty}) + \gamma ln(k)$
+Empirical and theoretical studies (@lorenzen2000allometry) have shown that natural mortality (M) decrease with fish length, but increase with the von Bertalanffy parameters of curvature $k$ and asymptotic length $L_{\infty}$.
+Therefore, M can be derived from life history parameters, e.g.
+$M=1.5kL{\infty}/L$) (@POPE2021105952) or M=1.5k @jensen1985comparison.
+@lorenzen2008fish @lorenzen2022natural showed that natural mortality rates vary with body size and age, often by orders of magnitude over the life cycle and support the existence of an allometric relationship between M and body mass (W) at-age of the form:
 
-where $\alpha=0.65$, $\beta=0.91$, $\gamma=0.87$
+$M=M_uW^d$
 
-Length-at-age ($L_a$) can be obtained from the length-weight relationship
+Where $M_u$ is $M$ at unit weight, and $d$ is the allometric exponent from empirical relationships shown to range from -0.3 to -0.37.
+$d$ is approximately equal to -1/3, which implies that $M$ is inversely proportional to body length, since mass is approximately proportional to the third power of length.
+This is consistent with [@gislason2008coexistence], who proposed a relationship for natural mortality based on length
 
-$L_a=(W/a)^{1/b}$
+$M={\alpha}k(L_{\infty}χ)L\beta$
 
-giving
-
-$ln(M_1) =(W/a)^{1/b}(\alpha + \beta ln(L_{\infty}) + \gamma ln(k))$
-
-For DD there is a relationship between Mass at age $M_a$ and total biomass ($B$)
-
-<!-- Mackerel: -->
-
-<!-- $M = 0.0232B + 0.0524$ -->
+The plot below shows the equilibrium values by age for simulations initiated with different levels of biomass relative to $B_{MSY}$.
+The red line indicates the values from the current assessment (for natural mortality that varies by mass and hence age).
 
@@
   name    = "sim_dd3", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in mass-at-age, maturity at age and Natural mortality at age for different levels of biomass relative to Bmsy")
+  caption = "Simulation of density dependence in mass-at-age, maturity-at-age and Natural mortality-at-age for different levels of biomass relative to $B_{MSY}$.
+Red line is without density dependence.")
 
@@
 
-_`r fig_nums("sim_dd3")`_
+*`r fig_nums("sim_dd3")`*
 
-Comparison of equilibrium yield plots for simulation with and without density-dependent processes
+From the theoretical life-history parameters and simulations, we can compare the equilibrium total biomass and yield from an assessment without density dependence, and then adding, step-by-step, the density dependent processes for mass-at-age, maturity-at-age and natural mortality-at-age.
+Comparison of equilibrium yield plots for simulation with and without density-dependent processes indicates that the expected values for $B_{MSY}$ tend to become lower when density dependence is included and that MSY becomes somewhat higher.
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "500px",  fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
   name    = "sim_yield", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to Bmsy")
+  caption = "Simulation of density dependence in M-at-age for different levels of biomass relative to $B_{MSY}$")
 
-knitr::include_graphics(
-  path=file.path(figuresdir, paste(section,mystk,"eq_yield.jpg",sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk,"eq_yield.jpg",sep="_")))
 
@@
 
-Stock assessments with SAM (non-DD), density-dependent SAM, VPA based on SAM (non-DD) and density-dependent VPA.
-The VPA reconstructions of the SAM assessment, shows a strongly divergent pattern in recruitment and also some differences in the trend of SSB.
-Mean fishing mortality is somewhat more variable in the VPA reconstruction compared to the SAM assessment.
-Density-dependent processes (on weight, maturity and M) tend to lead to lower SSB in the latter part of the time series.
+The operating models for this study are derived from the 2022 stock assessment.
+SAM is a statistical catch-at-age model that estimates error in the catch-at-age.
+In the case of  `r mystkname`, the assessment gives a higher weight to the recruitment index than to the catch data for ages 0 and 1.
+This means that the estimates of recruitment in the assessment can be very different from the actual number of fish that are recruited to the fishery.
+To ensure the MSE projections and the historical estimates are comparable, a Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
+In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimic the VPA assessment with fixed M-at-age.
+The figure below compares the SAM assessment with the VPA and the VPAM assessment.
+A main difference is in the recruitment estimates.
+Using age-varying M tends to flatten the recruitment pattern.
+The vpaM model was used as the base case to be compared against the density-dependent simulations.
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+```{r, echo=FALSE, out.width = "600px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+When applying density dependent processes to mass-at-age (i.e. lower mass-at-age for higher biomass), $B_{MSY}$ is expected to be somewhat lower. The expected increase in MSY is small.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
-**Density dependence in mass-at-age and maturity at age**
+**Density dependence in mass-at-age and maturity-at-age**
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+When applying density dependent processes to both mass-at-age and maturity-at-age, $B_{MSY}$ is expected to be lower again and MSY is expected to increase.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
   name    = "om_VPADDMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age and Maturity at age")
+  caption = "Simulation of density dependence in Mass-at-age and Maturity-at-age. Red lines are the non-density dependent values.")
 
@@
 
-**Density dependence in mass-at-age, maturity at age and M at age**
+**Density dependence in mass-at-age, maturity-at-age and M-at-age**
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+When applying density dependent processes to all three variables, $B_{MSY}$ is expected to be lower again and MSY is expected to increase a bit more.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
   name    = "om_VPADDMMM", level = 1, display = FALSE,
-  caption = "Simulation of density dependence in Mass-at-age, maturity at age and M at age")
+  caption = "Simulation of density dependence in Mass-at-age, maturity-at-age and M-at-age. Red lines are the non-density dependent values.")
 
@@
 
-```{r, echo=FALSE, fig.width=0.6, fig.asp=0.6, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 30%, $F_{MSY}$ is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
+So while, inclusion of density-dependent processes, has a substantial impact on biological reference points, the resulting change in equilibrium catch is expected to be small.
+
+```{r, echo=FALSE, out.width = "500px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
 
@@
 
+**Reference points**
+
+Reference points estimated for the different scenarios are in the table below.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+tab_nums(
+  name    = "refpoints", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Reference points for the different scenarios"))
+
+cat(readLines(file.path(tablesdir, paste(section,mystk, "refpoints.txt", sep="_"))), sep = '\n')
+
+```
+
+*`r tab_nums("refpoints")`*
+
 <!-- =============================================================================================================== -->
+
 <!-- Section 5: Running forward projections                                                                          -->
+
 <!-- =============================================================================================================== -->
@@
 
-[explanations here]
+Forward projections were run in two modes.
+Firstly in deterministic mode and secondly in probabilistic mode.
+We used the $F_{MSY}$ to project forward from 2020 until 2050.
+We used both the $F_{MSY}$ that is applicable to the density dependent scenario, but also evaluated what would happen if the 'wrong $F_{MSY}$' would be used with a particular scenario.
+
+The plot below shows the deterministic runs with the appropriate $F_{MSY}$ value being used to project forward.
+$F_{MSY}$ is highest when density dependence is taken into account for all three processes.
+Nevertheless, expected catches are somewhat lower when density dependence is taken into account.
 
@@
 
-**Deterministic projections at Fmsy for different scenarios: trends in weight, maturity and M at age**
+**Deterministic projections at** $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age
+
+How can we understand the results of the deterministic runs?
+Below is a plot of the trends in stock mass-at-age, maturity-at-age and natural mortality-at-age, for the most relevant ages (1-6) and for the four different scenarios (Base, DDM, DDMM and DDMMM).
+The density dependent effect of maturity is strongest when both stock weight and maturity are density dependent.
+When natural mortality is also density dependent, this affects the stock size and hence also the density dependence for stock weight and maturity, leading to a smaller density dependent effects for those variables.
 
@@
 
-**Stochastic projections at Fmsy for different scenarios**
+**Stochastic projections at** $F_{MSY}$ for different scenarios
+
+Using stochastic projections by including random processes for recruitment deviations and the $F_{MSY}$ values that are consistent with the different scenarios, leads to the overview below.
+SSB and total biomass are highest in the case of no density dependence.
+Catches are different in the first years of 'switching' to the new reference points, but afterwards, the catches for the different scenarios are highly comparable.
 
@@
 **Histograms of probability distributions in 2050 for different scenarios**
+
+The histograms of the probability distributions in 2050 for different scenarios, again demonstrate the expected larger stock size without taking density dependence into account, and only marginal differences in catch.
 
@@
 
-**Stochastic projections at all Fmsy's from all scenarios**
+**Stochastic projections at all $F_{MSY}$'s from all scenarios**
+
+Misspecification of Fmsy, i.e.
+using an Fmsy from one scenario and applying it to a different scenario has only limited impacts on the overall outcomes in terms of recruitment, biomass or yield.
+In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colors.
+For example a valid combination is the Base case with the red curves (Fmsy=0.33) or the M scenario with the green curve.
+Overall the misspecification does not results in noticeable differences in catch although they do have difference for biomass.
 
@@
 
-_`r fig_nums("mse_eq2050")`_
+*`r fig_nums("mse_sim2")`*
+
+\newpage
+
+For the four scenarios, we plotted the Kobe phase plot for 2050 with the associated density functions for B/Bmsy and Catch/MSY.
+The results confirm the results that equilibrium catch is not really affected by the DD assumptions of the Fmsy applied, but that biomass is sensitive to the assumptions.
+
+```{r, echo=FALSE, out.width = "300px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "kobe", level = 1, display = FALSE,
+  caption = "Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario). Topleft: base, topright: M, bottomleft: MM, bottomright: MMM")
+
+knitr::include_graphics(path=
+                    c(file.path(figuresdir,paste("05",mystk,"kobe base.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe M.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe MM.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe MMM.jpg",sep="_"))
+                      )
+                    )
+
+```
+
+*`r fig_nums("kobe")`*
+
+Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colors).
+The points fall on the equilibrium curve, thereby confirming the consistency between the two approaches.
+
+```{r, echo=FALSE, out.width = "500px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "eq_check", level = 1, display = FALSE,
+  caption = "Check of equilibrium curves against end points of simulated time series at $F_{MSY}$'s from all scenarios")
+
+  knitr::include_graphics(path=
+                            file.path(figuresdir,paste(section,mystk,"equilibrium_check.jpg",sep="_")))
+
+```
+
+*`r fig_nums("eq_check")`*
 
 <!-- =============================================================================================================== -->
+
 <!-- Section 6: Discussion and conclusions                                                                           -->
+
 <!-- =============================================================================================================== -->
@@
 
-The quality shall live up to that demanded by the International Council for the Exploration of the Sea (ICES) and potentially be used in ICES advice to fisheries managers.
-The Consultant shall be interacting with the leadership of the project Henrik Sparholt, or his representative, in the process of developing the software.
-The software shall be made publicly available via Github or on alternative platform if so directed by Henrik Sparholt or his representative.
+The current management of fish stocks is based on achieving a fishing mortality ($F_{MSY}$) that is expected to achieve the maximum sustainable yield (MSY).
+Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms.
+Density dependence, however, may also operate in processes such as growth, sexual maturity, and natural mortality.
+A main aim of this study was to condition Operating Models (OMs) on alternative hypotheses about density dependence for use in Management Strategy Evaluation (MSE).
+The objective is to evaluate the robustness of the assessments conducted to support the ICES advice framework.
+To do this, OMs were developed that include density dependence in growth, maturity, natural mortality, individually and combined.
+These OMs were then used to evaluate the consequences of including or not including density dependence in management advice.
+
+Initial explorations of density dependent effect were conducted based on life history theory, as used by the International Council for the Exploration of the Sea (ICES) to develop advice conducting MSE for data poor stocks [@fischer2020ga].
+**Figure 10** shows the the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
+These allow the relationships between the MSY reference points and virgin biomass to be seen.
+
+Estimating density-dependence from observations was only possible for mass-at-age and maturity-at-age.
+Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
+
+In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which would mimic the VPA assessment with fixed M-at-age.
+The figure below compares the SAM assessment with the VPA and the VPAM assessment.
+A main difference is in the recruitment estimates.
+Using age-varying M tends to flatten the recruitment pattern.
+The vpaM model was used as the base case to be compared against the density-dependent simulations.
+
+Next, the operating models were constructed. This was done in a number of steps.
+
+1.
+The ICES (SAM) assessment was taken from the ICES WGWIDE files (2022).
+SAM is a statistical catch-at-age model that estimates error in the catch-at-age.
+In the case of `r mystkname`, the assessment gives a higher weight to the recruitment index than to the catch data for ages 0 and 1.
+This means that the estimates of recruitment in the assessment can be very different from the actual number of fish that are recruited to the fishery.
+
+2.
+Next, a VPA model model was estimated.
+This was done to ensure the MSE projections and the historical estimates are comparable.
+The Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
+
+3.
+The third step was to convert the VPA model with fixed natural mortality at age and by year into an age-varying natural mortality vector.
+We applied a natural mortality-at-age vector derived from the Lorenzen approach (@lorenzen2000allometry) which would mimic the VPA assessment with fixed M-at-age.
+This assessment was then treated as the base case against which the density dependent operating models were compared.
+
+4.
+The density-dependent scenarios were constructed cumulative.
+First by adding density dependence in mass-at-age (M), then in mass and maturity-at-age (MM) and then in mass, maturity and M-at-age (MMM),
+
+Therefore, the four Operating Models were
+
+-   VPA Base Case, with varying M-at-age
+-   Density dependence in mass-at-age,
+-   Density dependence in mass, maturity-at-age,
+-   Density dependence in mass, maturity and M-at-age,
+
+The MSY reference points $B_{MSY}$ and $F_{MSY}$ were calculated based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship [@sissenwine1987alternative] or by conducting long-term projections under constant fishing moralities.
+
+Each of the four OMs were then projected for the values of $F_{MSY}$ based on the four assumptions about density dependence.
+This allows a comparison between the yield that could potentially be achieved and that forgone by making an incorrect assumption about density dependence.
+
+The OMs were first projected, without process error in recruitment, from the end of the historical estimates for a range of fishing mortality (spanning 0 to $F_{crash}$) for 30 years into the future.
+The estimates in the final year of the projections were then be used to construct equilibrium curves (**Figure 15**).
+The points on the curve correspond to the projected outcomes for the values of $F_{MSY}$ from each OM.
+This was run as a check, as the equilibrium and projections should agree, I.e.
+the points at MSY should be the same color as the curve.
+
+The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticty in the recruitment deviates.
+This allows the impact of misspecification of density dependence to be evaluated, i.e.
+if $F_{MSY}$ is underestimated then yield will be less than MSY, while if $F_{MSY}$ is overestimated yield will be lost due to overfishing.
+Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**
+
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, and **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield.
+
+Estimating density-dependence from observations was only possible for mass-at-age and maturity age.
+Therefore, density dependence in natural mortality can only be explored from theoretical considerations, because current assessment does not include age or time varying natural mortality.
+M is likely to vary by age and time in relation to processes such as density dependence.
+While the recent expansion in distribution area and change in weight-at-age could be associated with changes in natural mortality.
+Therefore in the OMs we modeled M as a function of mass-at-age.
+
+The impact of including density dependence was primarily on the biological reference points.
+Especially $F_{MSY}$ is much higher when including density dependence.
+However, the expected yields in equilibrium conditions are very similar.
+The main difference is in the initial step from the current $F_{MSY}$ to any new $F_{MSY}$ that would include DD, as the higher $F_{MSY}$ would be applied to a stock that is estimated without DD.
+
+Since the early 1980s, natural mortality ($M$) for `r mystkname` has been fixed within the assessment model at 0.2, for all ages and all assessment years.
+
+[@@This value was calculated based on estimates of total mortality derived from tagging data combined with catch data.]
+
+M must, however, vary both by age and over time due to changes in abundance of predators, feeding conditions, disease etc.
+[@siegfried2009review].
+Particularly any changes in `r mystkname` distribution and growth will affect natural mortality.
+
+Given the uncertainty in the assessment and the estimation of density dependence, it is likely that the Pretty Good Yield approach would ensure sustainable exploitation.
+Also, because we did not yet consider stochastic variability in future developments in mass, maturity and natural mortality which are likely to obscure the impacts of density dependence.
+Variations in these processes due to changes in climate and range expansion are also likely to occur.
+
+It is suggested that empirical harvest control rules that respond to changes in productivity could be evaluated in contrast to the highly complex harvest rules based on SAM stock assessment models.
+
+The quality of this analyses lives up to the quality standards that are required by ICES and so can potentially be used when providing advice to fisheries managers.
+The software is publicly available via Github (<https://github.com/martinpastoors/ddmse>).
+All data is derived from WGWIDE 2022 @ICES2022WGWIDE.
 
 <!-- =============================================================================================================== -->
+
 <!-- Funding                                                                                                         -->
+
 <!-- =============================================================================================================== -->
@@
 
+This work was funded under a consultancy agreement between the University of Copenhagen and `r ifelse(mystk=="mac", "Martin Pastoors F&F", "Sea++")` as part of the project “117957 Fmsy-project for six high profile fish stocks”.
+The current work focused on `r mystkname` and was carried out in the months of January-March 2023.
+A similar project was carried out for `r ifelse(mystk=="mac", "Blue whiting", mystkname)` by `r ifelse(mystk=="mac", "Sea++", "Martin Pastoors F&F")`.
+Both projects were carried in close coordination between Sea++ and Martin Pastoors F&F.
+
 <!-- =============================================================================================================== -->
+
 <!-- References                                                                                                      -->
+
 <!-- =============================================================================================================== -->
@@
 # References
-
-<!-- =============================================================================================================== -->
-<!-- Annex 1: OM conditioning                                                                           -->
-<!-- =============================================================================================================== -->
-
-# Annex 1: OM conditioning
-
-<!-- =============================================================================================================== -->
-<!-- Annex 2: Density dependent processes                                                                           -->
-<!-- =============================================================================================================== -->
-
-# Annex 2: Density dependent processes

### 2023-03-10 Rmd/report.R

@@ -76,8 +76,8 @@ myparams <- data.frame(
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-mystk     <- "mac";  
-# mystk     <- "whb";  
+# mystk     <- "mac";  
+mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   
   mystkname  <- myparams[myparams$stock==mystk,"mystkname"]
@@ -1419,16 +1419,8 @@ mystk     <- "mac";
               aes(ssb,catch,col=.id))+
     geom_point(data=subset(ts,year==2050),
                aes(ssb,catch,col=.id))
-    # geom_line(data=rfs,
-    #           aes(ssb,catch,col=.id))+
-    # geom_point(data=subset(prj_df,year==2050),
-    #            aes(ssb,catch,col=.id))
 
   ggplot()+
-    # geom_line(data=eqCurves,
-    #           aes(ssb,catch,col=.id))+
-    # geom_point(data=subset(ts,year==2050),
-    #            aes(ssb,catch,col=.id))
     geom_line(data=subset(ts,year==2050),
               aes(ssb,catch,col=.id))+
     geom_point(data=subset(prj_df,year==2050),
@@ -1505,30 +1497,6 @@ mystk     <- "mac";
     mutate(Fmsy=factor(F, levels=rfpts$Fmsy,labels=c("Base","M","MM","MMM"))) %>% 
     as_tibble()
 
-  # prjs_df <- tibble()
-  # for (i in 1:length(prjs)) {
-  #   prjs_df <-
-  #     bind_rows(
-  #       prjs_df,
-  #       ldply(prjs[[i]], function(x) {
-  #         model.frame(FLQuants(x,
-  #                              rec    =function(x) rec(x),
-  #                              biomass=function(x) stock(x),
-  #                              ssb    =function(x) ssb(  x),
-  #                              catch  =function(x) catch(x),
-  #                              f      =function(x) fbar( x)),drop=TRUE)})    
-  #     )
-  # }
-  # 
-  # nyears = length(unique(prjs_df$year))
-  # niters = length(unique(prjs_df$iter))
-  # nscen  = length(unique(prjs_df$.id))
-  # 
-  # prjs_df <- prjs_df %>% 
[…]

### 2023-03-10 Rmd/report whb.Rmd
\@\@ addresed

@@
 
-&nbsp;
-
 **Commissioned by:**
@@
 Denmark
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
-
-&nbsp;
 
@@
 
-[@@ The fixed maturity ogive for blue whiting is derived from ...]
+The fixed maturity ogive for blue whiting was taken from the ICES assessment, this did not vary by year.
 
@@
 
-The distribution of mass-at-age is shown in the figure below (left) and the relationship between mass-at-age and total biomass is shown to the right.
-
-[@@ Consistent negative slopes on the biomass-mass-at-age plots can be observed for ages 4 and beyond.
-The patterns for the younger ages are less clear.
-It should be remembered that the younger ages do not appear strongly in the catches.]
+The variation in mass-at-age is shown in the figure below (left) and the relationship between mass-at-age and total biomass  to the right.
+Consistent negative slopes on the biomass-mass-at-age plots are seen for ages 4 and beyond.
+The patterns for the younger ages, however, are less clear.
+The younger ages do not appear strongly in the catches, and so the catches may not be representative of the population.
 
@@
 
-[@@Negative slopes on the biomass-maturity-at-age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.]
+[\@ @Negative slopes on the biomass-maturity-at-age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.]
 
@@
 
-[@@This value was calculated based on estimates of total mortality derived from tagging data combined with catch data.]
+[\@ @This value was calculated based on estimates of total mortality derived from tagging data combined with catch data.]
 
@@
 
-This work was funded under a consultancy agreement between the University of Copenhagen and `r ifelse(mystk=="mac", "Martin Pastoors F&F", "Sea++")` as part of the project “117957 Fmsy-project for six high profile fish stocks”.
+This work was funded under a consultancy agreement between the University of Copenhagen and `r ifelse(mystk=="mac", "Martin Pastoors F&F", "Sea++")` as part of the project "117957 Fmsy-project for six high profile fish stocks".
 The current work focused on `r mystkname` and was carried out in the months of January-March 2023.

### 2023-03-10 Rmd/report whb.Rmd
WHB updates + refs

@@
 
-Trends in (stock) mass-at-age are shown in figure 1.
-Weight-at-age in the stock is calculated as the average of the weight-at-age in the three spawning components, weighted by the relative size of each component (as estimated in the egg surveys).
-The decreasing trend in weight-at-age observed since 2005 for fish of age 3 and older seems to have stopped in 2013 and values in the last 8 years show an increasing trend.
+Trends in (catch and stock) mass-at-age are shown in figure 1.
+Mean weight at ages 3-9 has generally decreased in the period 2010-2018, followed by an increase in the most recent years, for the most abundant ages in the catches.
+In 2022, a decrease in mean weight in almost all ages was observed.
 
@@
 The fixed maturity ogive for blue whiting was taken from the ICES assessment, this did not vary by year.
+Maturity-at-age was obtained by combining maturity ogives from the southern and northern areas, weighted by catch in numbers-at-age [@ICES1995BWWG, @ICES2022WHB].
 
@@
 
-Natural mortality is assumed to be 0.15 for all age groups and constant over time.This value was calculated based on estimates of total mortality derived from tagging data combined with catch data (@hamre1980biology; @ICES2022WGWIDE; @ICES2022MAC).
+Natural mortality is assumed to be 0.2 for all age groups and constant over time.
+The current M of 0.2 was derived from investigations undertaken in the 1980s that examined the age distribution of the stock before the industrial fishery started [@ICES1995BWWG].
+
+During the WKPELA benchmark [@ICES2012WKPELA] various methods were explored to estimate how M may vary over ages.
+The relationship between natural mortality and body weight was applied to the blue whiting data to determine a variable M by age.
+The values ranged from around 1.1 at age 0 to 0.7 at age 10, which was considerably higher than the values used so far.
+During the benchmark, it was considered that there was no new information to support a revision of the used estimate of M (=0.2).
 
@@
 The variation in mass-at-age is shown in the figure below (left) and the relationship between mass-at-age and total biomass  to the right.
-Consistent negative slopes on the biomass-mass-at-age plots are seen for ages 4 and beyond.
-The patterns for the younger ages, however, are less clear.
-The younger ages do not appear strongly in the catches, and so the catches may not be representative of the population.
+Consistent negative slopes on the biomass-mass-at-age plots are seen for ages 2-9.
 
@@
 weights are predicted from two variables: a smoother of total biomass/mean(total biomass) and age.
-The resulting factor is shown below, indicating a nearly linear decline in mass-at-age as total biomass increases.
+The resulting factor is shown below.
+There is no clear relationship between biomass and weight-at-age.
 
@@
 observed - mean values) is plotted against total biomass, this allows the relationship between growth and total biomass to be compared across all ages.
-It is seen that there is a common negative relationship for all ages.
+It is seen that there is no clear relationship between biomass and weight at age.
 A summary of the linear model is in the table below.
-The trend of growth with biomass is -0.18, this was then used to model mass-at-age as:
+
+**???The trend of growth with biomass is -0.18, this was then used to model mass-at-age as:**
 
@@
 The distribution of maturity-at-age is shown in the figure below (left) and the relationship between maturity-at-age and total biomass is shown to the right.
-
-[\@ @Negative slopes on the biomass-maturity-at-age plots can be observed for ages 2 and 3, which are the most relevant ages for maturity as all fish of age 4 and above are mature.]
+Obviously, as there is no time-dependent estimate of maturity, there is also no relationship with biomass.
 

### 2023-03-10 Rmd/report.R
section 5 fixed

@@ -1227,10 +1227,10 @@ mystk     <- "whb";
   # 5. MSE
   # ------------------------------------------------------------------------------
   
+  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
   section <- "05"
   
-  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
-  
+
   om=fwdWindow(window(vpaM,end=2020),end=2050,vpaM_eq)
   F    =rep(c(seq(0,                                    c(refpts(vpaM_eq)["msy",  "harvest"]),  length.out=51),
               seq(c(refpts(vpaM_eq)["msy",  "harvest"]),c(refpts(vpaM_eq)["crash","harvest"])*1.5,length.out=51)[-1]))
@@ -1758,5 +1758,3 @@ mystk     <- "whb";
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
-  section <- "05"
-  

### 2023-03-10 Rmd/report whb.Rmd
read thru

@@
 output:
-  word_document:
+  html_document:
     reference_docx: ../report_template_v1.1.dotx
@@
 
-All conditioning of the operating models and estimation of potential density dependent processes have been based on the ICES 2022 SAM assessment of `r mystkname`.
+All conditioning of the Operating Models and estimation of potential density dependent processes have been based on the ICES 2022 SAM assessment of `r mystkname`.
 
@@
 
-The operating models for this study are derived from the 2022 stock assessment.
+The Operating Models for this study are derived from the 2022 stock assessment.
 SAM is a statistical catch-at-age model that estimates error in the catch-at-age.
@@
 
-Estimating density-dependence from observations was only possible for mass-at-age and maturity-at-age.
-Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
+Estimating density-dependence from observations was only possible for mass-at-age but not maturity-at-age as the latter were assumed by ICES not to have varied over time.
+However, assuming that maturity is a function of mass-ata-ge allowed density deoendence to be modeeled.
+Natural mortality was assumed to be time and age invariant by ICES (i.e.
+0.2), therefore density dependence in natural mortality can only be explored from theoretical considerations.
 
-In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which would mimic the VPA assessment with fixed M-at-age.
-The figure below compares the SAM assessment with the VPA and the VPAM assessment.
-A main difference is in the recruitment estimates.
-Using age-varying M tends to flatten the recruitment pattern.
+In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which mimics the VPA assessment with fixed M-at-age.
+The figure below compares the SAM assessment with a Virtual Population Analysis (VPA) and a VPA with age varying M (VPAM).
+Trends and absolute levels swere similar across all runs.
 The vpaM model was used as the base case to be compared against the density-dependent simulations.
 
-Next, the operating models were constructed. This was done in a number of steps.
+Next, the Operating Models were constructed. This was done in a number of steps.
 
@@
 2.
-Next, a VPA model model was estimated.
-This was done to ensure the MSE projections and the historical estimates are comparable.
-The Virtual Population Analysis (VPA) was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
+Next, VPA was conducted.
+This was done to ensure the MSE projections, which assumed variable M-at-age, and the historical estimates are comparable.
+The VPA was conducted on the whole catch matrix and the survivors in the most recent year and oldest age.
 
@@
 We applied a natural mortality-at-age vector derived from the Lorenzen approach (@lorenzen2000allometry) which would mimic the VPA assessment with fixed M-at-age.
-This assessment was then treated as the base case against which the density dependent operating models were compared.
+This assessment was then treated as the base case against which the density dependent Operating Models were compared.
 
@@
 
--   VPA Base Case, with varying M-at-age
+-   Base Case, VPA with varying M-at-age
 -   Density dependence in mass-at-age,
@@
 The MSY reference points $B_{MSY}$ and $F_{MSY}$ were calculated based on equilibrium assumptions, either from combining yield per recruit and spawner per recruit relationship [@sissenwine1987alternative] or by conducting long-term projections under constant fishing moralities.
+Both should agree and so can be used as a check.
 
@@
 
-The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticty in the recruitment deviates.
+The OMs were also projected for the four estimates of  $F_{MSY}$, with stochasticty in the recruitment deviates.
 This allows the impact of misspecification of density dependence to be evaluated, i.e.
@@
 
-**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, and **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, and **Figure 21** shows phase plots of the final 2050 distributions of biomass and yield.
+THese are for the Base Case, then density-dependence in mass, mass and maturity, and mass, maturity and natural mortality, projected for the four estimates of $F_{MSY}$.
+The correctly specified $F_{MSY}$ for each model is indicated by the red points and marginal densities.
 
-Estimating density-dependence from observations was only possible for mass-at-age and maturity age.
-Therefore, density dependence in natural mortality can only be explored from theoretical considerations, because current assessment does not include age or time varying natural mortality.
-M is likely to vary by age and time in relation to processes such as density dependence.
-While the recent expansion in distribution area and change in weight-at-age could be associated with changes in natural mortality.
-Therefore in the OMs we modeled M as a function of mass-at-age.
+The main diofferences are seen in $SSB/B_{MSY}$, if no density dependence is assumed then in the operating model the estimate of $F_{MSY}$ is the lowest, therefore projreting for the higher estimates of $F_{MSY}$ assuming density dependence results in an overfished stock.
+Yileld are the same for the two values of $F_{MSY}$ that assume no density-dependence on density dependence in mass-at-age only.
+In the other two cases yield is reduced but only by about 10%.
+
+In the cases where density-dependence is assumed in the Operating Model, yields do not vary much, the main effect is seen in the estimates of $SSB/B_{MSY}$,
 
@@
 Since the early 1980s, natural mortality ($M$) for `r mystkname` has been fixed within the assessment model at 0.2, for all ages and all assessment years.
-
-[\@ @This value was calculated based on estimates of total mortality derived from tagging data combined with catch data.]
 

### 2023-03-12 Rmd/report.R
with kobePhase4

@@ -43,6 +43,7 @@ theme_set(theme_bw(16))
 source("R/dd.R")
 source("R/get_dropbox.r")
 source("R/pm.R")
+source("R/kobePhaseMar4.R")
 
 dropboxdir<-try(file.path(get_dropbox(), "DDMSE"))
 
@@ -1707,7 +1708,8 @@ mystk     <- "whb";
   sch=subset(sch,year==2050)
   sch=merge(sch,bmsy,by=".id")
   sch=merge(sch,msy,by=".id") 
-
+  sch=merge(sch,fmsy,by=".id") 
+  
   # Base
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe base.jpg", sep="_")),
        width=10, height=10, units="in", res=300)
@@ -1742,6 +1744,50 @@ mystk     <- "whb";
        width=10, height=10, units="in", res=300)
   kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                        xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red")) 
+  dev.off()
+  
+  ## Yield & F #################################################################
+  kobePhase=kobe:::kobePhase
+
+  # Base
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe FY base.jpg", sep="_")),
+       width=10, height=10, units="in", res=300)
+       kobePhaseMar4(subset(transmute(subset(sch,.id=="Base"),
+                                      stock  =f/Fmsy,
+                                      harvest=catch/msy,
+                                      run    =ac(f))),
+                quadcol=c("yellow","yellow","green","red"),
+                xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey"),
+                xlim=3) 
+  dev.off()
+  
+  # M
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe FY M.jpg", sep="_")),
+       width=10, height=10, units="in", res=300)
+       kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass, Mat"),
+                                      stock  =f/Fmsy,
+                                      harvest=catch/msy,
+                                      run    =ac(f))),
+                quadcol=c("yellow","yellow","green","red"),
[…]

### 2023-03-12 R/kobePhaseMar4.R
new

@@ -1,0 +1,86 @@
+
+
+kobePhaseMar4=function(pts,trks=NULL,mns=FALSE,size=1,
+                       xlab=expression(B:B[MSY]),
+                       ylab=expression(F:F[MSY]),
+                       xlim=2,ylim=xlim,
+                       quadcol=c("red","green","yellow","yellow"),
+                       col =colorRampPalette(c("orange","blue"),space="Lab"),
+                       shade=.5,col2=grey(shade),col3=grey(shade*1.1),
+                       layer=NULL,
+                       bref=1,
+                       fref=1){
+  
+  if (!("run" %in% names(pts)))
+    pts=cbind(pts,run=factor(1))
+  if (!is.null(trks) & !("run" %in% names(trks)))
+    trks=cbind(trks,run=factor(1))
+  
+  if ("function" %in% is(col))
+    col=col(length(unique(pts$run)))
+  
+  if (length(size)==1) size=rep(size,2)
+  
+  ##### Density plots   #############################################################################################
+  
+  # second density plot, oriented vertically (hence the 'coord_flip()' at the end
+  dH<-ggplot(pts) + 
+    #geom_density(aes(x = harvest, y =  ..count.., group=run), fill=col2, col=col3, position = "stack") + 
+    geom_density(aes(x = harvest, y = ..count..,               fill=run, alpha=0.4)) + 
+    geom_vline(xintercept=fref,col="red",data=data.frame(fref=fref))+
+    coord_cartesian(xlim=c(0,ylim))   +
+    scale_fill_manual(values=col)          +
+    xlab(xlab) + ylab(ylab)                +
+    theme(legend.position = "none", 
+          axis.title.x = element_text(colour ='NA'), 
+          axis.text.x  = element_text(colour ="NA"), 
+          axis.ticks.x = element_line(colour ="NA"),
+          axis.ticks =   element_line(colour ="NA"),
+          
+          axis.title.y = element_blank(), 
+          axis.text.y  = element_blank(), 
+          axis.ticks.y = element_blank(), 
+          
+          plot.margin = unit(c(0, 0, 1, 0), "lines"),
+          panel.background = element_rect(fill   ="NA", colour ="NA"), 
+          panel.border     = element_rect(fill   ="NA", colour ="NA"), 
[…]

### 2023-03-12 Rmd/report.R
rerun of mackerel

@@ -76,8 +76,8 @@ myparams <- data.frame(
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-# mystk     <- "mac";  
-mystk     <- "whb";  
+mystk     <- "mac";  
+# mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   
   mystkname  <- myparams[myparams$stock==mystk,"mystkname"]
@@ -660,7 +660,8 @@ mystk     <- "whb";
   rm(dat, df, p, p1, p2, p3, t, x)
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section3.RData", sep="_")))
-
+  # load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section3.RData", sep="_")))
+  
 
 
   
@@ -1209,6 +1210,7 @@ mystk     <- "whb";
   rm(dat, df, p, p1, p2, p3, p4, t, x, d1, d2, d3, d4, x)
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
+#  load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section4.RData", sep="_")))
   
   
   

### 2023-03-13 Rmd/report mac.Rmd
Included Kobe F-Yield

@@
 
-Version: 1.0 (07 March 2023)
+Version 1.0 (07 March 2023)
 
-Version: 1.1 (`r format(Sys.time(), '%d %B %Y')`)
+Version 1.1 (08 March 2023)
+
+Version 1.2 (10 March 2023)
+
+Version 1.3 (`r format(Sys.time(), '%d %B %Y')`)
 
@@
 
+We then plotted the Kobe phase plot for 2050 with the associated density functions for F/Fmsy and Catch/MSY.
+Because there is no uncertainty included in the fishing mortality during the simulations, the marginal distributions of F do not exist in this plot.
+
+```{r, echo=FALSE, out.width = "300px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "kobe2", level = 1, display = FALSE,
+  caption = "F Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario). Topleft: base, topright: M, bottomleft: MM, bottomright: MMM")
+
+knitr::include_graphics(path=
+                    c(file.path(figuresdir,paste("05",mystk,"kobe FY base.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe FY M.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe FY MM.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe FY MMM.jpg",sep="_"))
+                      )
+                    )
+
+```
+
+*`r fig_nums("kobe2")`*
+
 Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colors).
@@
 
-**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, and **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield and **Figure 22** the Kobe pltos for F and yield.
 

### 2023-03-13 Rmd/report.R

@@ -37,6 +37,7 @@ library(mgcv)
 library(mgcViz)
 
 library(tidyverse)
+library(grid)
 
 theme_set(theme_bw(16))
 
@@ -1749,23 +1750,36 @@ mystk     <- "mac";
   dev.off()
   
   ## Yield & F #################################################################
-  kobePhase=kobe:::kobePhase
-
+  
   # Base
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe FY base.jpg", sep="_")),
-       width=10, height=10, units="in", res=300)
-       kobePhaseMar4(subset(transmute(subset(sch,.id=="Base"),
-                                      stock  =f/Fmsy,
-                                      harvest=catch/msy,
-                                      run    =ac(f))),
+       width=10, height=8, units="in", res=300)
+  kobePhaseMar4(subset(transmute(subset(sch,.id=="Base"),
+                                 stock  =f/Fmsy,
+                                 harvest=catch/msy,
+                                 run    =ac(f))),
                 quadcol=c("yellow","yellow","green","red"),
-                xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey"),
+                xlab=expression(F/F[MSY]),
+                ylab=expression(Catch/MSY),
+                col=c("red","grey","grey","grey"),
+                xlim=3) 
+  dev.off()
+
+  # M
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe FY M.jpg", sep="_")),
+       width=10, height=8, units="in", res=300)
+  kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass"),
+                                 stock  =f/Fmsy,
+                                 harvest=catch/msy,
+                                 run    =ac(f))),
+                quadcol=c("yellow","yellow","green","red"),
+                xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey"),
                 xlim=3) 
   dev.off()
   
-  # M
-  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe FY M.jpg", sep="_")),
-       width=10, height=10, units="in", res=300)
+  # MM
+  jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe FY MM.jpg", sep="_")),
[…]

### 2023-03-13 Rmd/report.R
formatting kobe plots

@@ -39,7 +39,6 @@ library(mgcv)
 library(mgcViz)
 
 library(tidyverse)
-library(grid)
 
 theme_set(theme_bw(16))
 
@@ -1764,7 +1763,7 @@ mystk     <- "mac";
                 xlab=expression(F/F[MSY]),
                 ylab=expression(Catch/MSY),
                 col=c("red","grey","grey","grey"),
-                xlim=3) 
+                xlim=3, ylim=2) 
   dev.off()
 
   # M
@@ -1776,7 +1775,7 @@ mystk     <- "mac";
                                  run    =ac(f))),
                 quadcol=c("yellow","yellow","green","red"),
                 xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey"),
-                xlim=3) 
+                xlim=3, ylim=2) 
   dev.off()
   
   # MM
@@ -1788,7 +1787,7 @@ mystk     <- "mac";
                                       run    =ac(f))),
                 quadcol=c("yellow","yellow","green","red"),
                 xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey"),
-                xlim=3) 
+                xlim=3, ylim=2) 
   dev.off()
   
   # MMM
@@ -1800,7 +1799,7 @@ mystk     <- "mac";
                                  run    =ac(f))),
                 quadcol=c("yellow","yellow","green","red"),
                 xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red"),
-                xlim=3) 
+                xlim=3, ylim=2) 
   dev.off()
   
   
@@ -1818,3 +1817,6 @@ mystk     <- "mac";
   save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
   load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
+  section <- "05"
+  
+  

### 2023-03-13 Rmd/backUp/condWhb.Rmd
moving files to backup

--- 2023-02-09 Rmd/condWhb.Rmd
@@
 
-source("../R/get_dropbox.R")
+source("../R/get_dropbox.r")
 dropboxdir<-try(file.path(get_dropbox(), "pelagics"))
@@
 
-fn<-function(x,dat) 
-   sum((dat$mat-matFn(dat$wt,x))^2)
+matFn<-function(wt,x) 1/(1+exp(-x[1]*(wt-x[2])))
+
+fn<-function(x,dat) {
+  sum((dat$mat-matFn(dat$wt,x))^2)}
 

### 2023-03-13 Rmd/backUp/mseMac.Rmd
moving files to backup

--- 2023-03-09 Rmd/mseMac.Rmd
@@
 sch=merge(sch,msy,by=".id") 
+sch=merge(sch,fmsy,by=".id") 
 ```
@@
 ```{r, kobe-base, fig.height=8,fig.width=8}
-p1=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="Base"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                      xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey")) 
@@
 ```{r,kb-m, fig.height=8,fig.width=8}
-p2=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                      xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey")) 
@@
 ```{r,kobe-mm, fig.height=8,fig.width=8}
-p3=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                      xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey")) 
@@
 ```{r,kobe-mmm, fig.height=8,fig.width=8}
-p4=kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
+kobe:::kobePhaseMar2(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=ssb/bmsy,harvest=catch/msy,run=ac(f))),
                      xlab=expression(B/B[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red")) 
+```
+
+```{r}
+source("R/kobePhaseMar4.R", echo=TRUE)
+
+```
+
+```{r, kobe-base-fy, fig.height=8,fig.width=8}
+kobePhaseMar4(subset(transmute(subset(sch,.id=="Base"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
+                     quadcol=c("yellow","yellow","green","red"),
+                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("red","grey","grey","grey"),
+                     xlim=3) 
+```
+
+```{r,kb-m-fy, fig.height=8,fig.width=8}
+kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
+                     quadcol=c("yellow","yellow","green","red"),
+                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","red","grey","grey"),
+                     xlim=3) 
+```
+
+```{r,kobe-mm-fy, fig.height=8,fig.width=8}
+kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass, Mat"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
+                     quadcol=c("yellow","yellow","green","red"),
+                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","red","grey"),
+                     xlim=3) 
+```
+
+```{r,kobe-mmm-fy, fig.height=8,fig.width=8}
+kobePhaseMar4(subset(transmute(subset(sch,.id=="DD Mass, Mat, M"),stock=f/Fmsy,harvest=catch/msy,run=ac(f))),
+                     quadcol=c("yellow","yellow","green","red"),
+                     xlab=expression(F/F[MSY]),ylab=expression(Catch/MSY),col=c("grey","grey","grey","red"),
+                     xlim=3) 
 ```

### 2023-03-13 Rmd/backUp/mseWhb.Rmd
moving files to backup

@@
 ```{r, omCheck}
-om=fwdWindow(window(vpa,end=2020),end=2080,eqM)
-f =FLQuant(rep(seq(0,c(refpts(eqM)["crash","harvest"])*1.5,length.out=100),each=61),
+om=fwdWindow(window(om,end=2020),end=2080,eqM)
+f =FLQuant(rep(seq(0,c(refpts(eq)["crash","harvest"])*1.25,length.out=100),each=61),
            dimnames=dimnames(fbar(om)[,ac(2020:2080)]))
@@
   geom_line(aes(SSB,Yield),col="red",data=d1)
+
  ggarrange(p1,p2,p3,p4, ncol=2, nrow=2)
@@
 
-```{r, projectons-1}
-prj=FLStocks("1"=iter(oms[[1]],1:4),
-             "2"=iter(oms[[2]],1:4),
-             "3"=iter(oms[[3]],1:4),
-             "4"=iter(oms[[4]],1:4))
+```{r, projections-1}
+source("/home/laurie/Desktop/projects/ddmse/R/dd.R")
 
-F=fbar(prj[[1]][,ac(2020:2080)])%=%rep(rfpts$Fmsy,each=61)
+om=fwdWindow(window(iter(om,51),end=2020),end=2080,eqM)
+F =propagate(fbar(om[,ac(2020:2080)]),4)
+F[]=rep(rfpts$Fmsy,each=61)
+control=as(FLQuants("f"=F),"fwdControl")
 
-control=as(FLQuants("f"=F),"fwdControl")
-prj[[1]] =fwd(prj[[1]],control=control,sr=eqM)
+om =fwd(om,control=control,sr=eqM)
 
+prj=FLStocks("Base"     =om)
+prj[["DD Mass"]]        =om
+prj[["DD Mass, Mat"]]   =om
+prj[["DD Mass, Mat, M"]]=om
 for (i in ac(2020:2080)) {
-    prj[[2]] =ddFn(i,prj[[2]],par,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+    prj[[2]] =ddFn(i,prj[[2]],params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
     control=as(FLQuants("f"=F[,i]),"fwdControl")
@@
 ```{r}
-plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4) 
+ts=ldply(prj, function(x) model.frame(FLQuants(x, ssb  =function(x) ssb(  x), 
+                                                  f    =function(x) fbar(x), 
+                                                  catch=function(x) catch(x)),drop=T))
+ts=transform(ts,What=.id)
+ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
+
+#plot(prj,probs=0,worm=seq(4))+facet_grid(qname~stock,scale="free")
+
+ggplot()+
+  geom_line( aes(ssb,catch,col=.id),   data=eqCurves)+
+  geom_point(aes(ssb,catch,col=.id),data=subset(ts,year==2040))
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming no density dependence
+```{r}
+cast(subset(ts,year==2080),iter~What,value="catch")
+```
 
 ```{r}
-names(prj)=names(oms)
-plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)
+ ggarrange(plot(FLStocks(prj[[1]]),probs=c(0), worm=1:4)+ 
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(FLStocks(prj[[2]]),probs=c(0), worm=1:4)+ 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)+ 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)+ 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass-at-age only
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$
 
-```{r}
-plot(FLStocks(prj[[3]]),probs=c(0), worm=1:4)
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass and maturity-at-age only
-
-```{r}
-plot(FLStocks(prj[[4]]),probs=c(0), worm=1:4)
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at $F_{MSY}$ assuming density dependence in mass, maturity and natural mortality-at-age
-
-```{r, eval=FALSE}
-ldply(prj,ldply, function(x) model.frame(FLQuants(x,biomass=function(x) biomass(x)[,"2080"],
-                                                    ssb    =function(x) ssb(    x)[,"2080"],
-                                                    catch  =function(x) catch(  x)[,"2080"],
-                                                    f      =function(x) fbar(   x)[,"2080"]),drop=T)) 
-```
-
-```{r, projectons-2}
-devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2080),,,,1]%=%0,0.3)
-
-base=propagate(iter(prj[[1]],1),100)
+```{r, projections-2}
+base=propagate(iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1),100)
+devRec=rlnorm(100,rec(base)[,ac(2020:2080),,,,1]%=%0,0.3)
 
@@
 
+```{r}
+smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),  
+            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
+
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
+```
+
+```{r}
+ ggarrange(plot(prjs[[1]])+ 
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[2]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[3]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(prjs[[4]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
+```
+
+**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$
+
 ```{r, projectons-save}
@@
 
-```{r}
-plot(FLStocks(prjs[[1]])) 
+```{r, projections-3}
+base=iter(fwdWindow(window(prj[[1]],end=2020),end=2080,eqM),1)
+
+prjs=mlply(data.frame(F=rfpts$Fmsy), function(F) {
+
+    ## FMSY estimate by OM ######################################
+    F=FLQuant(F,dimnames=list(year=2020:2080))
+
+    ## No DD ####################################################
+    control=as(FLQuants("f"=F),"fwdControl")
+    x1     =fwd(iter(base,1),control=control,sr=eqM)
+
+    ## With DD ####################################################
+    x2=x1
+    x3=x1
+    x4=x1
+    for (iYr in ac(2020:2080)) {
+        control=as(FLQuants("f"=F[,iYr]),"fwdControl")
+
+          ## DD M ##################################################         
+          x2 =ddFn(iYr,x2,params,massFlag=TRUE,matFlag=FALSE,mFlag=FALSE)
+          x2 =fwd(x2,control=control,sr=eqM)
+
+          ## DD MM##################################################
+          x3 =ddFn(iYr,x3,params,massFlag=TRUE,matFlag=TRUE,mFlag=FALSE)
+          x3 =fwd(x3,control=control,sr=eqM)
+
+          ## DD MMM ################################################
+          x4 =ddFn(iYr,x4,params,massFlag=TRUE,matFlag=TRUE,mFlag=TRUE)
+          x4 =fwd(x4,control=control,sr=eqM)
+          }
+
+    FLStocks(list("Base"=x1,"M"=x2,"MM"=x3,"MMM"=x4))        
+    }) ## F loop
 ```
 
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, at Base $F_{MSY}$
+```{r}
+smry=mdply(names(prjs), function(x) mdply(names(prjs[[x]]),  
+            function(y) cbind(A=x,B=y,model.frame(FLQuants(prjs[[x]][[y]][,"2040"], F=fbar,SSB=ssb,Catch=catch),drop=T))))
+
+cast(ddply(smry,.(A,B), with, data.frame(F=median(F),SSB=median(SSB),Catch=median(Catch),F=median(F))),A~B,value="Catch")
+```
 
 ```{r}
-plot(FLStocks(prjs[[2]])) 
+ ggarrange(plot(prjs[[1]]) +
+               theme(axis.ticks.y = element_blank(),
+                     plot.margin = margin(r = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[2]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(r = 1, l = 1),
+                     strip.text.y = element_text(size=0)), 
+          plot(prjs[[3]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1),
+                     strip.text.y = element_text(size=0)),
+          plot(prjs[[4]]) + 
+               theme(axis.text.y = element_blank(),
+                     axis.ticks.y = element_blank(),
+                     axis.title.y = element_blank(),
+                     plot.margin = margin(l = 1)  ),
+          nrow = 1, common.legend = TRUE, widths=c(1.4,1,1,1)) 
 ```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models, with  $F_{MSY}$ for density dependence in mass-at-age
-
-```{r}
-plot(FLStocks(prjs[[3]])) 
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity-at-age
-
-```{r}
-plot(FLStocks(prjs[[4]])) 
-```
-
-**Figure `r iFig=iFig+1; iFig`** Projections for Operating Models,  with  $F_{MSY}$ for density dependence in mass, maturity, natural mortality-at-age
 

### 2023-03-13 Rmd/backUp/omWhb.Rmd
moving files to backup

--- 2023-02-13 Rmd/omWhb.Rmd
@@
 
+```{r}
+save(vpaM,vpaDDM,vpaDDMM,vpaDDMMM,par,file=file.path("/home/laurie/Desktop/projects/ddmse/data/om","omWhb.RData"))
+```
+
 # Funding

### 2023-03-13 Rmd/report mac.Rmd
updating reports

@@
 
-&nbsp;
-
 ```{r, echo=FALSE, out.width = "300px", fig.align="left", message=FALSE, warning=FALSE, cache=FALSE}
@@
 Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**.
-**Figures 20 and 21** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
+**Figures 20-22** summarizes the yields, SSBs and Fs obtained for each Operating Model under the correct and mis-specified form of density dependence.
 
-Conclusions
-
-Estimating density-dependence from observations only possible for mass-at-age and maturity age.
+Estimating density-dependence from observations was only possible for mass-at-age and maturity age.
 Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
 
-The impact of including density dependence: major impact on reference points.
+Including density dependence has a major impact on reference points.
 Especially $F_{MSY}$ is much higher when including DD.
@@
 
-**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield and **Figure 22** the Kobe pltos for F and yield.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield and **Figure 22** the Kobe plots for F and yield.
 

### 2023-03-13 Rmd/report whb.Rmd
updating reports

@@
 ---
+title: "Evaluating impacts of Density Dependent processes on Blue whiting via MSE simulations"
+author:
+- name: L.T. Kell
+  affiliation: Sea++
+- name: M.A. Pastoors
+  affiliation: MPFF
 output:
-  html_document:
+  word_document:
     reference_docx: ../report_template_v1.1.dotx
@@
 
-Version: 1.0 (`r format(Sys.time(), '%d %B %Y')`)
+Version 1.0 (`r format(Sys.time(), '%d %B %Y')`)
+
+&nbsp;
 
@@
 Denmark
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
+
+&nbsp;
 
@@
 Stochastic projections of the four Operating Models for the four estimates of $F_{MSY}$ are shown in **Figures 16-19**.
-**Figures 20 and 21** summarizes the yields and SSBs obtained for each Operating Model under the correct and mis-specified form of density dependence.
+**Figures 20-22** summarizes the yields, SSBs and Fs obtained for each Operating Model under the correct and mis-specified form of density dependence.
 
-Conclusions
+Estimating density-dependence from observations was not really possible for `r mystkname`.
+Therefore, density dependence in mass at age, maturity at age and natural mortality has only be explored from theoretical considerations in this report.
 
-Estimating density-dependence from observations only possible for mass-at-age and maturity age.
-Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).
-
-The impact of including density dependence: major impact on reference points.
+Including density dependence has a major impact on reference points.
 Especially $F_{MSY}$ is much higher when including DD.
@@
 A Generalized Additive Mixed Model (GAMM) was estimated to assess the relationship between mass-at-age and total biomass.
-
-weights are predicted from two variables: a smoother of total biomass/mean(total biomass) and age.
+Weights are predicted from two variables: a smoother of total biomass/mean(total biomass) and age.
 The resulting factor is shown below.
@@
 observed - mean values) is plotted against total biomass, this allows the relationship between growth and total biomass to be compared across all ages.
-It is seen that there is no clear relationship between biomass and weight at age.
+Here, we plotted both the GAMM model and the Linear Model variants.
+It is seen that there is still some weak information on density dependence in the Linear Model (slope = beta = -0.11).
 A summary of the linear model is in the table below.
-
-**???The trend of growth with biomass is -0.18, this was then used to model mass-at-age as:**
 
@@
 
-**Deterministic projections at** $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age
+**Deterministic projections at $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age**
 
@@
 The density dependent effect of maturity is strongest when both stock weight and maturity are density dependent.
-When natural mortality is also density dependent, this affects the stock size and hence also the density dependence for stock weight and maturity, leading to a smaller density dependent effects for those variables.
+When natural mortality is also density dependent, this has only minor impacts.
+
+**@Laurie: why would DD in natural mortality have such a minimal impact? Did you see that the M levels overall are lower than the assumed M in the assessment (=0.2)**
 
@@
 
-**Stochastic projections at** $F_{MSY}$ for different scenarios
+**Stochastic projections at $F_{MSY}$ for different scenarios**
 
@@
 SSB and total biomass are highest in the case of no density dependence.
+Biomass tends to be lower when density dependent processes are taken into account.
 Catches are different in the first years of 'switching' to the new reference points, but afterwards, the catches for the different scenarios are highly comparable.
@@
 In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colors.
-For example a valid combination is the Base case with the red curves (Fmsy=0.33) or the M scenario with the green curve.
-Overall the misspecification does not results in noticeable differences in catch although they do have difference for biomass.
+For example a valid combination is the Base case with the red curves (Fmsy=0.27) or the M scenario with the green curve.
+Overall the misspecification does not results in noticeable differences in catch although they do have important difference for biomass.
 
@@
 The results confirm the results that equilibrium catch is not really affected by the DD assumptions of the Fmsy applied, but that biomass is sensitive to the assumptions.
+Application of density dependent reference points when density dependency would not exist, is expected to lead to lower equilibrium biomass levels.
 
@@
 *`r fig_nums("kobe")`*
+
+\newpage
+
+We then plotted the Kobe phase plot for 2050 with the associated density functions for F/Fmsy and Catch/MSY.
+Because there is no uncertainty included in the fishing mortality during the simulations, the marginal distributions of F do not exist in this plot.
+
+```{r, echo=FALSE, out.width = "300px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+fig_nums(
+  name    = "kobe2", level = 1, display = FALSE,
+  caption = "F Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario). Topleft: base, topright: M, bottomleft: MM, bottomright: MMM")
+
+knitr::include_graphics(path=
+                    c(file.path(figuresdir,paste("05",mystk,"kobe FY base.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe FY M.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe FY MM.jpg",sep="_")),
+                      file.path(figuresdir,paste("05",mystk,"kobe FY MMM.jpg",sep="_"))
+                      )
+                    )
+
+```
+
+*`r fig_nums("kobe2")`*
+
+\newpage
 
@@
 
-**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, and **Figure 21** shows phase plots of the final 2050 distributions of biomass and yield.
-THese are for the Base Case, then density-dependence in mass, mass and maturity, and mass, maturity and natural mortality, projected for the four estimates of $F_{MSY}$.
+**Figure 20** summarizes the yields and SSBs obtained for each Operating Model under the correct and misspecified form of density dependence and the associated reference points, **Figure 21** shows the Kobe plots of the final 2050 distributions of biomass and yield and **Figure 22** the Kobe plots for F and yield.
+These are for the Base Case, then density-dependence in mass, mass and maturity, and mass, maturity and natural mortality, projected for the four estimates of $F_{MSY}$.
 The correctly specified $F_{MSY}$ for each model is indicated by the red points and marginal densities.
 
-The main diofferences are seen in $SSB/B_{MSY}$, if no density dependence is assumed then in the operating model the estimate of $F_{MSY}$ is the lowest, therefore projreting for the higher estimates of $F_{MSY}$ assuming density dependence results in an overfished stock.
-Yileld are the same for the two values of $F_{MSY}$ that assume no density-dependence on density dependence in mass-at-age only.
-In the other two cases yield is reduced but only by about 10%.
+The main differences are seen in $SSB/B_{MSY}$, if no density dependence is assumed then in the operating model the estimate of $F_{MSY}$ is the lowest, therefore projecting for the higher estimates of $F_{MSY}$ assuming density dependence results in an overfished stock.
+Yield are the same for the two values of $F_{MSY}$ that assume no density-dependence on density dependence in mass-at-age only.
+In the other two cases yield is reduced by about 10%.
 
@@
 
-Since the early 1980s, natural mortality ($M$) for `r mystkname` has been fixed within the assessment model at 0.2, for all ages and all assessment years.
-
-M must, however, vary both by age and over time due to changes in abundance of predators, feeding conditions, disease etc.
+Since the early 1980s, natural mortality ($M$) for `r mystkname` has been fixed within the assessment model at 0.2, for all ages and all assessment years.Natural mortality must, however, vary both by age and over time due to changes in abundance of predators, feeding conditions, disease etc.
 [@siegfried2009review].

### 2023-03-13 Rmd/report.R

@@ -79,8 +79,8 @@ myparams <- data.frame(
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
-mystk     <- "mac";  
-# mystk     <- "whb";  
+# mystk     <- "mac";  
+mystk     <- "whb";  
 # for (mystk in c("mac","whb")) {
   
   mystkname  <- myparams[myparams$stock==mystk,"mystkname"]
@@ -264,7 +264,7 @@ mystk     <- "mac";
   # run gammV  
   gmr=mgcViz::gammV(log(data) ~ s(tb/mean(tb), bs="tp") + age, 
                     data=wt2)
-  
+
   # plot gammV
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "gamm.jpg", sep="_")),
        width=10, height=8, units="in", res=300)
@@ -284,7 +284,7 @@ mystk     <- "mac";
   # plot(exp(coefficients(gmr)[2:nn]))  # NOT NEEDED?
   
   # connect prediction to wt dataset  
-  wt2=cbind(wt2,hat=predict(gmr))
+  wt2=cbind(wt2,hat =predict(gmr))
   wt2=merge(wt2,data.frame(age=1:nn,
                            mean=exp(c(0,coefficients(gmr)[2:nn]))))
   
@@ -292,6 +292,7 @@ mystk     <- "mac";
   p <-
     ggplot(wt2)+
     geom_line(aes(tb,exp(hat)/mean,col=age))+
+    geom_smooth(aes(tb,exp(hat)/mean),method="lm", colour="blue", linewidth=0.5, se=FALSE)+
     geom_point(aes(tb,data/mean,col=age))+
     xlab("Total biomass")+ylab("Mass-Mean") 
   
@@ -1714,6 +1715,10 @@ mystk     <- "mac";
   sch=merge(sch,msy,by=".id") 
   sch=merge(sch,fmsy,by=".id") 
   
+  if (mystk == "mac") {
+    xlim = 2; ylim=2} else {
+    xlim = 3; ylim=2} 
+  
   # Base
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "kobe base.jpg", sep="_")),
        width=10, height=10, units="in", res=300)
@@ -1723,7 +1728,8 @@ mystk     <- "mac";
                                         run=ac(f))),
                        xlab=expression(B/B[MSY]),
                        ylab=expression(Catch/MSY),
[…]

### 2023-03-15 Rmd/report whb.Rmd
Updated WHB + report; Second GAMM

@@
 
-&nbsp;
+**Laurie: why would DD in natural mortality have such a minimal impact? Did you see that the M levels overall are lower than the assumed M in the assessment (=0.2)**
 
@@
 
-Estimating density-dependence from observations was not really possible for `r mystkname`.
-Therefore, density dependence in mass at age, maturity at age and natural mortality has only be explored from theoretical considerations in this report.
+Estimating density-dependence from observations was not really possible for `r mystkname` because there was no clear relationship between mass-at-age and total biomass.
+When a second GAMM model was fitted with weight in relation to total biomass, year and age, we found that the mass-at-age is better described by a year effect than a biomass effect.
+In order to still explore the impacts of density dependence on the performance of the MSE, we used the density dependent factor for Atlantic mackerel (beta=-0.18) to parameterize the weight-biomass relationship.
 
@@
 
-**GAMM-at-age**
+**GAMM-at-age: biomass only**
 
@@
 
-**Linear model of density dependence in mass-at-age**
+**GAMM-at-age: biomass and year smoothers**
 
-The change in relative mass-at-age (i.e.
-observed - mean values) is plotted against total biomass, this allows the relationship between growth and total biomass to be compared across all ages.
-Here, we plotted both the GAMM model and the Linear Model variants.
-It is seen that there is still some weak information on density dependence in the Linear Model (slope = beta = -0.11).
-A summary of the linear model is in the table below.
+Because no clear relationship between mass-at-age and total biomass was established, we fitted a second GAMM model with weight in relation to total biomass, year and age.
+This demonstrated that the mass-at-age is better described by the year effect than the biomass effect (lower p-value).
+
+In order to still explore the impacts of density dependence on the performance of the MSE, we used the density dependent factor for Atlantic mackerel (beta=-0.18) to parameterize the following relationship:
 
@@
 fig_nums(
-  name    = "cond_lmwaa", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". LM analysis of density dependence and mass-at-age."))
+  name    = "cond_gammwaa2", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". GAMM analysis of mass-at-age against biomass, year and age"))
 
-knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "lm_waa.jpg", sep="_")))
+knitr::include_graphics(path=file.path(figuresdir, paste(section,mystk, "gamm2.jpg", sep="_")))
 
@@
 
-*`r fig_nums("cond_lmwaa")`*
+*`r fig_nums("cond_gammwaa2")`*
 
@@
 tab_nums(
-  name    = "cond_lmsummary", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". LM summary of density dependence and mass-at-age."))
+  name    = "cond_gammsummary2", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". GAMM summary of mass-at-age against biomass, year and age"))
 
-cat(readLines(file.path(tablesdir, paste(section,mystk, "lm summary.txt", sep="_"))), sep = '\n')
+cat(readLines(file.path(tablesdir, paste(section,mystk, "gamm2 summary.txt", sep="_"))), sep = '\n')
 
@@
 
-*`r tab_nums("cond_lmsummary")`*
+*`r tab_nums("cond_gammsummary2")`*
 
@@
 
-**@Laurie: why would DD in natural mortality have such a minimal impact? Did you see that the M levels overall are lower than the assumed M in the assessment (=0.2)**
-
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
@@
 
-Estimating density-dependence from observations was only possible for mass-at-age but not maturity-at-age as the latter were assumed by ICES not to have varied over time.
-However, assuming that maturity is a function of mass-ata-ge allowed density deoendence to be modeeled.
-Natural mortality was assumed to be time and age invariant by ICES (i.e.
-0.2), therefore density dependence in natural mortality can only be explored from theoretical considerations.
+Estimating density-dependence from observations was not really possible for `r mystkname` because there was no clear relationship between mass-at-age and total biomass.
+When a second GAMM model was fitted with weight in relation to total biomass, year and age, we found that the mass-at-age is better described by a year effect than a biomass effect.
+In order to still explore the impacts of density dependence on the performance of the MSE, we used the density dependent factor for Atlantic mackerel (beta=-0.18) to parameterize the weight-biomass relationship.
 

### 2023-03-15 Rmd/report.R

@@ -65,8 +65,8 @@ myparams <- data.frame(
   btrig   = c(2580000                      , 2250000          ),
   bmsy    = c(3500000                      , 4000000          ), # guesstimates
   alpha   = c(-0.2                         , -0.2             ),
-  delta   = c(-0.2                         , -0.2             ),
-  matk    = c(0.2                           , 0.1              ), # source: conditioning
+  delta   = c(-0.2                         , -0.1             ),  # delta = beta !!!! For DD weight
+  matk    = c(0.2                           , 0.2              ), # source: conditioning
   wt1     = c(0.01                         , NA               ), # for resetting weight at age 1
   minage  = c(0                            , 1                ),
   maxage  = c(12                           , 10               ),
@@ -258,17 +258,18 @@ mystk     <- "whb";
   # GAMM model and diagnostics of weight vs Total biomass -----------------------
   
   # drop age 0 if existing
-  wt2 = subset(wt,age!=0)
+  wt2 = subset(wt,age!=0) %>% mutate(age = stringr::str_pad(age, width=2, pad="0"))
   nn  = length(unique(wt2$age))
   
-  # run gammV  
+  # run gammV with weight against biomass and age
   gmr=mgcViz::gammV(log(data) ~ s(tb/mean(tb), bs="tp") + age, 
                     data=wt2)
 
-  # plot gammV
+  # plot gammV - only with age
   jpeg(filename=file.path(figuresdir, paste(section,mystk, "gamm.jpg", sep="_")),
-       width=10, height=8, units="in", res=300)
-  plot(gmr) + labs(x="Total biomass", y="Weight deviation")
+       width=10, height=6, units="in", res=300)
+  # plot(gmr, allTerms=TRUE) + labs(x="Total biomass", y="Weight deviation")
+  print(plot(gmr) + labs(title = NULL), pages = 1)
   dev.off()  
   
   # summary table
@@ -285,7 +286,7 @@ mystk     <- "whb";
   
   # connect prediction to wt dataset  
   wt2=cbind(wt2,hat =predict(gmr))
-  wt2=merge(wt2,data.frame(age=1:nn,
+  wt2=merge(wt2,data.frame(age=stringr::str_pad(1:nn, width=2, pad="0"),
                            mean=exp(c(0,coefficients(gmr)[2:nn]))))
   
   # Linear   
[…]

### 2023-03-15 Rmd/report.R
kobe with curves

@@ -1753,7 +1753,8 @@ mystk     <- "whb";
                        xlab=expression(B/B[MSY]),
                        ylab=expression(Catch/MSY),
                        col=c("red","grey","grey","grey"),
-                       xlim=xlim, ylim=ylim) 
+                       xlim=xlim, ylim=ylim,
+                       layer=geom_line(aes(ssb/bmsy,catch/msy),data=subset(rfs,.id=="Base")))  
   dev.off()
 
   # M
@@ -1766,7 +1767,8 @@ mystk     <- "whb";
                        xlab=expression(B/B[MSY]),
                        ylab=expression(Catch/MSY),
                        col=c("grey","red","grey","grey"),
-                       xlim=xlim, ylim=ylim) 
+                       xlim=xlim, ylim=ylim,
+                       layer=geom_line(aes(ssb/bmsy,catch/msy),data=subset(rfs,.id=="DD Mass")))  
   dev.off()
 
   # MM
@@ -1776,7 +1778,8 @@ mystk     <- "whb";
                        xlab=expression(B/B[MSY]),
                        ylab=expression(Catch/MSY),
                        col=c("grey","grey","red","grey"),
-                       xlim=xlim, ylim=ylim) 
+                       xlim=xlim, ylim=ylim,
+                       layer=geom_line(aes(ssb/bmsy,catch/msy),data=subset(rfs,.id=="DD Mass, Mat")))  
   dev.off()
 
   # MMM
@@ -1786,7 +1789,8 @@ mystk     <- "whb";
                        xlab=expression(B/B[MSY]),
                        ylab=expression(Catch/MSY),
                        col=c("grey","grey","grey","red"),
-                       xlim=xlim, ylim=ylim) 
+                       xlim=xlim, ylim=ylim,
+                       layer=geom_line(aes(ssb/bmsy,catch/msy),data=subset(rfs,.id=="DD Mass, Mat, M")))  
   dev.off()
   
   ## Yield & F #################################################################
@@ -1806,7 +1810,8 @@ mystk     <- "whb";
                 xlab=expression(F/F[MSY]),
                 ylab=expression(Catch/MSY),
                 col=c("red","grey","grey","grey"),
-                xlim=xlim, ylim=ylim) 
+                xlim=xlim, ylim=ylim,
[…]

### 2023-03-15 Rmd/report.R
kobe with curves

@@ -1739,6 +1739,8 @@ mystk     <- "whb";
   sch=merge(sch,msy,by=".id") 
   sch=merge(sch,fmsy,by=".id") 
   
+  rfs=merge(merge(merge(eqCurves,fmsy),bmsy),msy)
+  
   if (mystk == "mac") {
     xlim = 2; ylim=2} else {
     xlim = 3; ylim=2} 

### 2023-03-17 Rmd/report mac.Rmd
Fixed error in figure 15

@@
 
-Version 1.3 (`r format(Sys.time(), '%d %B %Y')`)
+Version 1.3 (13 March 2023)
+
+Version 1.4 (16 March 2023; fixed error in reference point table)
+
+Version 1.5 (17 March 2023; fixed error in figure 15)
 
@@
 Denmark
-
-&nbsp;
-
-&nbsp;
 
@@
 
-**Comparison of equilibrium curves for density dependence analyses: biomass against yield and F**
+<!-- =============================================================================================================== -->
 
-The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 30%, $F_{MSY}$ is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
+<!-- Section 5: Running forward projections                                                                          -->
+
+<!-- =============================================================================================================== -->
+
+# Forward projections without feedback
+
+Forward projections were run in two modes.
+Firstly in deterministic mode and secondly in probabilistic mode.
+We used the $F_{MSY}$ to project forward from 2020 until 2050.
+The reference points estimated for the different scenarios are in the table below.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+section <- "05"
+
+tab_nums(
+  name    = "refpoints", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Reference points for the different scenarios"))
+
+cat(readLines(file.path(tablesdir, paste(section,mystk, "refpoints.txt", sep="_"))), sep = '\n')
+
+```
+
+*`r tab_nums("refpoints")`*
+
+The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 35%, $F_{MSY}$ is expected to increase by factor of 2.5 and catch (MSY) is expected to increase by 9%.
 So while, inclusion of density-dependent processes, has a substantial impact on biological reference points, the resulting change in equilibrium catch is expected to be small.
@@
 
-**Reference points**
-
-Reference points estimated for the different scenarios are in the table below.
-
-```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-tab_nums(
-  name    = "refpoints", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Reference points for the different scenarios"))
-
-cat(readLines(file.path(tablesdir, paste(section,mystk, "refpoints.txt", sep="_"))), sep = '\n')
-
-```
-
-*`r tab_nums("refpoints")`*
-
-<!-- =============================================================================================================== -->
-
-<!-- Section 5: Running forward projections                                                                          -->
-
-<!-- =============================================================================================================== -->
-
-# Forward projections without feedback
-
-Forward projections were run in two modes.
-Firstly in deterministic mode and secondly in probabilistic mode.
-We used the $F_{MSY}$ to project forward from 2020 until 2050.
 We used both the $F_{MSY}$ that is applicable to the density dependent scenario, but also evaluated what would happen if the 'wrong $F_{MSY}$' would be used with a particular scenario.
@@
 ```{r, echo=FALSE, fig.asp=1.0, fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-section <- "05"
 

### 2023-03-17 Rmd/report whb.Rmd
Fixed error in figure 15

@@
 
-Version 1.0 (`r format(Sys.time(), '%d %B %Y')`)
+Version 1.0 (15 March 2023)
+
+Version 1.1 (16 March 2023, fixed error in reference point table)
+
+Version 1.2 (17 March 2023; fixed error in figure 15)
 
@@
 &nbsp;
-
-**Laurie: why would DD in natural mortality have such a minimal impact? Did you see that the M levels overall are lower than the assumed M in the assessment (=0.2)**
 
@@
 Initial evaluations were conducted based on life history theory, as used by the International Council for the Exploration of the Sea (ICES) to develop advice conducting MSE for data poor stocks [@fischer2020ga].
-**Figure 10** shows the the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
+**Figure 10** shows the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
 These allow the relationships between the MSY reference points and virgin biomass to be seen.
@@
 
-The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticty in the recruitment deviates.
+The OMs were also projected for the estimates of different estimates $F_{MSY}$, with stochasticity in the recruitment deviates.
 This allows the impact of misspecification of density dependence to be evaluated, i.e.
@@
 
-Four alternatives OMs have been included in the analyses:
+Four alternative OMs have been included in the analyses:
 
@@
 
-All MSE's have been developed within the FLR framework (<https://flr-project.org/>) and all the code is stored on github (<https://github.com/martinpastoors/ddmse>).
+All MSE's have been developed within the FLR framework (<https://flr-project.org/>) and all the code is stored on GitHub (<https://github.com/martinpastoors/ddmse>).
 
@@
 
-Mass-at-age ($W_a$) is modeled as
+Mass-at-age ($W_a$) is modelled as
 
@@
 
-Maturity is then modeled as a logistic function of mass-at-age i.e.
+Maturity is then modelled as a logistic function of mass-at-age i.e.
 
@@
 
+<!-- =============================================================================================================== -->
+
+<!-- Section 5: Running forward projections                                                                          -->
+
+<!-- =============================================================================================================== -->
+
+# Forward projections without feedback
+
+Forward projections were run in two modes.
+Firstly in deterministic mode and secondly in probabilistic mode.
+We used the $F_{MSY}$ to project forward from 2020 until 2050.
+Reference points estimated for the different scenarios are in the table below.
+
+```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
+
+section <- "05"
+
+tab_nums(
+  name    = "refpoints", level = 1, display = FALSE,
+  caption = paste0(mystkname, ". Reference points for the different scenarios"))
+
+cat(readLines(file.path(tablesdir, paste(section,mystk, "refpoints.txt", sep="_"))), sep = '\n')
+
+```
+
+*`r tab_nums("refpoints")`*
+
 \newpage
 
-**Comparison of equilibrium curves for density dependence analyses: biomass against yield and F**
-
-The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 30%, $F_{MSY}$ is expected to increase by 113% and catch (MSY) is expected to increase by 5%.
-So while, inclusion of density-dependent processes, has a substantial impact on biological reference points, the resulting change in equilibrium catch is expected to be small.
+The comparison of the equilibrium curves, indicate that B_MSY is expected to decrease by 45%, $F_{MSY}$ is expected to increase by a factor of 5 and catch (MSY) is expected to increase by 8%.
+So, while inclusion of density-dependent processes has a substantial impact on biological reference points, the resulting change in equilibrium catch is expected to be small.
 
@@
 
-\newpage
-
-**Reference points**
-
-Reference points estimated for the different scenarios are in the table below.
-
-```{r, echo=FALSE, out.width = "400px", fig.align="center", message=FALSE, warning=FALSE, cache=FALSE}
-
-tab_nums(
-  name    = "refpoints", level = 1, display = FALSE,
-  caption = paste0(mystkname, ". Reference points for the different scenarios"))
-
-cat(readLines(file.path(tablesdir, paste(section,mystk, "refpoints.txt", sep="_"))), sep = '\n')
-
-```
-
-*`r tab_nums("refpoints")`*
-
-<!-- =============================================================================================================== -->
-
-<!-- Section 5: Running forward projections                                                                          -->
-
-<!-- =============================================================================================================== -->
-
-# Forward projections without feedback
-
-Forward projections were run in two modes.
-Firstly in deterministic mode and secondly in probabilistic mode.
-We used the $F_{MSY}$ to project forward from 2020 until 2050.
 We used both the $F_{MSY}$ that is applicable to the density dependent scenario, but also evaluated what would happen if the 'wrong $F_{MSY}$' would be used with a particular scenario.
@@
 using an Fmsy from one scenario and applying it to a different scenario has only limited impacts on the overall outcomes in terms of recruitment, biomass or yield.
-In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colors.
-For example a valid combination is the Base case with the red curves (Fmsy=0.27) or the M scenario with the green curve.
-Overall the misspecification does not results in noticeable differences in catch although they do have important difference for biomass.
+In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colours.
+For example, a valid combination is the Base case with the red curves (Fmsy=0.27) or the M scenario with the green curve.
+Overall, the misspecification does not results in noticeable differences in catch although they do have important difference for biomass.
 
@@
 
-Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colors).
+Finally, as a check, we assessed the outcomes of the equilibrium curves from simulating over a wide F range, with the equilibrium values (points) in 2050 from the forward projections at different values of Fmsy (colours).
 The points fall on the equilibrium curve, thereby confirming the consistency between the two approaches.
@@
 Initial explorations of density dependent effect were conducted based on life history theory, as used by the International Council for the Exploration of the Sea (ICES) to develop advice conducting MSE for data poor stocks [@fischer2020ga].
-**Figure 10** shows the the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
+**Figure 10** shows the equilibrium curves of catch v SSB assuming a) no density dependence, and density dependence in b) mass-at-age, c) mass & maturity-at-age, and d) mass, maturity and natural mortality-at-age.
 These allow the relationships between the MSY reference points and virgin biomass to be seen.
@@
 The figure below compares the SAM assessment with a Virtual Population Analysis (VPA) and a VPA with age varying M (VPAM).
-Trends and absolute levels swere similar across all runs.
+Trends and absolute levels were similar across all runs.
 The vpaM model was used as the base case to be compared against the density-dependent simulations.
@@
 
-The OMs were also projected for the four estimates of  $F_{MSY}$, with stochasticty in the recruitment deviates.
+The OMs were also projected for the four estimates of  $F_{MSY}$, with stochasticity in the recruitment deviates.
 This allows the impact of misspecification of density dependence to be evaluated, i.e.
@@
 The quality of this analyses lives up to the quality standards that are required by ICES and so can potentially be used when providing advice to fisheries managers.
-The software is publicly available via Github (<https://github.com/martinpastoors/ddmse>).
+The software is publicly available via GitHub (<https://github.com/martinpastoors/ddmse>).
 All data is derived from WGWIDE 2022 @ICES2022WGWIDE.

### 2023-03-17 Rmd/report.R

@@ -65,14 +65,14 @@ myparams <- data.frame(
   btrig   = c(2580000                      , 2250000          ),
   bmsy    = c(3500000                      , 4000000          ), # guesstimates
   alpha   = c(-0.2                         , -0.2             ),
-  delta   = c(-0.2                         , -0.1             ),  # delta = beta !!!! For DD weight
+  delta   = c(-0.2                         , -0.2             ),  # delta = beta !!!! For DD weight
   matk    = c(0.2                           , 0.2              ), # source: conditioning
   wt1     = c(0.01                         , NA               ), # for resetting weight at age 1
   minage  = c(0                            , 1                ),
   maxage  = c(12                           , 10               ),
   minyear = c(1991                         , 2000             ),
   maxyear = c(2050                         , 2050             ),
-  m1scaler= c(8                            , 10               ),
+  m1scaler= c(8                            , 10                ), #10
   w50     = c(0.166                        , 0.08             ),
   matk2   = c(28                           , 63               ),
   steepness=c(0.8                          , 0.5               ))
@@ -92,7 +92,11 @@ mystk     <- "whb";
   
   figuresdir <- file.path(dropboxdir, "results", mystk, "figures")
   tablesdir  <- file.path(dropboxdir, "results", mystk, "tables")
-
+  
+  # empty directories
+  file.remove(dir(figuresdir, full.names = TRUE))
+  file.remove(dir(tablesdir, full.names = TRUE))
+  
   # load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
   
   
@@ -635,53 +639,6 @@ mystk     <- "whb";
   print(p) 
   dev.off()
   
-  # Check 1 --------------------------------------------------------------------
-
-  # ptm <- proc.time()
-  # 
-  # # generate FLquants with different F values
-  # f = FLQuant(c(rep(seq(0, 1,length.out=51)    ,each=101),
[…]

### 2023-03-17 Rmd/report mac.Rmd
Added extra scenarios

@@
 load(file=file.path(dropboxdir,paste0("data/om/",mystk,"matPar.RData")))
+load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
+load(file = file.path(dropboxdir, "results", mystk, paste(mystk,"section5.RData", sep="_")))
 
@@
 
-Version 1.5 (17 March 2023; fixed error in figure 15)
+Version 1.5 (17 March 2023; fixed error in figure 15; added two scenarios)
 
@@
 
+For the stochastic projections at all $F_{MSY}$'s, we added two F scenarios: one with F=0.1 and one with F=`r myparams[myparams$stock==mystk,"highf"]`.
+
 Misspecification of Fmsy, i.e.
 using an Fmsy from one scenario and applying it to a different scenario has only limited impacts on the overall outcomes in terms of recruitment, biomass or yield.
-In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colors.
-For example a valid combination is the Base case with the red curves (Fmsy=0.33) or the M scenario with the green curve.
-Overall the misspecification does not results in noticeable differences in catch although they do have difference for biomass.
+In the plot below, the scenarios for density dependence are shown in the columns and the F's are shown as the colours.
+For example, a valid combination is the Base case with the yellow-brownish curve (Fmsy=`r format(rfpts[1,"Fmsy"], digits=2)`) or the M (mass) scenario with the green curve.
+The misspecification of Fmsy does not results in noticeable differences in catch although they do have important difference for biomass.
+Application of a low F (=0.1) is expected to lead to lower catches in all scenarios.
 
@@
 
-For the four scenarios, we plotted the Kobe phase plot for 2050 with the associated density functions for B/Bmsy and Catch/MSY.
+For the four Fmsy-scenarios and the two additional F scenarios, we plotted the Kobe phase plot for 2050 with the associated density functions for B/Bmsy and Catch/MSY.
 The results confirm the results that equilibrium catch is not really affected by the DD assumptions of the Fmsy applied, but that biomass is sensitive to the assumptions.
@@
   name    = "kobe", level = 1, display = FALSE,
-  caption = "Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario). Topleft: base, topright: M, bottomleft: MM, bottomright: MMM")
+  caption = "Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario).
+Topleft: base, topright: M, bottomleft: MM, bottomright: MMM.
+White densities refer to the low and high F scenarios")
 
@@
   name    = "kobe2", level = 1, display = FALSE,
-  caption = "F Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario). Topleft: base, topright: M, bottomleft: MM, bottomright: MMM")
+  caption = "F Kobe plot for 2050 for base case (no DD) for all $F_{MSY}$'s (red is from the right scenario).
+Topleft: base, topright: M, bottomleft: MM, bottomright: MMM.
+White densities refer to the low and high F scenarios")
 

### 2023-03-17 Rmd/report whb.Rmd
Added extra scenarios

@@
 
-Version 1.2 (17 March 2023; fixed error in figure 15)
+Version 1.2 (17 March 2023; fixed error in figure 15; added two scenarios)
 
@@
 
-**Stochastic projections at all** $F_{MSY}$'s from all scenarios
+**Stochastic projections at all $F_{MSY}$'s from all scenarios**
+
+For the stochastic projections at all $F_{MSY}$'s, we added two F scenarios: one with F=0.1 and one with F=`r myparams[myparams$stock==mystk,"highf"]`.
 
@@
 using an Fmsy from one scenario and applying it to a different scenario has only limited impacts on the overall outcomes in terms of recruitment, biomass or yield.
-In the plot below, the scenarios for density dependence are shown in the columns and the Fmsy's are shown as the colours.
-For example, a valid combination is the Base case with the red curves (Fmsy=0.27) or the M scenario with the green curve.
-Overall, the misspecification does not results in noticeable differences in catch although they do have important difference for biomass.
+In the plot below, the scenarios for density dependence are shown in the columns and the F's are shown as the colours.
+For example, a valid combination is the Base case with the yellow-brownish curve (Fmsy=`r format(rfpts[1,"Fmsy"], digits=2)`) or the M (mass) scenario with the green curve.
+The misspecification of Fmsy does not results in noticeable differences in catch although they do have important difference for biomass.
+Application of a low F (=0.1) is expected to lead to lower catches in all scenarios.
 

### 2023-03-17 Rmd/report.R

@@ -75,7 +75,10 @@ myparams <- data.frame(
   m1scaler= c(8                            , 10                ), #10
   w50     = c(0.166                        , 0.08             ),
   matk2   = c(28                           , 63               ),
-  steepness=c(0.8                          , 0.5               ))
+  steepness=c(0.8                          , 0.5              ),
+  lowf     =c(0.1                          , 0.1              ),
+  highf    =c(1.5                          , 2.0              )
+)
 
 # save(myparams, file=file.path(dropboxdir, "data", "inputs", "myparams.RData"))
 
@@ -121,7 +124,7 @@ mystk     <- "whb";
     ts$tb=ts$tb/1e6
   }
   
-  save(.,
+  save(list=ls(),
        file = file.path(dropboxdir, "results", mystk, paste(mystk,"section0.RData", sep="_")))
   
   # ------------------------------------------------------------------------------
@@ -1460,17 +1463,17 @@ mystk     <- "whb";
   ts=transform(ts,.id=unique(eqCurves$.id)[an(iter)])
   
 
-  ggplot()+
-    geom_line(data=eqCurves,
-              aes(ssb,catch,col=.id))+
-    geom_point(data=subset(ts,year==2050),
-               aes(ssb,catch,col=.id))
+  # ggplot()+
+  #   geom_line(data=eqCurves,
+  #             aes(ssb,catch,col=.id))+
+  #   geom_point(data=subset(ts,year==2050),
+  #              aes(ssb,catch,col=.id))
 
-  ggplot()+
-    geom_line(data=subset(ts,year==2050),
-              aes(ssb,catch,col=.id))+
-    geom_point(data=subset(prj_df,year==2050),
-               aes(ssb,catch,col=.id))
+  # ggplot()+
+  #   geom_line(data=subset(ts,year==2050),
+  #             aes(ssb,catch,col=.id))+
+  #   geom_point(data=subset(prj_df,year==2050),
+  #              aes(ssb,catch,col=.id))
   
   
   
@@ -1481,8 +1484,11 @@ mystk     <- "whb";
   devRec=rlnorm(100,rec(prj[[1]])[,ac(2020:2050),,,,1]%=%0,0.3)
   base=fwdWindow(window(propagate(iter(prj[[1]],1),100),end=2020),end=2050,vpaM_eq)
   
-  prjs=mlply(data.frame(F=rfpts$Fmsy), 
[…]
